### 2015-09-03 tex/paper.tex
Added initial commit

\documentclass[a4paper,12pt]{article}
\usepackage{fullpage}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{authblk}
\usepackage{todonotes}
\usepackage{caption}
\usepackage{url}
\usepackage{float}
\usepackage{sectsty}
\usepackage{pdfpages}
\usepackage[section]{placeins}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}

\usepackage{setspace}
\usepackage[toc,page]{appendix}
\usepackage{framed}
\usepackage{geometry}

\usepackage{alltt}
\usepackage{subfig}

% Change section fonts
\allsectionsfont{\sffamily}

% For code box
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{%
  \parbox{\textwidth}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}\vskip-4pt}}
  \captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
  \lstset{frame=lrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}
% End code box

\usepackage{cite}

% General parameters, for ALL pages:
\renewcommand{\topfraction}{0.9}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
% Parameters for TEXT pages (not float pages):
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4} % 2 may work better
\setcounter{dbltopnumber}{2} % for 2-column pages

\addtolength{\topmargin}{0.5in}

\usepackage{fancyvrb}

\usepackage{tikz} \usetikzlibrary{trees}
\usepackage{hyperref} % should always be the last package

% useful colours (use sparingly!):
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\green}[1]{{\color{green}#1}}
\newcommand{\red}[1]{{\color{red}#1}}

% useful wrappers for algorithmic/Python notation:
\newcommand{\length}[1]{\text{len}(#1)}
\newcommand{\twodots}{\mathinner{\ldotp\ldotp}} % taken from clrscode3e.sty
\newcommand{\Oh}[1]{\mathcal{O}\left(#1\right)}

% useful (wrappers for) math symbols:
\newcommand{\Cardinality}[1]{\left\lvert#1\right\rvert}
\newcommand{\Ceiling}[1]{\left\lceil#1\right\rceil}
\newcommand{\Floor}[1]{\left\lfloor#1\right\rfloor}
\newcommand{\Iff}{\Leftrightarrow}
\newcommand{\Implies}{\Rightarrow}
\newcommand{\Intersect}{\cap}
\newcommand{\Sequence}[1]{\left[#1\right]}
\newcommand{\Set}[1]{\left\{#1\right\}}
\newcommand{\SetComp}[2]{\Set{#1\SuchThat#2}}
\newcommand{\SuchThat}{\mid}
\newcommand{\Tuple}[1]{\langle#1\rangle}
\newcommand{\Union}{\cup}
\usetikzlibrary{positioning,shapes,shadows,arrows}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords: }} #1}

\title{\textbf{NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection}}
\author{Lukas Klingsbo}

\begin{document}

%\maketitle
\includepdf[pages={1}]{front.pdf}
\thispagestyle{empty}
\newpage\null\thispagestyle{empty}\newpage

\pagenumbering{roman}
\setcounter{page}{2}

\includepdf[pages={1}]{abstract.pdf}
%\begin{abstract}
%Collecting and analysing data of analytical value is important for many service providers today. 
%Many make use of NoSQL databases for their larger software systems, what is less known is how to 
%effectively analyse and gather business intelligence from the data in these systems. 
%This paper suggests a method of separating the most valuable analytical data from the rest in 
%real time and at the same time providing an effective traditional database for the analyser. 
%
%In this paper we analyse our given data sets to decide whether big data tools are required and 
%then traditional databases are compared to see how well they fit the context. 
%A technique that makes use of an asynchronous logging system is used to insert the data from the 
%main system to the dedicated analytical database. 
%
%The tests show that our technique can efficiently be used with a traditional database even on large 
%data sets (>1000000 insertions/hour per database node) and still provide both historical data 
%and aggregate functions for the analyser.\\
%
%\keywords{NoSQL, data mining, data collection, DBMS benchmark, business intelligence, analytical logging}
%\end{abstract}

\newpage\null\thispagestyle{empty}\newpage

\setcounter{tocdepth}{3}
\tableofcontents

\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}
\section{Introduction to the\\"Big Data/NoSQL problem"}
In times when big data collection is growing rapidly~\cite{E1} companies and researchers alike face the problem of doing analysis of the data that has been collected.
The enormous amounts of data collected are often saved in distributed big data databases or key-value stores with fast insertion operations but with comparatively slow and complex (but highly parallelizable) query functions~\cite{RT1}.
This work tries to solve such a problem at Kivra~\cite{KIVRA}, a company working to digitalise and minimise the use and need of conventional paper mail. To gather analytical data from Kivra's distributed database today requires heavy Map/Reduce 
queries. To do day to day analysis of the data slowly becomes unbearable and non-scalable as it takes several hours or even days to run a Map/Reduced~\cite{G1} batch job to fetch the data that is going to be analysed.
The solution proposed and examined by this paper is to structure and asynchronously collect data after interesting events occur, that are of a high analytical value, into a smaller and more easily queryable database.

\newpage
\section{Related Terminology}

\subsection{Big data}
Big data~\cite{BD1} is a loose term for ways of collecting and storing unstructured complex data of such large volumes that normal methods of data manipulation and querying can not be used.

\subsection{NoSQL}
NoSQL~\cite{NOSQL} is a term that have not fully settled its meaning, it started by meaning no relational databases~\cite{NOSQL}, but now it has moved on to commonly be referred to as Not Only SQL. 
If a database calls itself NoSQL it usually means that is is non-relational, distributed, horizontally scalable, schema-free, eventually consistent and support large amounts of data~\cite{NOSQL2}.

\subsection{MapReduce}
MapReduce~\cite{G1} is referring to a programming model that makes parallel processing of large data sets possible. In simplified terms it consists of two functions, one called Map and one called Reduce.
The Map function is sent out to each node from a master node and the function is separately executed on each nodes data set and then each node assembles the result back together with the other nodes so that all 
input values with the same results are grouped together. Then the reduce function is executed, either in parallel or on a master node, and this function does calculations upon each separate result list with the data from all nodes that 
ran the Map function and it then combines the input from them to the answer that was originally inquired.

\subsection{Eventual consistency}
The expression eventual consistency is a term used for describing how data is processed in database systems. Eventual consistency means that the data may not be consistent in any one moment, 
but will eventually converge to being consistent~\cite{EVENTUAL}. This makes it possible for distributed databases to operate faster as data does not have to be locked during insertions, updates 
and deletion and thus does not have to be propagated to all nodes before more actions on the same data can be sent to a node. If one or more nodes operate on the same data as another node 
whose changes have not propagated through the system yet, the system will have to handle in which order the operations should be performed, this is usually called sibling or conflict resolution~\cite{RESOLUTION}.

\subsection{Riak}
Riak is a key-value datastore that implements NoSQL concepts such as eventual consistency. It is usually configured to run with at least 5 nodes to be as efficient as possible.
To query the datastore you either use the REST-ful API or MapReduce functionality that it provides.

Riak~\cite{RIAK} provides fault tolerance and high availability by replicating data to several nodes, the default is to have all data replicated to three nodes.

\subsection{Lager}
Basho~\cite{BASHO}, the company that developed Riak~\cite{RIAK}, has developed a modular logging framework called Lager~\cite{LAGER}. Lager is modular in the sense that it is possible to write separate backends 
which all can separately do asynchronous data collection. As the system performance can not be noticeably affected by adding a subsystem to collect analytically valued events it is of highest priority 
that they can be sent asynchronously so that the implication of implementing such a subsystem wont result in delays for the main system.

A problem that these form of logging systems can experience is that the events that they have to handle queue up faster than they can be processed. To solve this Lager~\cite{LAGER} has a built-in overload protection 
to make sure that its message queue will not overflow. It does this by swapping synchronous and asynchronous messaging for the separate backend.

\subsection{Abbreviations}

\subsubsection{RDBMS}
RDBMS~\cite{RDBMS} is an abbreviation for relational database management system. This is the most commonly used database management system and is usually referred to as a traditional database. Data is stored within tables and data can be in relation to each 
other across tables and such databases are therefore called relational.

\subsubsection{DDBMS}
DDBMS~\cite{DDBMS} is an abbreviation for distributed database management system. It is used to centrally manage a distributed database system and to give the user the impression and functionality as though the database system 
would run on only one node.

\newpage
\section{Background}
\subsection{About Kivra}
Kivra~\cite{KIVRA} is a company which aims to digitalise and eliminate the use of superfluous analogue paper mail. 
They do this by making it possible for companies and governments to send documents digitally to customers or citizens. 
By tying the users' accounts to their social security numbers in the service the sender can be sure that the document 
reaches the correct recipient.

Today Kivra has a continuously growing user base of more than 500.000 users, according to their website~\cite{KIVRA}.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.4]{kivraui}}
    \caption{The user interface of Kivra}
    \label{fig:ui}
\end{figure}

As can be seen in figure~\ref{fig:ui}, the interface reminds of the interface of any email provider, however the documents sent to the users can be interactive and have a more advanced layout and the main advantage over normal email is that every 
users identity is established in the registration process. 
The user can validate its identity either by using mobile BankID~\cite{BANKID} or by receiving an activation code to its registered place of residence. 
This advantage makes it possible for companies and organizations that otherwise would have had to send documents to a registered postal address to being able to send them digitally through Kivra.

\subsection{The current system}
Kivra's users interact with the backend system (this is the backend box represented in Figure~\ref{fig:before}) either through Kivra's web page or their app, where they can receive documents, pay bills, set whom they accept documents from and also set personal details. 
Users sign up either though mobile bankID or requests an activation code to be sent to the address registered to their social security number, this is to ensure the identity of the user. 
When a user receives a new document it is notified through email or SMS and can then log in to read the document.

The data which is of analytical interest is information about the users, meta data of the documents sent out to the users and whether a user accepts or does not accept (internally called \emph{send requests}) documents from a company or organization (internally called a \emph{tenants}).

\newpage
\subsection{Problem description}
The part of the current system that is being examined consists of Riak, a distributed key-value datastore, and a batch script containing MapReduce queries which are running each night to collect 
data to a PostgreSQL database which is later used for deeper analysis. This system architecture can be seen in Figure~\ref{fig:before}. 
The batch script is emptying all of the analytically interesting buckets in the Riak cluster,
currently there are two buckets being emptied every night. One of them contains send requests which is meta data telling the system if a user accepts documents for a specific tenant and the other bucket
just contains plain user information. In the future meta data regarding documents sent to users will be collected as well.

This solution does not scale and is also inefficient as you have to go through all the data in the growing buckets that are of interest every night, instead of accumulating the data to a separate database dedicated for analysis 
at the same time as you insert or change data in Riak.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{system_before}}
    \caption{The analytical system at Kivra}
    \label{fig:before}
\end{figure}

~\\
\newpage
\section{Methods for determining\\implementation details}
This chapter introduces the different methods used to determine how the new system should be implemented, which DBMS it should use and how the estimation of long term scaling was done.

\subsection{Data size and growth estimation}
The data size and growth was estimated with statistics collected from the current Riak cluster in Kivra's system. What the current insertion rate would be like with an event driven implementation was estimated 
by doing SQL queries measuring the number of insertions that could be made within certain time limits. 
To estimate the size and growth of the future database implementation both past estimations and the current internal estimated growth of the company were used. 
This, together with the scalability measurements, was required to make sure that the solution would be durable in the long run.

\subsection{Scalability measurement}
The scalability was estimated for the typical server hardware of Kivra with the different databases and with statistics of growth estimations from the current database. 
This was necessary to make sure that the system would scale with the expected data growth, within a reasonable time frame.

\subsection{DBMS comparison}
The database management systems were compared by testing required abilities and examining load and speed by running insertion jobs of different sizes. 
The DBMSs architecture and suitability were also compared to see which one suited the context best. 
These comparisons were vital to make sure that the chosen DBMS could handle the expected load of the system.

\newpage
\section{Solution for storing events}
When an event of analytical interest occurs, an asynchronous call with information about the event is sent to Lager. Lager then directs it to an in-house tailored Lager backend, which is responsible for 
making sure that the data is stored consistently and correctly. The database in which the data is stored can then be queried at any time to perform near to real time analysis of the system. 
After analysing internal database statistics an approach of this form is estimated to cause approximately 100-10000 inserts/hour depending on if it is during the peak hours or not. The estimation can not be averaged as 
the chosen database is required to be able to handle the peak hours without a problem. This approach will lead to a system with no updates and only reads during analytical calculations. This is due to the fact that 
the backend will handle each event with a strict progression of inserting the relevant metadata of the event with a timestamp and never have a need of updating, nor removing old data, as all of these 
chosen events are valuable to store to enable historical queries. This is similar to how Lager~\cite{LAGER} logs events to files, except that the analysis of the data will be easier due to the 
built-in aggregate functions of modern databases. In comparison to the current system it will be possible to perform near to instant analysis of the data already collected instead of having to re-collect all 
the data of analytical interest from the Riak cluster each time an up-to-date analysis of the data has to be performed.

\subsection{Data sets}
\subsubsection{Determining the type of the data sets}
Big data is one of the popular buzzwords in the computer industry today and a lot of revolutionising technology has been built around it. But it is not always efficient to use technology built for 
these data sets if the data set that is being operated on does not fit the definition of big data. But as there is a thriving hype of higher yield with existing data some companies and institutions use it 
just for the sake of being part of the trend without actually having a data set that objectively would be classified as big data~\cite{NOTBIG}.

To simplify the abstract definition of Big Data to determine whether a data set can be defined as such you can classify the data set by concluding the answers of 
three questions~\cite{BD1}, namely the following:
\begin{itemize}
    \item Is the data set large?
    \item Is the data set complex (maybe even unstructured)?
    \item Are methods optimized for other data sets inefficient?
\end{itemize}

If the data set satisfies these properties it can be a good idea to store the data in a Hadoop cluster or similar technology developed specifically for big data. 
If, however, the data set does not satisfy these criteria it is usually better to look for another storage solution~\cite{NOTBIG}.

At Kivra there are a few different data sets with information that is valuable for ongoing analysis. 
The three most important data sets are: 
\begin{itemize}
    \item User information - each users personal details.
    \item Send requests - connects a user to a tenant and specifies whether the user wants or does not want content from that specific tenant.
    \item Content metadata - contains meta data of content that has been sent to users. 
\end{itemize}
~\\
The tables representing this data would be quite simple and would not store too much information as analysis of the system would mostly regard metadata of some of the information stored in the Riak cluster.
See an example of how such tables could look like in listing~\ref{example_tables}.

\begin{lstlisting}[label=example_tables,caption=Example tables]
Example fields of the tables
+-------------+  +--------------+  +------------------+
| User        |  | Send request |  | Content Metadata |
+-------------+  +--------------+  +------------------+
| id          |  | user_id      |  | content_id       |
| name        |  | tenant_id    |  | tenant_id        |
| address     |  | accepted     |  | user_id          |
| city        |  | timestamp    |  | timestamp        |
| email       |  |              |  |                  |
| timestamp   |  |              |  |                  |
+-------------+  +--------------+  +------------------+
\end{lstlisting}

\subsubsection{Size of the data sets}
\label{DATASETSIZE}
As each record in the data sets is fairly small (under 200 bytes) it is enough to count the number of records plus the expected growth for the future. Today there are about 500 000 users with information~\cite{KIVRA}. 
The send requests records adds up to a little bit less than the number of users times the number of tenants and then there is also around 2 million content records. This results in a total of about 10 million records. 
All of the data sets have an expected growth of about 5\% per month according to internal statistics of Kivra collected from their Riak cluster. This is well within the limits of traditional relational databases~\cite{SQLPERF}.

\subsubsection{Complexity of the data sets}
None of the data sets have any complex structures, the data types are not varying and there are no big documents that requires to be stored in the analytical database as only the meta data of the content being 
sent to users has to be saved and not any attachments, images etc.

\subsubsection{Inefficiency of other methods}
There is no need for extremely fast reads for specific values which key-value stores~\cite{KEYVALUESTORE} offer or for extremely fast insertion rates which big data databases~\cite{BD1} can provide. 
As concluded in the section about the size of the data sets (~\ref{DATASETSIZE}) and the section called Choosing SQL database (~\ref{SQLDATABASES}) the insertion rate of 
traditional databases will be well above the demands. As there is no complex data the normal table structure of a RDBMS will be suitable for storage of the data.

\newpage
\subsection{Protocols for data transfer}
The requirements of the protocol and data structure used to transfer and translate data to the database differs highly on what type of data and from where the data is being sent, some data are required to have 
absolute transport security and reliability meanwhile other is of lower importance and rather trades transport reliability for speed. To maximize the speed and flexibility of the system as a whole, 
one or several APIs and web services that can accept several protocols and differently structured data being used towards it before translating it in order to store it in the database should be used, as can be seen in 
Figure~\ref{fig:api}.

The exact APIs, Web Services and data structures used for the implementation of this system is outside the scope for this work and will be left to a future implementer to decide. But examples of such protocols 
could be HTTP through a RESTful~\cite{REST} service, cURL~\cite{CURL} with one of its many supported protocols or simply whatever protocol that will be built in to the customised Lager~\cite{LAGER} backend.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{system_api}}
    \caption{API explanation}
    \label{fig:api}
\end{figure}

\subsection{Database}
\subsubsection{Requirements}
\label{REQUIREMENTS}
As the database will mostly be used for insertions, the insert operations will have to be as fast as possible with the chosen way to store the data. 
Read operations will not have to be very fast as they are not directly used by any response time critical system but rather by in-house users, a business intelligence system or a system providing an interface of generated 
graphs and tables for overviewing the information.

The requirements on the speed of updates and deletes are non essential as they will only occur when manually editing the database. 
The database will not have to be portable, it is not going to be replicated or moved around. 
For future scaling of the database it is needed to be easily configurable and preferably have wide support for distribution of the data over several nodes. 
The data is required to be accessible and easy to aggregate upon for future analysis.

\subsubsection{Why SQL over NoSQL?}
A traditional relational database system was chosen over a NoSQL document store or key-value store for a few different reasons.
The first reason was that the high insertion rate trade off that results in eventual consistency~\cite{EVENTUALCONS} given by NoSQL databases would simply not be needed as the amount of data insertions and updates would 
be well within the limits of a traditional RDBMS~\cite{SQLPERF} as can be seen in Section~\ref{SCALABILITYRESULT}.

The second reason was that the query speed of specific values, offered by key-value stores~\cite{KEYVALUESTORE}, was not required and the demand was rather for a full set of aggregate 
functions and easy access to complete subsets of the stored data. 
As many NoSQL databases use MapReduce for fetching data it would per definition be hard to write efficient joins and no built in aggregate functions would be provided~\cite{COMP}.

What was also considered was putting a layer on top of a regular NoSQL database that seamlessly provided this type of functionality. 
The layers considered were HIVE~\cite{HIVE}, PIG~\cite{PIG} and Scalding~\cite{SCALDING} for HADOOP~\cite{HADOOP}. But as the implementation did not need the special functionalities and speed trade-offs, 
which is deeper explained in Section~\ref{SCALABILITYRESULT} and Section~\ref{REQUIREMENTS}, that were offered by such layers the conclusion was to simply incorporate a traditional SQL database.

\subsubsection{Choosing SQL database}
\label{SQLDATABASES}
There are a large number of different traditional database systems being actively developed today. For this project the focus has been on popular open source DBMSs. 
The most popular open source relational database management systems~\cite{POPULARDBMS} today are MySQL, PostgreSQL and SQLite so the tests compare those towards each other and the requirements.

\newpage
\paragraph{MySQL}
~\\
MySQL\cite{MYSQL} is the most popular open source database and has several popular forks as well, with the biggest one~\cite{POPULARDBMS} called MariaDB~\cite{MARIADB}. MySQL does not aim at implementing 
the full SQL standard, but rather a subset of it with some extra features~\cite{POPULARDBMS}. MySQL is built in layers with one or several storage engines in the bottom and an optimizer, parser, 
query cache and connection handling layers built on top of that~\cite{HIMYSQL}, see Figure~\ref{fig:mysql}.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{mysql}}
    \caption{MySQL Architecture~\cite{HIMYSQL}}
    \label{fig:mysql}
\end{figure}
~\\

MySQL has a rich set of aggregate functions which are very useful when doing statistical analysis of the stored data. In total, MySQL has 17 aggregate functions in which the most common SQL standard 
aggregate functions AVG, COUNT, MAX, MIN, STD and SUM are included~\cite{MYSQLAGG}.

\FloatBarrier
\newpage
\paragraph{PostgreSQL}
~\\
PostgreSQL is one of the few modern SQL databases that implements the whole SQL standard (ANSI-SQL:2008) ~\cite{POSTGRESSTANDARD}. 
The architecture of PostgreSQL is quite straight forward, as can be seen in Figure~\ref{fig:postgres}. The postmaster handles new connections and spawns a PostgreSQL backend process that the client can directly communicate with. All communication 
between client and PostgreSQL goes through that process after it spawns. The postmaster stays active and awaits new incoming connections. Each backend that a client is connected to has its own separate 
parser, rewrite system, optimiser and executor. All of the active backends then uses the same storage engine~\cite{POSTGRES}. PostgreSQL, compared to MySQL, only has one very tightly integrated 
storage engine that in theory could be replaced but that has not been done in practice yet~\cite{POSTGRESQLPAPER}.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{postgres}}
    \caption{PostgreSQL Architecture~\cite{POSTGRES}}
    \label{fig:postgres}
\end{figure}
~\\

PostgreSQL have a comparatively large set of aggregate functions including 14 general purpose functions and 18 functions mainly used for statistics, totalling in 32 functions~\cite{POSTGRESAGG}.

\newgeometry{top=2.0cm}
\paragraph{SQLite}
~\\
SQLite as opposed to MySQL and PostgreSQL is not built as a traditional two tier architecture which can be seen in Figure~\ref{fig:sqlite}.
This leads to that each call to it is done through function calls to a linked SQLite library~\cite{SQLITEDEV}. SQLite is quite small (658KiB) and has therefore become widely used as an embedded database 
for local storage~\cite{SQLITE}. SQLite is bundled with two storage engines, one of them is built with a log-structured merge-tree and optimised for persistent 
databases and the other one is built on a binary tree which is kept in memory which makes it suitable for temporary databases. The architecture of SQLite has made it possible to design and interchange 
new storage engines with the built-in ones~\cite{SQLITEENGINE}.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{sqlite}}
    \caption{SQLite Architecture~\cite{SQLITE}}
    \label{fig:sqlite}
\end{figure}
~\\

SQLite has 9 of the most basic aggregate functions and lacks the more statistical functionality such as for example standard deviation and covariance that PostgreSQL and MySQL implement~\cite{SQLITEAGG}. 
SQLite's architecture is very focused on portability~\cite{SQLITE} and have made some trade-offs in order to achieve that, as can be seen in its simplistic architecture in Figure~\ref{fig:sqlite}. 
One result of these trade-offs is that insertions become very slow, $\sim60$ insertions per minute with transactions enabled~\cite{SQLITESLOW}.

As portability is not required for the implementation of the system regarded in this paper, because it is missing many important aggregate functions and because of its slow insertion rate, 
SQLite will be excluded from the tests in the next section.

\newpage
\subsubsection{Performance tests}
As explained in section~\ref{SQLDATABASES}, SQLite was excluded from the tests as its architecture was deemed unfitting for the task at hand and for the lack of many aggregate functions.
The statistics were generated on an average consumer computer with the following specifications:

\vspace{0.25in}
\textbf{CPU:} Intel Core2 Duo T7700 2.40GHz

\textbf{RAM:} SODIMM DDR2 Synchronous 667 MHz 2x2GB

\textbf{Disk:} KINGSTON SV300S3 120GB SSD partitioned with EXT4

\textbf{OS:} Debian Jessie (Testing)

~\\
PostgreSQL version 9.4 and MySQL version 5.6 were used during the tests and the packages were installed from the default Debian testing repositories.

~\\
The database and table were created in the same manner on both systems as can be seen in listing~\ref{database_creation}.
\begin{lstlisting}[label=database_creation,caption=Database and table creation]
create database testdb
create table test_table 
    (data varchar(100), 
    moddate timestamp default current_timestamp);
\end{lstlisting}

The tests were conducted with a series of simple insertions containing random data, as shown in listing~\ref{test_table}.
\begin{lstlisting}[label=test_table,caption=Example insertion]
insert into test_table (data) 
values ('010100100101010100...');
\end{lstlisting}

The execution times were measured by using the time~\cite{TIME} utility of the GNU project, it measures the time of execution and resource usage of the specified operation. 
The different files that were used in the tests contained 1000, 10 000, 100 000 and \\1 000 000 insertions of random data. 
When deploying this system the production servers will most likely be a lot more powerful than the consumer laptop that these test were run on.
These tests non the less gives a correct representation of how well the different databases will scale. 
To reproduce or verify the results, see listings 3 and 4 in Appendix~\ref{appendix:a}.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{linegraph2}}
    \caption{Performance graph}
    \label{fig:vs}
\end{figure}

As can be seen in Figure~\ref{fig:vs} both of the systems have a very close to linear performance and any small disturbances in the pattern can be concluded to be from other processes using system resources.
Both of the systems perform over the preconditions for an implementation with requirements similar to the ones in this context. But as insertion rate and aggregate functions are of high concern for such 
an implementation and both systems comply to the rest of the requirements, PostgreSQL will be recommended to use for a future implementation.

\newpage
\subsection{Storage requirements}
As the data comes in as events, there is never a need to update old records in the database, as can be seen in Figure~\ref{fig:storage}. This structure is slightly harder to do complicated 
queries (Listing~\ref{fetch_example}) towards, as you always have to take the timestamp of the event into consideration. 
But it also makes it possible to configure the DBMS not to lock tables during inserts, which will increase the speed of insertion of the database. The lack of locks will be especially useful if it is later 
decided to use a DDBMS for an eventual requirement of scaling up the system in the future. As everything is logged with timestamps it is particularly suitable for business intelligence systems and the likes of such, as you can query 
time ranges in the past as well as the current state of the system to gather analytics and make comparisons.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.6]{system_dbms}}
    \caption{Flow of data}
    \label{fig:storage}
\end{figure}

The only time inserts occur is during the time when the customised Lager~\cite{LAGER} backend or later added data sources sends events (1, see Figure~\ref{fig:storage}) and the only times reads are performed 
is when a manual query (2, see Figure~\ref{fig:storage}) is executed or when an eventual business intelligence system or an automated analyse system (3, see Figure~\ref{fig:storage}) that automatically queries the system is used. 
It will never be necessary to perform updates and no regular deletes will have to be performed, as explained in Section~\ref{REQUIREMENTS}.

The actual database size should not be a problem as the servers are usually equipped with a few terabytes of disk space. An average insertion of Kivra is estimated to around 200 bytes by looking at meta data formed by data currently in the Riak cluster.
This means that $5*10{^9}$ insertions would fit on a disk with 1TB of space. A rough estimation of an average of 5000 insertions per hour means that a 1TB would last for more than 113 years, see Figure~\ref{fig:storage2}.

\begin{figure}[h!]
\begin{framed}
\begin{enumerate}
\item $1TB/200B \thickapprox 5*10^{9}$ insertions
\item $5*10^{9}/5000/24/365 \thickapprox $ 113.5 years
\end{enumerate}
\end{framed}
\caption{Estimation of disk space usage}
\label{fig:storage2}
\end{figure}

\FloatBarrier
\newgeometry{top=2.0cm}
\newpage
\section{Resulting system}
The result of an implementation of the suggested solution of this paper would lead to a core system that would have a data flow as described in Figure~\ref{fig:result}.
In this flow, when data of analytical interest would be inserted into the Riak cluster it would also be sent to Lager which would insert it into a RDBMS dedicated for 
analytical data, instead of fetching data in Riak after it has been inserted as was it done in the previous systems data flow that can be seen in Figure~\ref{fig:before} and explained in Section 3.2.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{system_after}}
    \caption{Resulting system}
    \label{fig:result}
\end{figure}

What is not shown in Figure~\ref{fig:result}, is that separate data sources for the analytical database can be used. These can connect with different protocols and with differently 
structured data to an API that will be receiving data for the database as described in Section 5.2.

\subsection{Scalability}
\label{SCALABILITYRESULT}
If an unoptimised PostgreSQL database on an average consumer computer is used and with the assumption that little to no data transfer latency will occur during insertion of data the insertion rate can 
be increased by a factor of at least 100 if following the upper boundary of the estimation (Section 5) of 10000 insertions per hour. As can be seen in the load test in 5.3.4, PostgreSQL can handle a little bit over 
1 million insertions per hour. The expected data growth rate is 5\% per month, with this growth the same setup could be used for almost 8 years, see Figure~\ref{fig:scalability}.

\begin{figure}[h!]
\begin{framed}
\begin{enumerate}
\item $10000*1.05^{94} \thickapprox $ 981282 insertions/hour 
\item $94/12 \thickapprox $ 7.83 years 
\end{enumerate}
\end{framed}
\caption{Estimation of data transfer scaling}
\label{fig:scalability}
\end{figure}

In the unlikelihood that the same setup with the same hardware and PostgreSQL version would be used in 8 years the PostgreSQL database could easily be sharded, clustered or optimised to reach a lot higher 
insertion rates. As all he data is independent and has no defined relations to other data, clustering with a pool of independent databases could be used without any requirement of the databases to synchronise with each 
other, this would mean that you can linearly add 1000000 insertions/hour for each database node that you add to the cluster.

\subsection{Practical Example}
\subsubsection{Storing an Event}
After an implementation of a solution like this at Kivra the following would be the flow of a typical change triggered by a user. 
The system would update the send request bucket of Riak in Figure~\ref{fig:practical}, but changes to other buckets would look exactly the same.

Figure~\ref{fig:practical} is a practical example of what would happen if a user updated his or her preferences of which tenants it wishes to receive documents from.

\begin{figure}[h!]
    \vspace{0.25in}
    \centerline{\includegraphics[scale=0.7]{practical_flow}}
    \caption{Practical example of a send request changing}
    \label{fig:practical}
\end{figure}

\newpage
\subsubsection{Querying Data}
To query data stored in the analytical database a query that takes the timestamps into consideration has to be performed. The query in listing~\ref{fetch_example} is an example of how to query which 
tenants all users wish and do not wish to receive documents from. Queries for other data sets would look similar.

\begin{lstlisting}[label=fetch_example,caption=Example query]
SELECT SRQ.user_id, SRQ.tenant_id, SRQ.timestamp, SRQ.accepted 
FROM sendrequest SRQ 
INNER JOIN 
  (SELECT user_id, tenant_id, max(timestamp) as latest
   FROM sendrequest GROUP BY user_id, tenant_id) SRQ1 
ON SRQ.user_id = SRQ1.user_id 
AND SRQ.tenant_id = SRQ1.tenant_id 
AND SRQ.timestamp = SRQ1.latest;
\end{lstlisting}

The query in listing~\ref{fetch_example} doesn't take into consideration if the same name and tenant combination exists with the same timestamp. 
It is however highly unlikely that it would happen as the timestamps are stored with a micro second precision. 
To fully avoid that event, add an auto incremented ID column to the table and only consider the highest ID in the case of a clash.

\newpage
\section{Discussion}
\subsection{Event-driven data collection}
The suggested solution introduced by this paper was named event-driven data collection. The name has been used before for slightly similar purposes with the first recorded case 
being in 1974~\cite{EVENTDRIVEN}. It has also been used in a different manner in quite a few papers on the subject wireless sensor networks~\cite{EVENTDRIVEN2} and some regarding the collection of natural 
ionosphere scintillations~\cite{EVENTDRIVEN3}.

\subsection{Flat file database}
A flat file database was not compared in this paper because of the precondition of having pre-built APIs and to give easier access to querying and aggregation of data. A flat file database can also be complicated to scale out 
to several servers for an eventual future requirement.

\newpage
\newgeometry{top=4.0cm}
\section{Summary}
\subsection{Conclusions}
The system that has been designed in this paper can be used as an effective tool for logging and analysing events. This form of approach is especially valuable if the main system is using some form of MapReduce 
database which can be hard to do ongoing analysis of aggregated data on. If the main system is already using a SQL database as its primary form of information storage this type of solution would be quite inefficient 
due to duplication of data and possibly having to serve another DBMS\@. In the benchmarking of the different database management systems the standard version without modifications have been used, with modifications 
and optimisations the outcome would have been very different.

The unorthodox design of the database schema in this paper provides a full historical log of events which makes it possible to compare data from different time ranges, 
which can be very valuable when analysing data from business systems to for example detect trends and tendencies within the user base.

Another contribution of this work is the use of a logging system to effectively put in an analytical systems 
to an already existing backend with minimal impact on the performance of the existing system.

\subsection{Future work}
To finalise the architecture of this system, APIs and web service structures will have to be researched or developed to be put on top of the database. For larger flows of data other solutions would have 
to be considered or alternatively the used DBMS has to be heavily optimised by for example removing insertion locks. Scaling out to a clustered database could also be an alternative.

\newpage
\newgeometry{top=2.0cm}
\begin{appendices}
\section{Benchmarks}
\label{appendix:a}
\begin{lstlisting}[label=some-code,caption=PostgreSQL insertion test]
1 000 000 insertions
$ time psql --quiet -f test4.sql
psql --quiet -f test4.sql
40.42s user 23.08s system 2% cpu 31:05.03 total

100 000 insertions
$ time psql --quiet -f test.sql
psql --quiet -f test.sql
4.32s user 2.38s system 2% cpu 5:02.94 total

10 000 insertions
$ time psql --quiet -f test2.sql
psql --quiet -f test2.sql
0.51s user 0.22s system 2% cpu 30.172 total

1000 insertions
$ time psql --quiet -f test3.sql
psql --quiet -f test3.sql 
0.11s user 0.02s system 4% cpu 3.077 total
\end{lstlisting}
\begin{lstlisting}[label=some-code,caption=MySQL insertion test]
1 000 000 insertions
$ time mysql -u root -p********** testdb < test4.sql
mysql -u root -p********** testdb <   
39.78s user 23.36s system 1% cpu 1:36:56.31 total

100 000 insertions
$ time mysql -u root -p********** testdb < test.sql
mysql -u root -p********** testdb < 
3.88s user 2.83s system 1% cpu 9:22.67 total

10 000 insertions
$ time mysql -u root -p********** testdb < test2.sql 
mysql -u root -p********** testdb < 
0.31s user 0.39s system 1% cpu 53.174 total

1000 insertions
$ time mysql -u root -p********** testdb < test3.sql 
mysql -u root -p********** testdb < 
0.08s user 0.00s system 1% cpu 5.318 total
\end{lstlisting}
\end{appendices}
\restoregeometry

\newpage
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}

### 2015-09-16 tex/spec.tex
Added specification

--- 2015-09-03 tex/paper.tex
@@
 
-\title{\textbf{NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection}}
+\title{\textbf{Specification: Providing Access Control in Content Distribution Networks}}
 \author{Lukas Klingsbo}
@@
 
-%\maketitle
-\includepdf[pages={1}]{front.pdf}
-\thispagestyle{empty}
-\newpage\null\thispagestyle{empty}\newpage
-
-\pagenumbering{roman}
-\setcounter{page}{2}
-
-\includepdf[pages={1}]{abstract.pdf}
-%\begin{abstract}
-%Collecting and analysing data of analytical value is important for many service providers today. 
-%Many make use of NoSQL databases for their larger software systems, what is less known is how to 
-%effectively analyse and gather business intelligence from the data in these systems. 
-%This paper suggests a method of separating the most valuable analytical data from the rest in 
-%real time and at the same time providing an effective traditional database for the analyser. 
-%
-%In this paper we analyse our given data sets to decide whether big data tools are required and 
-%then traditional databases are compared to see how well they fit the context. 
-%A technique that makes use of an asynchronous logging system is used to insert the data from the 
-%main system to the dedicated analytical database. 
-%
-%The tests show that our technique can efficiently be used with a traditional database even on large 
-%data sets (>1000000 insertions/hour per database node) and still provide both historical data 
-%and aggregate functions for the analyser.\\
-%
-%\keywords{NoSQL, data mining, data collection, DBMS benchmark, business intelligence, analytical logging}
-%\end{abstract}
-
-\newpage\null\thispagestyle{empty}\newpage
+\maketitle
 
@@
 \setcounter{page}{1}
-\section{Introduction to the\\"Big Data/NoSQL problem"}
-In times when big data collection is growing rapidly~\cite{E1} companies and researchers alike face the problem of doing analysis of the data that has been collected.
-The enormous amounts of data collected are often saved in distributed big data databases or key-value stores with fast insertion operations but with comparatively slow and complex (but highly parallelizable) query functions~\cite{RT1}.
-This work tries to solve such a problem at Kivra~\cite{KIVRA}, a company working to digitalise and minimise the use and need of conventional paper mail.
-To gather analytical data from Kivra's distributed database today requires heavy Map/Reduce
-queries.
-To do day to day analysis of the data slowly becomes unbearable and non-scalable as it takes several hours or even days to run a Map/Reduced~\cite{G1} batch job to fetch the data that is going to be analysed.
-The solution proposed and examined by this paper is to structure and asynchronously collect data after interesting events occur, that are of a high analytical value, into a smaller and more easily queryable database.
+\section{Background}
+This work will be done at the company Uprise (Dragarbrunnsgatan 36C, Uppsala). 
+The goal of this project is to explore, and creating, the possibility to have 
+access control in a content distribution network.
 
-\newpage
-\section{Related Terminology}
-
-\subsection{Big data}
-Big data~\cite{BD1} is a loose term for ways of collecting and storing unstructured complex data of such large volumes that normal methods of data manipulation and querying can not be used.
-
-\subsection{NoSQL}
-NoSQL~\cite{NOSQL} is a term that have not fully settled its meaning, it started by meaning no relational databases~\cite{NOSQL}, but now it has moved on to commonly be referred to as Not Only SQL. 
-If a database calls itself NoSQL it usually means that is is non-relational, distributed, horizontally scalable, schema-free, eventually consistent and support large amounts of data~\cite{NOSQL2}.
-
-\subsection{MapReduce}
-MapReduce~\cite{G1} is referring to a programming model that makes parallel processing of large data sets possible.
-In simplified terms it consists of two functions, one called Map and one called Reduce.
-The Map function is sent out to each node from a master node and the function is separately executed on each nodes data set and then each node assembles the result back together with the other nodes so that all
-input values with the same results are grouped together.
-Then the reduce function is executed, either in parallel or on a master node, and this function does calculations upon each separate result list with the data from all nodes that
-ran the Map function and it then combines the input from them to the answer that was originally inquired.
-
-\subsection{Eventual consistency}
-The expression eventual consistency is a term used for describing how data is processed in database systems. Eventual consistency means that the data may not be consistent in any one moment, 
-but will eventually converge to being consistent~\cite{EVENTUAL}. This makes it possible for distributed databases to operate faster as data does not have to be locked during insertions, updates 
-and deletion and thus does not have to be propagated to all nodes before more actions on the same data can be sent to a node. If one or more nodes operate on the same data as another node 
-whose changes have not propagated through the system yet, the system will have to handle in which order the operations should be performed, this is usually called sibling or conflict resolution~\cite{RESOLUTION}.
-
-\subsection{Riak}
-Riak is a key-value datastore that implements NoSQL concepts such as eventual consistency. It is usually configured to run with at least 5 nodes to be as efficient as possible.
-To query the datastore you either use the REST-ful API or MapReduce functionality that it provides.
-
-Riak~\cite{RIAK} provides fault tolerance and high availability by replicating data to several nodes, the default is to have all data replicated to three nodes.
-
-\subsection{Lager}
-Basho~\cite{BASHO}, the company that developed Riak~\cite{RIAK}, has developed a modular logging framework called Lager~\cite{LAGER}.
-Lager is modular in the sense that it is possible to write separate backends
-which all can separately do asynchronous data collection.
-As the system performance can not be noticeably affected by adding a subsystem to collect analytically valued events it is of highest priority
-that they can be sent asynchronously so that the implication of implementing such a subsystem wont result in delays for the main system.
-
-A problem that these form of logging systems can experience is that the events that they have to handle queue up faster than they can be processed.
-To solve this Lager~\cite{LAGER} has a built-in overload protection
-to make sure that its message queue will not overflow. It does this by swapping synchronous and asynchronous messaging for the separate backend.
-
-\subsection{Abbreviations}
-
-\subsubsection{RDBMS}
-RDBMS~\cite{RDBMS} is an abbreviation for relational database management system.
-This is the most commonly used database management system and is usually referred to as a traditional database.
-Data is stored within tables and data can be in relation to each
-other across tables and such databases are therefore called relational.
-
-\subsubsection{DDBMS}
-DDBMS~\cite{DDBMS} is an abbreviation for distributed database management system.
-It is used to centrally manage a distributed database system and to give the user the impression and functionality as though the database system
-would run on only one node.
-
-\newpage
-\section{Background}
-\subsection{About Kivra}
-Kivra~\cite{KIVRA} is a company which aims to digitalise and eliminate the use of superfluous analogue paper mail. 
-They do this by making it possible for companies and governments to send documents digitally to customers or citizens. 
-By tying the users' accounts to their social security numbers in the service the sender can be sure that the document 
-reaches the correct recipient.
-
-Today Kivra has a continuously growing user base of more than 500.000 users, according to their website~\cite{KIVRA}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.4]{kivraui}}
-    \caption{The user interface of Kivra}
-    \label{fig:ui}
-\end{figure}
-
-As can be seen in figure~\ref{fig:ui}, the interface reminds of the interface of any email provider, however the documents sent to the users can be interactive and have a more advanced layout and the main advantage over normal email is that every
-users identity is established in the registration process. 
-The user can validate its identity either by using mobile BankID~\cite{BANKID} or by receiving an activation code to its registered place of residence. 
-This advantage makes it possible for companies and organizations that otherwise would have had to send documents to a registered postal address to being able to send them digitally through Kivra.
-
-\subsection{The current system}
-Kivra's users interact with the backend system (this is the backend box represented in Figure~\ref{fig:before}) either through Kivra's web page or their app, where they can receive documents, pay bills, set whom they accept documents from and also set personal details.
-Users sign up either though mobile bankID or requests an activation code to be sent to the address registered to their social security number, this is to ensure the identity of the user. 
-When a user receives a new document it is notified through email or SMS and can then log in to read the document.
-
-The data which is of analytical interest is information about the users, meta data of the documents sent out to the users and whether a user accepts or does not accept (internally called \emph{send requests}) documents from a company or organization (internally called a \emph{tenants}).
-
-\newpage
-\subsection{Problem description}
-The part of the current system that is being examined consists of Riak, a distributed key-value datastore, and a batch script containing MapReduce queries which are running each night to collect 
-data to a PostgreSQL database which is later used for deeper analysis. This system architecture can be seen in Figure~\ref{fig:before}. 
-The batch script is emptying all of the analytically interesting buckets in the Riak cluster,
-currently there are two buckets being emptied every night.
-One of them contains send requests which is meta data telling the system if a user accepts documents for a specific tenant and the other bucket
-just contains plain user information. In the future meta data regarding documents sent to users will be collected as well.
-
-This solution does not scale and is also inefficient as you have to go through all the data in the growing buckets that are of interest every night, instead of accumulating the data to a separate database dedicated for analysis
-at the same time as you insert or change data in Riak.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_before}}
-    \caption{The analytical system at Kivra}
-    \label{fig:before}
-\end{figure}
-
-~\\
-\newpage
-\section{Methods for determining\\implementation details}
-This chapter introduces the different methods used to determine how the new system should be implemented, which DBMS it should use and how the estimation of long term scaling was done.
-
-\subsection{Data size and growth estimation}
-The data size and growth was estimated with statistics collected from the current Riak cluster in Kivra's system.
-What the current insertion rate would be like with an event driven implementation was estimated
-by doing SQL queries measuring the number of insertions that could be made within certain time limits. 
-To estimate the size and growth of the future database implementation both past estimations and the current internal estimated growth of the company were used. 
-This, together with the scalability measurements, was required to make sure that the solution would be durable in the long run.
-
-\subsection{Scalability measurement}
-The scalability was estimated for the typical server hardware of Kivra with the different databases and with statistics of growth estimations from the current database. 
-This was necessary to make sure that the system would scale with the expected data growth, within a reasonable time frame.
-
-\subsection{DBMS comparison}
-The database management systems were compared by testing required abilities and examining load and speed by running insertion jobs of different sizes. 
-The DBMSs architecture and suitability were also compared to see which one suited the context best. 
-These comparisons were vital to make sure that the chosen DBMS could handle the expected load of the system.
-
-\newpage
-\section{Solution for storing events}
-When an event of analytical interest occurs, an asynchronous call with information about the event is sent to Lager.
-Lager then directs it to an in-house tailored Lager backend, which is responsible for
-making sure that the data is stored consistently and correctly. The database in which the data is stored can then be queried at any time to perform near to real time analysis of the system. 
-After analysing internal database statistics an approach of this form is estimated to cause approximately 100-10000 inserts/hour depending on if it is during the peak hours or not.
-The estimation can not be averaged as
-the chosen database is required to be able to handle the peak hours without a problem.
-This approach will lead to a system with no updates and only reads during analytical calculations.
-This is due to the fact that
-the backend will handle each event with a strict progression of inserting the relevant metadata of the event with a timestamp and never have a need of updating, nor removing old data, as all of these 
-chosen events are valuable to store to enable historical queries. This is similar to how Lager~\cite{LAGER} logs events to files, except that the analysis of the data will be easier due to the 
-built-in aggregate functions of modern databases.
-In comparison to the current system it will be possible to perform near to instant analysis of the data already collected instead of having to re-collect all
-the data of analytical interest from the Riak cluster each time an up-to-date analysis of the data has to be performed.
-
-\subsection{Data sets}
-\subsubsection{Determining the type of the data sets}
-Big data is one of the popular buzzwords in the computer industry today and a lot of revolutionising technology has been built around it. But it is not always efficient to use technology built for 
-these data sets if the data set that is being operated on does not fit the definition of big data.
-But as there is a thriving hype of higher yield with existing data some companies and institutions use it
-just for the sake of being part of the trend without actually having a data set that objectively would be classified as big data~\cite{NOTBIG}.
-
-To simplify the abstract definition of Big Data to determine whether a data set can be defined as such you can classify the data set by concluding the answers of 
-three questions~\cite{BD1}, namely the following:
+\section{Description}
+The project should research and possibly develop a system where it is possible 
+to set different types of access to different files in a CDN. 
+For example, the access control could be divided into three groups:
 \begin{itemize}
-    \item Is the data set large?
-    \item Is the data set complex (maybe even unstructured)?
-    \item Are methods optimized for other data sets inefficient?
+  \item Very secret - Only accessible from white listed IP addresses
+  \item Secret - Only accessible from authenticated users with the correct access
+  \item Public - Accessible by anybody
 \end{itemize}
 
-If the data set satisfies these properties it can be a good idea to store the data in a Hadoop cluster or similar technology developed specifically for big data. 
-If, however, the data set does not satisfy these criteria it is usually better to look for another storage solution~\cite{NOTBIG}.
+When a file is set to a belong to one of these groups it should not be static, 
+it should be possible to, at a later stage, assign the file to another group.
 
-At Kivra there are a few different data sets with information that is valuable for ongoing analysis. 
-The three most important data sets are: 
+\section{Methods}
 \begin{itemize}
-    \item User information - each users personal details.
-    \item Send requests - connects a user to a tenant and specifies whether the user wants or does not want content from that specific tenant.
-    \item Content metadata - contains meta data of content that has been sent to users. 
+  \item The prototype/system will be developed in Scala. 
+  \item A literature study will be done once the specification is accepted. 
+  \item Results will be evaluated from a security and scalability point of view. 
+  \item A work station and desk space will be provided by Uprise.
 \end{itemize}
-~\\
-The tables representing this data would be quite simple and would not store too much information as analysis of the system would mostly regard metadata of some of the information stored in the Riak cluster.
-See an example of how such tables could look like in listing~\ref{example_tables}.
 
-\begin{lstlisting}[label=example_tables,caption=Example tables]
-Example fields of the tables
-+-------------+  +--------------+  +------------------+
-| User        |  | Send request |  | Content Metadata |
-+-------------+  +--------------+  +------------------+
-| id          |  | user_id      |  | content_id       |
-| name        |  | tenant_id    |  | tenant_id        |
-| address     |  | accepted     |  | user_id          |
-| city        |  | timestamp    |  | timestamp        |
-| email       |  |              |  |                  |
-| timestamp   |  |              |  |                  |
-+-------------+  +--------------+  +------------------+
-\end{lstlisting}
+\section{Relevant Courses}
+\begin{itemize}
+  \item Computer Networks I (1DT052)
+  \item Computer Networks II (1DT074)
+  \item Distributed Systems (1DT064)
+  \item Secure Computer Systems I (1DT072)
+  \item Secure Computer Systems II (1DT073)
+  \item Process Oriented Programming (1DT083)
+\end{itemize}
 
-\subsubsection{Size of the data sets}
-\label{DATASETSIZE}
-As each record in the data sets is fairly small (under 200 bytes) it is enough to count the number of records plus the expected growth for the future.
-Today there are about 500 000 users with information~\cite{KIVRA}.
-The send requests records adds up to a little bit less than the number of users times the number of tenants and then there is also around 2 million content records.
-This results in a total of about 10 million records.
-All of the data sets have an expected growth of about 5\% per month according to internal statistics of Kivra collected from their Riak cluster.
-This is well within the limits of traditional relational databases~\cite{SQLPERF}.
+\section{Delimitations}
+If the project is easier than expected it could be expanded by looking into 
+automatically doing simpler image analysis to find images that are about to 
+be assigned to the incorrect category, for example a ``Very secret'' image 
+that is about to be published as ``Public''.
 
-\subsubsection{Complexity of the data sets}
-None of the data sets have any complex structures, the data types are not varying and there are no big documents that requires to be stored in the analytical database as only the meta data of the content being
-sent to users has to be saved and not any attachments, images etc.
+It is possible to do delimitation of the core research question as the project 
+proceeds, but it will most likely not be needed as the research question is 
+quite concise.
 
-\subsubsection{Inefficiency of other methods}
-There is no need for extremely fast reads for specific values which key-value stores~\cite{KEYVALUESTORE} offer or for extremely fast insertion rates which big data databases~\cite{BD1} can provide. 
-As concluded in the section about the size of the data sets (~\ref{DATASETSIZE}) and the section called Choosing SQL database (~\ref{SQLDATABASES}) the insertion rate of 
-traditional databases will be well above the demands. As there is no complex data the normal table structure of a RDBMS will be suitable for storage of the data.
+\section{Time Plan}
+Note that this is just a rough time plan that will change during the course of 
+the project when the circumstances change.
+There will rarely be one task at a time, but rather a lot of the tasks will be 
+overlapping.
+\begin{itemize}
+  \item A few days - Properly setting up the work station
+  \item 2 weeks - Literature study and obtaining relevant books etc
+  \item 20 weeks - Writing the report should be done pretty much 
+  throughout the whole project
+  \item 1 week - Analyze the current system for managing the CDN assets
+  \item 1 week - Analyze what has been done earlier regarding the 
+  subject and determine if anything is useful for this project
+  \item 4 weeks - Implement a prototype
+  \item 1 week - Scalability testing
+  \item 3 weeks - Security testing
+\end{itemize}
 
-\newpage
-\subsection{Protocols for data transfer}
-The requirements of the protocol and data structure used to transfer and translate data to the database differs highly on what type of data and from where the data is being sent, some data are required to have
-absolute transport security and reliability meanwhile other is of lower importance and rather trades transport reliability for speed. To maximize the speed and flexibility of the system as a whole, 
-one or several APIs and web services that can accept several protocols and differently structured data being used towards it before translating it in order to store it in the database should be used, as can be seen in
-Figure~\ref{fig:api}.
+\section{References}
+One possibility to develop a solution from:
 
-The exact APIs, Web Services and data structures used for the implementation of this system is outside the scope for this work and will be left to a future implementer to decide.
-But examples of such protocols
-could be HTTP through a RESTful~\cite{REST} service, cURL~\cite{CURL} with one of its many supported protocols or simply whatever protocol that will be built in to the customised Lager~\cite{LAGER} backend.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_api}}
-    \caption{API explanation}
-    \label{fig:api}
-\end{figure}
-
-\subsection{Database}
-\subsubsection{Requirements}
-\label{REQUIREMENTS}
-As the database will mostly be used for insertions, the insert operations will have to be as fast as possible with the chosen way to store the data. 
-Read operations will not have to be very fast as they are not directly used by any response time critical system but rather by in-house users, a business intelligence system or a system providing an interface of generated
-graphs and tables for overviewing the information.
-
-The requirements on the speed of updates and deletes are non essential as they will only occur when manually editing the database. 
-The database will not have to be portable, it is not going to be replicated or moved around. 
-For future scaling of the database it is needed to be easily configurable and preferably have wide support for distribution of the data over several nodes. 
-The data is required to be accessible and easy to aggregate upon for future analysis.
-
-\subsubsection{Why SQL over NoSQL?}
-A traditional relational database system was chosen over a NoSQL document store or key-value store for a few different reasons.
-The first reason was that the high insertion rate trade off that results in eventual consistency~\cite{EVENTUALCONS} given by NoSQL databases would simply not be needed as the amount of data insertions and updates would
-be well within the limits of a traditional RDBMS~\cite{SQLPERF} as can be seen in Section~\ref{SCALABILITYRESULT}.
-
-The second reason was that the query speed of specific values, offered by key-value stores~\cite{KEYVALUESTORE}, was not required and the demand was rather for a full set of aggregate 
-functions and easy access to complete subsets of the stored data. 
-As many NoSQL databases use MapReduce for fetching data it would per definition be hard to write efficient joins and no built in aggregate functions would be provided~\cite{COMP}.
-
-What was also considered was putting a layer on top of a regular NoSQL database that seamlessly provided this type of functionality. 
-The layers considered were HIVE~\cite{HIVE}, PIG~\cite{PIG} and Scalding~\cite{SCALDING} for HADOOP~\cite{HADOOP}.
-But as the implementation did not need the special functionalities and speed trade-offs,
-which is deeper explained in Section~\ref{SCALABILITYRESULT} and Section~\ref{REQUIREMENTS}, that were offered by such layers the conclusion was to simply incorporate a traditional SQL database.
-
-\subsubsection{Choosing SQL database}
-\label{SQLDATABASES}
-There are a large number of different traditional database systems being actively developed today. For this project the focus has been on popular open source DBMSs. 
-The most popular open source relational database management systems~\cite{POPULARDBMS} today are MySQL, PostgreSQL and SQLite so the tests compare those towards each other and the requirements.
-
-\newpage
-\paragraph{MySQL}
-~\\
-MySQL\cite{MYSQL} is the most popular open source database and has several popular forks as well, with the biggest one~\cite{POPULARDBMS} called MariaDB~\cite{MARIADB}.
-MySQL does not aim at implementing
-the full SQL standard, but rather a subset of it with some extra features~\cite{POPULARDBMS}. MySQL is built in layers with one or several storage engines in the bottom and an optimizer, parser, 
-query cache and connection handling layers built on top of that~\cite{HIMYSQL}, see Figure~\ref{fig:mysql}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{mysql}}
-    \caption{MySQL Architecture~\cite{HIMYSQL}}
-    \label{fig:mysql}
-\end{figure}
-~\\
-
-MySQL has a rich set of aggregate functions which are very useful when doing statistical analysis of the stored data. In total, MySQL has 17 aggregate functions in which the most common SQL standard 
-aggregate functions AVG, COUNT, MAX, MIN, STD and SUM are included~\cite{MYSQLAGG}.
-
-\FloatBarrier
-\newpage
-\paragraph{PostgreSQL}
-~\\
-PostgreSQL is one of the few modern SQL databases that implements the whole SQL standard (ANSI-SQL:2008) ~\cite{POSTGRESSTANDARD}. 
-The architecture of PostgreSQL is quite straight forward, as can be seen in Figure~\ref{fig:postgres}.
-The postmaster handles new connections and spawns a PostgreSQL backend process that the client can directly communicate with.
-All communication
-between client and PostgreSQL goes through that process after it spawns.
-The postmaster stays active and awaits new incoming connections.
-Each backend that a client is connected to has its own separate
-parser, rewrite system, optimiser and executor. All of the active backends then uses the same storage engine~\cite{POSTGRES}. PostgreSQL, compared to MySQL, only has one very tightly integrated 
-storage engine that in theory could be replaced but that has not been done in practice yet~\cite{POSTGRESQLPAPER}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{postgres}}
-    \caption{PostgreSQL Architecture~\cite{POSTGRES}}
-    \label{fig:postgres}
-\end{figure}
-~\\
-
-PostgreSQL have a comparatively large set of aggregate functions including 14 general purpose functions and 18 functions mainly used for statistics, totalling in 32 functions~\cite{POSTGRESAGG}.
-
-\newgeometry{top=2.0cm}
-\paragraph{SQLite}
-~\\
-SQLite as opposed to MySQL and PostgreSQL is not built as a traditional two tier architecture which can be seen in Figure~\ref{fig:sqlite}.
-This leads to that each call to it is done through function calls to a linked SQLite library~\cite{SQLITEDEV}.
-SQLite is quite small (658KiB) and has therefore become widely used as an embedded database
-for local storage~\cite{SQLITE}. SQLite is bundled with two storage engines, one of them is built with a log-structured merge-tree and optimised for persistent 
-databases and the other one is built on a binary tree which is kept in memory which makes it suitable for temporary databases.
-The architecture of SQLite has made it possible to design and interchange
-new storage engines with the built-in ones~\cite{SQLITEENGINE}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{sqlite}}
-    \caption{SQLite Architecture~\cite{SQLITE}}
-    \label{fig:sqlite}
-\end{figure}
-~\\
-
-SQLite has 9 of the most basic aggregate functions and lacks the more statistical functionality such as for example standard deviation and covariance that PostgreSQL and MySQL implement~\cite{SQLITEAGG}.
-SQLite's architecture is very focused on portability~\cite{SQLITE} and have made some trade-offs in order to achieve that, as can be seen in its simplistic architecture in Figure~\ref{fig:sqlite}. 
-One result of these trade-offs is that insertions become very slow, $\sim60$ insertions per minute with transactions enabled~\cite{SQLITESLOW}.
-
-As portability is not required for the implementation of the system regarded in this paper, because it is missing many important aggregate functions and because of its slow insertion rate, 
-SQLite will be excluded from the tests in the next section.
-
-\newpage
-\subsubsection{Performance tests}
-As explained in section~\ref{SQLDATABASES}, SQLite was excluded from the tests as its architecture was deemed unfitting for the task at hand and for the lack of many aggregate functions.
-The statistics were generated on an average consumer computer with the following specifications:
-
-\vspace{0.25in}
-\textbf{CPU:} Intel Core2 Duo T7700 2.40GHz
-
-\textbf{RAM:} SODIMM DDR2 Synchronous 667 MHz 2x2GB
-
-\textbf{Disk:} KINGSTON SV300S3 120GB SSD partitioned with EXT4
-
-\textbf{OS:} Debian Jessie (Testing)
-
-~\\
-PostgreSQL version 9.4 and MySQL version 5.6 were used during the tests and the packages were installed from the default Debian testing repositories.
-
-~\\
-The database and table were created in the same manner on both systems as can be seen in listing~\ref{database_creation}.
-\begin{lstlisting}[label=database_creation,caption=Database and table creation]
-create database testdb
-create table test_table 
-    (data varchar(100), 
-    moddate timestamp default current_timestamp);
-\end{lstlisting}
-
-The tests were conducted with a series of simple insertions containing random data, as shown in listing~\ref{test_table}.
-\begin{lstlisting}[label=test_table,caption=Example insertion]
-insert into test_table (data) 
-values ('010100100101010100...');
-\end{lstlisting}
-
-The execution times were measured by using the time~\cite{TIME} utility of the GNU project, it measures the time of execution and resource usage of the specified operation. 
-The different files that were used in the tests contained 1000, 10 000, 100 000 and \\1 000 000 insertions of random data. 
-When deploying this system the production servers will most likely be a lot more powerful than the consumer laptop that these test were run on.
-These tests non the less gives a correct representation of how well the different databases will scale. 
-To reproduce or verify the results, see listings 3 and 4 in Appendix~\ref{appendix:a}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{linegraph2}}
-    \caption{Performance graph}
-    \label{fig:vs}
-\end{figure}
-
-As can be seen in Figure~\ref{fig:vs} both of the systems have a very close to linear performance and any small disturbances in the pattern can be concluded to be from other processes using system resources.
-Both of the systems perform over the preconditions for an implementation with requirements similar to the ones in this context.
-But as insertion rate and aggregate functions are of high concern for such
-an implementation and both systems comply to the rest of the requirements, PostgreSQL will be recommended to use for a future implementation.
-
-\newpage
-\subsection{Storage requirements}
-As the data comes in as events, there is never a need to update old records in the database, as can be seen in Figure~\ref{fig:storage}. This structure is slightly harder to do complicated 
-queries (Listing~\ref{fetch_example}) towards, as you always have to take the timestamp of the event into consideration. 
-But it also makes it possible to configure the DBMS not to lock tables during inserts, which will increase the speed of insertion of the database.
-The lack of locks will be especially useful if it is later
-decided to use a DDBMS for an eventual requirement of scaling up the system in the future.
-As everything is logged with timestamps it is particularly suitable for business intelligence systems and the likes of such, as you can query
-time ranges in the past as well as the current state of the system to gather analytics and make comparisons.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.6]{system_dbms}}
-    \caption{Flow of data}
-    \label{fig:storage}
-\end{figure}
-
-The only time inserts occur is during the time when the customised Lager~\cite{LAGER} backend or later added data sources sends events (1, see Figure~\ref{fig:storage}) and the only times reads are performed
-is when a manual query (2, see Figure~\ref{fig:storage}) is executed or when an eventual business intelligence system or an automated analyse system (3, see Figure~\ref{fig:storage}) that automatically queries the system is used.
-It will never be necessary to perform updates and no regular deletes will have to be performed, as explained in Section~\ref{REQUIREMENTS}.
-
-The actual database size should not be a problem as the servers are usually equipped with a few terabytes of disk space.
-An average insertion of Kivra is estimated to around 200 bytes by looking at meta data formed by data currently in the Riak cluster.
-This means that $5*10{^9}$ insertions would fit on a disk with 1TB of space.
-A rough estimation of an average of 5000 insertions per hour means that a 1TB would last for more than 113 years, see Figure~\ref{fig:storage2}.
-
-\begin{figure}[h!]
-\begin{framed}
-\begin{enumerate}
-\item $1TB/200B \thickapprox 5*10^{9}$ insertions
-\item $5*10^{9}/5000/24/365 \thickapprox $ 113.5 years
-\end{enumerate}
-\end{framed}
-\caption{Estimation of disk space usage}
-\label{fig:storage2}
-\end{figure}
-
-\FloatBarrier
-\newgeometry{top=2.0cm}
-\newpage
-\section{Resulting system}
-The result of an implementation of the suggested solution of this paper would lead to a core system that would have a data flow as described in Figure~\ref{fig:result}.
-In this flow, when data of analytical interest would be inserted into the Riak cluster it would also be sent to Lager which would insert it into a RDBMS dedicated for 
-analytical data, instead of fetching data in Riak after it has been inserted as was it done in the previous systems data flow that can be seen in Figure~\ref{fig:before} and explained in Section 3.2.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_after}}
-    \caption{Resulting system}
-    \label{fig:result}
-\end{figure}
-
-What is not shown in Figure~\ref{fig:result}, is that separate data sources for the analytical database can be used. These can connect with different protocols and with differently 
-structured data to an API that will be receiving data for the database as described in Section 5.2.
-
-\subsection{Scalability}
-\label{SCALABILITYRESULT}
-If an unoptimised PostgreSQL database on an average consumer computer is used and with the assumption that little to no data transfer latency will occur during insertion of data the insertion rate can
-be increased by a factor of at least 100 if following the upper boundary of the estimation (Section 5) of 10000 insertions per hour.
-As can be seen in the load test in 5.3.4, PostgreSQL can handle a little bit over
-1 million insertions per hour. The expected data growth rate is 5\% per month, with this growth the same setup could be used for almost 8 years, see Figure~\ref{fig:scalability}.
-
-\begin{figure}[h!]
-\begin{framed}
-\begin{enumerate}
-\item $10000*1.05^{94} \thickapprox $ 981282 insertions/hour 
-\item $94/12 \thickapprox $ 7.83 years 
-\end{enumerate}
-\end{framed}
-\caption{Estimation of data transfer scaling}
-\label{fig:scalability}
-\end{figure}
-
-In the unlikelihood that the same setup with the same hardware and PostgreSQL version would be used in 8 years the PostgreSQL database could easily be sharded, clustered or optimised to reach a lot higher
-insertion rates.
-As all he data is independent and has no defined relations to other data, clustering with a pool of independent databases could be used without any requirement of the databases to synchronise with each
-other, this would mean that you can linearly add 1000000 insertions/hour for each database node that you add to the cluster.
-
-\subsection{Practical Example}
-\subsubsection{Storing an Event}
-After an implementation of a solution like this at Kivra the following would be the flow of a typical change triggered by a user. 
-The system would update the send request bucket of Riak in Figure~\ref{fig:practical}, but changes to other buckets would look exactly the same.
-
-Figure~\ref{fig:practical} is a practical example of what would happen if a user updated his or her preferences of which tenants it wishes to receive documents from.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{practical_flow}}
-    \caption{Practical example of a send request changing}
-    \label{fig:practical}
-\end{figure}
-
-\newpage
-\subsubsection{Querying Data}
-To query data stored in the analytical database a query that takes the timestamps into consideration has to be performed. The query in listing~\ref{fetch_example} is an example of how to query which 
-tenants all users wish and do not wish to receive documents from. Queries for other data sets would look similar.
-
-\begin{lstlisting}[label=fetch_example,caption=Example query]
-SELECT SRQ.user_id, SRQ.tenant_id, SRQ.timestamp, SRQ.accepted 
-FROM sendrequest SRQ 
-INNER JOIN 
-  (SELECT user_id, tenant_id, max(timestamp) as latest
-   FROM sendrequest GROUP BY user_id, tenant_id) SRQ1 
-ON SRQ.user_id = SRQ1.user_id 
-AND SRQ.tenant_id = SRQ1.tenant_id 
-AND SRQ.timestamp = SRQ1.latest;
-\end{lstlisting}
-
-The query in listing~\ref{fetch_example} doesn't take into consideration if the same name and tenant combination exists with the same timestamp. 
-It is however highly unlikely that it would happen as the timestamps are stored with a micro second precision. 
-To fully avoid that event, add an auto incremented ID column to the table and only consider the highest ID in the case of a clash.
-
-\newpage
-\section{Discussion}
-\subsection{Event-driven data collection}
-The suggested solution introduced by this paper was named event-driven data collection. The name has been used before for slightly similar purposes with the first recorded case 
-being in 1974~\cite{EVENTDRIVEN}.
-It has also been used in a different manner in quite a few papers on the subject wireless sensor networks~\cite{EVENTDRIVEN2} and some regarding the collection of natural
-ionosphere scintillations~\cite{EVENTDRIVEN3}.
-
-\subsection{Flat file database}
-A flat file database was not compared in this paper because of the precondition of having pre-built APIs and to give easier access to querying and aggregation of data.
-A flat file database can also be complicated to scale out
-to several servers for an eventual future requirement.
-
-\newpage
-\newgeometry{top=4.0cm}
-\section{Summary}
-\subsection{Conclusions}
-The system that has been designed in this paper can be used as an effective tool for logging and analysing events.
-This form of approach is especially valuable if the main system is using some form of MapReduce
-database which can be hard to do ongoing analysis of aggregated data on.
-If the main system is already using a SQL database as its primary form of information storage this type of solution would be quite inefficient
-due to duplication of data and possibly having to serve another DBMS\@.
-In the benchmarking of the different database management systems the standard version without modifications have been used, with modifications
-and optimisations the outcome would have been very different.
-
-The unorthodox design of the database schema in this paper provides a full historical log of events which makes it possible to compare data from different time ranges, 
-which can be very valuable when analysing data from business systems to for example detect trends and tendencies within the user base.
-
-Another contribution of this work is the use of a logging system to effectively put in an analytical systems 
-to an already existing backend with minimal impact on the performance of the existing system.
-
-\subsection{Future work}
-To finalise the architecture of this system, APIs and web service structures will have to be researched or developed to be put on top of the database.
-For larger flows of data other solutions would have
-to be considered or alternatively the used DBMS has to be heavily optimised by for example removing insertion locks. Scaling out to a clustered database could also be an alternative.
-
-\newpage
-\newgeometry{top=2.0cm}
-\begin{appendices}
-\section{Benchmarks}
-\label{appendix:a}
-\begin{lstlisting}[label=some-code,caption=PostgreSQL insertion test]
-1 000 000 insertions
-$ time psql --quiet -f test4.sql
-psql --quiet -f test4.sql
-40.42s user 23.08s system 2% cpu 31:05.03 total
-
-100 000 insertions
-$ time psql --quiet -f test.sql
-psql --quiet -f test.sql
-4.32s user 2.38s system 2% cpu 5:02.94 total
-
-10 000 insertions
-$ time psql --quiet -f test2.sql
-psql --quiet -f test2.sql
-0.51s user 0.22s system 2% cpu 30.172 total
-
-1000 insertions
-$ time psql --quiet -f test3.sql
-psql --quiet -f test3.sql 
-0.11s user 0.02s system 4% cpu 3.077 total
-\end{lstlisting}
-\begin{lstlisting}[label=some-code,caption=MySQL insertion test]
-1 000 000 insertions
-$ time mysql -u root -p********** testdb < test4.sql
-mysql -u root -p********** testdb <   
-39.78s user 23.36s system 1% cpu 1:36:56.31 total
-
-100 000 insertions
-$ time mysql -u root -p********** testdb < test.sql
-mysql -u root -p********** testdb < 
-3.88s user 2.83s system 1% cpu 9:22.67 total
-
-10 000 insertions
-$ time mysql -u root -p********** testdb < test2.sql 
-mysql -u root -p********** testdb < 
-0.31s user 0.39s system 1% cpu 53.174 total
-
-1000 insertions
-$ time mysql -u root -p********** testdb < test3.sql 
-mysql -u root -p********** testdb < 
-0.08s user 0.00s system 1% cpu 5.318 total
-\end{lstlisting}
-\end{appendices}
-\restoregeometry
-
-\newpage
-\bibliographystyle{ieeetr}
-\bibliography{references}
+\url{http://docs.aws.amazon.com/AmazonCloudFront/latest/DeveloperGuide/private-content-signed-cookies.html}
 

### 2015-09-21 tex/paper.tex
Added some introduction

@@
 
-\title{\textbf{NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection}}
+\title{\textbf{Providing an Access Control Layer for Content Distribution Networks}}
 \author{Lukas Klingsbo}
@@
 
-%\maketitle
-\includepdf[pages={1}]{front.pdf}
-\thispagestyle{empty}
-\newpage\null\thispagestyle{empty}\newpage
+\maketitle
+%\includepdf[pages={1}]{front.pdf}
+%\thispagestyle{empty}
+%\newpage\null\thispagestyle{empty}\newpage
 
@@
 
-\includepdf[pages={1}]{abstract.pdf}
-%\begin{abstract}
-%Collecting and analysing data of analytical value is important for many service providers today. 
-%Many make use of NoSQL databases for their larger software systems, what is less known is how to 
-%effectively analyse and gather business intelligence from the data in these systems. 
-%This paper suggests a method of separating the most valuable analytical data from the rest in 
-%real time and at the same time providing an effective traditional database for the analyser. 
-%
-%In this paper we analyse our given data sets to decide whether big data tools are required and 
-%then traditional databases are compared to see how well they fit the context. 
-%A technique that makes use of an asynchronous logging system is used to insert the data from the 
-%main system to the dedicated analytical database. 
-%
-%The tests show that our technique can efficiently be used with a traditional database even on large 
-%data sets (>1000000 insertions/hour per database node) and still provide both historical data 
-%and aggregate functions for the analyser.\\
-%
-%\keywords{NoSQL, data mining, data collection, DBMS benchmark, business intelligence, analytical logging}
-%\end{abstract}
+%\includepdf[pages={1}]{abstract.pdf}
+\begin{abstract}
+    TODO: Abstract
+
+\keywords{}
+\end{abstract}
 
@@
 \setcounter{page}{1}
-\section{Introduction to the\\"Big Data/NoSQL problem"}
-In times when big data collection is growing rapidly~\cite{E1} companies and researchers alike face the problem of doing analysis of the data that has been collected.
-The enormous amounts of data collected are often saved in distributed big data databases or key-value stores with fast insertion operations but with comparatively slow and complex (but highly parallelizable) query functions~\cite{RT1}.
-This work tries to solve such a problem at Kivra~\cite{KIVRA}, a company working to digitalise and minimise the use and need of conventional paper mail.
-To gather analytical data from Kivra's distributed database today requires heavy Map/Reduce
-queries.
-To do day to day analysis of the data slowly becomes unbearable and non-scalable as it takes several hours or even days to run a Map/Reduced~\cite{G1} batch job to fetch the data that is going to be analysed.
-The solution proposed and examined by this paper is to structure and asynchronously collect data after interesting events occur, that are of a high analytical value, into a smaller and more easily queryable database.
+\section{Introduction}
+Developing larger projects containing static content usually involves using a Content Distribution Network 
+to be able to scale to a large user base. The commercial Content Distribution Networks are usually fairly easy to use, 
+the content that is to be used in a project is usually simply uploaded and then directly available to the public. 
+For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
+access control on content and groups of content in the form of views. A system was developed to make the underlying 
+theory work in practice.
 
@@
 \section{Related Terminology}
+\subsection{Technologies}
+\subsubsection{React}
+React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and the browser's, 
+this makes it able to efficiently update dynamic web pages after a change of state through comparing the old virtual 
+DOM with the resulting virtual DOM after the state change and then only update the browser's DOM according to the 
+difference between the virtual DOMs~\cite{REACT}.
 
-\subsection{Big data}
-Big data~\cite{BD1} is a loose term for ways of collecting and storing unstructured complex data of such large volumes that normal methods of data manipulation and querying can not be used.
+\subsubsection{Flux}
 
-\subsection{NoSQL}
-NoSQL~\cite{NOSQL} is a term that have not fully settled its meaning, it started by meaning no relational databases~\cite{NOSQL}, but now it has moved on to commonly be referred to as Not Only SQL. 
-If a database calls itself NoSQL it usually means that is is non-relational, distributed, horizontally scalable, schema-free, eventually consistent and support large amounts of data~\cite{NOSQL2}.
+\subsubsection{Scala}
+Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared to Java it supports 
+most functional programming features at the same time as it supports object oriented programming~\cite{SCALA}.
 
-\subsection{MapReduce}
-MapReduce~\cite{G1} is referring to a programming model that makes parallel processing of large data sets possible.
-In simplified terms it consists of two functions, one called Map and one called Reduce.
-The Map function is sent out to each node from a master node and the function is separately executed on each nodes data set and then each node assembles the result back together with the other nodes so that all
-input values with the same results are grouped together.
-Then the reduce function is executed, either in parallel or on a master node, and this function does calculations upon each separate result list with the data from all nodes that
-ran the Map function and it then combines the input from them to the answer that was originally inquired.
-
-\subsection{Eventual consistency}
-The expression eventual consistency is a term used for describing how data is processed in database systems. Eventual consistency means that the data may not be consistent in any one moment, 
-but will eventually converge to being consistent~\cite{EVENTUAL}. This makes it possible for distributed databases to operate faster as data does not have to be locked during insertions, updates 
-and deletion and thus does not have to be propagated to all nodes before more actions on the same data can be sent to a node. If one or more nodes operate on the same data as another node 
-whose changes have not propagated through the system yet, the system will have to handle in which order the operations should be performed, this is usually called sibling or conflict resolution~\cite{RESOLUTION}.
-
-\subsection{Riak}
-Riak is a key-value datastore that implements NoSQL concepts such as eventual consistency. It is usually configured to run with at least 5 nodes to be as efficient as possible.
-To query the datastore you either use the REST-ful API or MapReduce functionality that it provides.
-
-Riak~\cite{RIAK} provides fault tolerance and high availability by replicating data to several nodes, the default is to have all data replicated to three nodes.
-
-\subsection{Lager}
-Basho~\cite{BASHO}, the company that developed Riak~\cite{RIAK}, has developed a modular logging framework called Lager~\cite{LAGER}.
-Lager is modular in the sense that it is possible to write separate backends
-which all can separately do asynchronous data collection.
-As the system performance can not be noticeably affected by adding a subsystem to collect analytically valued events it is of highest priority
-that they can be sent asynchronously so that the implication of implementing such a subsystem wont result in delays for the main system.
-
-A problem that these form of logging systems can experience is that the events that they have to handle queue up faster than they can be processed.
-To solve this Lager~\cite{LAGER} has a built-in overload protection
-to make sure that its message queue will not overflow. It does this by swapping synchronous and asynchronous messaging for the separate backend.
-
+\subsubsection{JSON-RPC}
+Over http? WebSockets? 
+\subsubsection{TODO: Insert persistent storage here}
+    TODO: Write down related terminology, if any
 \subsection{Abbreviations}
 
-\subsubsection{RDBMS}
-RDBMS~\cite{RDBMS} is an abbreviation for relational database management system.
-This is the most commonly used database management system and is usually referred to as a traditional database.
-Data is stored within tables and data can be in relation to each
-other across tables and such databases are therefore called relational.
+\subsubsection{CDN}
+Content Distribution Network - Replicates content to several servers, usually spread out geographically. Once a 
+request is made, the network serves content from the server closest to the requester.
 
-\subsubsection{DDBMS}
-DDBMS~\cite{DDBMS} is an abbreviation for distributed database management system.
-It is used to centrally manage a distributed database system and to give the user the impression and functionality as though the database system
-would run on only one node.
-
-\newpage
 \section{Background}
-\subsection{About Kivra}
-Kivra~\cite{KIVRA} is a company which aims to digitalise and eliminate the use of superfluous analogue paper mail. 
-They do this by making it possible for companies and governments to send documents digitally to customers or citizens. 
-By tying the users' accounts to their social security numbers in the service the sender can be sure that the document 
-reaches the correct recipient.
-
-Today Kivra has a continuously growing user base of more than 500.000 users, according to their website~\cite{KIVRA}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.4]{kivraui}}
-    \caption{The user interface of Kivra}
-    \label{fig:ui}
-\end{figure}
-
-As can be seen in figure~\ref{fig:ui}, the interface reminds of the interface of any email provider, however the documents sent to the users can be interactive and have a more advanced layout and the main advantage over normal email is that every
-users identity is established in the registration process. 
-The user can validate its identity either by using mobile BankID~\cite{BANKID} or by receiving an activation code to its registered place of residence. 
-This advantage makes it possible for companies and organizations that otherwise would have had to send documents to a registered postal address to being able to send them digitally through Kivra.
+\subsection{About Uprise}
+Uprise is a company based in Uppsala, Sweden.
 
 \subsection{The current system}
-Kivra's users interact with the backend system (this is the backend box represented in Figure~\ref{fig:before}) either through Kivra's web page or their app, where they can receive documents, pay bills, set whom they accept documents from and also set personal details.
-Users sign up either though mobile bankID or requests an activation code to be sent to the address registered to their social security number, this is to ensure the identity of the user. 
-When a user receives a new document it is notified through email or SMS and can then log in to read the document.
+Maybe write about battlebinary?
 
-The data which is of analytical interest is information about the users, meta data of the documents sent out to the users and whether a user accepts or does not accept (internally called \emph{send requests}) documents from a company or organization (internally called a \emph{tenants}).
+\subsection{Problem description}
+Having
 
-\newpage
-\subsection{Problem description}
-The part of the current system that is being examined consists of Riak, a distributed key-value datastore, and a batch script containing MapReduce queries which are running each night to collect 
-data to a PostgreSQL database which is later used for deeper analysis. This system architecture can be seen in Figure~\ref{fig:before}. 
-The batch script is emptying all of the analytically interesting buckets in the Riak cluster,
-currently there are two buckets being emptied every night.
-One of them contains send requests which is meta data telling the system if a user accepts documents for a specific tenant and the other bucket
-just contains plain user information. In the future meta data regarding documents sent to users will be collected as well.
+Battle Binary
 
-This solution does not scale and is also inefficient as you have to go through all the data in the growing buckets that are of interest every night, instead of accumulating the data to a separate database dedicated for analysis
-at the same time as you insert or change data in Riak.
+What is needed
+* Security layers
+* Views
+* Virtual file structure
+* Versioning of content
+* Multi project support
+* Auth and audit logs
+* Users
 
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_before}}
-    \caption{The analytical system at Kivra}
-    \label{fig:before}
-\end{figure}
-
-~\\
-\newpage
 \section{Methods for determining\\implementation details}
@@
 
-\subsection{Data size and growth estimation}
-The data size and growth was estimated with statistics collected from the current Riak cluster in Kivra's system.
-What the current insertion rate would be like with an event driven implementation was estimated
-by doing SQL queries measuring the number of insertions that could be made within certain time limits. 
-To estimate the size and growth of the future database implementation both past estimations and the current internal estimated growth of the company were used. 
-This, together with the scalability measurements, was required to make sure that the solution would be durable in the long run.
+\section{Security of the system}
+\subsection{Authorization}
+\subsection{Audit logs}
 
-\subsection{Scalability measurement}
-The scalability was estimated for the typical server hardware of Kivra with the different databases and with statistics of growth estimations from the current database. 
-This was necessary to make sure that the system would scale with the expected data growth, within a reasonable time frame.
+\section{Resulting system}
+\subsection{Scalability}
 
-\subsection{DBMS comparison}
-The database management systems were compared by testing required abilities and examining load and speed by running insertion jobs of different sizes. 
-The DBMSs architecture and suitability were also compared to see which one suited the context best. 
-These comparisons were vital to make sure that the chosen DBMS could handle the expected load of the system.
+\section{Discussion}
 
-\newpage
-\section{Solution for storing events}
-When an event of analytical interest occurs, an asynchronous call with information about the event is sent to Lager.
-Lager then directs it to an in-house tailored Lager backend, which is responsible for
-making sure that the data is stored consistently and correctly. The database in which the data is stored can then be queried at any time to perform near to real time analysis of the system. 
-After analysing internal database statistics an approach of this form is estimated to cause approximately 100-10000 inserts/hour depending on if it is during the peak hours or not.
-The estimation can not be averaged as
-the chosen database is required to be able to handle the peak hours without a problem.
-This approach will lead to a system with no updates and only reads during analytical calculations.
-This is due to the fact that
-the backend will handle each event with a strict progression of inserting the relevant metadata of the event with a timestamp and never have a need of updating, nor removing old data, as all of these 
-chosen events are valuable to store to enable historical queries. This is similar to how Lager~\cite{LAGER} logs events to files, except that the analysis of the data will be easier due to the 
-built-in aggregate functions of modern databases.
-In comparison to the current system it will be possible to perform near to instant analysis of the data already collected instead of having to re-collect all
-the data of analytical interest from the Riak cluster each time an up-to-date analysis of the data has to be performed.
-
-\subsection{Data sets}
-\subsubsection{Determining the type of the data sets}
-Big data is one of the popular buzzwords in the computer industry today and a lot of revolutionising technology has been built around it. But it is not always efficient to use technology built for 
-these data sets if the data set that is being operated on does not fit the definition of big data.
-But as there is a thriving hype of higher yield with existing data some companies and institutions use it
-just for the sake of being part of the trend without actually having a data set that objectively would be classified as big data~\cite{NOTBIG}.
-
-To simplify the abstract definition of Big Data to determine whether a data set can be defined as such you can classify the data set by concluding the answers of 
-three questions~\cite{BD1}, namely the following:
-\begin{itemize}
-    \item Is the data set large?
-    \item Is the data set complex (maybe even unstructured)?
-    \item Are methods optimized for other data sets inefficient?
-\end{itemize}
-
-If the data set satisfies these properties it can be a good idea to store the data in a Hadoop cluster or similar technology developed specifically for big data. 
-If, however, the data set does not satisfy these criteria it is usually better to look for another storage solution~\cite{NOTBIG}.
-
-At Kivra there are a few different data sets with information that is valuable for ongoing analysis. 
-The three most important data sets are: 
-\begin{itemize}
-    \item User information - each users personal details.
-    \item Send requests - connects a user to a tenant and specifies whether the user wants or does not want content from that specific tenant.
-    \item Content metadata - contains meta data of content that has been sent to users. 
-\end{itemize}
-~\\
-The tables representing this data would be quite simple and would not store too much information as analysis of the system would mostly regard metadata of some of the information stored in the Riak cluster.
-See an example of how such tables could look like in listing~\ref{example_tables}.
-
-\begin{lstlisting}[label=example_tables,caption=Example tables]
-Example fields of the tables
-+-------------+  +--------------+  +------------------+
-| User        |  | Send request |  | Content Metadata |
-+-------------+  +--------------+  +------------------+
-| id          |  | user_id      |  | content_id       |
-| name        |  | tenant_id    |  | tenant_id        |
-| address     |  | accepted     |  | user_id          |
-| city        |  | timestamp    |  | timestamp        |
-| email       |  |              |  |                  |
-| timestamp   |  |              |  |                  |
-+-------------+  +--------------+  +------------------+
-\end{lstlisting}
-
-\subsubsection{Size of the data sets}
-\label{DATASETSIZE}
-As each record in the data sets is fairly small (under 200 bytes) it is enough to count the number of records plus the expected growth for the future.
-Today there are about 500 000 users with information~\cite{KIVRA}.
-The send requests records adds up to a little bit less than the number of users times the number of tenants and then there is also around 2 million content records.
-This results in a total of about 10 million records.
-All of the data sets have an expected growth of about 5\% per month according to internal statistics of Kivra collected from their Riak cluster.
-This is well within the limits of traditional relational databases~\cite{SQLPERF}.
-
-\subsubsection{Complexity of the data sets}
-None of the data sets have any complex structures, the data types are not varying and there are no big documents that requires to be stored in the analytical database as only the meta data of the content being
-sent to users has to be saved and not any attachments, images etc.
-
-\subsubsection{Inefficiency of other methods}
-There is no need for extremely fast reads for specific values which key-value stores~\cite{KEYVALUESTORE} offer or for extremely fast insertion rates which big data databases~\cite{BD1} can provide. 
-As concluded in the section about the size of the data sets (~\ref{DATASETSIZE}) and the section called Choosing SQL database (~\ref{SQLDATABASES}) the insertion rate of 
-traditional databases will be well above the demands. As there is no complex data the normal table structure of a RDBMS will be suitable for storage of the data.
-
-\newpage
-\subsection{Protocols for data transfer}
-The requirements of the protocol and data structure used to transfer and translate data to the database differs highly on what type of data and from where the data is being sent, some data are required to have
-absolute transport security and reliability meanwhile other is of lower importance and rather trades transport reliability for speed. To maximize the speed and flexibility of the system as a whole, 
-one or several APIs and web services that can accept several protocols and differently structured data being used towards it before translating it in order to store it in the database should be used, as can be seen in
-Figure~\ref{fig:api}.
-
-The exact APIs, Web Services and data structures used for the implementation of this system is outside the scope for this work and will be left to a future implementer to decide.
-But examples of such protocols
-could be HTTP through a RESTful~\cite{REST} service, cURL~\cite{CURL} with one of its many supported protocols or simply whatever protocol that will be built in to the customised Lager~\cite{LAGER} backend.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_api}}
-    \caption{API explanation}
-    \label{fig:api}
-\end{figure}
-
-\subsection{Database}
-\subsubsection{Requirements}
-\label{REQUIREMENTS}
-As the database will mostly be used for insertions, the insert operations will have to be as fast as possible with the chosen way to store the data. 
-Read operations will not have to be very fast as they are not directly used by any response time critical system but rather by in-house users, a business intelligence system or a system providing an interface of generated
-graphs and tables for overviewing the information.
-
-The requirements on the speed of updates and deletes are non essential as they will only occur when manually editing the database. 
-The database will not have to be portable, it is not going to be replicated or moved around. 
-For future scaling of the database it is needed to be easily configurable and preferably have wide support for distribution of the data over several nodes. 
-The data is required to be accessible and easy to aggregate upon for future analysis.
-
-\subsubsection{Why SQL over NoSQL?}
-A traditional relational database system was chosen over a NoSQL document store or key-value store for a few different reasons.
-The first reason was that the high insertion rate trade off that results in eventual consistency~\cite{EVENTUALCONS} given by NoSQL databases would simply not be needed as the amount of data insertions and updates would
-be well within the limits of a traditional RDBMS~\cite{SQLPERF} as can be seen in Section~\ref{SCALABILITYRESULT}.
-
-The second reason was that the query speed of specific values, offered by key-value stores~\cite{KEYVALUESTORE}, was not required and the demand was rather for a full set of aggregate 
-functions and easy access to complete subsets of the stored data. 
-As many NoSQL databases use MapReduce for fetching data it would per definition be hard to write efficient joins and no built in aggregate functions would be provided~\cite{COMP}.
-
-What was also considered was putting a layer on top of a regular NoSQL database that seamlessly provided this type of functionality. 
-The layers considered were HIVE~\cite{HIVE}, PIG~\cite{PIG} and Scalding~\cite{SCALDING} for HADOOP~\cite{HADOOP}.
-But as the implementation did not need the special functionalities and speed trade-offs,
-which is deeper explained in Section~\ref{SCALABILITYRESULT} and Section~\ref{REQUIREMENTS}, that were offered by such layers the conclusion was to simply incorporate a traditional SQL database.
-
-\subsubsection{Choosing SQL database}
-\label{SQLDATABASES}
-There are a large number of different traditional database systems being actively developed today. For this project the focus has been on popular open source DBMSs. 
-The most popular open source relational database management systems~\cite{POPULARDBMS} today are MySQL, PostgreSQL and SQLite so the tests compare those towards each other and the requirements.
-
-\newpage
-\paragraph{MySQL}
-~\\
-MySQL\cite{MYSQL} is the most popular open source database and has several popular forks as well, with the biggest one~\cite{POPULARDBMS} called MariaDB~\cite{MARIADB}.
-MySQL does not aim at implementing
-the full SQL standard, but rather a subset of it with some extra features~\cite{POPULARDBMS}. MySQL is built in layers with one or several storage engines in the bottom and an optimizer, parser, 
-query cache and connection handling layers built on top of that~\cite{HIMYSQL}, see Figure~\ref{fig:mysql}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{mysql}}
-    \caption{MySQL Architecture~\cite{HIMYSQL}}
-    \label{fig:mysql}
-\end{figure}
-~\\
-
-MySQL has a rich set of aggregate functions which are very useful when doing statistical analysis of the stored data. In total, MySQL has 17 aggregate functions in which the most common SQL standard 
-aggregate functions AVG, COUNT, MAX, MIN, STD and SUM are included~\cite{MYSQLAGG}.
-
-\FloatBarrier
-\newpage
-\paragraph{PostgreSQL}
-~\\
-PostgreSQL is one of the few modern SQL databases that implements the whole SQL standard (ANSI-SQL:2008) ~\cite{POSTGRESSTANDARD}. 
-The architecture of PostgreSQL is quite straight forward, as can be seen in Figure~\ref{fig:postgres}.
-The postmaster handles new connections and spawns a PostgreSQL backend process that the client can directly communicate with.
-All communication
-between client and PostgreSQL goes through that process after it spawns.
-The postmaster stays active and awaits new incoming connections.
-Each backend that a client is connected to has its own separate
-parser, rewrite system, optimiser and executor. All of the active backends then uses the same storage engine~\cite{POSTGRES}. PostgreSQL, compared to MySQL, only has one very tightly integrated 
-storage engine that in theory could be replaced but that has not been done in practice yet~\cite{POSTGRESQLPAPER}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{postgres}}
-    \caption{PostgreSQL Architecture~\cite{POSTGRES}}
-    \label{fig:postgres}
-\end{figure}
-~\\
-
-PostgreSQL have a comparatively large set of aggregate functions including 14 general purpose functions and 18 functions mainly used for statistics, totalling in 32 functions~\cite{POSTGRESAGG}.
-
-\newgeometry{top=2.0cm}
-\paragraph{SQLite}
-~\\
-SQLite as opposed to MySQL and PostgreSQL is not built as a traditional two tier architecture which can be seen in Figure~\ref{fig:sqlite}.
-This leads to that each call to it is done through function calls to a linked SQLite library~\cite{SQLITEDEV}.
-SQLite is quite small (658KiB) and has therefore become widely used as an embedded database
-for local storage~\cite{SQLITE}. SQLite is bundled with two storage engines, one of them is built with a log-structured merge-tree and optimised for persistent 
-databases and the other one is built on a binary tree which is kept in memory which makes it suitable for temporary databases.
-The architecture of SQLite has made it possible to design and interchange
-new storage engines with the built-in ones~\cite{SQLITEENGINE}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{sqlite}}
-    \caption{SQLite Architecture~\cite{SQLITE}}
-    \label{fig:sqlite}
-\end{figure}
-~\\
-
-SQLite has 9 of the most basic aggregate functions and lacks the more statistical functionality such as for example standard deviation and covariance that PostgreSQL and MySQL implement~\cite{SQLITEAGG}.
-SQLite's architecture is very focused on portability~\cite{SQLITE} and have made some trade-offs in order to achieve that, as can be seen in its simplistic architecture in Figure~\ref{fig:sqlite}. 
-One result of these trade-offs is that insertions become very slow, $\sim60$ insertions per minute with transactions enabled~\cite{SQLITESLOW}.
-
-As portability is not required for the implementation of the system regarded in this paper, because it is missing many important aggregate functions and because of its slow insertion rate, 
-SQLite will be excluded from the tests in the next section.
-
-\newpage
-\subsubsection{Performance tests}
-As explained in section~\ref{SQLDATABASES}, SQLite was excluded from the tests as its architecture was deemed unfitting for the task at hand and for the lack of many aggregate functions.
-The statistics were generated on an average consumer computer with the following specifications:
-
-\vspace{0.25in}
-\textbf{CPU:} Intel Core2 Duo T7700 2.40GHz
-
-\textbf{RAM:} SODIMM DDR2 Synchronous 667 MHz 2x2GB
-
-\textbf{Disk:} KINGSTON SV300S3 120GB SSD partitioned with EXT4
-
-\textbf{OS:} Debian Jessie (Testing)
-
-~\\
-PostgreSQL version 9.4 and MySQL version 5.6 were used during the tests and the packages were installed from the default Debian testing repositories.
-
-~\\
-The database and table were created in the same manner on both systems as can be seen in listing~\ref{database_creation}.
-\begin{lstlisting}[label=database_creation,caption=Database and table creation]
-create database testdb
-create table test_table 
-    (data varchar(100), 
-    moddate timestamp default current_timestamp);
-\end{lstlisting}
-
-The tests were conducted with a series of simple insertions containing random data, as shown in listing~\ref{test_table}.
-\begin{lstlisting}[label=test_table,caption=Example insertion]
-insert into test_table (data) 
-values ('010100100101010100...');
-\end{lstlisting}
-
-The execution times were measured by using the time~\cite{TIME} utility of the GNU project, it measures the time of execution and resource usage of the specified operation. 
-The different files that were used in the tests contained 1000, 10 000, 100 000 and \\1 000 000 insertions of random data. 
-When deploying this system the production servers will most likely be a lot more powerful than the consumer laptop that these test were run on.
-These tests non the less gives a correct representation of how well the different databases will scale. 
-To reproduce or verify the results, see listings 3 and 4 in Appendix~\ref{appendix:a}.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{linegraph2}}
-    \caption{Performance graph}
-    \label{fig:vs}
-\end{figure}
-
-As can be seen in Figure~\ref{fig:vs} both of the systems have a very close to linear performance and any small disturbances in the pattern can be concluded to be from other processes using system resources.
-Both of the systems perform over the preconditions for an implementation with requirements similar to the ones in this context.
-But as insertion rate and aggregate functions are of high concern for such
-an implementation and both systems comply to the rest of the requirements, PostgreSQL will be recommended to use for a future implementation.
-
-\newpage
-\subsection{Storage requirements}
-As the data comes in as events, there is never a need to update old records in the database, as can be seen in Figure~\ref{fig:storage}. This structure is slightly harder to do complicated 
-queries (Listing~\ref{fetch_example}) towards, as you always have to take the timestamp of the event into consideration. 
-But it also makes it possible to configure the DBMS not to lock tables during inserts, which will increase the speed of insertion of the database.
-The lack of locks will be especially useful if it is later
-decided to use a DDBMS for an eventual requirement of scaling up the system in the future.
-As everything is logged with timestamps it is particularly suitable for business intelligence systems and the likes of such, as you can query
-time ranges in the past as well as the current state of the system to gather analytics and make comparisons.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.6]{system_dbms}}
-    \caption{Flow of data}
-    \label{fig:storage}
-\end{figure}
-
-The only time inserts occur is during the time when the customised Lager~\cite{LAGER} backend or later added data sources sends events (1, see Figure~\ref{fig:storage}) and the only times reads are performed
-is when a manual query (2, see Figure~\ref{fig:storage}) is executed or when an eventual business intelligence system or an automated analyse system (3, see Figure~\ref{fig:storage}) that automatically queries the system is used.
-It will never be necessary to perform updates and no regular deletes will have to be performed, as explained in Section~\ref{REQUIREMENTS}.
-
-The actual database size should not be a problem as the servers are usually equipped with a few terabytes of disk space.
-An average insertion of Kivra is estimated to around 200 bytes by looking at meta data formed by data currently in the Riak cluster.
-This means that $5*10{^9}$ insertions would fit on a disk with 1TB of space.
-A rough estimation of an average of 5000 insertions per hour means that a 1TB would last for more than 113 years, see Figure~\ref{fig:storage2}.
-
-\begin{figure}[h!]
-\begin{framed}
-\begin{enumerate}
-\item $1TB/200B \thickapprox 5*10^{9}$ insertions
-\item $5*10^{9}/5000/24/365 \thickapprox $ 113.5 years
-\end{enumerate}
-\end{framed}
-\caption{Estimation of disk space usage}
-\label{fig:storage2}
-\end{figure}
-
-\FloatBarrier
-\newgeometry{top=2.0cm}
-\newpage
-\section{Resulting system}
-The result of an implementation of the suggested solution of this paper would lead to a core system that would have a data flow as described in Figure~\ref{fig:result}.
-In this flow, when data of analytical interest would be inserted into the Riak cluster it would also be sent to Lager which would insert it into a RDBMS dedicated for 
-analytical data, instead of fetching data in Riak after it has been inserted as was it done in the previous systems data flow that can be seen in Figure~\ref{fig:before} and explained in Section 3.2.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{system_after}}
-    \caption{Resulting system}
-    \label{fig:result}
-\end{figure}
-
-What is not shown in Figure~\ref{fig:result}, is that separate data sources for the analytical database can be used. These can connect with different protocols and with differently 
-structured data to an API that will be receiving data for the database as described in Section 5.2.
-
-\subsection{Scalability}
-\label{SCALABILITYRESULT}
-If an unoptimised PostgreSQL database on an average consumer computer is used and with the assumption that little to no data transfer latency will occur during insertion of data the insertion rate can
-be increased by a factor of at least 100 if following the upper boundary of the estimation (Section 5) of 10000 insertions per hour.
-As can be seen in the load test in 5.3.4, PostgreSQL can handle a little bit over
-1 million insertions per hour. The expected data growth rate is 5\% per month, with this growth the same setup could be used for almost 8 years, see Figure~\ref{fig:scalability}.
-
-\begin{figure}[h!]
-\begin{framed}
-\begin{enumerate}
-\item $10000*1.05^{94} \thickapprox $ 981282 insertions/hour 
-\item $94/12 \thickapprox $ 7.83 years 
-\end{enumerate}
-\end{framed}
-\caption{Estimation of data transfer scaling}
-\label{fig:scalability}
-\end{figure}
-
-In the unlikelihood that the same setup with the same hardware and PostgreSQL version would be used in 8 years the PostgreSQL database could easily be sharded, clustered or optimised to reach a lot higher
-insertion rates.
-As all he data is independent and has no defined relations to other data, clustering with a pool of independent databases could be used without any requirement of the databases to synchronise with each
-other, this would mean that you can linearly add 1000000 insertions/hour for each database node that you add to the cluster.
-
-\subsection{Practical Example}
-\subsubsection{Storing an Event}
-After an implementation of a solution like this at Kivra the following would be the flow of a typical change triggered by a user. 
-The system would update the send request bucket of Riak in Figure~\ref{fig:practical}, but changes to other buckets would look exactly the same.
-
-Figure~\ref{fig:practical} is a practical example of what would happen if a user updated his or her preferences of which tenants it wishes to receive documents from.
-
-\begin{figure}[h!]
-    \vspace{0.25in}
-    \centerline{\includegraphics[scale=0.7]{practical_flow}}
-    \caption{Practical example of a send request changing}
-    \label{fig:practical}
-\end{figure}
-
-\newpage
-\subsubsection{Querying Data}
-To query data stored in the analytical database a query that takes the timestamps into consideration has to be performed. The query in listing~\ref{fetch_example} is an example of how to query which 
-tenants all users wish and do not wish to receive documents from. Queries for other data sets would look similar.
-
-\begin{lstlisting}[label=fetch_example,caption=Example query]
-SELECT SRQ.user_id, SRQ.tenant_id, SRQ.timestamp, SRQ.accepted 
-FROM sendrequest SRQ 
-INNER JOIN 
-  (SELECT user_id, tenant_id, max(timestamp) as latest
-   FROM sendrequest GROUP BY user_id, tenant_id) SRQ1 
-ON SRQ.user_id = SRQ1.user_id 
-AND SRQ.tenant_id = SRQ1.tenant_id 
-AND SRQ.timestamp = SRQ1.latest;
-\end{lstlisting}
-
-The query in listing~\ref{fetch_example} doesn't take into consideration if the same name and tenant combination exists with the same timestamp. 
-It is however highly unlikely that it would happen as the timestamps are stored with a micro second precision. 
-To fully avoid that event, add an auto incremented ID column to the table and only consider the highest ID in the case of a clash.
-
-\newpage
-\section{Discussion}
-\subsection{Event-driven data collection}
-The suggested solution introduced by this paper was named event-driven data collection. The name has been used before for slightly similar purposes with the first recorded case 
-being in 1974~\cite{EVENTDRIVEN}.
-It has also been used in a different manner in quite a few papers on the subject wireless sensor networks~\cite{EVENTDRIVEN2} and some regarding the collection of natural
-ionosphere scintillations~\cite{EVENTDRIVEN3}.
-
-\subsection{Flat file database}
-A flat file database was not compared in this paper because of the precondition of having pre-built APIs and to give easier access to querying and aggregation of data.
-A flat file database can also be complicated to scale out
-to several servers for an eventual future requirement.
-
-\newpage
-\newgeometry{top=4.0cm}
 \section{Summary}
 \subsection{Conclusions}
-The system that has been designed in this paper can be used as an effective tool for logging and analysing events.
-This form of approach is especially valuable if the main system is using some form of MapReduce
-database which can be hard to do ongoing analysis of aggregated data on.
-If the main system is already using a SQL database as its primary form of information storage this type of solution would be quite inefficient
-due to duplication of data and possibly having to serve another DBMS\@.
-In the benchmarking of the different database management systems the standard version without modifications have been used, with modifications
-and optimisations the outcome would have been very different.
-
-The unorthodox design of the database schema in this paper provides a full historical log of events which makes it possible to compare data from different time ranges, 
-which can be very valuable when analysing data from business systems to for example detect trends and tendencies within the user base.
-
-Another contribution of this work is the use of a logging system to effectively put in an analytical systems 
-to an already existing backend with minimal impact on the performance of the existing system.
 
 \subsection{Future work}
-To finalise the architecture of this system, APIs and web service structures will have to be researched or developed to be put on top of the database.
-For larger flows of data other solutions would have
-to be considered or alternatively the used DBMS has to be heavily optimised by for example removing insertion locks. Scaling out to a clustered database could also be an alternative.
 
-\newpage
-\newgeometry{top=2.0cm}
-\begin{appendices}
-\section{Benchmarks}
-\label{appendix:a}
-\begin{lstlisting}[label=some-code,caption=PostgreSQL insertion test]
-1 000 000 insertions
-$ time psql --quiet -f test4.sql
-psql --quiet -f test4.sql
-40.42s user 23.08s system 2% cpu 31:05.03 total
-
-100 000 insertions
-$ time psql --quiet -f test.sql
-psql --quiet -f test.sql
-4.32s user 2.38s system 2% cpu 5:02.94 total
-
-10 000 insertions
-$ time psql --quiet -f test2.sql
-psql --quiet -f test2.sql
-0.51s user 0.22s system 2% cpu 30.172 total
-
-1000 insertions
-$ time psql --quiet -f test3.sql
-psql --quiet -f test3.sql 
-0.11s user 0.02s system 4% cpu 3.077 total
-\end{lstlisting}
-\begin{lstlisting}[label=some-code,caption=MySQL insertion test]
-1 000 000 insertions
-$ time mysql -u root -p********** testdb < test4.sql
-mysql -u root -p********** testdb <   
-39.78s user 23.36s system 1% cpu 1:36:56.31 total
-
-100 000 insertions
-$ time mysql -u root -p********** testdb < test.sql
-mysql -u root -p********** testdb < 
-3.88s user 2.83s system 1% cpu 9:22.67 total
-
-10 000 insertions
-$ time mysql -u root -p********** testdb < test2.sql 
-mysql -u root -p********** testdb < 
-0.31s user 0.39s system 1% cpu 53.174 total
-
-1000 insertions
-$ time mysql -u root -p********** testdb < test3.sql 
-mysql -u root -p********** testdb < 
-0.08s user 0.00s system 1% cpu 5.318 total
-\end{lstlisting}
-\end{appendices}
-\restoregeometry
+Stuff to write about:
+Modular design, every piece should be interchangable
+LDAP - why it was used as standard AUTH
 

### 2015-09-21 tex/spec.tex
Added some introduction

@@
 
-\title{\textbf{Specification: Providing Access Control in Content Distribution Networks}}
+\title{\textbf{Specification: Providing an Access Control Layer in Content Distribution Networks}}
 \author{Lukas Klingsbo}
@@
 The goal of this project is to explore, and creating, the possibility to have 
-access control in a content distribution network.
+an local access control layer for files in a content distribution network.
+This will make it easier to have control over the security settings of the 
+projects and will also make it possible to enforce local (as in not on the CDN) 
+auth and audit logging.
 
@@
 
+It should also be possible to create so called views and projects which a group 
+of files are added to, which then have their own security settings.
+
 \section{Methods}
 \begin{itemize}
-  \item The prototype/system will be developed in Scala. 
+  \item The prototype/system will be developed in Scala and React. 
   \item A literature study will be done once the specification is accepted. 
@@
 \section{References}
-One possibility to develop a solution from:
+One possibility to connect the backend to:
 

### 2015-09-23 tex/spec.tex
Added details to specification as requested

@@
 
-\title{\textbf{Specification: Providing an Access Control Layer in Content Distribution Networks}}
+\title{\textbf{Specification: Designing a Virtual Security Layer for Cloud Content}}
 \author{Lukas Klingsbo}
@@
 This work will be done at the company Uprise (Dragarbrunnsgatan 36C, Uppsala). 
-The goal of this project is to explore, and creating, the possibility to have 
-an local access control layer for files in a content distribution network.
-This will make it easier to have control over the security settings of the 
-projects and will also make it possible to enforce local (as in not on the CDN) 
-auth and audit logging.
+The goal of this project is to design and develop a local access control layer 
+for files, projects and views in a content distribution network.
+This should make handling of content safer and make it easier to have control 
+over the security settings of projects and groups of files and will also make 
+it possible to enforce local (as in not on the CDN) auth and audit logging.
 
 \section{Description}
-The project should research and possibly develop a system where it is possible 
-to set different types of access to different files in a CDN. 
+The project should design and possibly develop a system where it is possible 
+to set different types of access to different clusters of virtual files. 
 For example, the access control could be divided into three groups:
@@
 \end{itemize}
+The security rules put in place should not be hard coded, but possible to update 
+and extend. When a file is set to a belong to one of these groups it should not be static, 
+it should be possible to, at a later stage, assign it to another group.
 
-When a file is set to a belong to one of these groups it should not be static, 
-it should be possible to, at a later stage, assign the file to another group.
+A virtual file, as described in the beginning of the description, is a form of pointer 
+connected to some meta-data that points at an actual file in the filesystem or remote 
+service. One ``physical'' file can be referenced by many virtual files.
 
@@
 of files are added to, which then have their own security settings.
+A file can be a part of several views that then in their turn have different 
+security settings. The system should then be able to determine what the correct 
+security setting for the actual file is.
 
+For example, which security rules should take precedence if a file, lets call it 
+``Test.jpg'', is uploaded to a CDN once, but is contained in two views or projects 
+with different security settings. If then Test.jpg is changed in one of the projects, 
+but is not supposed to be changed in the other, how is that handled and how is 
+it determined when a change should propagate through the whole system or when it 
+should just create a new version of it in the current virtual view or project.
+\vspace{12pt}
+\\
+More things to examine:
+\begin{itemize}
+    \item A way to dynamically switch a single or a group of files security settings has 
+to be examined as the storage method of the files usually will be distributed.
+    \item To make sure you are handling the correct file previews has to be shown in the system.
+    \item Access to the system should be configured in a modular sense, one module could be 
+LDAP.
+\end{itemize}
+
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.5]{overview.pdf}}
+    \caption{System Overview}
+    \label{fig:overview}
+\end{figure}
+If looking at Figure~\ref{fig:overview}, the component that this work is supposed to specifically 
+design is the Security Layer in the middle, everything else should be interchangeable. The persistent 
+storage will most likely be a database which keeps track of the files and security settings. 
+The CDN could be any service with a rougher security system which needs to be more 
+fine grained for assets. The LDAP server in the figure can be any form of authentication service so that 
+wont need to be handled in the security layer. The React front-end is for handling files and 
+security settings in the back-end, it should be possible to write any front-end for it.
+
+\newpage
 \section{Methods}
 \begin{itemize}
+  \item The design will be well evaluated in the sense of security and technical usability.
   \item The prototype/system will be developed in Scala and React. 
@@
 
+\newpage
 \section{Time Plan}
@@
 \section{References}
-One possibility to connect the backend to:
+    One possibility to connect the security layer to:
 

### 2015-09-29 tex/paper.tex
Added content about copy-on-write

@@
 
-\title{\textbf{Providing an Access Control Layer for Content Distribution Networks}}
+\title{\textbf{Designing a Virtual Security Layer for Cloud Content}}
 \author{Lukas Klingsbo}
@@
 
-\subsubsection{JSON-RPC}
+\subsubsection{REST}
 Over http? WebSockets? 
@@
 request is made, the network serves content from the server closest to the requester.
+
+\section{Related Work}
+\subsection{Copy-on-Write}
+This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach kernel~\cite{COPYONWRITE}. 
+Today Copy-on-Write is used in everything from file systems~\cite{BTRFS} to desktop compositors~\cite{COMPOSITOR}.
+
+Its principle is that when processes share data in between each other, the data is not copied until one of the processes 
+does changes to it. This is an optimisation as the processes does not have to send all of the related data that is in memory, 
+rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure can be built up, 
+but it is possible to solve that structure~\cite{COPYONWRITE2}.
 
@@
 
+\section{Model}
+\subsection{Entities}
+%\subsubsection*{View} for not having section numbers
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.8]{relation.pdf}}
+    \caption{Entity Relationships}
+    \label{fig:relation}
+\end{figure}
+
+\subsubsection{View}
+A view is a group of files which can be changed independently of the view's parent.
+A view can have subviews which are either automatically dependant on its parent and 
+updated when the parent is updated or optionally updated or forked from its parent 
+once their content diverge.
+
+\subsubsection{Project}
+A project is just a name for the view which is a root of a tree of views.
+The system can contain several projects and their trees are completely disjoint.
+
+\subsubsection{File}
+A file refers to an actual physical file in the file system.
+
+\subsubsection{Content}
+Content is an asset contained in a view, it is a virtual file.
+It could be described as a link between the view and the real file. 
+It can be for example an image, video or binary blob combined with meta-data.
+
+\subsubsection{User}
+A user is the structure that handles people who has been granted access to the system.
+Access to the system is handled by a separate service, like LDAP.
+
+\section{Copy-on-Write}
+
 \section{Methods for determining\\implementation details}
-This chapter introduces the different methods used to determine how the new system should be implemented, which DBMS it should use and how the estimation of long term scaling was done.
+This chapter introduces the different methods used to determine how the new system should be implemented, 
+which DBMS it should use and how the estimation of long term scaling was done.
 
@@
 \section{Resulting system}
+\subsection{REST Endpoints}
+For the frontend to communicate with the backend REST is used, the following endpoints were configured:
+
+\begin{itemize}
+  \item projects
+      \subitem GET - list all projects
+      \subitem POST - create new project
+  \item projects/\{id\}
+      \subitem GET - get specific project
+      \subitem PUT - update existing project
+      \subitem DELETE - delete existing project
+  \item projects/\{id\}/content
+      \subitem GET - list all content in a specific project
+      \subitem POST - create new content in a specific project 
+  \item projects/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific project
+      \subitem PUT - update existing content in a specific project
+      \subitem DELETE - delete existing content in a specific project
+
+  \item projects/\{id\}/views
+      \subitem GET - list all views in a specific project
+      \subitem POST - create new view in a specific project 
+  \item projects/\{id\}/views/\{id\}
+      \subitem GET - get specific view in a specific project
+      \subitem PUT - update existing view
+      \subitem DELETE - delete existing view
+  \item projects/\{id\}/views/\{id\}/content
+      \subitem GET - list all content in a specific view
+      \subitem POST - create new content in a specific view 
+  \item projects/\{id\}/views/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific view
+      \subitem PUT - update existing content in a specific view
+      \subitem DELETE - delete existing content in a specific view
+\end{itemize}
+
 \subsection{Scalability}

### 2015-09-29 tex/spec.tex
Added content about copy-on-write

@@
 The goal of this project is to design and develop a local access control layer 
-for files, projects and views in a content distribution network.
+for files and views in a content distribution network.
 This should make handling of content safer and make it easier to have control 
-over the security settings of projects and groups of files and will also make 
+over the security settings of groups of files and will also make 
 it possible to enforce local (as in not on the CDN) auth and audit logging.
@@
 
-It should also be possible to create so called views and projects which a group 
+It should also be possible to create so called views which a group 
 of files are added to, which then have their own security settings.
+A view is a subset of files, and a view can be a view of another view and then 
+only contain files which is in contained in the parent view. 
 A file can be a part of several views that then in their turn have different 
 security settings. The system should then be able to determine what the correct 
-security setting for the actual file is.
+security setting for the actual file is or if it needs to create several versions 
+of the file.
 
 For example, which security rules should take precedence if a file, lets call it 
-``Test.jpg'', is uploaded to a CDN once, but is contained in two views or projects 
-with different security settings. If then Test.jpg is changed in one of the projects, 
+``Test.jpg'', is uploaded to a CDN once, but is contained in two views 
+with different security settings. If then Test.jpg is changed in one of the views, 
 but is not supposed to be changed in the other, how is that handled and how is 
 it determined when a change should propagate through the whole system or when it 
-should just create a new version of it in the current virtual view or project.
+should just create a new version of it in the current virtual view.
 \vspace{12pt}
@@
 \begin{itemize}
-  \item The design will be well evaluated in the sense of security and technical usability.
-  \item The prototype/system will be developed in Scala and React. 
-  \item A literature study will be done once the specification is accepted. 
-  \item Results will be evaluated from a security and scalability point of view. 
-  \item A work station and desk space will be provided by Uprise.
+  \item The design will be well evaluated in the sense of security and technical usability
+  \item The prototype/system will be developed in Scala and React
+  \item A literature study will be done to find more relevant litterature once the specification is accepted 
+  \item The proof should be an informal paper proof or if time is sufficient a proper proof should be done in a model checker
+  \item Results from the informal proof will be evaluated and documented
+  \item The implementation will be evaluated with existing security tools, it needs to be researched which these tools will be
+      \item Results from these tools and that are relevant to the report will be added in an appendix
+  \item Everything relevant in the work process, from problems that occured to final results, will be documented in the report
+  \item The code will have its own documentation separated from the report 
+  \item The prototype/system will be released under an open source license
+  \item A work station and desk space will be provided by Uprise
 \end{itemize}
@@
 quite concise.
-
 \newpage
@@
   subject and determine if anything is useful for this project
-  \item 4 weeks - Implement a prototype
-  \item 1 week - Scalability testing
-  \item 3 weeks - Security testing
+  \item 3 weeks - Define system mechanism and model
+  \item 3 weeks - Analyse system mechanism and model (this item will overlap with the definition)
+  \item 3 weeks - Informal proof (or formal if time is sufficient) of the systems mechanism and model
+  \item 8 weeks - Implement a prototype/system
+      \subitem 1 week - defining and developing backend API
+      \subitem 2 weeks - define a basic set of extendable security rules
+      \subitem <1 week - investigate and implement "physical" file handler
+      \subitem <1 week - implement simple persistant storage for meta data
+      \subitem 1-2 weeks - implement authentication interface
+      \subitem <1 week - implement CDN interface
+      \subitem 2 weeks - implement user interface
+  \item 1 weeks - Test security of implementation, if only a prototype is developed the security should only be analysed in theory
 \end{itemize}
+
+The interfaces does not have to have anything real connected to them, it just needs 
+to be possible to hook up services to them.
 

### 2015-10-08 tex/paper.tex
Added section about REST

@@
 * Users
-
+\newpage 
 \section{Model}
+\newpage 
 \subsection{Entities}
@@
 
-\subsubsection{View}
-A view is a group of files which can be changed independently of the view's parent.
-A view can have subviews which are either automatically dependant on its parent and 
-updated when the parent is updated or optionally updated or forked from its parent 
-once their content diverge.
+\subsubsection{Content}
+Content is an asset contained in a project and/or view, it is a form of virtual file.
+It could be described as a link between the view or project and the real file. 
+    The content can be for example an image, video or binary blob, combined with meta-data.
 
 \subsubsection{Project}
-A project is just a name for the view which is a root of a tree of views.
-The system can contain several projects and their trees are completely disjoint.
+A project is what is created to contain all content related to a real project. Files can be changed within a project and 
+the system can contain several projects and their virtual content are completely disjoint.
+
+\subsubsection{View}
+A view is a subset of content, in a project, from the state which they were in when the view was created. A project can 
+contain many views. A view can have content which either is automatically dependant on its parent project and updated 
+when the parent is updated or stay in the state which it was in when the view was created. This results in that the content 
+within a view can not be updated independently from its project.
+
+\subsubsection{Folder}
+A folder is a virtual folder within the projects which contains content.
 
@@
 
-\subsubsection{Content}
-Content is an asset contained in a view, it is a virtual file.
-It could be described as a link between the view and the real file. 
-It can be for example an image, video or binary blob combined with meta-data.
-
 \subsubsection{User}
-A user is the structure that handles people who has been granted access to the system.
+A user is the structure that handles people who have been granted access to the system.
 Access to the system is handled by a separate service, like LDAP.
 
+\newpage 
 \section{Copy-on-Write}
@@
 \section{Resulting system}
-\subsection{REST Endpoints}
+\subsection{Persistent storage}
+\subsubsection{MongoDB}
+
+\subsection{API}
+REST was chosen as the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON 
+format which is used by RESTful services~\cite{JSON}.
+
+\subsubsection{REST Endpoints}
 For the frontend to communicate with the backend REST is used, the following endpoints were configured:
@@
       \subitem DELETE - delete existing content in a specific view
+
+  \item projects/\{id\}/folders
+      \subitem GET - list all folders in a specific project
+      \subitem POST - create new folder in a specific project 
+  \item projects/\{id\}/folder/\{id\}
+      \subitem GET - get specific folder in a specific project
+      \subitem PUT - update existing folder
+      \subitem DELETE - delete existing folder
+  \item projects/\{id\}/folder/\{id\}/content
+      \subitem GET - list all content in a specific folder
+      \subitem POST - create new content in a specific folder 
+  \item projects/\{id\}/folder/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific folder
+      \subitem PUT - update existing content in a specific folder
+      \subitem DELETE - delete existing content in a specific folder
 \end{itemize}

### 2015-10-14 tex/paper.tex
Added JPF results

@@
 \subsubsection{REST}
-Over http? WebSockets? 
-\subsubsection{TODO: Insert persistent storage here}
+\subsubsection{MongoDB}
     TODO: Write down related terminology, if any
 \subsection{Abbreviations}
+
+\subsubsection{JPF}
+Java Path Finder - It was developed by NASA and in 2005 they released it under an open source licence, which made more 
+people contribute to the project. JPF is usually used for doing model checking of concurrent programs to easily find 
+for example race conditions and dead locks.
 
@@
 This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach kernel~\cite{COPYONWRITE}.
-Today Copy-on-Write is used in everything from file systems~\cite{BTRFS} to desktop compositors~\cite{COMPOSITOR}.
+
+Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, snapshot functionality in both file 
+systems~\cite{FSCOW} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
+types in programming languages\cite{LANGCOW}.
 
@@
 \section{Model}
-\newpage 
 \subsection{Entities}
@@
 
+\subsection{JPF}
+Java path finder was used to show that the model and plan of how to build the system was sound. The model was built in Java 
+with the objective of being as reduced and simple as possible, without loosing any of the cases that needs to be covered 
+by the model checker.
+
+Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type. 
+Each client was represented by a thread and each action taken by the client was randomised. The id hashes which MongoDB is 
+using for each entity was imported from the mongo-java-driver-2.13.3 and each object had its own id, generated in the same 
+fashion as the real implementation is using, randomly generated by the ObjectId class to minimise collisions that is. 
+Further more no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.
+
+JPF checked each permutation of states that the threads can end up in, the result of the run can be seen in Listing~\ref{JPFRESULT}.
+
+\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
+elapsed time:       14:26:53
+states:             new=160853259,
+                    visited=451102505,
+                    backtracked=611955764,
+                    end=21640
+search:             maxDepth=380,
+                    constraints=0
+choice generators:  thread=160853255 
+                    (signal=0,
+                    lock=3603938,
+                    sharedRef=146989208,
+                    threadApi=3,
+                    reschedule=10260106), 
+                    data=0
+
+heap:               new=676056850,
+                    released=435060996,
+                    maxLive=655,
+                    gcCycles=523950061
+
+instructions:       11917045758
+max memory:         6256MB
+loaded code:        classes=111,
+                    methods=2179
+\end{lstlisting}
+
 \newpage 
 \section{Copy-on-Write}
+As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
+clients are working on the same data set at the same time. Such problems could be race conditions,
 
@@
 \subsubsection{MongoDB}
+MongoDB was chosen as the persistent storage because of its internal storage format called BSON, which is very similar to JSON 
+which the API is using. As the formats are similar, the process of marshalling and unmarshalling becomes quite easy between 
+the core code, MongoDB instance and REST interface. The second reason was that if the system needs to scale in the future it 
+is very easy to distribute MongoDB and if needed the system can easily be migrated to Reactive Mongo, which is asynchronous and 
+non-blocking and can therefore scale even further~\cite{REACTIVEMONGO}.
 

### 2015-10-15 tex/paper.tex
Changed snapshot definition

@@
 For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
-access control on content and groups of content in the form of views. A system was developed to make the underlying 
+access control on content and groups of content in the form of snapshots. A system was developed to make the underlying 
 theory work in practice.
@@
 \subsubsection{CDN}
-Content Distribution Network - Replicates content to several servers, usually spread out geographically. Once a 
+Content Distribution/Delivery Network - Replicates content to several servers, usually spread out geographically. Once a 
 request is made, the network serves content from the server closest to the requester.
@@
 \subsection{The current system}
-Maybe write about battlebinary?
+Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files, mostly images, to content 
+delivery networks. The current system does not make use out of the security features that the CDN's are offering, instead it 
+uses a form of security by obscurity. When a file is uploaded to a CDN it is open for the public, but its filename is 
+composed out of its original filename concatenated with a part of the MD5 hash of the content of the file, which makes it an 
+extremely hard process to access the file on the CDN without access to the original file or a reference to the URI.
+
+In the current system you can only upload a file once as there will be a collision in the upload otherwise, as the old and 
+the new file will have the same MD5 hash.
 
 \subsection{Problem description}
-Having
+As the current system does not offer proper security measurements, is lacking a lot of features that is needed and does not 
+scale very well, a new system should be developed. This work is about examining a way of implementing Copy-on-Write in a 
+high level system, which should solve the scalability problem and makes it possible to implement the wanted features, like 
+snapshots and concurrent modifications of content.
 
+What to write about
 Battle Binary
+Security features of CDN's
 
@@
 * Security layers
-* Views
+* snapshots
 * Virtual file structure
@@
 \subsection{Entities}
-%\subsubsection*{View} for not having section numbers
+%\subsubsection*{snapshot} for not having section numbers
 \begin{figure}[htp] \centering{
@@
 \subsubsection{Content}
-Content is an asset contained in a project and/or view, it is a form of virtual file.
-It could be described as a link between the view or project and the real file. 
+Content is an asset contained in a project and/or snapshot, it is a form of virtual file.
+It could be described as a link between the snapshot or project and the real file. 
 The content can be for example an image, video or binary blob, combined with meta-data.
@@
 the system can contain several projects and their virtual content are completely disjoint.
+snapshot
+\subsubsection{Snapshot}
+A snapshot is a read-only container from the state which the container the was in when the snapshot was created. A snapshot can 
+have content which either is automatically dependant on the container which it was created from and updated when that container 
+is updated or stay in the state which it was in when the snapshot was created. This results in that the content within a snapshot 
+can not be updated independently from the container which it was created from.
 
-\subsubsection{View}
-A view is a subset of content, in a project, from the state which they were in when the view was created. A project can 
-contain many views. A view can have content which either is automatically dependant on its parent project and updated 
-when the parent is updated or stay in the state which it was in when the view was created. This results in that the content 
-within a view can not be updated independently from its project.
-
-\subsubsection{Folder}
-A folder is a virtual folder within the projects which contains content.
+\subsubsection{Container}
+A container is a virtual folder within the projects which contains content and other containers.
 
@@
 with the objective of being as reduced and simple as possible, without loosing any of the cases that needs to be covered 
-by the model checker.
+by the model checker. As the users are mainly going to be handled by external systems they were not included in the model.
 
@@
 Further more no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.
+
+ConcurrentHashMap had to be used in favour of the normal HashMap as the normal HashMaps can't be iterated over concurrently.
 
@@
 
-  \item projects/\{id\}/views
-      \subitem GET - list all views in a specific project
-      \subitem POST - create new view in a specific project 
-  \item projects/\{id\}/views/\{id\}
-      \subitem GET - get specific view in a specific project
-      \subitem PUT - update existing view
-      \subitem DELETE - delete existing view
-  \item projects/\{id\}/views/\{id\}/content
-      \subitem GET - list all content in a specific view
-      \subitem POST - create new content in a specific view 
-  \item projects/\{id\}/views/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific view
-      \subitem PUT - update existing content in a specific view
-      \subitem DELETE - delete existing content in a specific view
+  \item projects/\{id\}/snapshots
+      \subitem GET - list all snapshots in a specific project
+      \subitem POST - create new snapshot in a specific project 
+  \item projects/\{id\}/snapshots/\{id\}
+      \subitem GET - get specific snapshot in a specific project
+      \subitem PUT - update existing snapshot
+      \subitem DELETE - delete existing snapshot
+  \item projects/\{id\}/snapshots/\{id\}/content
+      \subitem GET - list all content in a specific snapshot
+      \subitem POST - create new content in a specific snapshot 
+  \item projects/\{id\}/snapshots/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific snapshot
+      \subitem PUT - update existing content in a specific snapshot
+      \subitem DELETE - delete existing content in a specific snapshot
 
-  \item projects/\{id\}/folders
-      \subitem GET - list all folders in a specific project
-      \subitem POST - create new folder in a specific project 
-  \item projects/\{id\}/folder/\{id\}
-      \subitem GET - get specific folder in a specific project
-      \subitem PUT - update existing folder
-      \subitem DELETE - delete existing folder
-  \item projects/\{id\}/folder/\{id\}/content
-      \subitem GET - list all content in a specific folder
-      \subitem POST - create new content in a specific folder 
-  \item projects/\{id\}/folder/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific folder
-      \subitem PUT - update existing content in a specific folder
-      \subitem DELETE - delete existing content in a specific folder
+  \item projects/\{id\}/containers
+      \subitem GET - list all containers in a specific project
+      \subitem POST - create new container in a specific project 
+  \item projects/\{id\}/containers/\{id\}
+      \subitem GET - get specific container in a specific project
+      \subitem PUT - update existing container
+      \subitem DELETE - delete existing container
+  \item projects/\{id\}/containers/\{id\}/content
+      \subitem GET - list all content in a specific container
+      \subitem POST - create new content in a specific container 
+  \item projects/\{id\}/containers/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific container
+      \subitem PUT - update existing content in a specific container
+      \subitem DELETE - delete existing content in a specific container
 \end{itemize}

### 2015-10-27 tex/paper.tex
Added dirty presentation folder

@@
 For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
-access control on content and groups of content in the form of snapshots. A system was developed to make the underlying 
-theory work in practice.
+access control on content and groups of content in the form of folders and snapshots. A system was developed to make 
+the underlying theory work in practice.
 
@@
 DOM with the resulting virtual DOM after the state change and then only update the browser's DOM according to the 
-difference between the virtual DOMs~\cite{REACT}.
+delta between the virtual DOMs~\cite{REACT}. React can be seen as the system for handling views in front-ends 
+implementing a MVC (Model-View-Controller) architecture.
 
-\subsubsection{Flux}
+\subsubsection{Reflux}
+Explain difference to MVC etc.
 
@@
 \subsubsection{REST}
+REST stands for representational state transfer, it is an architectural idea for writing stateless services. These 
+services usually use URIs to identify specific resources and HTTP to modify or query these resources~\cite{REST}. 
+In this work we use a RESTful API, which means that it is an API following the REST architecture.
+
 \subsubsection{MongoDB}
-    TODO: Write down related terminology, if any
+MongoDB is a document-oriented database which means that it does not have the concept of rows as normal relational 
+databases has. Instead each entity in the database is stored as a document which is not fixed to a predefined table 
+structure~\cite{MONGODB}. MongoDB lacks support for joins too improve its possibility to scale, which can be a big 
+down side to some applications containing the need for such logic.
+
 \subsection{Abbreviations}
@@
 Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, snapshot functionality in both file 
-systems~\cite{FSCOW} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
+systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
 types in programming languages\cite{LANGCOW}.
 
-Its principle is that when processes share data in between each other, the data is not copied until one of the processes 
-does changes to it. This is an optimisation as the processes does not have to send all of the related data that is in memory, 
-rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure can be built up, 
-but it is possible to solve that structure~\cite{COPYONWRITE2}.
+Its principle is that when processes or nodes share data in between each other, the data is not copied until one of the 
+processes does changes to it. This is an optimisation as the processes does not have to send or copy all of the related 
+data that is in memory, rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure 
+can be built up, but it is possible to solve that structure~\cite{COPYONWRITE2}.
+
+TODO: Polish and extend
 
@@
 scale very well, a new system should be developed. This work is about examining a way of implementing Copy-on-Write in a 
-high level system, which should solve the scalability problem and makes it possible to implement the wanted features, like 
-snapshots and concurrent modifications of content.
+high level system like this, which should solve the scalability problem and make it possible to implement wanted features
+like snapshots, cloning and concurrent modifications of content.
 
-What to write about
-Battle Binary
-Security features of CDN's
+%What to write about
+%Battle Binary
+%Security features of CDN's
+%What is needed by the logic, like what happens after two consecutive deletes
+%
+%What is needed
+%* Security layers
+%* snapshots
+%* Virtual file structure
+%* Versioning of content
+%* Multi project support
+%* Auth and audit logs
+%* Users
 
-What is needed
-* Security layers
-* snapshots
-* Virtual file structure
-* Versioning of content
-* Multi project support
-* Auth and audit logs
-* Users
+%Stuff to write about:
+%Modular design, every piece should be interchangeable
+%LDAP - why it was used as standard AUTH
+
+%Using copy-on-write as the sole update strategy has pros and cons. The upside is
+%that it is simple to guarantee operation atomicity, and data-structure integrity. The
+%downside is that performance relies on the ability to maintain large extents of free
+%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
+%destroying sequentiality. A good defragmentation algorithm is required; this is
+%described in Section 5
+
 \newpage 
@@
 \subsubsection{Content}
-Content is an asset contained in a project and/or snapshot, it is a form of virtual file.
-It could be described as a link between the snapshot or project and the real file. 
-The content can be for example an image, video or binary blob, combined with meta-data.
+Content is meta data about a file and is stored in a container, it is a form of virtual file.
+The content can refer to for example an image, video or binary blob.
 
 \subsubsection{Project}
-A project is what is created to contain all content related to a real project. Files can be changed within a project and 
-the system can contain several projects and their virtual content are completely disjoint.
-snapshot
-\subsubsection{Snapshot}
-A snapshot is a read-only container from the state which the container the was in when the snapshot was created. A snapshot can 
-have content which either is automatically dependant on the container which it was created from and updated when that container 
-is updated or stay in the state which it was in when the snapshot was created. This results in that the content within a snapshot 
-can not be updated independently from the container which it was created from.
+A project is what is created to contain all content and containers related to a real project. Files can be changed 
+within a project and the system can contain several projects and their virtual content are completely disjoint.
 
 \subsubsection{Container}
-A container is a virtual folder within the projects which contains content and other containers.
+A container is a virtual folder within a project which can contain content and other containers.
+
+\subsubsection{Snapshot}
+A snapshot is a read-only container from the state which the container the was in when the snapshot was created. 
+A snapshot can not be updated and can only be deleted from the root of the snapshot. Snapshots are by default stored 
+as siblings to the container which they were made from, but they can be contained by any container.
 
 \subsubsection{File}
-A file refers to an actual physical file in the file system.
+A file refers to an actual physical file. Files are stored in the database to make backup, deployment and migration easier.
 
@@
 Java path finder was used to show that the model and plan of how to build the system was sound. The model was built in Java 
-with the objective of being as reduced and simple as possible, without loosing any of the cases that needs to be covered 
+with the objective of being as reduced and simple as possible, without loosing any of the cases that needed to be covered 
 by the model checker. As the users are mainly going to be handled by external systems they were not included in the model.
@@
 fashion as the real implementation is using, randomly generated by the ObjectId class to minimise collisions that is. 
-Further more no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.
+Furthermore no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.
 
@@
 
+\newpage
 \begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
@@
 As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
-clients are working on the same data set at the same time. Such problems could be race conditions,
+clients are working on the same data set at the same time. Such problems could be race conditions and determining the 
+happened-before relation. In this work this problem is solved by implementing Copy-on-Write.
+
+\subsection{Comparison of Copy-on-Write system implementations}
+\subsubsection{BTRFS}
+Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do efficient writeable 
+snapshots and clones. It also supports cloning of subtrees without having to actually copy the whole subtree, 
+this is due to the Copy-on-Write effect. As several nodes in the tree can refer to the same node each node keeps 
+track of how many parents it has by a reference counter so that the node can be deallocated once the node does not 
+have any parents any more. The reference counter is not stored in the nodes themselves but rather in a separate data 
+structure so that a nodes counter can be modified without modifying the node itself and therefore eludes the 
+Copy-on-Write that would have to occur.
+
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.4]{newtree.png}}
+    \caption{Cloning mechanism of Btrfs~\cite{BTRFS}}
+    \label{fig:btrfs_tree}
+\end{figure}
+
+\subsubsection{Mach kernel}
+In the mid 80s when the development of the Mach kernel started, there was problems with that physically copying memory 
+was too slow. Too minimise the copying of memory, Copy-on-Write was implemented. It was implemented so that virtual 
+copy operations could be done and so that tasks could share read-write memory~\cite{MACH}.
+
+TODO: Insert more systems here
+
+\subsubsection{Perius}
+Perius is the implementation that was done to solve the problem at hand at Uprise. This implementation is far from as 
+efficient as the others in most aspects, but more efficient in some. As the implementation is built upon MongoDB as 
+persistent storage and not a pure tree structure, single nodes can be fetched in O(1) but when querying for subtrees 
+they need to be built first, which takes O(log(n)), where n is the number of nodes in the subtree.
+
+In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in 
+Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is instead stored in a 
+flat storage space, such operation would be too computationally expensive as trees would have to be merged when 
+collisions occur, due to the non-blocking nature of the application. Instead this implementation makes a full copy of 
+the meta-data of the tree, but still refers to the same binary files until they are changed, which results in the 
+creation of a new node.
 
@@
 \subsubsection{MongoDB}
-MongoDB was chosen as the persistent storage because of its internal storage format called BSON, which is very similar to JSON 
-which the API is using. As the formats are similar, the process of marshalling and unmarshalling becomes quite easy between 
-the core code, MongoDB instance and REST interface. The second reason was that if the system needs to scale in the future it 
-is very easy to distribute MongoDB and if needed the system can easily be migrated to Reactive Mongo, which is asynchronous and 
-non-blocking and can therefore scale even further~\cite{REACTIVEMONGO}.
+MongoDB was chosen as the persistent storage because of its quick lookups and because of its internal storage 
+format called BSON, which is very similar to JSON which the API is using. As the formats are similar, the process of 
+marshalling and unmarshalling becomes quite easy between the core code, MongoDB instance and REST interface. 
+The second reason was that if the system needs to scale in the future it is very easy to distribute MongoDB and if 
+needed the system can easily be migrated to Reactive Mongo, which is an asynchronous and 
+non-blocking driver for MongoDB and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
 
 \subsection{API}
-REST was chosen as the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON 
-format which is used by RESTful services~\cite{JSON}.
+REST was chosen as only basic CRUD operations needs to be performed and because the BSON format which is used in 
+MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is used by RESTful services~\cite{JSON}.
 
 \subsubsection{REST Endpoints}
-For the frontend to communicate with the backend REST is used, the following endpoints were configured:
+For the front-end to communicate with the back-end, a RESTful service is implemented.
+The following endpoints were configured:
 
@@
   \item projects/\{id\}/content
-      \subitem GET - list all content in a specific project
       \subitem POST - create new content in a specific project 
@@
   \item projects/\{id\}/snapshots
-      \subitem GET - list all snapshots in a specific project
       \subitem POST - create new snapshot in a specific project
-  \item projects/\{id\}/snapshots/\{id\}
-      \subitem GET - get specific snapshot in a specific project
-      \subitem PUT - update existing snapshot
-      \subitem DELETE - delete existing snapshot
-  \item projects/\{id\}/snapshots/\{id\}/content
-      \subitem GET - list all content in a specific snapshot
-      \subitem POST - create new content in a specific snapshot 
-  \item projects/\{id\}/snapshots/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific snapshot
-      \subitem PUT - update existing content in a specific snapshot
-      \subitem DELETE - delete existing content in a specific snapshot
 
   \item projects/\{id\}/containers
-      \subitem GET - list all containers in a specific project
       \subitem POST - create new container in a specific project 
@@
   \item projects/\{id\}/containers/\{id\}/content
-      \subitem GET - list all content in a specific container
       \subitem POST - create new content in a specific container 
@@
 \end{itemize}
+
+As can be seen several expected endpoints are missing, this is intentional as the operations missing can be performed in 
+a more efficient way. Such endpoint is for example $GET projects/\{id\}/containers$ as all containers exist in 
+$GET projects/\{id\}$ and the interface should present a file structure where both content and containers are shown.
 
@@
 
-Stuff to write about:
-Modular design, every piece should be interchangable
-LDAP - why it was used as standard AUTH
-
 \newpage

### 2015-11-12 tex/paper.tex
Added subsubsubsection functionality

@@
 \usepackage{fancyvrb}
+\usepackage{titlesec}
 
@@
 \usepackage{hyperref} % should always be the last package
+
+% subsubsubsection
+\titleclass{\subsubsubsection}{straight}[\subsection]
+
+\newcounter{subsubsubsection}[subsubsection]
+\renewcommand\thesubsubsubsection{\thesubsubsection.\arabic{subsubsubsection}}
+\renewcommand\theparagraph{\thesubsubsubsection.\arabic{paragraph}} % optional; useful if paragraphs are to be numbered
+
+\titleformat{\subsubsubsection}
+{\normalfont\normalsize\bfseries}{\thesubsubsubsection}{1em}{}
+\titlespacing*{\subsubsubsection}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+
+\makeatletter
+\renewcommand\paragraph{\@startsection{paragraph}{5}{\z@}%
+    {3.25ex \@plus1ex \@minus.2ex}%
+    {-1em}%
+{\normalfont\normalsize\bfseries}}
+\renewcommand\subparagraph{\@startsection{subparagraph}{6}{\parindent}%
+    {3.25ex \@plus1ex \@minus .2ex}%
+    {-1em}%
+{\normalfont\normalsize\bfseries}}
+\def\toclevel@subsubsubsection{4}
+\def\toclevel@paragraph{5}
+\def\toclevel@paragraph{6}
+\def\l@subsubsubsection{\@dottedtocline{4}{7em}{4em}}
+\def\l@paragraph{\@dottedtocline{5}{10em}{5em}}
+\def\l@subparagraph{\@dottedtocline{6}{14em}{6em}}
+\makeatother
+
+\setcounter{secnumdepth}{4}
+\setcounter{tocdepth}{4}
 
@@
 
-%What to write about
-%Battle Binary
-%Security features of CDN's
-%What is needed by the logic, like what happens after two consecutive deletes
-%
-%What is needed
-%* Security layers
-%* snapshots
-%* Virtual file structure
-%* Versioning of content
-%* Multi project support
-%* Auth and audit logs
-%* Users
-
-%Stuff to write about:
-%Modular design, every piece should be interchangeable
-%LDAP - why it was used as standard AUTH
-
-%Using copy-on-write as the sole update strategy has pros and cons. The upside is
-%that it is simple to guarantee operation atomicity, and data-structure integrity. The
-%downside is that performance relies on the ability to maintain large extents of free
-%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
-%destroying sequentiality. A good defragmentation algorithm is required; this is
-%described in Section 5
-
 \newpage 
@@
 TODO: Insert more systems here
-
-\subsubsection{Perius}
+\section{Implementation}
+\subsection{Perius}
 Perius is the implementation that was done to solve the problem at hand at Uprise. This implementation is far from as 
@@
 
-\section{Methods for determining\\implementation details}
+\subsection{Methods for determining\\implementation details}
 This chapter introduces the different methods used to determine how the new system should be implemented, 
@@
 
-\section{Security of the system}
-\subsection{Authorization}
-\subsection{Audit logs}
-
-\section{Resulting system}
-\subsection{Persistent storage}
-\subsubsection{MongoDB}
+\subsection{Resulting system}
+\subsubsection{Persistent storage}
+\subsubsubsection{MongoDB}
 MongoDB was chosen as the persistent storage because of its quick lookups and because of its internal storage 
@@
 
-\subsection{API}
+\subsubsubsection{API}
 REST was chosen as only basic CRUD operations needs to be performed and because the BSON format which is used in 
@@
 
-\subsection{Scalability}
+\subsection{Findings}
+\subsubsection{Scalability}
+
+\subsubsection{Security of the system}
+\paragraph{Authorization}
+\paragraph{Audit logs}
 
@@
 \end{document}
+
+%What to write about
+%Battle Binary
+%Security features of CDN's
+%What is needed by the logic, like what happens after two consecutive deletes
+%
+%What is needed
+%* Security layers
+%* snapshots
+%* Virtual file structure
+%* Versioning of content
+%* Multi project support
+%* Auth and audit logs
+%* Users
+
+%Stuff to write about:
+%Modular design, every piece should be interchangeable
+%LDAP - why it was used as standard AUTH
+
+%Using copy-on-write as the sole update strategy has pros and cons. The upside is
+%that it is simple to guarantee operation atomicity, and data-structure integrity. The
+%downside is that performance relies on the ability to maintain large extents of free
+%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
+%destroying sequentiality. A good defragmentation algorithm is required; this is
+%described in Section 5

### 2015-11-24 tex/paper.tex
Written some basic implementation introduction

@@
 
+The research question that this report answers is whether it is practically feasible to use Copy-on-Write 
+for a high-level system like the one that is implemented. The work also thoroughly compares different types of lazy 
+snapshots to figure out which is the most
+
 \newpage
@@
 \subsubsection{Reflux}
-Explain difference to MVC etc.
+Reflux~\cite{REFLUX} is an idea and a simple library of how to structure your application. It features a unidirectional
+dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example Flux~\cite{FLUX}, when using a 
+functional reactive programming style.
+
+\begin{figure}[htp] 
+    \centering
+    \includegraphics[scale=0.4]{reflux.png}
+    \caption{Reflux unidirectional dataflow}
+    \label{fig:reflux}
+\end{figure}
 
 \subsubsection{Scala} 
-Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared to Java it supports 
-most functional programming features at the same time as it supports object oriented programming~\cite{SCALA}.
+Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared 
+to Java it supports most functional programming features at the same time as it supports object oriented 
+programming~\cite{SCALA}.
 
-\subsubsection{REST}
-REST stands for representational state transfer, it is an architectural idea for writing stateless services. These 
-services usually use URIs to identify specific resources and HTTP to modify or query these resources~\cite{REST}. 
+\subsubsection{REST} REST stands for representational state transfer, it is an architectural 
+idea for writing stateless services. These services usually use URIs to identify specific resources and HTTP to 
+modify or query these resources~\cite{REST}.
+
 In this work we use a RESTful API, which means that it is an API following the REST architecture.
@@
 \subsection{About Uprise}
-Uprise is a company based in Uppsala, Sweden.
+Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming experiences, that means 
+that they are not focussed on the actual gameplay which other EA studios like DICE is.
+
+At the moment they are for example very involved with producing the Star Wars Battlefront game by making menu systems and 
+developing the companion app, Battlefront companion.
 
@@
 \section{Copy-on-Write}
+\label{sec:copy-on-write}
 As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
@@
 \subsection{Perius}
-Perius is the implementation that was done to solve the problem at hand at Uprise. This implementation is far from as 
-efficient as the others in most aspects, but more efficient in some. As the implementation is built upon MongoDB as 
-persistent storage and not a pure tree structure, single nodes can be fetched in O(1) but when querying for subtrees 
-they need to be built first, which takes O(log(n)), where n is the number of nodes in the subtree.
+Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
+back-end written in Scala and a front-end written in ES6 (Javascript), but they are both interchangeable. 
+The back-end has a REST API running, which is how the front-end communicates with the back-end.
+
+TODO: Picture of the front-end
+
+The service features a virtual file structure over the assets that has been stored, snapshots,
+security management of whole containers as well as individual files, audit and access logging, multi
+project support and a modular design for persistent storage.
+
+The front-end is written in ES6 with React and Reflux, the styling is done with the help of
+Bootstrap.
+
+\subsection{Snapshots}
+This implementation is far from as 
+efficient as the other Copy-on-Write systems described in Section~\ref{sec:copy-on-write} in most aspects, but more efficient 
+in some. As the implementation is built upon MongoDB as persistent storage and not a pure tree structure, single nodes 
+can be fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)), where n is the 
+number of nodes in the subtree.
 

### 2015-11-25 tex/paper.tex
Aligned text and started writing about snapshots

@@
 \section{Introduction}
-Developing larger projects containing static content usually involves using a Content Distribution Network 
-to be able to scale to a large user base. The commercial Content Distribution Networks are usually fairly easy to use, 
-the content that is to be used in a project is usually simply uploaded and then directly available to the public. 
-For secret content this can be a problem and that is what this thesis is about. This work examines ways of enforcing 
-access control on content and groups of content in the form of folders and snapshots. A system was developed to make 
-the underlying theory work in practice.
+Developing larger projects containing static content usually involves using a Content Distribution
+Network to be able to scale to a large user base. The commercial Content Distribution Networks are
+usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
+and then directly available to the public.  For secret content this can be a problem and that is
+what this thesis is about. This work examines ways of enforcing access control on content and groups
+of content in the form of folders and snapshots. A system was developed to make the underlying
+theory work in practice.
 
-The research question that this report answers is whether it is practically feasible to use Copy-on-Write 
-for a high-level system like the one that is implemented. The work also thoroughly compares different types of lazy 
-snapshots to figure out which is the most
+The research question that this report answers is whether it is practically feasible to use
+Copy-on-Write for a high-level system like the one that is implemented. The work also thoroughly
+compares different types of lazy snapshots to figure out which is the most
 
@@
 \subsubsection{React}
-React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and the browser's, 
-this makes it able to efficiently update dynamic web pages after a change of state through comparing the old virtual 
-DOM with the resulting virtual DOM after the state change and then only update the browser's DOM according to the 
-delta between the virtual DOMs~\cite{REACT}. React can be seen as the system for handling views in front-ends 
-implementing a MVC (Model-View-Controller) architecture.
+React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
+the browser's, this makes it able to efficiently update dynamic web pages after a change of state
+through comparing the old virtual DOM with the resulting virtual DOM after the state change and then
+only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}. React
+can be seen as the system for handling views in front-ends implementing a MVC
+(Model-View-Controller) architecture.
 
 \subsubsection{Reflux}
-Reflux~\cite{REFLUX} is an idea and a simple library of how to structure your application. It features a unidirectional
-dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example Flux~\cite{FLUX}, when using a 
-functional reactive programming style.
+Reflux~\cite{REFLUX} is an idea and a simple library of how to structure your application. It
+features a unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than
+for example Flux~\cite{FLUX}, when using a functional reactive programming style.
 
@@
 \subsubsection{Scala} 
-Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared 
-to Java it supports most functional programming features at the same time as it supports object oriented 
-programming~\cite{SCALA}.
+Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared to
+Java it supports most functional programming features at the same time as it supports object
+oriented programming~\cite{SCALA}.
 
-\subsubsection{REST} REST stands for representational state transfer, it is an architectural 
-idea for writing stateless services. These services usually use URIs to identify specific resources and HTTP to 
-modify or query these resources~\cite{REST}.
+\subsubsection{REST} 
+REST stands for representational state transfer, it is an architectural idea for writing stateless
+services. These services usually use URIs to identify specific resources and HTTP to modify or query
+these resources~\cite{REST}.
 
@@
 \subsubsection{MongoDB}
-MongoDB is a document-oriented database which means that it does not have the concept of rows as normal relational 
-databases has. Instead each entity in the database is stored as a document which is not fixed to a predefined table 
-structure~\cite{MONGODB}. MongoDB lacks support for joins too improve its possibility to scale, which can be a big 
-down side to some applications containing the need for such logic.
+MongoDB is a document-oriented database which means that it does not have the concept of rows as
+normal relational databases has. Instead each entity in the database is stored as a document which
+is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks support for joins too
+improve its possibility to scale, which can be a big down side to some applications containing the
+need for such logic.
 
@@
 \subsubsection{JPF}
-Java Path Finder - It was developed by NASA and in 2005 they released it under an open source licence, which made more 
-people contribute to the project. JPF is usually used for doing model checking of concurrent programs to easily find 
-for example race conditions and dead locks.
+Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
+licence, which made more people contribute to the project. JPF is usually used for doing model
+checking of concurrent programs to easily find for example race conditions and dead locks.
 
 \subsubsection{CDN}
-Content Distribution/Delivery Network - Replicates content to several servers, usually spread out geographically. Once a 
-request is made, the network serves content from the server closest to the requester.
+Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
+geographically. Once a request is made, the network serves content from the server closest to the
+requester.
 
@@
 \subsection{Copy-on-Write}
-This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach kernel~\cite{COPYONWRITE}.
+This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach
+kernel~\cite{COPYONWRITE}.
 
-Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, snapshot functionality in both file 
-systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW}, and as an optimisation technique for objects and 
-types in programming languages\cite{LANGCOW}.
+Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, for snapshot
+functionality in both file systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW},
+and as an optimisation technique for objects and types in programming languages\cite{LANGCOW}.
 
-Its principle is that when processes or nodes share data in between each other, the data is not copied until one of the 
-processes does changes to it. This is an optimisation as the processes does not have to send or copy all of the related 
-data that is in memory, rather they only have to send pointers to the data. After many Copy-on-Write's a complex structure 
-can be built up, but it is possible to solve that structure~\cite{COPYONWRITE2}.
+Its principle is that when processes or nodes share data in between each other, the data is not
+copied until one of the processes does changes to it. This is an optimisation as the processes does
+not have to send or copy all of the related data that is in memory, rather they only have to send
+pointers to the data. After many Copy-on-Write's a complex structure can be built up, but it is
+possible to solve that structure~\cite{COPYONWRITE2}.
 
@@
 \subsection{About Uprise}
-Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming experiences, that means 
-that they are not focussed on the actual gameplay which other EA studios like DICE is.
+Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
+experiences, that means that they are not focussed on the actual gameplay which other EA studios
+like DICE is.
 
-At the moment they are for example very involved with producing the Star Wars Battlefront game by making menu systems and 
-developing the companion app, Battlefront companion.
+At the moment they are for example very involved with producing the Star Wars Battlefront game by
+making menu systems and developing the companion app, Battlefront companion.
 
 \subsection{The current system}
-Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files, mostly images, to content 
-delivery networks. The current system does not make use out of the security features that the CDN's are offering, instead it 
-uses a form of security by obscurity. When a file is uploaded to a CDN it is open for the public, but its filename is 
-composed out of its original filename concatenated with a part of the MD5 hash of the content of the file, which makes it an 
-extremely hard process to access the file on the CDN without access to the original file or a reference to the URI.
+Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
+mostly images, to content delivery networks. The current system does not make use out of the
+security features that the CDN's are offering, instead it uses a form of security by obscurity. When
+a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
+original filename concatenated with a part of the MD5 hash of the content of the file, which makes
+it an extremely hard process to access the file on the CDN without access to the original file or a
+reference to the URI.
 
-In the current system you can only upload a file once as there will be a collision in the upload otherwise, as the old and 
-the new file will have the same MD5 hash.
+In the current system you can only upload a file once as there will be a collision in the upload
+otherwise, as the old and the new file will have the same MD5 hash.
 
 \subsection{Problem description}
-As the current system does not offer proper security measurements, is lacking a lot of features that is needed and does not 
-scale very well, a new system should be developed. This work is about examining a way of implementing Copy-on-Write in a 
-high level system like this, which should solve the scalability problem and make it possible to implement wanted features
-like snapshots, cloning and concurrent modifications of content.
+As the current system does not offer proper security measurements, is lacking a lot of features that
+is needed and does not scale very well, a new system should be developed. This work is about
+examining a way of implementing Copy-on-Write in a high level system like this, which should solve
+the scalability problem and make it possible to implement wanted features like snapshots, cloning
+and concurrent modifications of content.
 
@@
 \subsubsection{Content}
-Content is meta data about a file and is stored in a container, it is a form of virtual file.
-The content can refer to for example an image, video or binary blob.
+Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
+content can refer to for example an image, video or binary blob.
 
 \subsubsection{Project}
-A project is what is created to contain all content and containers related to a real project. Files can be changed 
-within a project and the system can contain several projects and their virtual content are completely disjoint.
+A project is what is created to contain all content and containers related to a real project. Files
+can be changed within a project and the system can contain several projects and their virtual
+content are completely disjoint.
 
@@
 \subsubsection{Snapshot}
-A snapshot is a read-only container from the state which the container the was in when the snapshot was created. 
-A snapshot can not be updated and can only be deleted from the root of the snapshot. Snapshots are by default stored 
-as siblings to the container which they were made from, but they can be contained by any container.
+A snapshot is a read-only container from the state which the container the was in when the snapshot
+was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
+Snapshots are by default stored as siblings to the container which they were made from, but they can
+be contained by any container.
 
 \subsubsection{File}
-A file refers to an actual physical file. Files are stored in the database to make backup, deployment and migration easier.
+A file refers to an actual physical file. Files are stored in the database to make backup,
+deployment and migration easier.
 
 \subsubsection{User}
-A user is the structure that handles people who have been granted access to the system.
-Access to the system is handled by a separate service, like LDAP.
+A user is the structure that handles people who have been granted access to the system.  Access to
+the system is handled by a separate service, like LDAP.
 
 \subsection{JPF}
-Java path finder was used to show that the model and plan of how to build the system was sound. The model was built in Java 
-with the objective of being as reduced and simple as possible, without loosing any of the cases that needed to be covered 
-by the model checker. As the users are mainly going to be handled by external systems they were not included in the model.
+Java path finder was used to show that the model and plan of how to build the system was sound. The
+model was built in Java with the objective of being as reduced and simple as possible, without
+loosing any of the cases that needed to be covered by the model checker. As the users are mainly
+going to be handled by external systems they were not included in the model.
 
 Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
-Each client was represented by a thread and each action taken by the client was randomised. The id hashes which MongoDB is 
-using for each entity was imported from the mongo-java-driver-2.13.3 and each object had its own id, generated in the same 
-fashion as the real implementation is using, randomly generated by the ObjectId class to minimise collisions that is. 
-Furthermore no locking or transactions were used and the threads were running fully concurrently, without any sleep statements.
+Each client was represented by a thread and each action taken by the client was randomised. The id
+hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
+each object had its own id, generated in the same fashion as the real implementation is using,
+randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
+transactions were used and the threads were running fully concurrently, without any sleep
+statements.
 
-ConcurrentHashMap had to be used in favour of the normal HashMap as the normal HashMaps can't be iterated over concurrently.
+ConcurrentHashMap had to be used in favour of the normal HashMap as the normal HashMaps can't be
+iterated over concurrently.
 
-JPF checked each permutation of states that the threads can end up in, the result of the run can be seen in Listing~\ref{JPFRESULT}.
+JPF checked each permutation of states that the threads can end up in, the result of the run can be
+seen in Listing~\ref{JPFRESULT}.
 
@@
 \newpage 
-\section{Copy-on-Write}
+\section{Snapshots}
+Write what snapshots are
+\subsection{Copy-on-Write}
 \label{sec:copy-on-write}
-As the persistent storage does not implement transactions or locks a lot of different problems can occur when several 
-clients are working on the same data set at the same time. Such problems could be race conditions and determining the 
-happened-before relation. In this work this problem is solved by implementing Copy-on-Write.
+To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
+snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
+Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
+not removed, see Figure~\ref{fig:btrfs_tree}.
+
+As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
+implement transactions or locks a lot of different problems can occur when several clients are
+working on the same data set at the same time. Such problems could be race conditions and
+determining the happened-before relation. In this work this problem is solved by implementing
+Copy-on-Write.
+
+\subsection{Full Copy}
 
@@
 \subsubsection{BTRFS}
-Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do efficient writeable 
-snapshots and clones. It also supports cloning of subtrees without having to actually copy the whole subtree, 
-this is due to the Copy-on-Write effect. As several nodes in the tree can refer to the same node each node keeps 
-track of how many parents it has by a reference counter so that the node can be deallocated once the node does not 
-have any parents any more. The reference counter is not stored in the nodes themselves but rather in a separate data 
-structure so that a nodes counter can be modified without modifying the node itself and therefore eludes the 
+Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
+efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
+actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
+tree can refer to the same node each node keeps track of how many parents it has by a reference
+counter so that the node can be deallocated once the node does not have any parents any more. The
+reference counter is not stored in the nodes themselves but rather in a separate data structure so
+that a nodes counter can be modified without modifying the node itself and therefore eludes the
 Copy-on-Write that would have to occur.
@@
 \subsubsection{Mach kernel}
-In the mid 80s when the development of the Mach kernel started, there was problems with that physically copying memory 
-was too slow. Too minimise the copying of memory, Copy-on-Write was implemented. It was implemented so that virtual 
-copy operations could be done and so that tasks could share read-write memory~\cite{MACH}.
+In the mid 80s when the development of the Mach kernel started, there was problems with that
+physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
+implemented. It was implemented so that virtual copy operations could be done and so that tasks
+could share read-write memory~\cite{MACH}.
 
@@
 Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
-back-end written in Scala and a front-end written in ES6 (Javascript), but they are both interchangeable. 
-The back-end has a REST API running, which is how the front-end communicates with the back-end.
+back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
+interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
+the back-end.
 
@@
 
-The front-end is written in ES6 with React and Reflux, the styling is done with the help of
+The front-end is written in ES6 with React and Reflux, and the styling is done with the help of
 Bootstrap.
@@
 \subsection{Snapshots}
-This implementation is far from as 
-efficient as the other Copy-on-Write systems described in Section~\ref{sec:copy-on-write} in most aspects, but more efficient 
-in some. As the implementation is built upon MongoDB as persistent storage and not a pure tree structure, single nodes 
-can be fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)), where n is the 
-number of nodes in the subtree.
+This implementation is far from as efficient as the other Copy-on-Write systems described in
+Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
+is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
+fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
+where n is the number of nodes in the subtree.
 
 In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
-Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is instead stored in a 
-flat storage space, such operation would be too computationally expensive as trees would have to be merged when 
-collisions occur, due to the non-blocking nature of the application. Instead this implementation makes a full copy of 
-the meta-data of the tree, but still refers to the same binary files until they are changed, which results in the 
-creation of a new node.
+Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
+instead stored in a flat storage space, such operation would be too computationally expensive as
+trees would have to be merged when collisions occur, due to the non-blocking nature of the
+application. Instead this implementation makes a full copy of the meta-data of the tree, but still
+refers to the same binary files until they are changed, which results in the creation of a new node.
 
 \subsection{Methods for determining\\implementation details}
-This chapter introduces the different methods used to determine how the new system should be implemented, 
-which DBMS it should use and how the estimation of long term scaling was done.
+This chapter introduces the different methods used to determine how the new system should be
+implemented, which DBMS it should use and how the estimation of long term scaling was done.
 
@@
 \subsubsection{Persistent storage}
+\label{persistent_storage}
 \subsubsubsection{MongoDB}
-MongoDB was chosen as the persistent storage because of its quick lookups and because of its internal storage 
-format called BSON, which is very similar to JSON which the API is using. As the formats are similar, the process of 
-marshalling and unmarshalling becomes quite easy between the core code, MongoDB instance and REST interface. 
-The second reason was that if the system needs to scale in the future it is very easy to distribute MongoDB and if 
-needed the system can easily be migrated to Reactive Mongo, which is an asynchronous and 
-non-blocking driver for MongoDB and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
+MongoDB was chosen as the persistent storage because of its quick lookups and because of its
+internal storage format called BSON, which is very similar to JSON which the API is using. As the
+formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
+core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
+scale in the future it is very easy to distribute MongoDB and if needed the system can easily be
+migrated to Reactive Mongo, which is an asynchronous and non-blocking driver for MongoDB and can
+therefore make the system scale even further~\cite{REACTIVEMONGO}.
 
 \subsubsubsection{API}
-REST was chosen as only basic CRUD operations needs to be performed and because the BSON format which is used in 
-MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is used by RESTful services~\cite{JSON}.
+REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
+which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
+used by RESTful services~\cite{JSON}.
 
@@
 
-As can be seen several expected endpoints are missing, this is intentional as the operations missing can be performed in 
-a more efficient way. Such endpoint is for example $GET projects/\{id\}/containers$ as all containers exist in 
-$GET projects/\{id\}$ and the interface should present a file structure where both content and containers are shown.
+As can be seen several expected endpoints are missing, this is intentional as the operations missing
+can be performed in a more efficient way. Such endpoint is for example $GET
+projects/\{id\}/containers$ as all containers exist in $GET projects/\{id\}$ and the interface
+should present a file structure where both content and containers are shown.
 

### 2015-11-25 tex/paper.tex
Restructured the sections

@@
+%{{{
 \documentclass[a4paper,12pt]{article}
@@
 \captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
-
 \usepackage{setspace}
@@
 
+%}}}
+
+%{{{
 \begin{document}
@@
 %\includepdf[pages={1}]{abstract.pdf}
+
+%}}}
 \begin{abstract}
@@
 
+%{{{
 \newpage\null\thispagestyle{empty}\newpage
@@
 \setcounter{page}{1}
+
+%}}}
 \section{Introduction}
@@
 \newpage 
-\section{Model}
-\subsection{Entities}
-%\subsubsection*{snapshot} for not having section numbers
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.8]{relation.pdf}}
-    \caption{Entity Relationships}
-    \label{fig:relation}
-\end{figure}
-
-\subsubsection{Content}
-Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
-content can refer to for example an image, video or binary blob.
-
-\subsubsection{Project}
-A project is what is created to contain all content and containers related to a real project. Files
-can be changed within a project and the system can contain several projects and their virtual
-content are completely disjoint.
-
-\subsubsection{Container}
-A container is a virtual folder within a project which can contain content and other containers.
-
-\subsubsection{Snapshot}
-A snapshot is a read-only container from the state which the container the was in when the snapshot
-was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
-Snapshots are by default stored as siblings to the container which they were made from, but they can
-be contained by any container.
-
-\subsubsection{File}
-A file refers to an actual physical file. Files are stored in the database to make backup,
-deployment and migration easier.
-
-\subsubsection{User}
-A user is the structure that handles people who have been granted access to the system.  Access to
-the system is handled by a separate service, like LDAP.
-
-\subsection{JPF}
-Java path finder was used to show that the model and plan of how to build the system was sound. The
-model was built in Java with the objective of being as reduced and simple as possible, without
-loosing any of the cases that needed to be covered by the model checker. As the users are mainly
-going to be handled by external systems they were not included in the model.
-
-Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
-Each client was represented by a thread and each action taken by the client was randomised. The id
-hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
-each object had its own id, generated in the same fashion as the real implementation is using,
-randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
-transactions were used and the threads were running fully concurrently, without any sleep
-statements.
-
-ConcurrentHashMap had to be used in favour of the normal HashMap as the normal HashMaps can't be
-iterated over concurrently.
-
-JPF checked each permutation of states that the threads can end up in, the result of the run can be
-seen in Listing~\ref{JPFRESULT}.
-
-\newpage
-\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
-elapsed time:       14:26:53
-states:             new=160853259,
-                    visited=451102505,
-                    backtracked=611955764,
-                    end=21640
-search:             maxDepth=380,
-                    constraints=0
-choice generators:  thread=160853255 
-                    (signal=0,
-                    lock=3603938,
-                    sharedRef=146989208,
-                    threadApi=3,
-                    reschedule=10260106), 
-                    data=0
-
-heap:               new=676056850,
-                    released=435060996,
-                    maxLive=655,
-                    gcCycles=523950061
-
-instructions:       11917045758
-max memory:         6256MB
-loaded code:        classes=111,
-                    methods=2179
-\end{lstlisting}
-
 \newpage 
@@
 
-\subsection{Snapshots}
-This implementation is far from as efficient as the other Copy-on-Write systems described in
-Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
-is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
-fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
-where n is the number of nodes in the subtree.
+\subsection{Model}
+\subsubsection{Entities}
+%\subsubsection*{snapshot} for not having section numbers
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.8]{relation.pdf}}
+    \caption{Entity Relationships}
+    \label{fig:relation}
+\end{figure}
 
+\subsubsubsection{Content}
+Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
+content can refer to for example an image, video or binary blob.
+
+\subsubsubsection{Project}
+A project is what is created to contain all content and containers related to a real project. Files
+can be changed within a project and the system can contain several projects and their virtual
+content are completely disjoint.
+
+\subsubsubsection{Container}
+A container is a virtual folder within a project which can contain content and other containers.
+
+\subsubsubsection{Snapshot}
+A snapshot is a read-only container from the state which the container the was in when the snapshot
+was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
+Snapshots are by default stored as siblings to the container which they were made from, but they can
+be contained by any container.
+
+\subsubsubsection{File}
+A file refers to an actual physical file. Files are stored in the database to make backup,
+deployment and migration easier.
+
+\subsubsubsection{User}
+A user is the structure that handles people who have been granted access to the system.  Access to
+the system is handled by a separate service, like LDAP.
+
+\subsubsection{JPF}
+Java path finder was used to show that the model and plan of how to build the system was sound. The
+model was built in Java with the objective of being as reduced and simple as possible, without
+loosing any of the cases that needed to be covered by the model checker. As the users are mainly
+going to be handled by external systems they were not included in the model.
+
+Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
+Each client was represented by a thread and each action taken by the client was randomised. The id
+hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
+each object had its own id, generated in the same fashion as the real implementation is using,
+randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
+transactions were used and the threads were running fully concurrently, without any sleep
+statements.
+
+ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can't be
+iterated over concurrently.
+
+JPF checked each permutation of states that the threads can end up in, the result of the run can be
+seen in Listing~\ref{JPFRESULT}.
+
+\newpage
+\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
+elapsed time:       14:26:53
+states:             new=160853259,
+                    visited=451102505,
+                    backtracked=611955764,
+                    end=21640
+search:             maxDepth=380,
+                    constraints=0
+choice generators:  thread=160853255 
+                    (signal=0,
+                    lock=3603938,
+                    sharedRef=146989208,
+                    threadApi=3,
+                    reschedule=10260106), 
+                    data=0
+
+heap:               new=676056850,
+                    released=435060996,
+                    maxLive=655,
+                    gcCycles=523950061
+
+instructions:       11917045758
+max memory:         6256MB
+loaded code:        classes=111,
+                    methods=2179
+\end{lstlisting}
+
+\subsection{Snapshot functionality}
 In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
@@
 \subsection{Resulting system}
+This implementation is far from as efficient as the other Copy-on-Write systems described in
+Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
+is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
+fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
+where n is the number of nodes in the subtree.
+
 \subsubsection{Persistent storage}
@@
 
+%{{{
+
 %What to write about
@@
 %described in Section 5
+
+%}}}

### 2015-11-27 tex/paper.tex
Reshuffled Models to be outside of implementation, again!

--- 2015-11-25 tex/paper.tex
@@
 compares different types of lazy snapshots to figure out which is the most
+
+\subsection{Prior Work}
+\subsubsection{Copy-on-Write}
+This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach
+kernel~\cite{COPYONWRITE} as it can be used to efficiently create snapshots.
+
+Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, for snapshot
+functionality in both file systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW},
+and as an optimisation technique for objects and types in programming languages\cite{LANGCOW}.
+
+Its principle is that when processes or nodes share data in between each other, the data is not
+copied until one of the processes does changes to it. This is an optimisation as the processes does
+not have to send or copy all of the related data that is in memory, rather they only have to send
+pointers to the data. After many Copy-on-Write's a complex structure can be built up, but it is
+possible to solve that structure~\cite{COPYONWRITE2}.
+
+TODO: Polish and extend
 
@@
 
-\section{Related Work}
-\subsection{Copy-on-Write}
-This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach
-kernel~\cite{COPYONWRITE}.
-
-Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, for snapshot
-functionality in both file systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW},
-and as an optimisation technique for objects and types in programming languages\cite{LANGCOW}.
-
-Its principle is that when processes or nodes share data in between each other, the data is not
-copied until one of the processes does changes to it. This is an optimisation as the processes does
-not have to send or copy all of the related data that is in memory, rather they only have to send
-pointers to the data. After many Copy-on-Write's a complex structure can be built up, but it is
-possible to solve that structure~\cite{COPYONWRITE2}.
-
-TODO: Polish and extend
+\subsection{Terms}
+\subsubsection{Snapshot}
+A snapshot is a way to record the full state of a system at a specific time. The term comes from
+photography where a photo can be seen as the state of what the photo is of at a certain time.
+Snapshots should not be confused with full copies of a system as full copies can be used as backups
+meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
+not effective against data corruption as snapshots usually still refer to unchanged data that is
+still a part of the system~\cite{SNAPSHOT}.
 
@@
 \newpage
+
 \section{Model}
+The model for this work should make sure that data can not be accessed by unauthorized users in the
+Perius system (See section TODO: fill in). There could also be another model done to make sure that
+content can not be accessed by unauthorized viewers once the content is uploaded to the CDN, but as
+that should already have been thoroughly checked by the CDN providers this work can focus on the
+internal users of the management system.
+
 \subsection{Entities}
@@
     \includegraphics[scale=0.8]{relation.pdf}}
-    \caption{Entity Relationships}
+    \caption{High Level Entity Relationships}
     \label{fig:relation}
@@
 
-ConcurrentHashMap had to be used in favour of the normal HashMap as the normal HashMaps can't be
+ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can't be
 iterated over concurrently.
@@
 
-\newpage 
-\section{Snapshots}
-Write what snapshots are
+\section{Snapshot systems}
 \subsection{Copy-on-Write}
@@
 
-\subsection{Snapshots}
-This implementation is far from as efficient as the other Copy-on-Write systems described in
-Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
-is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
-fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
-where n is the number of nodes in the subtree.
-
+\subsection{Snapshot functionality}
 In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in

### 2015-12-01 tex/paper.tex
Restructured sections

@@
 \newcommand{\Union}{\cup}
+
+\newcommand{\fix}{\colorbox{yellow!30}{TODO:}}
+
 \usetikzlibrary{positioning,shapes,shadows,arrows}
@@
 
+%}}}
 \title{\textbf{Designing a Virtual Security Layer for Cloud Content}}
 \author{Lukas Klingsbo}
-
-%}}}
 
@@
 \pagenumbering{roman}
-\setcounter{page}{2}
+\setcounter{page}{1}
 
@@
 \begin{abstract}
-    TODO: Abstract
+    \fix Abstract
 
@@
 usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
-and then directly available to the public.  For secret content this can be a problem and that is
-what this thesis is about. This work examines ways of enforcing access control on content and groups
-of content in the form of folders and snapshots. A system was developed to make the underlying
-theory work in practice.
+and then distributed over the globe when the public requests it. For secret content this can be a
+problem and an inconvenience, and that is what this thesis is about. This work examines ways of
+enforcing virtual access control on content and groups of content, in the form of folders and
+snapshots. A system was developed to make the underlying theory work in practice.
 
-The research question that this report answers is whether it is practically feasible to use
-Copy-on-Write for a high-level system like the one that is implemented. The work also thoroughly
-compares different types of lazy snapshots to figure out which is the most
+The research question that this report answers is how and whether it is practically feasible to use
+Copy-on-Write for a high-level system like the one that is implemented.
+
+\fix Clarify problem definition and research questions
 
@@
 \subsubsection{Copy-on-Write}
-This work relies heavily on the Copy-on-Write principle which was founded and used in the Mach
-kernel~\cite{COPYONWRITE} as it can be used to efficiently create snapshots.
+This work relies heavily on the Copy-on-Write principle, which was founded and used in the Mach
+kernel~\cite{COPYONWRITE}, as it can be used to efficiently create snapshots and help solving
+concurrency problems that otherwise can occur.
 
-Copy-on-Write is used in virtual memory management systems~\cite{VIRTCOW}, for snapshot
-functionality in both file systems~\cite{BTRFS} and logical volume management systems~\cite{LVMCOW},
-and as an optimisation technique for objects and types in programming languages\cite{LANGCOW}.
+Copy-on-Write is used in for example virtual memory management systems~\cite{VIRTCOW}, for snapshot
+and as an optimisation technique for objects and types in several programming
+languages\cite{LANGCOW}.
 
@@
 not have to send or copy all of the related data that is in memory, rather they only have to send
-pointers to the data. After many Copy-on-Write's a complex structure can be built up, but it is
-possible to solve that structure~\cite{COPYONWRITE2}.
+pointers to the data. After many Copy-on-Write's a complex tree structure can be built up, but
+optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.
 
-TODO: Polish and extend
+\fix Polish and extend
 
@@
 \section{Related Terminology}
-\subsection{Technologies}
+\subsection{Abbreviations}
+\subsubsection{JPF}
+Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
+licence, which made more people contribute to the project. JPF is usually used for doing model
+checking of concurrent programs to easily find for example race conditions and dead locks.
+
+\subsubsection{CDN}
+Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
+geographically. Once a request is made, the network serves content from the server closest to the
+requester.
+
+\subsection{Terms}
+\subsubsection{Snapshot}
+A snapshot is a way to record the full state of a system at a specific time. The term comes from
+photography where a photo can be seen as the state of what the photo is of, at a certain time.
+Snapshots should not be confused with full copies of a system as full copies can be used as backups
+meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
+not effective against data corruption as snapshots usually still refer to unchanged data that is
+still a part of the system~\cite{SNAPSHOT}.
+
+\newpage
+
+\section{Model}
+The model for this work should show how the data can not be accessed or modified by unauthorized
+users and how the integrity of the data is always kept in the Perius system.
+
+There could also be another relevant model done to show that content can not be accessed by
+unauthorized viewers once the content is uploaded to a CDN, but as that should already have been
+thoroughly checked by the CDN providers this work can focus solely on the internal users and content
+of the management system.
+
+\subsection{Related work}
+
+\subsection{Approach}
+
+\subsection{Design}
+\fix Mathematical formal description of how the system works
+
+\subsection{Findings}
+
+\newpage
+\section{Implementation}
+\subsection{Background}
+\subsubsection{About Uprise}
+Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
+experiences, which means that they are not focussed on the actual gameplay which other EA studios
+like DICE is.
+
+At the moment they are for example very involved with producing the Star Wars Battlefront game by
+making menu systems and developing the companion app, Battlefront companion.
+
+\subsubsection{The current system}
+Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
+mostly images, to content delivery networks. The current system does not make use out of the
+security features that the CDN's are offering, instead it uses a form of security by obscurity. When
+a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
+original filename concatenated with a part of the MD5 hash of the content of the file, which makes
+it an extremely hard process to access the file on the CDN without access to the original file or a
+reference to the URI.
+
+In the current system you can only upload a file once as there will be a collision in the upload
+otherwise, as the old and the new file will have the same MD5 hash.
+
+\subsubsection{Problem description}
+As the current system does not offer proper security measurements, is lacking a lot of features that
+is needed and does not scale very well, a new system should be developed. This work is about
+examining a way of implementing Copy-on-Write in a high level system like this, which should solve
+the scalability problem and make it possible to implement wanted features like snapshots, cloning
+and concurrent modifications of content.
+
+\subsection{Related Technologies}
 \subsubsection{React}
@@
 
-In this work we use a RESTful API, which means that it is an API following the REST architecture.
-
 \subsubsection{MongoDB}
@@
 
-\subsection{Abbreviations}
+\subsection{Methods for determining\\implementation details}
+This chapter introduces the different methods used to determine how the new system should be
+implemented, which DBMS it should use and how the estimation of long term scaling was done.
 
-\subsubsection{JPF}
-Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
-licence, which made more people contribute to the project. JPF is usually used for doing model
-checking of concurrent programs to easily find for example race conditions and dead locks.
+\subsection{Snapshot functionality}
+\fix Structure to compare snapshot systems and conclude how Perius snapshot system was designed
+\subsubsection{Copy-on-Write}
+\label{sec:copy-on-write}
+To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
+snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
+Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
+not removed, see Figure~\ref{fig:btrfs_tree}.
 
-\subsubsection{CDN}
-Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
-geographically. Once a request is made, the network serves content from the server closest to the
-requester.
+As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
+implement transactions or locks a lot of different problems can occur when several clients are
+working on the same data set at the same time. Such problems could be race conditions and
+determining the happened-before relation. In this work this problem is solved by implementing
+Copy-on-Write.
+\fix Move last paragraph
 
-\subsection{Terms}
-\subsubsection{Snapshot}
-A snapshot is a way to record the full state of a system at a specific time. The term comes from
-photography where a photo can be seen as the state of what the photo is of at a certain time.
-Snapshots should not be confused with full copies of a system as full copies can be used as backups
-meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
-not effective against data corruption as snapshots usually still refer to unchanged data that is
-still a part of the system~\cite{SNAPSHOT}.
+\subsubsection{Full Copy}
+\fix Maybe a short description of full copy, as opposed to copy-on-write
 
-\section{Background}
-\subsection{About Uprise}
-Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
-experiences, that means that they are not focussed on the actual gameplay which other EA studios
-like DICE is.
+\subsubsection{Comparison of Copy-on-Write system implementations}
+\subsubsubsection{BTRFS}
+Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
+efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
+actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
+tree can refer to the same node each node keeps track of how many parents it has by a reference
+counter so that the node can be deallocated once the node does not have any parents any more. The
+reference counter is not stored in the nodes themselves but rather in a separate data structure so
+that a nodes counter can be modified without modifying the node itself and therefore eludes the
+Copy-on-Write that would have to occur.
 
-At the moment they are for example very involved with producing the Star Wars Battlefront game by
-making menu systems and developing the companion app, Battlefront companion.
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.4]{newtree.png}}
+    \caption{Cloning mechanism of Btrfs~\cite{BTRFS}}
+    \label{fig:btrfs_tree}
+\end{figure}
 
-\subsection{The current system}
-Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
-mostly images, to content delivery networks. The current system does not make use out of the
-security features that the CDN's are offering, instead it uses a form of security by obscurity. When
-a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
-original filename concatenated with a part of the MD5 hash of the content of the file, which makes
-it an extremely hard process to access the file on the CDN without access to the original file or a
-reference to the URI.
+\subsubsubsection{Mach kernel}
+In the mid 80's when the development of the Mach kernel started, there was problems with that
+physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
+implemented. It was implemented so that virtual copy operations could be done and so that tasks
+could share read-write memory~\cite{MACH}.
 
-In the current system you can only upload a file once as there will be a collision in the upload
-otherwise, as the old and the new file will have the same MD5 hash.
+\fix Insert more systems here
 
-\subsection{Problem description}
-As the current system does not offer proper security measurements, is lacking a lot of features that
-is needed and does not scale very well, a new system should be developed. This work is about
-examining a way of implementing Copy-on-Write in a high level system like this, which should solve
-the scalability problem and make it possible to implement wanted features like snapshots, cloning
-and concurrent modifications of content.
+\subsubsection{Snapshot functionality of Perius}
+In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
+Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
+instead stored in a flat storage space, such operation would be too computationally expensive as
+trees would have to be merged when collisions occur, due to the non-blocking nature of the
+application. Instead this implementation makes a full copy of the meta-data of the tree, but still
+refers to the same binary files until they are changed, which results in the creation of a new node.
 
-\newpage
+\fix Relate section more to comparison
 
-\section{Model}
-The model for this work should make sure that data can not be accessed by unauthorized users in the
-Perius system (See section TODO: fill in). There could also be another model done to make sure that
-content can not be accessed by unauthorized viewers once the content is uploaded to the CDN, but as
-that should already have been thoroughly checked by the CDN providers this work can focus on the
-internal users of the management system.
+\subsection{Resulting system}
+\subsubsection{Perius}
+Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
+back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
+interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
+the back-end.
 
+\fix Picture of the front-end
+
+The service features a virtual file structure over the assets that has been stored, snapshots,
+security management of whole containers as well as individual files, audit and access logging, multi
+project support and a modular design for persistent storage.
+
+The front-end is written in ES6 with React and Reflux, and the styling is done with the help of
+Bootstrap.
+
+\subsubsection{Copy-on-Write}
+\fix Rename section and move
+
+This implementation is far from as efficient as the other Copy-on-Write systems described in
+Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
+is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
+fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
+where n is the number of nodes in the subtree.
+
+\subsubsection{Persistent storage}
+\label{persistent_storage}
+\subsubsubsection{MongoDB}
+MongoDB was chosen as the persistent storage because of its quick lookups and because of its
+internal storage format called BSON, which is very similar to JSON which the API is using. As the
+formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
+core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
+scale in the future it is very easy to distribute MongoDB and if needed the system can easily be
+migrated to Reactive Mongo, which is an asynchronous and non-blocking driver for MongoDB and can
+therefore make the system scale even further~\cite{REACTIVEMONGO}.
+
+\subsubsubsection{API}
+In this work a RESTful API was implemented and used for back-end $\Leftrightarrow$ front-end
+communication.
+
+REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
+which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
+used by RESTful services~\cite{JSON}.
+
+\subsubsection{REST Endpoints}
+For the front-end to communicate with the back-end, a RESTful service is implemented.
+The following endpoints were configured:
+
+\begin{itemize}
+  \item projects
+      \subitem GET - list all projects
+      \subitem POST - create new project
+  \item projects/\{id\}
+      \subitem GET - get specific project
+      \subitem PUT - update existing project
+      \subitem DELETE - delete existing project
+  \item projects/\{id\}/content
+      \subitem POST - create new content in a specific project 
+  \item projects/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific project
+      \subitem PUT - update existing content in a specific project
+      \subitem DELETE - delete existing content in a specific project
+
+  \item projects/\{id\}/snapshots
+      \subitem POST - create new snapshot in a specific project
+
+  \item projects/\{id\}/containers
+      \subitem POST - create new container in a specific project 
+  \item projects/\{id\}/containers/\{id\}
+      \subitem GET - get specific container in a specific project
+      \subitem PUT - update existing container
+      \subitem DELETE - delete existing container
+\end{itemize}
+
+As can be seen several expected endpoints are missing, this is intentional as the operations missing
+can be performed in a more efficient way. Such endpoint is for example $GET
+projects/\{id\}/containers$ as all containers exist in $GET projects/\{id\}$ and the interface
+should present a file structure where both content and containers are shown.
+
+\subsection{Findings}
+\subsubsection{Scalability}
+
+\subsubsection{Security of the system}
+\paragraph{Authorization}
+\paragraph{Audit logs}
+
+\section{JPF}
+\fix Not sure where to put this section or if it's even relevant any more, maybe skip JPF and put
+the entities under implementation -> design
 \subsection{Entities}
-%\subsubsection*{snapshot} for not having section numbers
 \begin{figure}[htp] \centering{
@@
 
-\subsection{JPF}
+\subsection{Execution}
 Java path finder was used to show that the model and plan of how to build the system was sound. The
@@
 
-\newpage
 \begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
@@
 
-\section{Snapshot systems}
-\subsection{Copy-on-Write}
-\label{sec:copy-on-write}
-To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
-snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
-Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
-not removed, see Figure~\ref{fig:btrfs_tree}.
-
-As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
-implement transactions or locks a lot of different problems can occur when several clients are
-working on the same data set at the same time. Such problems could be race conditions and
-determining the happened-before relation. In this work this problem is solved by implementing
-Copy-on-Write.
-
-\subsection{Full Copy}
-
-\subsection{Comparison of Copy-on-Write system implementations}
-\subsubsection{BTRFS}
-Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
-efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
-actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
-tree can refer to the same node each node keeps track of how many parents it has by a reference
-counter so that the node can be deallocated once the node does not have any parents any more. The
-reference counter is not stored in the nodes themselves but rather in a separate data structure so
-that a nodes counter can be modified without modifying the node itself and therefore eludes the
-Copy-on-Write that would have to occur.
-
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.4]{newtree.png}}
-    \caption{Cloning mechanism of Btrfs~\cite{BTRFS}}
-    \label{fig:btrfs_tree}
-\end{figure}
-
-\subsubsection{Mach kernel}
-In the mid 80s when the development of the Mach kernel started, there was problems with that
-physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
-implemented. It was implemented so that virtual copy operations could be done and so that tasks
-could share read-write memory~\cite{MACH}.
-
-TODO: Insert more systems here
-\section{Implementation}
-\subsection{Perius}
-Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
-back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
-interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
-the back-end.
-
-TODO: Picture of the front-end
-
-The service features a virtual file structure over the assets that has been stored, snapshots,
-security management of whole containers as well as individual files, audit and access logging, multi
-project support and a modular design for persistent storage.
-
-The front-end is written in ES6 with React and Reflux, and the styling is done with the help of
-Bootstrap.
-
-\subsection{Snapshot functionality}
-In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
-Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
-instead stored in a flat storage space, such operation would be too computationally expensive as
-trees would have to be merged when collisions occur, due to the non-blocking nature of the
-application. Instead this implementation makes a full copy of the meta-data of the tree, but still
-refers to the same binary files until they are changed, which results in the creation of a new node.
-
-\subsection{Methods for determining\\implementation details}
-This chapter introduces the different methods used to determine how the new system should be
-implemented, which DBMS it should use and how the estimation of long term scaling was done.
-
-\subsection{Resulting system}
-This implementation is far from as efficient as the other Copy-on-Write systems described in
-Section~\ref{sec:copy-on-write} in most aspects, but more efficient in some. As the implementation
-is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
-fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
-where n is the number of nodes in the subtree.
-
-\subsubsection{Persistent storage}
-\label{persistent_storage}
-\subsubsubsection{MongoDB}
-MongoDB was chosen as the persistent storage because of its quick lookups and because of its
-internal storage format called BSON, which is very similar to JSON which the API is using. As the
-formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
-core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
-scale in the future it is very easy to distribute MongoDB and if needed the system can easily be
-migrated to Reactive Mongo, which is an asynchronous and non-blocking driver for MongoDB and can
-therefore make the system scale even further~\cite{REACTIVEMONGO}.
-
-\subsubsubsection{API}
-REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
-which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
-used by RESTful services~\cite{JSON}.
-
-\subsubsection{REST Endpoints}
-For the front-end to communicate with the back-end, a RESTful service is implemented.
-The following endpoints were configured:
-
-\begin{itemize}
-  \item projects
-      \subitem GET - list all projects
-      \subitem POST - create new project
-  \item projects/\{id\}
-      \subitem GET - get specific project
-      \subitem PUT - update existing project
-      \subitem DELETE - delete existing project
-  \item projects/\{id\}/content
-      \subitem POST - create new content in a specific project 
-  \item projects/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific project
-      \subitem PUT - update existing content in a specific project
-      \subitem DELETE - delete existing content in a specific project
-
-  \item projects/\{id\}/snapshots
-      \subitem POST - create new snapshot in a specific project
-
-  \item projects/\{id\}/containers
-      \subitem POST - create new container in a specific project 
-  \item projects/\{id\}/containers/\{id\}
-      \subitem GET - get specific container in a specific project
-      \subitem PUT - update existing container
-      \subitem DELETE - delete existing container
-  \item projects/\{id\}/containers/\{id\}/content
-      \subitem POST - create new content in a specific container 
-  \item projects/\{id\}/containers/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific container
-      \subitem PUT - update existing content in a specific container
-      \subitem DELETE - delete existing content in a specific container
-\end{itemize}
-
-As can be seen several expected endpoints are missing, this is intentional as the operations missing
-can be performed in a more efficient way. Such endpoint is for example $GET
-projects/\{id\}/containers$ as all containers exist in $GET projects/\{id\}$ and the interface
-should present a file structure where both content and containers are shown.
-
-\subsection{Findings}
-\subsubsection{Scalability}
-
-\subsubsection{Security of the system}
-\paragraph{Authorization}
-\paragraph{Audit logs}
-
 \section{Discussion}
@@
 %What to write about
-%Battle Binary
+%Battle Binar
+y
 %Security features of CDN's

### 2015-12-01 tex/paper.tex
Added some equations that might or might not be correct

@@
 
+\begin{itemize}
+    \item[] c - Container
+        \subitem Represents a structure that symbolises a folder in a virtual file system
+    \item[] f - File
+        \subitem A file in the system 
+    \item[] p - Pointer
+        \subitem A pointer to a file in the system 
+    \item[] u - User
+        \subitem A user in the system
+    \item[] a - Access property
+        \subitem Granted access to container
+    \item[] A[u,c] - Access matrix
+        \subitem Describes what containers the users in the system can access 
+\end{itemize}
+
+\subsection{Access rights}
+\begin{equation}
+    \begin{split}
+        u \in U \text{ can read } c \in C & \Iff u \in A[u,c] \\
+        u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        u \in U \text{ can delete } p \in c & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        \forall c \in C, \quad \exists u \in U & \mid a \in A[u,c] 
+    \end{split}
+\end{equation}
+
 \subsection{Findings}
@@
 
-\fix Picture of the front-end
+\fix Picture of the newest front-end
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.4]{frontend.png}}
+    \caption{Front-end~\cite{BTRFS}}
+    \label{fig:frontend}
+\end{figure}
 
@@
 \subsubsubsection{API}
-In this work a RESTful API was implemented and used for back-end $\Leftrightarrow$ front-end
+In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
 communication.

### 2015-12-10 tex/paper.tex
Added some biba sections

--- 2015-12-01 tex/paper.tex
@@
 
-\subsection{Design}
+\subsection{Elements of the Model}
 \fix Mathematical formal description of how the system works
+
+\begin{center}
+    \begin{tabular}{ | l | l | l | p{5cm} |}
+        \hline
+        Set & Elements & semantics \\ \hline
+        C   & $c_0\dots c_n$                & Containers; folders in the virtual file system\\ \hline
+        F   & $f_0\dots f_n$                & Files; files, images, videos\\ \hline
+        P   & $p_0\dots p_n$                & Content; Meta-data for files\\ \hline
+        U   & $u_0\dots u_n$                & Users; registered users in the system\\ \hline
+        A   & $A[u_0,c_0]\dots A[u_n, c_n]$ & Access matrix; describes what containers users have access to\\ \hline
+    \end{tabular}
+\end{center}
+
+\subsection{Access rights}
+\begin{equation}
+    \begin{split}
+        u \in U \text{ can read } c \in C & \Iff u \in A[u,c] \\
+        u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        u \in U \text{ can delete } p \in c & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        u \nexists U \text{ can delete } f \in F \\
+        \forall c \in C, \quad \exists u \in U & \mid a \in A[u,c] 
+    \end{split}
+\end{equation}
+
+\subsection{Data integrity}
+A computer system or subsystem is defined as possessing the property of integrity if it behaves consistently
+according to a defined standard. This implies that a subsystem possessing the property of integrity
+does not guarantee an absolute behaviour of the system, but rather that it performs according to
+what its creator intended ~\cite{BIBA}.
+
+\subsection{Initial Assumptions}
+To create an integrity model, some initial assumptions have to be made about what the correct behaviour
+of the system is, which the model then can be shown to follow. In this work unintentional behaviour
+as the result of data modification is the main concern, which could be used for sabotage or simply be the
+effect unintentional unfortunate race conditions etc.
+
+\subsection{Integrity Threats}
+According to Biba et. al~\cite{BIBA} one can consider two threat sources, namely subsystem external
+and subsystem internal. The external sources could be another system calling the subsystem with
+faulty data or trying to make inaccurate calls to program functions, it could also be somebody
+trying to tamper with the exposed functions of the program. Threats that are internal could be a
+malicious part of the subsystem or simply an incorrect part of the subsystem, which does not behave
+according to specification.
+
+In this work external threats are handled as threats that can occur from what has been exposed by
+the API (See Section~\ref{sec:API} and internal threats as incorrect implementation. As the server
+and its system are assumed safe malicious subsystems are not considered.
+
+\fix Translate to mathematical expressions
+
+\begin{itemize}
+    \item if a user wants to update a file in a content, the file is copied and the original is intact
+    \item if a user reads content and then writes to it, the content is directly changed
+    \item if a user copies content, a new content is created at the destination with reference to the same
+          file
+    \item if a user reads from a container and then writes to it, the container is directly changed
+          (Which means last write wins, which doesn't matter)
+    \item if a user copies a container, the full container tree is re created with new ids at the destination,
+          its content still refers to the same files.
+\end{itemize}
 
@@
 \subsubsubsection{API}
-In this work a RESTful API was implemented and used for back-end $\Leftrightarrow$ front-end
+\label{sec:API}
+In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
 communication.
@@
 %What to write about
-%Battle Binar
-y
+%Battle Binary
 %Security features of CDN's

### 2016-01-25 tex/paper.tex
Added full copy and JPF introduction

@@
 
+\label{sec:model}
 \section{Model}
@@
           (Which means last write wins, which doesn't matter)
-    \item if a user copies a container, the full container tree is re created with new ids at the destination,
-          its content still refers to the same files.
+    \item if a user creates a snapshot of a container, the full container tree is re created with 
+          new ids at the destination, its content still refers to the same files.
 \end{itemize}
@@
 \subsubsection{Full Copy}
-\fix Maybe a short description of full copy, as opposed to copy-on-write
+Full copy or deep copy, as opposed to copy-on-write, is a copy where everything is copied directly
+and not only when an object is changed. This is easier to implement but is in most cases more
+inefficient as more disk space will have to be used and if used with for example certain tree 
+structures the part of the tree that needs to be copied will have to be traversed.
 
@@
 As can be seen several expected endpoints are missing, this is intentional as the operations missing
-can be performed in a more efficient way. Such endpoint is for example $GET
-projects/\{id\}/containers$ as all containers exist in $GET projects/\{id\}$ and the interface
+can be performed in a more efficient way. Such endpoint is for example \textit{GET 
+projects/\{id\}/containers} as all containers exist in \textit{GET projects/\{id\}} and the interface
 should present a file structure where both content and containers are shown.
@@
 \paragraph{Authorization}
+The authorization of the system is currently being handled by LDAP
+
 \paragraph{Audit logs}
@@
 \section{JPF}
-\fix Not sure where to put this section or if it's even relevant any more, maybe skip JPF and put
-the entities under implementation -> design
+\fix Not sure where to put this section
+JPF was used to test the idea and state transitioning of the application. Its a very simplified
+version of the real system that still contains all the important Copy-on-Write core concepts and the 
+assumptions that have been made for the model. This simplified version could then be automatically
+tested for soundness. It is not a proof that the model works, but it is very exhaustive in its
+testing.
+
 \subsection{Entities}
@@
 \subsection{Future work}
+Full access control was not implemented according to the model described in ~\ref{sec:model}, it
+was only implemented to check whether a user should have access to the system as a whole or not, the
+implementation did not set or check any specific access rights to certain contents or containers.
 

### 2016-01-27 tex/paper.tex
Added scalability and GridFS discussion

@@
 %}}}
-\title{\textbf{Designing a Virtual Security Layer for Cloud Content}}
+\title{\textbf{Designing a Virtual Security Management Layer for Cloud Content}}
 \author{Lukas Klingsbo}
@@
 Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
-experiences, which means that they are not focussed on the actual gameplay which other EA studios
-like DICE is.
+experiences, which means that they are mostly not focussed on the actual gameplay which other EA 
+studios like DICE is.
 
@@
 normal relational databases has. Instead each entity in the database is stored as a document which
-is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks support for joins too
+is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
 improve its possibility to scale, which can be a big down side to some applications containing the
@@
 Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
-back-end written in Scala and a front-end written in ES6 (Javascript), but they are both
+back-end written in Scala and a front-end written in Javascript (ES6), but they are both
 interchangeable.  The back-end has a REST API running, which is how the front-end communicates with
@@
 
+All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
+according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
+that backups of the Perius state is easily done through a database backup, no separate files 
+needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
+ranges of a file, although that advantage is not needed in the Perius implementation. The 
+disadvantage of using this approach is that when using a non-distributed database it us slower 
+to read and write to the database than reading or writing directly to the filesystem. Another 
+disadvantage is that to access the files it is needed to go through the database layer in some 
+way, instead of accessing the filesystem.
+
 \subsubsubsection{API}
@@
 which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
-used by RESTful services~\cite{JSON}.
+usually used by RESTful services~\cite{JSON}.
 
@@
 \subsubsection{Scalability}
+When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
+the application has no limits of how much load and users it can handle as the hardware and nodes 
+can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
+implementation, it is harder to scale to the enourmous amounts of load which ReactiveMongo can 
+support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
+is offered by ReactiveMongo is not needed as the load will not reach the peak (\fix relate to 
+graph) for what Casbah can handle on a single server.
+
+\fix Add graphs of load testing etc
 
@@
 \fix Not sure where to put this section
+
 JPF was used to test the idea and state transitioning of the application. Its a very simplified

### 2016-01-28 tex/paper.tex
Added load testing text

@@
 \subsection{Access rights}
+The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the requested enitity of
+the matrix the corresponding user to that entity has full access to the corresponding object.
+
 \begin{equation}
@@
 
-\subsubsubsection{API}
+\subsubsection{API}
 \label{sec:API}
@@
 
-\subsubsection{REST Endpoints}
+\subsubsubsection{REST Endpoints}
 For the front-end to communicate with the back-end, a RESTful service is implemented.
@@
 can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
-implementation, it is harder to scale to the enourmous amounts of load which ReactiveMongo can 
+implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
 support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
@@
 graph) for what Casbah can handle on a single server.
+
+\subsubsubsection{Load Testing}
+Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
+create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
+RAM. The results that will be analysed are latency and load of the server as that will give a fair
+judgement of how well the system can perform on any given machine.
+
+\textbf{GET requests}
+To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
+requests} are performed. The first one tries to maximise the number of requests that spray-can (HTTP
+server) can handle with the given specifications and therefore the server just returns 200 (OK). The
+second request requests the simplest API call which involves getting all currently stored projects
+from the MongoDB instance. As this involves the server answering with some payload, the third
+request is simply the same amount of static payload as the second request but without involving any
+connections to MongoDB.
+
+\textbf{POST requests}
 
@@
 \paragraph{Authorization}
-The authorization of the system is currently being handled by LDAP
+The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius is mainly
+focussed at being deployed in internal networks which usually has an LDAP service enabled. LDAP also
+makes it easier for the user to login as no separate account is needed for Perius and thus the user
+can use the same account as for all the LDAP connected services on the internal network.
 

### 2016-01-28 tex/paper.tex
Added GET request load tests

@@
 create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
-RAM. The results that will be analysed are latency and load of the server as that will give a fair
-judgement of how well the system can perform on any given machine.
+RAM. The results that will be analysed are latency and throughput of the server as that will give a 
+fair judgement of how well the system can perform.
 
-\textbf{GET requests}
+\paragraph{GET requests}
+\par
 To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
@@
 connections to MongoDB.
+\par
+The typical requests with Wrk in this work will look similar to this:
+\begin{lstlisting}[frame=single]
+wrk -t100 -c100 -d10s http://perius:8000/ok
+\end{lstlisting}
+where -t100 means that it uses 100 threads, -c100 that it emulates 100 clients requesting over and
+over and -d100s is the amount of time that the load test will run, in this case 100 seconds.
+
+\begin{lstlisting}[label=OKREQUEST,caption=Result of OK requests]
+wrk -t100 -c100 -d10s http://perius:8000/ok
+
+Running 2m test @ http://perius:8000/ok
+  100 threads and 100 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency     1.22ms    3.20ms 166.78ms   97.75%
+    Req/Sec     1.04k   152.50     2.25k    79.18%
+  10345746 requests in 1.67m, 1.69GB read
+Requests/sec: 103353.85
+Transfer/sec:     17.25MB
+\end{lstlisting}
+
+\begin{lstlisting}[label=DBREQUEST,caption=Result of MongoDB requests]
+wrk -t100 -c100 -d10s http://perius:8000/projects
+
+Running 2m test @ http://perius:8000/projects
+  100 threads and 100 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency    14.52ms   13.34ms 375.51ms   99.31%
+    Req/Sec    72.85      6.91   232.00     60.20%
+  726351 requests in 1.67m, 196.03MB read
+Requests/sec:   7256.38
+Transfer/sec:      1.96MB
+\end{lstlisting}
+
+\begin{lstlisting}[label=STATICREQUEST,caption=Result of static text requests]
+wrk -t100 -c100 -d10s http://perius:8000/static
+
+Running 2m test @ http://perius:8000/static
+  100 threads and 100 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency     6.42ms   28.99ms 290.37ms   96.26%
+    Req/Sec     0.94k   203.30     2.00k    83.83%
+  9199737 requests in 1.67m, 2.19GB read
+Requests/sec:  91902.48
+Transfer/sec:     22.44MB
+\end{lstlisting}
+
+Listing~\ref{OKREQUEST} shows that the HTTP server can answer about 100K requests/s when not
+involving any payload other than status code 200 (OK). When comparing that to the request which
+involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST} one can 
+see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when comparing 
+that to the requests that needed database access it is obvious that the HTTP server can handle all 
+the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents in a 
+collection (in this case only one) is dramatically slower, the throughput was only \textasciitilde7000
+requests/s.
+
+\par
+These are very simple requests, to evaluate how much load the application can handle in its current
+state a more complex but common request has to be analysed. The most complicated and still common
+request that system is receiving is requests of a full project tree. This request is computationally
+expensive as the tree is not stored directly in the database but has to be built from the id and
+parent id of each container and content, a simulated project tree of similar size to the ones stored 
+in the current management system reaches around 25KB in uncompressed size.
+
+\begin{lstlisting}[label=TREEREQUEST,caption=Result of project tree requests]
+wrk -t10 -c10 -d10s http://perius:8000/projects/56a5f...
+
+Running 10s test @ http://perius:8000/projects/56a5f...
+  10 threads and 10 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency   247.54ms   21.64ms 323.78ms   96.76%
+    Req/Sec     3.80      1.22    20.00     99.00%
+  401 requests in 10.07s, 9.67MB read
+Requests/sec:     39.82
+Transfer/sec:      0.96MB
+\end{lstlisting}
+
+The result in Listing~\ref{TREEREQUEST} shows an average of 40 requests/s, which means this is the
+real bottleneck in the application. There are two solutions to fix this bottleneck, as the project
+tree wont be modified nearly as often as it is requested the first solution would be to cache the
+results of the tree requests and invalidate those caches when the tree is modified. The second
+solution would be to make a more computationally feasible way of building and fetching the whole 
+tree from the database. This is left to implement if the need comes for it in the future. Even
+though this massive bottleneck is present, it wont be a problem in a small scale production
+environment as the tree is fetched approximately once every 30 seconds by active users, which means
+that the application still could support at least 1200 very active users.
+
+\begin{lstlisting}[label=CACHEREQUEST,caption=Result of cached project tree requests]
+wrk -t100 -c100 -d100s 
+  http://perius:8000/projects/cached/56a5f...
+
+Running 2m test @ 
+  http://perius:8000/projects/cached/56a5f...
+  100 threads and 100 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency     1.46ms    3.62ms 180.68ms   97.87%
+    Req/Sec   838.58    120.95     1.74k    79.12%
+  8343565 requests in 1.67m, 17.03GB read
+Requests/sec:  83352.77
+Transfer/sec:    174.17MB
+\end{lstlisting}
+\par
+With caching turned on, on a single back-end, the application can theoretically support several 
+million active users which are not posting any content, this is simulated in 
+Listing~\ref{CACHEREQUEST}.
 

### 2016-01-29 tex/paper.tex
Added result of adding indices

@@
 
-\paragraph{GET requests}
-\par
+\paragraph{GET requests}\mbox{}\\
 To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
@@
 
-\textbf{POST requests}
+\par
+Another simple optimisation that was added was indices on the parent id's, which are heavily queried
+when building the project trees.
 
-\fix Add graphs of load testing etc
+\begin{lstlisting}[label=INDEXREQUEST,caption=Result of indexed project tree requests]
+wrk -t100 -c100 -d100s 
+  http://perius:8000/projects/56a5f...
+
+Running 10s test @ http://perius:8000/projects/56a...
+  100 threads and 100 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency    47.27ms    3.03ms  86.54ms   93.79%
+    Req/Sec    21.12      3.50    60.00     87.51%
+  21271 requests in 10.10s, 9.76MB read
+Requests/sec:   2105.82
+Transfer/sec:      0.97MB
+\end{lstlisting}
+
+After adding indices, the building of the project tree was able to finish 50 times faster and the
+server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}.
+
+\paragraph{POST requests}\mbox{}\\
+Two different POST requests will be tested, creation of containers and creation of content together
+with file uploads.
+
+\begin{lstlisting}[label=CONTAINERPOSTREQUEST,caption=Result of container creation]
+wrk -c24 -t12 -d4s -s post.lua 
+  http://perius:8000/projects/56ab.../containers
+
+Running 4s test @ 
+  http://perius:8000/projects/56ab.../containers
+  12 threads and 24 connections
+  Thread Stats   Avg      Stdev     Max   +/- Stdev
+    Latency     5.92ms    1.25ms  19.58ms   91.67%
+    Req/Sec   339.86     76.68     1.28k    95.65%
+  16357 requests in 4.10s, 5.04MB read
+Requests/sec:   3990.12
+Transfer/sec:      1.23MB
+\end{lstlisting}
+
+As can be seen in Listing~\ref{CONTAINERPOSTREQUEST} the server can handle around 4000 container
+POST requests per second. The problem that this test resulted in was not the amount of containers
+that could be posted, but rather loading all those containers in the interface or API afterwards.
+When requesting the project that the containers were posted to the web server timed out the request,
+as it took too long for the back-end to build the project tree. In this work, the web server wont 
+be configured to accept longer data processing time, as trees as large as this for a single project
+wont be used in production.
+
+\par
+As the project tree is built by querying documents parent id one optimisation that could be done was
+to create an index of the documents parent ids. This drastically improved the size of the tree that
+the server was able to handle within the time out, going from being able to handle 4000 documents to
+at least 50000 documents, which the browser instead will have problems displaying represented in the 
+UI without lag.
 

### 2016-02-01 tex/paper.tex
Added equations to the model

@@
 \def\l@subparagraph{\@dottedtocline{6}{14em}{6em}}
+\newenvironment{custommargins}[2]%
+  {\addtolength{\leftskip}{#1}\addtolength{\rightskip}{#2}}{\par}
 \makeatother
@@
 \subsection{Related work}
+\fix Write about Bell-Lapadula et. al
+The model for this paper is based on the work that was done by Bell, D Elliott, LaPadula and Leonard
+J in their papers Secure computer systems: Mathematical foundations~\cite{BLP1} and Secure computer
+systems: A mathematical model~\cite{BLP2}. In these papers the foundation is laid for how to model
+a computer system to be able to analyse the security of a system. Furthermore this paper was also 
+inspired by Biba, Integrity considerations for secure computer systems~\cite{BIBA}, where many of 
+the points made by him was taken into consideration when inspecting that the integrity of the data 
+was always sound.
 
 \subsection{Approach}
+\fix Write about how the approach is similar to BLP and that it is not a proof etc
+As this work only presents an informal model of how the system is designed it can not be regarded as
+a proof for the actual implementation of the system to be flawless. The model should how ever give a 
+strong idea of the soundness of the designed system.
 
@@
         \hline
-        Set & Elements & semantics \\ \hline
+        Set & Elements & Semantics \\ \hline
         C   & $c_0\dots c_n$                & Containers; folders in the virtual file system\\ \hline
         F   & $f_0\dots f_n$                & Files; files, images, videos\\ \hline
-        P   & $p_0\dots p_n$                & Content; Meta-data for files\\ \hline
+        M   & $m_0\dots m_n$                & Content; Meta-data for files\\ \hline
         U   & $u_0\dots u_n$                & Users; registered users in the system\\ \hline
@@
 \subsection{Access rights}
-The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the requested enitity of
+The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the requested entity of
 the matrix the corresponding user to that entity has full access to the corresponding object.
 
+\fix Add content, m
 \begin{equation}
@@
         u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
-        u \in U \text{ can delete } p \in c & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        u \in U \text{ can delete } m \in c & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
         u \nexists U \text{ can delete } f \in F \\
@@
 
+\begin{center}
+    \begin{tabular}{ | l | l | l | p{5cm} |}
+        \hline
+        Action & Semantics \\ \hline
+        u, read(o)            & User u reads object o\\ \hline
+        u, write(o)           & User u writes object o\\ \hline
+        u, copy(o)            & User u copies object o and the copy gets a new ID\\ \hline
+        u, change(o)          & User u locally changes object o \\ \hline
+        u, modify(o)          & User u globally modifies object o\\ \hline
+        u, modify(o', o)      & User u globally modifies object o based on object o'\\ \hline
+        u, delete(o)          & User u deletes object o\\ \hline
+        u, snapshot(o)        & User u takes a snapshot of object o \\ \hline
+    \end{tabular}
+\end{center}
+
 \fix Translate to mathematical expressions
+\subsubsection{Consequences of modification}
+\begin{equation} \label{eq:fileupdate}
+    \begin{split}
+        m' & = u, read(m) \land \\
+        x  & = f \in m' \land \\
+        f' & = u, change(x) \land \\
+        u, & write(f') \land \\
+        u, & modify(f', m) \Implies f \in F \land f' \in F
+    \end{split}
+\end{equation}
+If a user wants to update a file in a content, the file is copied and the original is intact.
 
-\begin{itemize}
-    \item if a user wants to update a file in a content, the file is copied and the original is intact
-    \item if a user reads content and then writes to it, the content is directly changed
-    \item if a user copies content, a new content is created at the destination with reference to the same
-          file
-    \item if a user reads from a container and then writes to it, the container is directly changed
-          (Which means last write wins, which doesn't matter)
-    \item if a user creates a snapshot of a container, the full container tree is re created with 
-          new ids at the destination, its content still refers to the same files.
-\end{itemize}
+\begin{equation} \label{eq:contentupdate}
+    \begin{split}
+        m' & = u, read(m) \land \\
+        m'' & = u, change(m') \land \\
+        m''' & = u, modify(m'', m) \Implies \\
+        m''' & \in M \land m \not \in M
+    \end{split}
+\end{equation}
+If a user reads content and then writes to it, the content is directly changed
+
+\begin{equation} \label{eq:containerupdate}
+    \begin{split}
+        c' & = u, read(c) \land \\
+        c'' & = u, change(c') \land \\
+        c''' & = u, modify(c'', c) \Implies \\
+        c''' & \in C \land c \not \in C
+    \end{split}
+\end{equation}
+If a user reads a container and then writes to it, the container is directly changed
+
+\begin{equation} \label{eq:copy}
+    \begin{split}
+        & f' = f \in m \\
+        & m' = u, copy(m) \Implies \\
+        & m' \neq m \land \\
+        & f' \in m' \land f' \in m
+    \end{split}
+\end{equation}
+If a user copies a content, the new content is not equals the old (because of its new ID), however
+they both refer to the same file.
+
+\begin{equation} \label{eq:snapshot}
+    \begin{split}
+        s = u, snapshot(c') \Implies \\
+        \forall c \in c', u, copy(c) \in s \land \\
+        \forall m \in c', u, copy(m) \in s \land \\
+        (\forall f \in c' \Implies f \in s)
+    \end{split}
+\end{equation}
+If a user creates a snapshot of a container the full container tree is re-created with new ids and
+its content still refers to the same files.
 
@@
 
-\subsection{Findings}
-\subsubsection{Scalability}
-When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
-the application has no limits of how much load and users it can handle as the hardware and nodes 
-can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
-implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
-support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
-is offered by ReactiveMongo is not needed as the load will not reach the peak (\fix relate to 
-graph) for what Casbah can handle on a single server.
-
-\subsubsubsection{Load Testing}
+\newpage
+\subsection{Load Testing}
 Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
@@
 
-\paragraph{GET requests}\mbox{}\\
+\subsubsection{GET requests}
+\label{sec:GET_REQUESTS}
 To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
@@
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=OKREQUEST,caption=Result of OK requests]
@@
 \end{lstlisting}
+\end{minipage}
 
+\begin{custommargins}{0cm}{-2cm}
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=DBREQUEST,caption=Result of MongoDB requests]
@@
 \end{lstlisting}
+\end{minipage}
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=STATICREQUEST,caption=Result of static text requests]
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 in the current management system reaches around 25KB in uncompressed size.
+\end{custommargins}
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=TREEREQUEST,caption=Result of project tree requests]
@@
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=CACHEREQUEST,caption=Result of cached project tree requests]
@@
 \end{lstlisting}
+\end{minipage}
+
 \par
@@
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=INDEXREQUEST,caption=Result of indexed project tree requests]
@@
 
-\paragraph{POST requests}\mbox{}\\
+\subsubsection{POST requests}
 Two different POST requests will be tested, creation of containers and creation of content together
@@
 
+\begin{minipage}{\linewidth-1cm}
 \begin{lstlisting}[label=CONTAINERPOSTREQUEST,caption=Result of container creation]
@@
 the server was able to handle within the time out, going from being able to handle 4000 documents to
-at least 50000 documents, which the browser instead will have problems displaying represented in the 
-UI without lag.
+at least 50000 documents, which the browser instead will have problems displaying, represented in the 
+UI, without lag.
+
+\subsection{Findings}
+\subsubsection{Scalability}
+When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
+the application has no limits of how much load and users it can handle as the hardware and nodes 
+can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
+implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
+support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
+is offered by ReactiveMongo is not needed as the load will not reach the peak, as discussed in 
+Section~\ref{sec:GET_REQUESTS}, for what Casbah can handle on a single server.
 

### 2016-02-02 tex/paper.tex
Added some consequences of multiple accesses

@@
 
-Copy-on-Write is used in for example virtual memory management systems~\cite{VIRTCOW}, for snapshot
-and as an optimisation technique for objects and types in several programming
+Copy-on-Write is used in for example virtual memory management systems~\cite{VIRTCOW}, snapshot
+algorithms and as an optimisation technique for objects and types in several programming
 languages\cite{LANGCOW}.
@@
 Its principle is that when processes or nodes share data in between each other, the data is not
-copied until one of the processes does changes to it. This is an optimisation as the processes does
+copied until one of the processes makes changes to it. This is an optimisation as the processes does
 not have to send or copy all of the related data that is in memory, rather they only have to send
@@
 photography where a photo can be seen as the state of what the photo is of, at a certain time.
-Snapshots should not be confused with full copies of a system as full copies can be used as backups
-meanwhile snapshots are not very effective means of backups in the case of data corruption. It is
-not effective against data corruption as snapshots usually still refer to unchanged data that is
-still a part of the system~\cite{SNAPSHOT}.
+Snapshots should not be confused with a full copy of a system, or part of a system, as full copies 
+can be used as backups meanwhile snapshots are not very effective means of backup in the case of 
+data corruption. It is not effective against data corruption as snapshots usually still refer to 
+unchanged data that is still a part of the system~\cite{SNAPSHOT}.
 
@@
 
-There could also be another relevant model done to show that content can not be accessed by
-unauthorized viewers once the content is uploaded to a CDN, but as that should already have been
-thoroughly checked by the CDN providers this work can focus solely on the internal users and content
-of the management system.
+There could also be another relevant part included in the model, to show that content can not be 
+accessed by unauthorized viewers once the content is uploaded to a CDN. But as that should already 
+have been thoroughly checked by the CDN providers this work can focus solely on the internal users 
+and content of the management system.
 
 \subsection{Related work}
-\fix Write about Bell-Lapadula et. al
 The model for this paper is based on the work that was done by Bell, D Elliott, LaPadula and Leonard
 J in their papers Secure computer systems: Mathematical foundations~\cite{BLP1} and Secure computer
-systems: A mathematical model~\cite{BLP2}. In these papers the foundation is laid for how to model
+systems: A mathematical model~\cite{BLP2}. In these papers the foundation was laid for how to model
 a computer system to be able to analyse the security of a system. Furthermore this paper was also 
@@
         \hline
-        Set & Elements & Semantics \\ \hline
+        \textbf{Set} & \textbf{Elements} & \textbf{Semantics} \\ \hline
         C   & $c_0\dots c_n$                & Containers; folders in the virtual file system\\ \hline
@@
 \end{equation}
+
+% The last one means that no users can delete actual files
 
@@
 
+\subsection{Consequences of modification}
 \begin{center}
@@
         \hline
-        Action & Semantics \\ \hline
+        \textbf{Action} & \textbf{Semantics} \\ \hline
         u, read(o)            & User u reads object o\\ \hline
@@
         u, change(o)          & User u locally changes object o \\ \hline
-        u, modify(o)          & User u globally modifies object o\\ \hline
         u, modify(o', o)      & User u globally modifies object o based on object o'\\ \hline
@@
 
-\fix Translate to mathematical expressions
-\subsubsection{Consequences of modification}
 \begin{equation} \label{eq:fileupdate}
@@
 its content still refers to the same files.
+
+\begin{equation} \label{eq:delete}
+    \begin{split}
+        f' = f \in (u, read(c)),  \\
+        u, delete(c) \Implies \\
+        c \not \in C \land f' \in F
+    \end{split}
+\end{equation}
+If a user deletes content, the file referred to in the content is not deleted.
+
+\subsection{Interaction of Multiple Accesses}
+\fix Rename consequence?
+
+As a result of consequence~\ref{eq:contentupdate} and ~\ref{eq:containerupdate} there can be race
+conditions where the last write wins. This could be solved by locks or transactions, but as updates
+are not based on each other, the trade-off for this inconsistency to scalability and speed is
+intentional. However as can also be seen in consequence~\ref{eq:fileupdate}, if a file is
+removed from a content that file is not removed from the set of files, even though its modification
+process is not the last one to write to the content. This is necessary as a file can be referenced
+from several content and the removal process can not atomically ensure that the file is not referred
+to by another content. This results in a lot of files in the system which are not referenced from
+anywhere. This can be solved by either making each file keep track of how many times it is referred
+to and also make it possible to check this number and delete the file in an atomic fashion. An
+easier approach would be to have a garbage collector which locks parts of the system momentarily
+meanwhile it removes files which are not referenced.
 

### 2016-02-02 tex/paper.tex
Added some CDN security details

@@
 \fix Clarify problem definition and research questions
+\section{Background}
+\subsection{About Uprise}
+Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
+experiences, which means that they are mostly not focussed on the actual gameplay which other EA 
+studios like DICE is.
+
+At the moment they are for example very involved with producing the Star Wars Battlefront game by
+making menu systems and developing the companion app, Battlefront companion.
 
@@
 \fix Write about how the approach is similar to BLP and that it is not a proof etc
+
 As this work only presents an informal model of how the system is designed it can not be regarded as
@@
 \subsection{Elements of the Model}
-\fix Mathematical formal description of how the system works
-
 \begin{center}
@@
 
-\fix Add content, m
 \begin{equation}
@@
 
+\label{sec:conseq}
 \subsection{Consequences of modification}
+
 \begin{center}
@@
 \end{equation}
-If a user reads content and then writes to it, the content is directly changed
+If a user reads content and then writes to it, the content is directly changed.
 
@@
 \end{equation}
-If a user reads a container and then writes to it, the container is directly changed
+If a user reads a container and then writes to it, the container is directly changed.
 
@@
 
-As a result of consequence~\ref{eq:contentupdate} and ~\ref{eq:containerupdate} there can be race
+As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} there can be race
 conditions where the last write wins. This could be solved by locks or transactions, but as updates
@@
 
+There is also the chance to have content and containers that are not referenced from anywhere and
+result in being garbage in the system. This could happen as a result of consequence~\ref{eq:delete}
+in combination with any of the other consequences listed in Section~\ref{sec:conseq}. The basic idea
+is that an entity is deleted meanwhile another is created or modified to have that entity as its
+parent and thus creating a separate tree that can not be reached from the root node of the project.
+
 \subsection{Findings}
@@
 \subsection{Background}
-\subsubsection{About Uprise}
-Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
-experiences, which means that they are mostly not focussed on the actual gameplay which other EA 
-studios like DICE is.
-
-At the moment they are for example very involved with producing the Star Wars Battlefront game by
-making menu systems and developing the companion app, Battlefront companion.
-
 \subsubsection{The current system}
@@
 \par
-As the project tree is built by querying documents parent id one optimisation that could be done was
+As the project tree is built by querying documents parent id an optimisation that could be done was
 to create an index of the documents parent ids. This drastically improved the size of the tree that
-the server was able to handle within the time out, going from being able to handle 4000 documents to
-at least 50000 documents, which the browser instead will have problems displaying, represented in the 
-UI, without lag.
+the server was able to handle within the time out, going from being able to handle 4000 documents in 
+one tree to at least 50000 documents, which the browser instead will have problems displaying, 
+represented in the UI, without lag.
 
@@
 \subsubsection{Security of the system}
-\paragraph{Authorization}
-The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius is mainly
-focussed at being deployed in internal networks which usually has an LDAP service enabled. LDAP also
-makes it easier for the user to login as no separate account is needed for Perius and thus the user
-can use the same account as for all the LDAP connected services on the internal network.
+\subsubsubsection{Authorization}
+The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
+is mainly focussed at being deployed in internal networks which usually has an LDAP service enabled. 
+LDAP also makes it easier for the user to login as no separate account is needed for Perius and 
+thus the user can use the same account as for all the LDAP connected services on the internal 
+network.
 
-\paragraph{Audit logs}
+\subsubsubsection{Audit logs}
+
+\subsubsubsection{CDN Connections}
+\fix Describe how the CDN connections are encrypted and secure.
+
+When uploading files to the CDN networks their TLS/SSL protected API's are used to ensure that no
+data leaks through packet sniffing etc. The files are grouped corresponding to their parent
+containers in Perius, which makes it possible to handle the security settings for all files in a
+group at once, without having to traverse through the tree in Perius.
+
+\fix Describe the different CDN mechanisms, like signed cookies and urls
 

### 2016-02-03 tex/paper.tex
Added introduction to serving private content

@@
 \subsection{Background}
-\subsubsection{The current system}
+\subsubsection{The current system} \label{sec:current_system}
 Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
@@
 
+\subsection{Security of the system}
+\subsubsection{Authorization}
+The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
+is mainly focussed at being deployed in internal networks which usually has an LDAP service enabled. 
+LDAP also makes it easier for the user to login as no separate account is needed for Perius and 
+thus the user can use the same account as for all the LDAP connected services on the internal 
+network.
+
+\subsubsection{Audit logs}
+
+\subsubsection{CDN Connections}
+\fix Describe how the CDN connections are encrypted and secure.
+
+When uploading files to the CDN networks, their TLS/SSL protected API's are used to ensure that no
+data leaks through packet sniffing etc. The files are grouped corresponding to their parent
+containers in Perius, which makes it possible to handle the security settings for all files in a
+group at once, without having to traverse through the tree in Perius.
+
+\subsubsubsection{Serving private content}
+\fix Describe the different CDN mechanisms, like signed cookies and urls
+
+There are three different ways, that this implementation make use of, to make sure that 
+the CDNs serve content in a private manner. The first two is signed URLs and signed cookies, and 
+the third is a mixture between one of the first two together with a range of IP
+addresses.~\cite{AWSPRIVATE} 
+These mechanisms are needed, as explained in Section~\ref{sec:current_system}, to ensure that
+assets are not leaked if an attacker manages to find out the URL for the asset.
+
+\par \textbf{Signed URLs} \\
+test test
+
+\par \textbf{Signed cookies} \\
+test test
+
+\par \textbf{IP range restriction} \\
+test test
+
 \subsection{Findings}
@@
 Section~\ref{sec:GET_REQUESTS}, for what Casbah can handle on a single server.
-
-\subsubsection{Security of the system}
-\subsubsubsection{Authorization}
-The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
-is mainly focussed at being deployed in internal networks which usually has an LDAP service enabled. 
-LDAP also makes it easier for the user to login as no separate account is needed for Perius and 
-thus the user can use the same account as for all the LDAP connected services on the internal 
-network.
-
-\subsubsubsection{Audit logs}
-
-\subsubsubsection{CDN Connections}
-\fix Describe how the CDN connections are encrypted and secure.
-
-When uploading files to the CDN networks their TLS/SSL protected API's are used to ensure that no
-data leaks through packet sniffing etc. The files are grouped corresponding to their parent
-containers in Perius, which makes it possible to handle the security settings for all files in a
-group at once, without having to traverse through the tree in Perius.
-
-\fix Describe the different CDN mechanisms, like signed cookies and urls
 

### 2016-02-03 tex/paper.tex
Added some future work

@@
 \begin{abstract}
-    \fix Abstract
+    \fix Abstract is written last
 
@@
 problem and an inconvenience, and that is what this thesis is about. This work examines ways of
-enforcing virtual access control on content and groups of content, in the form of folders and
+enforcing virtual access control on content and groups of content, in the form of containers and
 snapshots. A system was developed to make the underlying theory work in practice.
@@
 \fix Clarify problem definition and research questions
+\newpage
 \section{Background}
@@
 \newpage
-
 \label{sec:model}
@@
 a proof for the actual implementation of the system to be flawless. The model should how ever give a 
-strong idea of the soundness of the designed system.
+strong idea of the soundness of the design of the system.
 
@@
 
-\label{sec:conseq}
-\subsection{Consequences of modification}
+\subsection{Consequences of modification} \label{sec:conseq}
 
@@
 
-\subsection{Interaction of Multiple Accesses}
+\subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
 \fix Rename consequence?
-
+\\
 As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} there can be race
@@
 meanwhile it removes files which are not referenced.
-
+\\
 There is also the chance to have content and containers that are not referenced from anywhere and
@@
 parent and thus creating a separate tree that can not be reached from the root node of the project.
+These type of inconveniences should also be cleaned up by the garbage collector.
 
 \subsection{Findings}
+\fix Add more findings
+\subsubsection{Garbage Collection}
+For the system to reach the decided eventual consistency~\cite{KLINGSBO}, garbage collection will be
+needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section ~\ref{sec:conseq} 
+there will be objects in the system that can not be reached and should therefore be cleaned up for 
+them not to cause negative performance effects on the system.
 
@@
 \subsubsubsection{Serving private content}
-\fix Describe the different CDN mechanisms, like signed cookies and urls
-
 There are three different ways, that this implementation make use of, to make sure that 
-the CDNs serve content in a private manner. The first two is signed URLs and signed cookies, and 
+the CDNs serve content in a private manner. The first two are signed URLs and signed cookies, and 
 the third is a mixture between one of the first two together with a range of IP
@@
 \par \textbf{Signed URLs} \\
-test test
+Signed URLs work by writing a policy statement that specifies the restrictions that should apply for
+the asset the URL is referring to. There are two different types of these policies; canned and
+custom. In the canned policy there is only an option to specify the date for when the URL is no 
+longer valid. In the custom policy it is possible to also specify the date when the asset should be
+made available, ranges of IP addresses (which is discussed more in a later paragraph) and inclusion
+of the base64 version of the policy in the URL~\cite{AWSSIGNED}. For custom policies it is also 
+possible to reuse the policy and have it refer to multiple assets.\\
 
 \par \textbf{Signed cookies} \\
-test test
+Signed cookies work very similarly to signed URLs, the same type of canned and custom policies can
+be set, except for the policy to include the base64 version of the policy in the URL. Signed cookies
+are to be used when the URL should remain static even though the policies change~\cite{AWSCOOKIES}.
+\\
 
 \par \textbf{IP range restriction} \\
-test test
+The IP range restriction is presented as a third option in Perius, even though it in fact is just a
+part of a Signed Cookie or URL policy. The IP range restriction makes it possible to restrict which
+IP addresses that can access the asset, this option can be fitting when for example an office or
+companies internet access is based on a set of public static IP addresses and thus restricts the
+content from ever being accessed outside of that controlled network.
 
@@
 
+\subsubsection{Security}
+\fix Write what security implications that can be seen
+
+\newpage
 \section{JPF}
@@
 
+\newpage
 \section{Discussion}
-
+\subsection{Persistent Storage}
+More research could have been done in the choosing of persistent storage, as mentioned in ``NoSQL:
+Moving from MapReduce Batch Jobs to Event-Driven Data Collection''~\cite{KLINGSBO} many applications
+that choose NoSQL databases as their persistent storage actually don't need it. This is most likely
+true for Perius too, it would have been efficient enough generating the project trees from an
+indexed SQL database with foreign keys. On the other hand it was quite nice having the BSON
+documents for insertion as they were so similar the accepted JSON format from the REST service.
+As Perius most common operations involve modifying the project tree, a graph database should have 
+been researched too.
+
+\newpage
 \section{Summary}
@@
 \subsection{Future work}
+\subsubsection{Access Control}
 Full access control was not implemented according to the model described in ~\ref{sec:model}, it
@@
 implementation did not set or check any specific access rights to certain contents or containers.
+
+\subsubsection{Front-end Refactor}
+The application could be made substantially more efficient by rewriting the front-end to update
+itself according to the REST response after a modification of the project tree, instead of 
+re-fetching the full project tree every time a change is made.
 

### 2016-02-04 tex/paper.tex
Modified introduction

@@
 \section{Introduction}
-Developing larger projects containing static content usually involves using a Content Distribution
-Network to be able to scale to a large user base. The commercial Content Distribution Networks are
+Developing large projects containing static content usually involves using a Content Distribution
+Network to be able to scale to a larger user base. The commercial Content Distribution Networks are
 usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
@@
 \subsection{About Uprise}
-Uprise is a company based in Uppsala, Sweden. It is an EA studio focussing on creating great gaming
-experiences, which means that they are mostly not focussed on the actual gameplay which other EA 
-studios like DICE is.
-
-At the moment they are for example very involved with producing the Star Wars Battlefront game by
-making menu systems and developing the companion app, Battlefront companion.
+Uprise (formerly known as ESN) is a company based in Uppsala, Sweden. It is an EA studio focussing 
+on creating great gaming experiences, which means that they are mostly not focussed on the actual 
+gameplay which other EA studios like DICE is. Currently Uprise has a lot of focus on developing 
+companion apps and a new form of menu systems.
 
@@
 strong idea of the soundness of the design of the system.
+
+The copy-on-write system is only used for files as all of the other objects in the system is
+basically just meta data and not classed as important, from an data integrity point of view, as the 
+actual files themselves.
 
@@
 needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
-ranges of a file, although that advantage is not needed in the Perius implementation. The 
-disadvantage of using this approach is that when using a non-distributed database it us slower 
-to read and write to the database than reading or writing directly to the filesystem. Another 
-disadvantage is that to access the files it is needed to go through the database layer in some 
-way, instead of accessing the filesystem.
+ranges of a file, although that advantage is not needed in the Perius implementation. For
+scalability this could be used to retrieve different parts of a file from different servers, 
+normal load balancing would probably work better in a system like Perius where no extremely large 
+files are expected to be stored.
+
+The disadvantage of using the GridFS approach is that when using a non-distributed database 
+it is slower to read and write to the database than reading or writing directly to the filesystem. 
+Another disadvantage is that to access the files it is needed to go through the database layer 
+in some way, instead of accessing the filesystem.
 

### 2016-02-09 tex/paper.tex
Added some conclusions

@@
 \newpage
-\section{Implementation}
+\section{Implementation} \label{sec:implementation}
 \subsection{Background}
@@
 \newpage
-\subsection{Load Testing}
+\subsection{Load Testing} \label{sec:load_testing}
 Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
@@
 
-\subsubsection{CDN Connections}
+\subsubsection{CDN Connections} \label{sec:cdn_connections}
 \fix Describe how the CDN connections are encrypted and secure.
@@
 
-\subsubsubsection{Serving private content}
+\subsubsubsection{Serving private content} \label{sec:private_content}
 There are three different ways, that this implementation make use of, to make sure that 
@@
 \subsection{Conclusions}
+The goal of this thesis was to see whether it was feasible to use copy-on-write in a high level
+application. As the implementation (see Section~\ref{sec:implementation}) made as a part of this 
+thesis project is already replacing its predecessor the simple conclusion is that it definitely is 
+possible. In the beginning of the project thoughts were on having every element being operated upon
+in a copy-on-write fashion but this was later narrowed down to only have the most important part,
+the files, as copy-on-write. This was due to that the conclusion that it did not matter if the 
+other elements were resolved in a last-write-wins manner when modified.
+
+\par If this application would be distributed with several persistent storage nodes like MongoDB, the
+application would not always be in a global consistent state as there would not be any global locks.
+This could theoretically cause some inconvenience for the user but in all real world tests no 
+users have noticed it. The theoretical inconvenience for the user was a trade-off made so that the
+application could scale on the width almost endlessly, especially if the issue with building the
+project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
+
+\par The positive effects of deploying Perius instead of its predecessor was not only on about
+scaling, it also features a more secure way of handling private assets 
+(Section~\ref{sec:current_system},~\ref{sec:private_content}) and reduces the
+administration needed to control the security settings of large groups of assets
+(Section~\ref{sec:cdn_connections}) at the same time.
 

### 2016-02-10 tex/paper.tex
Added security findings

@@
 %}}}
-\title{\textbf{Designing a Virtual Security Management Layer for Cloud Content}}
+\title{\textbf{Designing a Security Management Layer for CDN Assets}}
 \author{Lukas Klingsbo}
@@
 
-The research question that this report answers is how and whether it is practically feasible to use
-Copy-on-Write for a high-level system like the one that is implemented.
+The research question that this report tries to answers is how and whether it is practically 
+feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
+it can effect the systems scalability.
 
@@
 
-When uploading files to the CDN networks, their TLS/SSL protected API's are used to ensure that no
+When uploading files to the CDN networks, their TLS/SSL protected APIs are used to ensure that no
 data leaks through packet sniffing etc. The files are grouped corresponding to their parent
 containers in Perius, which makes it possible to handle the security settings for all files in a
-group at once, without having to traverse through the tree in Perius.
+group at once, without having to manually traverse through the tree in Perius.
 
@@
 \subsubsection{Security}
-\fix Write what security implications that can be seen
+The greatest security improvement of Perius compared to its predecessor Battlebinary is not the
+security of the system itself but rather how Perius handles the contents security settings on the
+CDN providers. Perius uses several of the security features for private content that the CDN
+providers offer (see Section~\ref{sec:private_content}) to make sure that the content only is
+accessible from where it should be. In Battlebinary the content was secured by adding a part of the
+hash of the file (see Section~\ref{sec:current_system}) to the filename that it was uploaded with, 
+which resulted in a URL on the CDN that was not guessable but if the link somehow leaked the 
+content that the link referred to would be open to anyone.
 
@@
 
-\par The positive effects of deploying Perius instead of its predecessor was not only on about
+\par The positive effects of deploying Perius instead of its predecessor was not only about
 scaling, it also features a more secure way of handling private assets 

### 2016-02-12 tex/paper.tex
Added discussion about COW scalability

@@
 
+\subsubsection{Copy-on-Write's effect on Scalability}
+As no comparison with solutions that were not Copy-on-Write based were made, no scientific
+conclusions can be drawn from how its effect was on the scalability of the system. However some
+intuitively drawn conclusions can be suggested. If comparing to a system which uses a SQL database
+the Copy-on-write system would have an easier time to scale on the width as no global locks or 
+transactions has to be used, which with high probability would slow down the system, especially 
+when widely distributed. 
+The gain of having the locks or transactions would be that the system would always be
+in a fully consistent state. The loss would be that all of the database nodes would need to 
+constantly have to be fully aware of the state of the other nodes.
+
+\par A more interesting solution would be to not implement any Copy-on-Write system and simply use 
+a persistent storage that scales with eventual consistency. That implementation would most likely be
+more efficient than the current solution, with MongoDB as a persistent storage and the Copy-on-Write
+handling in the back-end code, as a lot of optimisations could be done in the database engine
+instead.
+
 \newpage

### 2016-02-15 tex/paper.tex
Added Implementation introduction

@@
 \label{sec:model}
-\section{Model}
+\section{Model} \label{sec:model}
 The model for this work should show how the data can not be accessed or modified by unauthorized
-users and how the integrity of the data is always kept in the Perius system.
+users and how the defined integrity of the data is always kept in the Perius system.
 
@@
 have been thoroughly checked by the CDN providers this work can focus solely on the internal users 
-and content of the management system.
+and content in the management system.
 
 \subsection{Related work}
-The model for this paper is based on the work that was done by Bell, D Elliott, LaPadula and Leonard
-J in their papers Secure computer systems: Mathematical foundations~\cite{BLP1} and Secure computer
-systems: A mathematical model~\cite{BLP2}. In these papers the foundation was laid for how to model
-a computer system to be able to analyse the security of a system. Furthermore this paper was also 
-inspired by Biba, Integrity considerations for secure computer systems~\cite{BIBA}, where many of 
-the points made by him was taken into consideration when inspecting that the integrity of the data 
-was always sound.
+The model for this paper is based on the work that was done by Bell, D Elliott, LaPadula and 
+Leonard J in their papers \textit{Secure computer systems: Mathematical foundations}~\cite{BLP1} 
+and \textit{Secure computer systems: A mathematical model}~\cite{BLP2}. 
+In these papers the foundation was laid for how to model a computer system to be able to analyse 
+the security of a system. Furthermore this paper was also inspired by Biba, \textit{Integrity 
+considerations for secure computer systems}~\cite{BIBA}, where many of the points made by him was 
+taken into consideration when inspecting that the integrity of the data was always sound.
 
@@
 \section{Implementation} \label{sec:implementation}
+This chapter deals with the second part of this work, namely the implementation of the Perius 
+system based which is based on the model described in Section~\ref{sec:model}. As this work is not
+intended as a instructions manual more information about how to deploy and use Perius can be found
+on \url{http://perius.se}. This chapter's focus is instead on how the decisions for the
+implementation were made, how well the system would scale and the resulting security implications.
+
 \subsection{Background}

### 2016-02-15 tex/paper.tex
Added finding for the model

@@
 \section{Related Terminology}
+In this section concepts and abbreviations, that are recurring throughout the paper, which the
+reader needs to be familiar with are explained.
+
 \subsection{Abbreviations}
@@
 
+\subsubsection{Eventual Consistency}
+The expression eventual consistency is a term used for describing how data is processed. It is often
+used in the context of database systems. Eventual consistency means that stored data is guaranteed 
+to be consistent in any one moment, but at some point the data will eventually converge to being 
+consistent~\cite{EVENTUAL}.
+
+The effect of this is that persistent storages, especially distributed databases, can
+operate faster as there is no need for locks and transactions for any operations. As no locks or
+transactions are used, operations on the same entities can be done before earlier operations have
+propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
+that needs to be handled, for example two operations that wants to change the same data at once. To
+solve this the system needs to have defined which operation that will take precedence and this is
+usually called conflict or sibling resolution.
+
 \newpage
-\label{sec:model}
 \section{Model} \label{sec:model}
@@
 and \textit{Secure computer systems: A mathematical model}~\cite{BLP2}. 
-In these papers the foundation was laid for how to model a computer system to be able to analyse 
-the security of a system. Furthermore this paper was also inspired by Biba, \textit{Integrity 
+In these papers the foundation was laid for how to model computer systems to be able to analyse 
+the security of them. Furthermore this paper was also inspired by Biba, \textit{Integrity 
 considerations for secure computer systems}~\cite{BIBA}, where many of the points made by him was 
@@
 
-The copy-on-write system is only used for files as all of the other objects in the system is
+The Copy-on-Write system is only used for files as all of the other objects in the system is
 basically just meta data and not classed as important, from an data integrity point of view, as the 
@@
 In this work external threats are handled as threats that can occur from what has been exposed by
-the API (See Section~\ref{sec:API} and internal threats as incorrect implementation. As the server
-and its system are assumed safe malicious subsystems are not considered.
+the API (See Section~\ref{sec:API}) and internal threats as incorrect implementation. As the server
+and its system are assumed, safe malicious subsystems are not considered.
 
 \subsection{Consequences of modification} \label{sec:conseq}
+
+In this section the consequences of different operations are stated in an informal mathematical
+manner. These rules are what should be expected to be followed by the implementation in
+Section~\ref{sec:implementation}.
 
@@
 \\
-As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} there can be race
-conditions where the last write wins. This could be solved by locks or transactions, but as updates
-are not based on each other, the trade-off for this inconsistency to scalability and speed is
-intentional. However as can also be seen in consequence~\ref{eq:fileupdate}, if a file is
-removed from a content that file is not removed from the set of files, even though its modification
-process is not the last one to write to the content. This is necessary as a file can be referenced
-from several content and the removal process can not atomically ensure that the file is not referred
-to by another content. This results in a lot of files in the system which are not referenced from
-anywhere. This can be solved by either making each file keep track of how many times it is referred
-to and also make it possible to check this number and delete the file in an atomic fashion. An
-easier approach would be to have a garbage collector which locks parts of the system momentarily
-meanwhile it removes files which are not referenced.
-\\
+As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
+Section~\ref{sec:conseq} there can be race conditions where the last write wins. This could be
+solved by locks or transactions, but as updates are not based on each other, the trade-off for this
+inconsistency to scalability and speed is intentional. However as can also be seen in
+consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
+the set of files, even though its modification process is not the last one to write to the content.
+This is necessary as a file can be referenced from several content and the removal process can not
+atomically ensure that the file is not referred to by another content. This results in a lot of
+files in the system which are not referenced from anywhere. This can be solved by either making each
+file keep track of how many times it is referred to and also make it possible to check this number
+and delete the file in an atomic fashion. An easier approach would be to have a garbage collector
+which locks parts of the system momentarily meanwhile it removes files which are not referenced.
+
+\par
 There is also the chance to have content and containers that are not referenced from anywhere and
@@
 them not to cause negative performance effects on the system.
+
+\subsubsection{Conflict Resolution}
+If the persistent storage would be distributed the different database nodes would have to have rules
+set up for conflict resolution as this model does not consider what to do if the same meta data is
+modified on the nodes before they have time to synchronise. The easiest way to handle this is to
+have the clocks on the nodes synchronised and timestamp each change and when synchronising the nodes
+only consider the newest data if there is a conflict. This will have the same result as concurrent
+modification with a single database node with this model, namely last write
+wins~\cite{LASTWRITEWINS}.
 
@@
 
-As the persistent storage, used in this implementation (Section~\ref{persistent_storage}), does not
+As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does not
 implement transactions or locks a lot of different problems can occur when several clients are
@@
 
-\subsubsection{Persistent storage}
-\label{persistent_storage}
+\subsubsection{Persistent storage} \label{sec:persistent_storage}
 \subsubsubsection{MongoDB}
@@
 
-\subsubsection{GET requests}
-\label{sec:GET_REQUESTS}
+\subsubsection{GET requests} \label{sec:GET_REQUESTS}
 To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
@@
 \subsubsection{Audit logs}
+Audit logs are used in the system to keep track of which users that perform which actions. In the
+event that somebody, consciously or unconsciously, are leaking private CDN content from within the
+organisation using Perius the audit trail can be used to find which LDAP account that was
+responsible for the leak.
 
 \subsubsection{CDN Connections} \label{sec:cdn_connections}
-\fix Describe how the CDN connections are encrypted and secure.
-
 When uploading files to the CDN networks, their TLS/SSL protected APIs are used to ensure that no
@@
 \subsubsubsection{Serving private content} \label{sec:private_content}
-There are three different ways, that this implementation make use of, to make sure that 
-the CDNs serve content in a private manner. The first two are signed URLs and signed cookies, and 
-the third is a mixture between one of the first two together with a range of IP
-addresses.~\cite{AWSPRIVATE} 
-These mechanisms are needed, as explained in Section~\ref{sec:current_system}, to ensure that
-assets are not leaked if an attacker manages to find out the URL for the asset.
+There are three different ways, that this implementation make use of, to make sure that the CDNs
+serve content in a private manner. The first two are signed URLs and signed cookies, and the third
+is a mixture between one of the first two together with a range of IP addresses.~\cite{AWSPRIVATE}
+These mechanisms are needed, as explained in Section~\ref{sec:current_system}, to ensure that assets
+are not leaked if an attacker manages to find out the URL for the asset.
 
@@
 \subsection{Persistent Storage}
-More research could have been done in the choosing of persistent storage, as mentioned in ``NoSQL:
-Moving from MapReduce Batch Jobs to Event-Driven Data Collection''~\cite{KLINGSBO} many applications
-that choose NoSQL databases as their persistent storage actually don't need it. This is most likely
-true for Perius too, it would have been efficient enough generating the project trees from an
-indexed SQL database with foreign keys. On the other hand it was quite nice having the BSON
-documents for insertion as they were so similar the accepted JSON format from the REST service.
-As Perius most common operations involve modifying the project tree, a graph database should have 
-been researched too.
+More research could have been done in the choosing of persistent storage, as mentioned in 
+\textit{``NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection''}
+~\cite{KLINGSBO} many applications that choose NoSQL databases as their persistent storage 
+actually don't need it. This is most likely true for Perius too, it would have been efficient 
+enough generating the project trees from an indexed SQL database with foreign keys. On the 
+other hand it was quite nice having the BSON documents for insertion as they were so similar 
+the accepted JSON format from the REST service. As Perius most common operations involve 
+modifying the project tree, a graph database should have been researched too.
 

### 2016-02-16 tex/paper.tex
Added some introductions

@@
 \section{Background}
+This section gives the reader some background of the company which the thesis was done at and
+explanation of the core concepts which are exerted.
+
 \subsection{About Uprise}
 Uprise (formerly known as ESN) is a company based in Uppsala, Sweden. It is an EA studio focussing 
-on creating great gaming experiences, which means that they are mostly not focussed on the actual 
-gameplay which other EA studios like DICE is. Currently Uprise has a lot of focus on developing 
+on creating great gaming experiences, which means that they are not focussed on the actual 
+gameplay, which other EA studios like DICE are. Currently Uprise has a lot of focus on developing 
 companion apps and a new form of menu systems.
@@
 \newpage
-\section{Related Terminology}
+\subsection{Related Terminology}
 In this section concepts and abbreviations, that are recurring throughout the paper, which the
@@
 
-\subsection{Abbreviations}
-\subsubsection{JPF}
+\subsubsection{Abbreviations}
+\subsubsubsection{JPF}
 Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
@@
 
-\subsubsection{CDN}
+\subsubsubsection{CDN}
 Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
@@
 
-\subsection{Terms}
-\subsubsection{Snapshot}
+\subsubsection{Terms}
+\subsubsubsection{Snapshot}
 A snapshot is a way to record the full state of a system at a specific time. The term comes from
@@
 
-\subsubsection{Eventual Consistency}
+\subsubsubsection{Eventual Consistency}
 The expression eventual consistency is a term used for describing how data is processed. It is often
@@
 usually called conflict or sibling resolution.
+
+%\subsubsubsection{Perius}
+%Perius is the name of the implementation that was made during this project. It is an anagram of
+%Uprise and also a type of butterfly.
 
@@
 \subsection{Approach}
-\fix Write about how the approach is similar to BLP and that it is not a proof etc
-
 As this work only presents an informal model of how the system is designed it can not be regarded as
@@
 \section{Implementation} \label{sec:implementation} 
-This chapter deals with the second part of this work, namely the implementation of the Perius 
-system based which is based on the model described in Section~\ref{sec:model}. As this work is not
-intended as a instructions manual more information about how to deploy and use Perius can be found
-on \url{http://perius.se}. This chapter's focus is instead on how the decisions for the
-implementation were made, how well the system would scale and the resulting security implications.
+This chapter deals with the second part of this work, namely the implementation of the Perius system
+based which is based on the model described in Section~\ref{sec:model}. As this work is not intended
+as an instructions manual, more information about how to deploy and use Perius can be found on
+\url{http://perius.se}. This chapter's focus is instead on how the decisions for the implementation
+were made, how well the system would scale and the resulting security implications.
 
@@
 Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
-mostly images, to content delivery networks. The current system does not make use out of the
-security features that the CDN's are offering, instead it uses a form of security by obscurity. When
+mostly images, to content delivery networks. The current system does not make use of the
+security features that the CDNs are offering, instead it uses a form of security by obscurity. When
 a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
@@
 In the current system you can only upload a file once as there will be a collision in the upload
-otherwise, as the old and the new file will have the same MD5 hash.
+otherwise, as the old and the new file will have the same MD5 hash, which is by design as duplicate
+content just wastes space on the CDN. The problem with this is that the current interface does not
+handle it very well as one file can not be shown or uploaded in two different places in the virtual
+filesystem.
 
@@
 
-\fix Insert more systems here
+\fix Insert more systems here and add a comparison of the systems
 
@@
 The service features a virtual file structure over the assets that has been stored, snapshots,
-security management of whole containers as well as individual files, audit and access logging, multi
+security management of whole containers as well as individual files, audit and auth logging, multi
 project support and a modular design for persistent storage.
@@
 \end{minipage}
+\end{custommargins}
 
@@
 involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST} one can
-see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when comparing 
-that to the requests that needed database access it is obvious that the HTTP server can handle all 
-the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents in a 
-collection (in this case only one) is dramatically slower, the throughput was only \textasciitilde7000
-requests/s.
-
-\par
-These are very simple requests, to evaluate how much load the application can handle in its current
-state a more complex but common request has to be analysed. The most complicated and still common
-request that system is receiving is requests of a full project tree. This request is computationally
-expensive as the tree is not stored directly in the database but has to be built from the id and
-parent id of each container and content, a simulated project tree of similar size to the ones stored 
-in the current management system reaches around 25KB in uncompressed size.
-\end{custommargins}
+see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when
+comparing that to the requests that needed database access it is obvious that the HTTP server can
+handle all the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents
+in a collection (in this case only one) is dramatically slower, the throughput was only
+\textasciitilde7000 requests/s.
+
+\par These are very simple requests, to evaluate how much load the application can handle in its
+current state a more complex but common request has to be analysed. The most complicated and still
+common request that system is receiving is requests of a full project tree. This request is
+computationally expensive as the tree is not stored directly in the database but has to be built
+from the id and parent id of each container and content, a simulated project tree of similar size to
+the ones stored in the current management system reaches around 25KB in uncompressed size.
 
@@
 \par
-Another simple optimisation that was added was indices on the parent id's, which are heavily queried
+Another simple optimisation that was added was indices on the parent ids, which are heavily queried
 when building the project trees.
@@
 These mechanisms are needed, as explained in Section~\ref{sec:current_system}, to ensure that assets
-are not leaked if an attacker manages to find out the URL for the asset.
+are not leaked if an attacker manages to find out the URL for the asset. \\
 
@@
 %Battle Binary
-%Security features of CDN's
+%Security features of CDNn's
 %What is needed by the logic, like what happens after two consecutive deletes

### 2016-02-17 tex/paper.tex
Added a conclusion

@@
 
+\par The conclusion that can be drawn from implementing Copy-on-Write as high up in the stack as it
+was in Perius is that the code would have to be largely rewritten for every application that wants
+to use a similar approach. The solution to this could be to use Copy-on-Write in the persistent
+storage instead of in the actual application, which also efficiently could handle conflict 
+resolution and every improvement made to it could be shared for every system that uses the same 
+type of system architecture and underlying persistent storage.
+
 \subsection{Future work}

### 2016-02-17 tex/paper.tex
Moved the JPF section

@@
 
-%\subsubsubsection{Perius}
-%Perius is the name of the implementation that was made during this project. It is an anagram of
-%Uprise and also a type of butterfly.
+\subsubsubsection{Perius}
+Perius is the name of the implementation that was made during this project. It is an anagram of
+Uprise and also a type of butterfly.
 
@@
 
+\subsection{JPF}
+\fix Not sure where to put this section
+
+JPF was used to test the idea and state transitioning of the application. It's a very simplified
+version of the real system that still contains all the important Copy-on-Write core concepts and the 
+assumptions that have been made in the model. This simplified version could then be automatically
+tested for soundness. It is not a proof that the model works, but it is very exhaustive in its
+testing.
+
+\subsubsection{Entities}
+\begin{figure}[htp] \centering{
+    \includegraphics[scale=0.8]{relation.pdf}}
+    \caption{High Level Entity Relationships}
+    \label{fig:relation}
+\end{figure}
+
+\par \textbf{Content} \\
+Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
+content can refer to for example an image, video or binary blob.\\
+
+\par \textbf{Project} \\
+A project is what is created to contain all content and containers related to a real project. Files
+can be changed within a project and the system can contain several projects and their virtual
+content are completely disjoint.\\
+
+\par \textbf{Container} \\
+A container is a virtual folder within a project which can contain content and other containers.\\
+
+\par \textbf{Snapshot} \\
+A snapshot is a read-only container from the state which the container the was in when the snapshot
+was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
+Snapshots are by default stored as siblings to the container which they were made from, but they can
+be contained by any container.\\
+
+\par \textbf{File} \\
+A file refers to an actual physical file. Files are stored in the database to make backup,
+deployment and migration easier.\\
+
+\par \textbf{User} \\
+A user is the structure that handles people who have been granted access to the system. Access to
+the system is handled by a separate service, like LDAP.\\
+
+\subsubsection{Execution}
+Java path finder was used to show that the model and plan of how to build the system was sound. The
+model was built in Java with the objective of being as reduced and simple as possible, without
+loosing any of the cases that needed to be covered by the model checker. As the users are mainly
+going to be handled by external systems they were not included in the model.
+
+Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
+Each client was represented by a thread and each action taken by the client was randomised. The id
+hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
+each object had its own id, generated in the same fashion as the real implementation is using,
+randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
+transactions were used and the threads were running fully concurrently, without any sleep
+statements.
+
+ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can't be
+iterated over concurrently.
+
+JPF checked each permutation of states that the threads can end up in, the result of the run can be
+seen in Listing~\ref{JPFRESULT}.
+
+\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
+elapsed time:       14:26:53
+states:             new=160853259,
+                    visited=451102505,
+                    backtracked=611955764,
+                    end=21640
+search:             maxDepth=380,
+                    constraints=0
+choice generators:  thread=160853255 
+                    (signal=0,
+                    lock=3603938,
+                    sharedRef=146989208,
+                    threadApi=3,
+                    reschedule=10260106), 
+                    data=0
+
+heap:               new=676056850,
+                    released=435060996,
+                    maxLive=655,
+                    gcCycles=523950061
+
+instructions:       11917045758
+max memory:         6256MB
+loaded code:        classes=111,
+                    methods=2179
+\end{lstlisting}
+
 \subsection{Findings}
@@
 \newpage
-\section{JPF}
-\fix Not sure where to put this section
-
-JPF was used to test the idea and state transitioning of the application. Its a very simplified
-version of the real system that still contains all the important Copy-on-Write core concepts and the 
-assumptions that have been made for the model. This simplified version could then be automatically
-tested for soundness. It is not a proof that the model works, but it is very exhaustive in its
-testing.
-
-\subsection{Entities}
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.8]{relation.pdf}}
-    \caption{High Level Entity Relationships}
-    \label{fig:relation}
-\end{figure}
-
-\subsubsection{Content}
-Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
-content can refer to for example an image, video or binary blob.
-
-\subsubsection{Project}
-A project is what is created to contain all content and containers related to a real project. Files
-can be changed within a project and the system can contain several projects and their virtual
-content are completely disjoint.
-
-\subsubsection{Container}
-A container is a virtual folder within a project which can contain content and other containers.
-
-\subsubsection{Snapshot}
-A snapshot is a read-only container from the state which the container the was in when the snapshot
-was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
-Snapshots are by default stored as siblings to the container which they were made from, but they can
-be contained by any container.
-
-\subsubsection{File}
-A file refers to an actual physical file. Files are stored in the database to make backup,
-deployment and migration easier.
-
-\subsubsection{User}
-A user is the structure that handles people who have been granted access to the system. Access to
-the system is handled by a separate service, like LDAP.
-
-\subsection{Execution}
-Java path finder was used to show that the model and plan of how to build the system was sound. The
-model was built in Java with the objective of being as reduced and simple as possible, without
-loosing any of the cases that needed to be covered by the model checker. As the users are mainly
-going to be handled by external systems they were not included in the model.
-
-Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
-Each client was represented by a thread and each action taken by the client was randomised. The id
-hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
-each object had its own id, generated in the same fashion as the real implementation is using,
-randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
-transactions were used and the threads were running fully concurrently, without any sleep
-statements.
-
-ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can't be
-iterated over concurrently.
-
-JPF checked each permutation of states that the threads can end up in, the result of the run can be
-seen in Listing~\ref{JPFRESULT}.
-
-\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
-elapsed time:       14:26:53
-states:             new=160853259,
-                    visited=451102505,
-                    backtracked=611955764,
-                    end=21640
-search:             maxDepth=380,
-                    constraints=0
-choice generators:  thread=160853255 
-                    (signal=0,
-                    lock=3603938,
-                    sharedRef=146989208,
-                    threadApi=3,
-                    reschedule=10260106), 
-                    data=0
-
-heap:               new=676056850,
-                    released=435060996,
-                    maxLive=655,
-                    gcCycles=523950061
-
-instructions:       11917045758
-max memory:         6256MB
-loaded code:        classes=111,
-                    methods=2179
-\end{lstlisting}
-
-\newpage
 \section{Discussion}
@@
 
-\subsubsection{Front-end Refactor}
+\subsubsection{Front-end Refactorization}
 The application could be made substantially more efficient by rewriting the front-end to update
 itself according to the REST response after a modification of the project tree, instead of 
-re-fetching the full project tree every time a change is made.
+re-fetching the full project tree every time a change is made. The front-end should also be fixed to
+properly follow the Reflux streamlining.
 

### 2016-02-18 tex/paper.tex
Fixed figures and added stack explanation

@@
 
-JPF was used to test the idea and state transitioning of the application. It's a very simplified
+JPF was used to test the idea and state transitioning of the model. It's a very simplified
 version of the real system that still contains all the important Copy-on-Write core concepts and the 
-assumptions that have been made in the model. This simplified version could then be automatically
-tested for soundness. It is not a proof that the model works, but it is very exhaustive in its
+assumptions that have been made in the model. This simplified version was then used to automatically
+test for unwanted quirks. It is not a proof that the model works, but it is very exhaustive in its
 testing.
@@
 \subsubsection{Entities}
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.8]{relation.pdf}}
+\begin{figure}[H] 
+    \centering{
+        \includegraphics[scale=0.8]{relation.pdf}
+    }
     \caption{High Level Entity Relationships}
@@
 
+\begin{figure}[H] 
+    \centering
+    \renewcommand{\arraystretch}{1.3}
+    \begin{tabular}{|r|l|}
+        \hline
+        Application &
+        \begin{tabular}{|r|l|}
+            \hline
+            Perius Front-end\\
+            Perius Back-end\\
+            \hline
+        \end{tabular}\\
+        Persistent Storage & MongoDB \\ \cline{2-2}
+        Filesystem & Ext4 \\ \cline{2-2}
+        OS & Arch Linux \\
+        \hline \hline
+        Hardware & \\
+        \hline
+    \end{tabular}
+    \renewcommand{\arraystretch}{1.0}
+    \caption{Software Stack}
+    \label{fig:stack}
+\end{figure}
+
+To clarify what is meant by Perius being high level software the stack in Figure~\ref{fig:stack} can
+be examined. Theoretically the Copy-on-Write features could be implemented in any of the separate
+parts of the stack to support the features that Perius has to achieve, in this work only the second
+highest layer is considered, the Perius back-end.
+
 \subsection{Related Technologies}
@@
 
-\begin{figure}[htp] 
+\begin{figure}[H] 
     \centering
@@
 
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.4]{newtree.png}}
+\begin{figure}[H] 
+    \centering{
+        \includegraphics[scale=0.4]{newtree.png}
+    }
     \caption{Cloning mechanism of Btrfs~\cite{BTRFS}}
@@
 \fix Picture of the newest front-end
-\begin{figure}[htp] \centering{
-    \includegraphics[scale=0.4]{frontend.png}}
+\begin{figure}[H] 
+    \centering{
+        \includegraphics[scale=0.4]{frontend.png}
+    }
     \caption{Front-end~\cite{BTRFS}}
@@
 The goal of this thesis was to see whether it was feasible to use copy-on-write in a high level
-application. As the implementation (see Section~\ref{sec:implementation}) made as a part of this 
-thesis project is already replacing its predecessor the simple conclusion is that it definitely is 
-possible. In the beginning of the project thoughts were on having every element being operated upon
-in a copy-on-write fashion but this was later narrowed down to only have the most important part,
-the files, as copy-on-write. This was due to that the conclusion that it did not matter if the 
-other elements were resolved in a last-write-wins manner when modified.
+application (See Figure~\ref{fig:stack}). As the implementation (see
+Section~\ref{sec:implementation}) made as a part of this thesis project is already replacing its
+predecessor the simple conclusion is that it definitely is possible. In the beginning of the project
+thoughts were on having every element being operated upon in a copy-on-write fashion but this was
+later narrowed down to only have the most important part, the files, as copy-on-write. This was due
+to that the conclusion that it did not matter if the other elements were resolved in a
+last-write-wins manner when modified.
 

### 2016-02-18 tex/paper.tex
Added conclusion

@@
 \subsection{Conclusions}
-The goal of this thesis was to see whether it was feasible to use copy-on-write in a high level
+The goal of this thesis was to see whether it was feasible to use Copy-on-Write in a high level
 application (See Figure~\ref{fig:stack}). As the implementation (see
@@
 thoughts were on having every element being operated upon in a copy-on-write fashion but this was
-later narrowed down to only have the most important part, the files, as copy-on-write. This was due
+later narrowed down to only have the most important part, the files, as Copy-on-Write. This was due
 to that the conclusion that it did not matter if the other elements were resolved in a
@@
 
-\par If this application would be distributed with several persistent storage nodes like MongoDB, the
-application would not always be in a global consistent state as there would not be any global locks.
-This could theoretically cause some inconvenience for the user but in all real world tests no 
-users have noticed it. The theoretical inconvenience for the user was a trade-off made so that the
-application could scale on the width almost endlessly, especially if the issue with building the
+\par If this application would be distributed with several persistent storage nodes like MongoDB,
+the application would not always be in a global consistent state as there would not be any global
+locks.  This could theoretically cause some inconvenience for the user but in all real world tests
+no users have noticed it. The theoretical inconvenience for the user was a trade-off made so that
+the application could scale on the width almost endlessly, especially if the issue with building the
 project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
@@
 resolution and every improvement made to it could be shared for every system that uses the same 
-type of system architecture and underlying persistent storage.
+type of system architecture and underlying persistent storage. On the other hand, by having all the
+Copy-on-Write specific code in the back-end instead of the persistent storage the choices of
+storage becomes wider as the database does not have to implement the named things. This could be
+positive when deploying Perius in environments with high requirements for the number of users
+that the system should be able to handle as the level of consistency needed could be fully tweaked
+to make a trade off for how many users the application would work for.
 

### 2016-02-19 tex/paper.tex
Added some run-time speculations

@@
 
+\subsubsection{Run-time of Operations}
+
+\par \textbf{GET single entity} \\
+When the back-end is handed a GET request it looks at what type it is and returns the document with
+the correct id from the corresponding collection. As the collections are hash indexed on id this
+operation can be made in constant time, O(1) and thus $\theta(1)$. \\
+
+\par \textbf{GET project tree} \\
+When requesting a project the full project tree with all entities except files has to be built as
+the entities not are stored as a tree in MongoDB. The tree is built layer by layer by fetching
+content with the same parent id, starting from the root of the project. Parent id's are also indexed
+by a hash index which makes it possible to get all nodes with the same parent id in constant time.
+This has to be done for every level with a different parent in the tree which makes the upper bound
+O(n) if all of the subtrees only have one node each. The tight bound can not be calculated as the
+trees balance is not determined by an algorithm but rather by a user, but from observation one 
+could argue that the trees usually have several nodes for each parent node in the tree, which then
+would give a logarithmic tight bound, $\theta(log(n))$.\\
+
+\par \textbf{POST entity} \\
+Creating a new entity in the system is a constant time operation for containers and content as they
+only have to be inserted as new documents in MongoDB, which is a O(1)
+operation~\cite{MONGOPERFOMANCE}. Even though the insertion itself has constant run-time the index 
+needs to be updated or some times even rebuilt, which can slow the time before the document is
+accessible in constant time. When inserting files they need to be chunked in to pieces and inserted
+into GridFS, this results in n/s insertions where n is the size of the file and s is the chunk size. 
+The insertion of files can then be regarded as O(n) if the size of the file is regarded, but if n is
+regarded as the number of files in the system the upper bound is constant as the insertion for files
+is made in the same way as for other documents.
+
 \subsubsection{Copy-on-Write}

### 2016-02-19 tex/paper.tex
Added more run-time explanations

@@
 O(n) if all of the subtrees only have one node each. The tight bound can not be calculated as the
-trees balance is not determined by an algorithm but rather by a user, but from observation one 
-could argue that the trees usually have several nodes for each parent node in the tree, which then
-would give a logarithmic tight bound, $\theta(log(n))$.\\
-
-\par \textbf{POST entity} \\
+trees balance is not determined by an algorithm but rather by a user, but from observation one could
+argue that the trees usually have several nodes for each parent node in the tree, which then would
+give a logarithmic tight bound, $\theta(log(n))$.\\
+
+\par \textbf{POST/Create entity} \\
 Creating a new entity in the system is a constant time operation for containers and content as they
@@
 regarded as the number of files in the system the upper bound is constant as the insertion for files
-is made in the same way as for other documents.
+is made in the same way as for other documents.\\
+
+\par \textbf{Snapshot creation} \\
+In the creation of a snapshot all meta data in the tree that the snapshot is taken of is duplicated,
+meanwhile the files remain the same, this results in n insertions where n is the number of
+containers and content entities in the subtree. A snapshot insertions runtime is therefore
+$\theta(n)$ and O(n). \\
+
+\par \textbf{Delete entity} \\
+Deleting an entity is a constant time operation when that entity does not have any children. When
+the entity has children all those children has to be deleted too which will be done in the same
+manner as the project tree is built and therefore results in the same runtime, O(n) as an upper
+bound, where n is the number of entities in the subtree.
+
+\par \textbf{PUT/Update entity} \\
+The update operation in MongoDB completely replaces a document with another in regards to their
+unique identifier. This could be seen as a deletion of the old document and an insertion of the new
+document, but with the same id, which makes it unnecessary to update the index for the unique id
+(other indices might have to be updated though). As both deletion and insertion are constant time
+operations, update is also O(1).\\
 

### 2016-02-19 tex/paper.tex
Added explaining figures for Mongo and project tree

@@
 
+\subsubsubsection{GUID}
+Global Unique Identifier - Usually in distributed environments normal incremented identifiers can
+not be used as there can be insertions on several nodes at the same time. A GUID is generated by a
+function that makes it impossible or extremely unlikely that the same identifier will be generated
+and used again.
+
 \subsubsection{Terms}
@@
 
-\subsubsection{Run-time of Operations}
+\subsubsection{Run-time Complexity of Operations}
+\begin{figure}[H] 
+    \centering{
+         \begin{tikzpicture}[nodes={draw, circle, fill=blue!20}, <-]
+             \node{R}
+             child{node{C} child{node{C}} child{node{C} child{node{M}} child{node{M}}}}
+             child{node{C}};
+         \end{tikzpicture}
+    }
+    \caption{Project Tree\\R - RootContainer, C - Container, M - Content}
+    \label{fig:project_tree}
+\end{figure}
+
+In the user interface the stored data is always represented and worked on as a tree, or a forest if
+all projects are considered. For a very small project that tree could look like the example in
+Figure~\ref{fig:project_tree}. This is however not what the actual storage of the data looks like.
+
+\begin{figure}[H] 
+    \centering
+    \renewcommand{\arraystretch}{1.3}
+    \begin{tabular}{|r|c|l|}
+        \hline
+        \textbf{Type} & \textbf{Id} & \textbf{Parent Id} \\
+        \hline
+        RootContainer & 1 & 1 \\
+        Container     & 2 & 1 \\
+        Container     & 3 & 1 \\
+        Container     & 4 & 2 \\
+        Container     & 5 & 2 \\
+        Content       & 6 & 5 \\
+        Content       & 7 & 5 \\
+        \hline
+    \end{tabular}
+    \renewcommand{\arraystretch}{1.0}
+    \caption{Simplified Representation of the Actual Storage}
+    \label{fig:storage}
+\end{figure}
+
+The actual storage of the data is represented by documents in a flat document storage, which is in
+this case MongoDB. A very simplified version of how the data is stored can be seen in
+Figure~\ref{fig:storage}. One important point that differs in the example and in the actual storage
+is that GUIDs are used for identifiers and not incremented numbers, as that would not work in
+distributed environments.
+
+As the data is not stored as a tree, but rather as separate documents loosely referring to each
+other, the run-time complexity is not obvious and that is what is discussed in this section.\\
 
@@
 manner as the project tree is built and therefore results in the same runtime, O(n) as an upper
-bound, where n is the number of entities in the subtree.
+bound, where n is the number of entities in the subtree.\\
 

### 2016-02-22 tex/paper.tex
Added abstract and clarified load test

@@
 \begin{abstract}
-    \fix Abstract is written last
-
-\keywords{}
+This work presents a technique for incorporating Copy-on-Write in the development of a high level
+system on top of an unaware persistent storage. The system in this case is a security management
+system, but the technique and conclusions drawn are general enough to be applicable to any system
+with vaguely similar consistency requirements. The scalability of the system is tested and the
+technique is evaluated and the technique is shown to be very efficient (scalable to >
+simultaneous users on a single node), but a better solution is also suggested in the
+conclusion.\\
+
+\keywords{CDN, Copy-on-Write, Eventual Consistency, MongoDB, Persistent Storage, Scalability, 
+Software Security}
 \end{abstract}
@@
 \subsubsection{Run-time Complexity of Operations}
+The runtime complexities of the operations in this section are quite intuitive, that is why they 
+are only explained, and not proven to fulfil the complexities which are claimed. 
+%To prove them, methods like the master theorem~\cite{ALGORITHMS} or \dots could be used.
+
 \begin{figure}[H] 
@@
 After adding indices, the building of the project tree was able to finish 50 times faster and the
-server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}.
+server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
+approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
+maximum number of clients that one node can handle will be >60000, without taking the web servers
+limits or content posting into consideration.
 

### 2016-02-23 tex/paper.tex
Modified abstract and fixed width

@@
 \begin{abstract}
-This work presents a technique for incorporating Copy-on-Write in the development of a high level
-system on top of an unaware persistent storage. The system in this case is a security management
-system, but the technique and conclusions drawn are general enough to be applicable to any system
-with vaguely similar consistency requirements. The scalability of the system is tested and the
-technique is evaluated and the technique is shown to be very efficient (scalable to >
-simultaneous users on a single node), but a better solution is also suggested in the
-conclusion.\\
+This work presents a technique for incorporating Copy-on-Write in a high level system on top of an
+unaware persistent storage. The system in this case is a security management system, but the
+technique and conclusions drawn are general enough to be applicable to any system with vaguely
+similar consistency requirements. The scalability of the system is tested and the technique is
+evaluated and the technique is shown to be quite efficient (scalable to >6000 simultaneous active
+users per node), but a better solution is also suggested in the conclusion.\\
 
@@
 
+\par
+As the project tree is built by querying documents parent id an optimisation that could be done was
+to create an index of the documents parent ids. This drastically improved the size of the tree that
+the server was able to handle within the time out, going from being able to handle 4000 documents in 
+one tree to at least 50000 documents, which the browser instead will have problems displaying, 
+represented in the UI, without lag.
+
+\par
 After adding indices, the building of the project tree was able to finish 50 times faster and the
@@
 \par
-As the project tree is built by querying documents parent id an optimisation that could be done was
-to create an index of the documents parent ids. This drastically improved the size of the tree that
-the server was able to handle within the time out, going from being able to handle 4000 documents in 
-one tree to at least 50000 documents, which the browser instead will have problems displaying, 
-represented in the UI, without lag.
-
+File uploads was not supported by WRK and therefore Apache Bench was used. As there is a lot of
+difference depending on the hard drive and file system when benchmarking uploads which are greater
+in size, which images usually are, no real conclusions can be drawn from this benchmark. It could
+perform about 200 POSTS/second with an image of 200KB, but it's unsure if the server could
+keep up that pace forever. If following the same example as earlier in the section with one POST per
+user every 30 seconds that would make the system able to serve 6000 active users.
+
+\newpage
 \subsection{Security of the system}
@@
 
-\subsubsection{Copy-on-Write's effect on Scalability}
+\subsection{Copy-on-Write's effect on Scalability}
 As no comparison with solutions that were not Copy-on-Write based were made, no scientific
@@
 instead.
+
+\subsection{GET/POST Estimation}
+The estimation of 30 seconds/action that was made in Section~\ref{sec:load_testing} was only based
+on observation, no logging of a system used in production (as no such system existed) could be made 
+to back up that estimation. In the future those numbers can easily be recalculated, when there are
+enough metrics from a fully deployed version of Perius.
 

### 2016-02-23 tex/paper.tex
Fixed the small paper changes

@@
 %}}}
-\title{\textbf{Designing a Security Management Layer for CDN Assets}}
+\title{\textbf{Designing a Security Management Layer with Copy-on-Write for CDN Assets}}
 \author{Lukas Klingsbo}
@@
 \section{Background}
-This section gives the reader some background of the company which the thesis was done at and
-explanation of the core concepts which are exerted.
+This section gives the reader some background of the company where the thesis was written and
+implementation was made. It also gives some brief explanation of the core concepts which are 
+exerted throughout this work.
 
@@
 not have to send or copy all of the related data that is in memory, rather they only have to send
-pointers to the data. After many Copy-on-Write's a complex tree structure can be built up, but
+pointers to the data. After many Copy-on-Writes a complex tree structure can be built up, but
 optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.
@@
 \subsubsection{Abbreviations}
-\subsubsubsection{JPF}
-Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
-licence, which made more people contribute to the project. JPF is usually used for doing model
-checking of concurrent programs to easily find for example race conditions and dead locks.
-
 \subsubsubsection{CDN}
@@
 
+\subsubsubsection{JPF}
+Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
+licence, which made more people contribute to the project. JPF is usually used for doing model
+checking of concurrent programs to easily find for example race conditions and dead locks.
+
 \subsubsection{Terms}
+\subsubsubsection{Eventual Consistency}
+The expression eventual consistency is a term used for describing how data is processed. It is often
+used in the context of database systems. Eventual consistency means that stored data is guaranteed 
+to be consistent in any one moment, but at some point the data will eventually converge to being 
+consistent~\cite{EVENTUAL}.
+
+The effect of this is that persistent storages, especially distributed databases, can
+operate faster as there is no need for locks and transactions for any operations. As no locks or
+transactions are used, operations on the same entities can be done before earlier operations have
+propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
+that needs to be handled, for example two operations that wants to change the same data at once. To
+solve this the system needs to have defined which operation that will take precedence, or how the
+concurrent operations are mixed, and this is usually called conflict or sibling resolution.
+
+\subsubsubsection{Perius}
+Perius is the name of the implementation that was made during this project. It is an anagram of
+Uprise and also a type of butterfly.
+
 \subsubsubsection{Snapshot}
@@
 
-\subsubsubsection{Eventual Consistency}
-The expression eventual consistency is a term used for describing how data is processed. It is often
-used in the context of database systems. Eventual consistency means that stored data is guaranteed 
-to be consistent in any one moment, but at some point the data will eventually converge to being 
-consistent~\cite{EVENTUAL}.
-
-The effect of this is that persistent storages, especially distributed databases, can
-operate faster as there is no need for locks and transactions for any operations. As no locks or
-transactions are used, operations on the same entities can be done before earlier operations have
-propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
-that needs to be handled, for example two operations that wants to change the same data at once. To
-solve this the system needs to have defined which operation that will take precedence and this is
-usually called conflict or sibling resolution.
-
-\subsubsubsection{Perius}
-Perius is the name of the implementation that was made during this project. It is an anagram of
-Uprise and also a type of butterfly.
-
 \newpage
 \section{Model} \label{sec:model}
-The model for this work should show how the data can not be accessed or modified by unauthorized
+The model for this work should show how the data can only be accessed or modified by authorized
 users and how the defined integrity of the data is always kept in the Perius system.
@@
 As this work only presents an informal model of how the system is designed it can not be regarded as
-a proof for the actual implementation of the system to be flawless. The model should how ever give a 
+a proof for the actual implementation of the system to be flawless. The model should however give a 
 strong idea of the soundness of the design of the system.
@@
 The Copy-on-Write system is only used for files as all of the other objects in the system is
-basically just meta data and not classed as important, from an data integrity point of view, as the 
+basically just meta data and not classed as important, from a data integrity point of view, as the 
 actual files themselves.
@@
 \subsection{Access rights}
-The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the requested entity of
-the matrix the corresponding user to that entity has full access to the corresponding object.
+The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the
+requested entity of the matrix the corresponding user to that entity has full access to the
+corresponding object.
 
@@
 
-% The last one means that no users can delete actual files
+\fix The last one means that no users can delete actual files, probably a bit confusingly written, 
+or even wrong?
 
@@
 solved by locks or transactions, but as updates are not based on each other, the trade-off for this
-inconsistency to scalability and speed is intentional. However as can also be seen in
+inconsistency to scalability and speed is intentional. However, as can also be seen in
 consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
-the set of files, even though its modification process is not the last one to write to the content.
+the set of files, even though its modification process is the last one to write to the content.
 This is necessary as a file can be referenced from several content and the removal process can not
@@
 JPF was used to test the idea and state transitioning of the model. It's a very simplified
-version of the real system that still contains all the important Copy-on-Write core concepts and the 
+version of the real system that still contains all the important Copy-on-Write concepts and the 
 assumptions that have been made in the model. This simplified version was then used to automatically
@@
 A project is what is created to contain all content and containers related to a real project. Files
-can be changed within a project and the system can contain several projects and their virtual
-content are completely disjoint.\\
+and meta-data can only be changed within a project and the system can contain several projects 
+that have their virtual content and containers completely disjoint.\\
 
@@
 randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
-transactions were used and the threads were running fully concurrently, without any sleep
+transactions were used and the threads were running fully concurrently and without any sleep
 statements.
@@
 
-JPF checked each permutation of states that the threads can end up in, the result of the run can be
-seen in Listing~\ref{JPFRESULT}.
+JPF checked each permutation of states, to a certain depth, that the threads can end up in and the 
+result of the run can be seen in Listing~\ref{JPFRESULT}.
 
@@
 \subsubsection{Garbage Collection}
-For the system to reach the decided eventual consistency~\cite{KLINGSBO}, garbage collection will be
-needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section ~\ref{sec:conseq} 
-there will be objects in the system that can not be reached and should therefore be cleaned up for 
-them not to cause negative performance effects on the system.
+For the system to reach a clean state and the decided eventual consistency~\cite{KLINGSBO}, garbage
+collection will be needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section
+~\ref{sec:conseq} there will be objects in the system that are unreferenced and can not be reached
+and should therefore be cleaned up for them not to cause negative performance effects on the system.
 
@@
 \section{Implementation} \label{sec:implementation} 
-This chapter deals with the second part of this work, namely the implementation of the Perius system
-based which is based on the model described in Section~\ref{sec:model}. As this work is not intended
-as an instructions manual, more information about how to deploy and use Perius can be found on
-\url{http://perius.se}. This chapter's focus is instead on how the decisions for the implementation
-were made, how well the system would scale and the resulting security implications.
+This chapter deals with the second part of this work, namely the implementation of the Perius
+system. The implementation is based on the model described in Section~\ref{sec:model}. As this work
+is not intended as an instructions manual, more information about how to deploy and use Perius can
+be found on \url{http://perius.se}. This chapter's focus is instead on how the decisions for the
+implementation were made, how well the system would scale and the resulting security implications.
 
@@
 content just wastes space on the CDN. The problem with this is that the current interface does not
-handle it very well as one file can not be shown or uploaded in two different places in the virtual
-filesystem.
+handle it very well as one file can not be shown or uploaded in two different places in Battle
+Binary's virtual filesystem.
 
@@
 As the current system does not offer proper security measurements, is lacking a lot of features that
-is needed and does not scale very well, a new system should be developed. This work is about
+are needed and does not scale very well, a new system should be developed. This work is about
 examining a way of implementing Copy-on-Write in a high level system like this, which should solve
@@
 React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
-the browser's, this makes it able to efficiently update dynamic web pages after a change of state
-through comparing the old virtual DOM with the resulting virtual DOM after the state change and then
-only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}. React
-can be seen as the system for handling views in front-ends implementing a MVC
+the browser's normal DOM, this makes it able to efficiently update dynamic web pages after a change
+of state through comparing the old virtual DOM with the resulting virtual DOM after the state change
+and then only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}.
+React can be seen as the system for handling views in front-ends implementing a MVC
 (Model-View-Controller) architecture.
@@
 \subsubsection{Scala} 
-Scala is a multi-paradigm programming language. It most commonly runs on the JVM and compared to
+Scala is a multi-paradigm programming language. It most commonly runs on the JVM and as apart from
 Java it supports most functional programming features at the same time as it supports object
@@
 \subsubsection{MongoDB}
-MongoDB is a document-oriented database which means that it does not have the concept of rows as
+MongoDB is a document-oriented database, which means that it does not have the concept of rows as
 normal relational databases has. Instead each entity in the database is stored as a document which
@@
 implemented, which DBMS it should use and how the estimation of long term scaling was done.
+\\
+\fix Fix this section, move things here and fill out with missing details
 
@@
 \label{sec:copy-on-write}
-To efficiently create snapshots of a system Copy-on-Write can be used to make it possible to create
-snapshots in O(1)\cite{BTRFS}, this is due to the fact that to create a snapshot in a system using
+To efficiently create snapshots of a system, Copy-on-Write can be used to make it possible to create
+snapshots in O(1)\cite{BTRFS}. This is due to the fact that to create a snapshot in a system using
 Copy-on-Write you only need to reference the current nodes in the tree and make sure that they are
@@
 implement transactions or locks a lot of different problems can occur when several clients are
-working on the same data set at the same time. Such problems could be race conditions and
+working on the same data set at the same time. Such problems could be race conditions or
 determining the happened-before relation. In this work this problem is solved by implementing
@@
 application. Instead this implementation makes a full copy of the meta-data of the tree, but still
-refers to the same binary files until they are changed, which results in the creation of a new node.
+refers to the same binary files until they are changed.
 
@@
 
-The service features a virtual file structure over the assets that has been stored, snapshots,
-security management of whole containers as well as individual files, audit and auth logging, multi
-project support and a modular design for persistent storage.
+The service features a virtual file structure over the assets and meta-data that have been stored,
+as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
+whole containers as well as individual files, audit and auth logging, multiple separate project 
+and a design which allows for an interchangeable persistent storage.
 
@@
 
-\subsubsection{Run-time Complexity of Operations}
+\subsubsection{Run-time Complexity of Operations} \label{sec:run_time}
 The runtime complexities of the operations in this section are quite intuitive, that is why they 
@@
 this case MongoDB. A very simplified version of how the data is stored can be seen in
-Figure~\ref{fig:storage}. One important point that differs in the example and in the actual storage
-is that GUIDs are used for identifiers and not incremented numbers, as that would not work in
-distributed environments.
-
-As the data is not stored as a tree, but rather as separate documents loosely referring to each
-other, the run-time complexity is not obvious and that is what is discussed in this section.\\
+Figure~\ref{fig:storage}, however in reality it not quite so table-like. One important point that 
+differs in the example and in the actual storage is that GUIDs are used for identifiers and not 
+incremented numbers, as that would not work in distributed environments.
+
+The data is not stored as a tree (see Example~\ref{fig:storage}, but rather as separate documents 
+loosely referring to each other. How the different operations perform their actions and what 
+implications these actions have on the operations run-time complexities are not obvious and that 
+is what is discussed in this section.\\
 
 \par \textbf{GET single entity} \\
-When the back-end is handed a GET request it looks at what type it is and returns the document with
-the correct id from the corresponding collection. As the collections are hash indexed on id this
-operation can be made in constant time, O(1) and thus $\theta(1)$. \\
+When the back-end is handed a GET request it looks at what type of entity that is requested and
+returns the document with the correct id from the corresponding collection. As the collections are
+hash indexed on id this operation can be made in constant time, O(1) and thus $\theta(1)$. \\
 
@@
 only have to be inserted as new documents in MongoDB, which is a O(1)
-operation~\cite{MONGOPERFOMANCE}. Even though the insertion itself has constant run-time the index
-needs to be updated or some times even rebuilt, which can slow the time before the document is
-accessible in constant time. When inserting files they need to be chunked in to pieces and inserted
+operation~\cite{MONGOPERFORMANCE}. Even though the insertion itself has constant run-time the index
+needs to be updated or sometimes even rebuilt, which can slow down the time before the document is
+accessible in constant time. When inserting files they need to be chunked into pieces and inserted
 into GridFS, this results in n/s insertions where n is the size of the file and s is the chunk size.
@@
 The update operation in MongoDB completely replaces a document with another in regards to their
-unique identifier. This could be seen as a deletion of the old document and an insertion of the new
-document, but with the same id, which makes it unnecessary to update the index for the unique id
-(other indices might have to be updated though). As both deletion and insertion are constant time
-operations, update is also O(1).\\
+unique identifier. This could be seen as a deletion of the old document and an insertion of a new
+one, but with the same id, which makes it unnecessary to update the index for the unique id. Other 
+indices might have to be updated though, if for example the parent of the document is changed. 
+As both deletion and insertion are constant time operations, update is also O(1).\\
 
@@
 is built upon MongoDB as persistent storage and not a pure tree structure, single nodes can be
-fetched in O(1) but when querying for subtrees they need to be built first, which takes O(log(n)),
-where n is the number of nodes in the subtree.
+fetched in O(1) but when querying for subtrees they need to be built first, which arguable (See
+Section~\ref{sec:run_time}) takes $\theta(log(n))$, where n is the number of nodes in the subtree.
 
@@
 core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
-scale in the future it is very easy to distribute MongoDB and if needed the system can easily be
-migrated to Reactive Mongo, which is an asynchronous and non-blocking driver for MongoDB and can
-therefore make the system scale even further~\cite{REACTIVEMONGO}.
+scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
+migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
+and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
 
@@
 according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
-that backups of the Perius state is easily done through a database backup, no separate files 
+that backups of the Perius system is easily done through a database backup, no separate files 
 needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
@@
 The disadvantage of using the GridFS approach is that when using a non-distributed database 
-it is slower to read and write to the database than reading or writing directly to the filesystem. 
+it is slower to read and write to the database than reading or writing directly to a filesystem. 
 Another disadvantage is that to access the files it is needed to go through the database layer 
-in some way, instead of accessing the filesystem.
+in some way, instead of accessing the filesystem directly.
 
@@
 
-REST was chosen as only basic CRUD operations needs to be performed and because the BSON format
-which is used in MongoDB is almost identical ~\cite{BSON} to the standardised JSON format which is
-usually used by RESTful services~\cite{JSON}.
+REST was chosen as all operations can be defined as basic CRUD operations needs to be performed and
+because the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the
+standardised JSON format which is usually used by RESTful services~\cite{JSON}.
 
@@
 create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
-RAM. The results that will be analysed are latency and throughput of the server as that will give a 
-fair judgement of how well the system can perform.
+RAM. The results that will be analysed are throughputs performed without breaking the set timeout 
+of the server, as that will give a fair judgement of how well the system can perform.
 
 \subsubsection{GET requests} \label{sec:GET_REQUESTS}
-To be able to determine what is the bottleneck of the back-end three different types of \textit{GET
+To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
 requests} are performed. The first one tries to maximise the number of requests that spray-can (HTTP
@@
 involving any payload other than status code 200 (OK). When comparing that to the request which
-involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST} one can
+involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST}) one can
 see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when
@@
 \par These are very simple requests, to evaluate how much load the application can handle in its
-current state a more complex but common request has to be analysed. The most complicated and still
-common request that system is receiving is requests of a full project tree. This request is
+current state, a more complex but common request has to be analysed. The most complicated and still
+common request that the system is receiving is requests of a full project tree. This request is
 computationally expensive as the tree is not stored directly in the database but has to be built
@@
 
-\par
-Another simple optimisation that was added was indices on the parent ids, which are heavily queried
-when building the project trees.
-
 \begin{minipage}{\linewidth-1cm}
@@
 
-\par
-As the project tree is built by querying documents parent id an optimisation that could be done was
+\par As a project tree is built by querying documents parent id an optimisation that could be done was
 to create an index of the documents parent ids. This drastically improved the size of the tree that
@@
 
-\par
-After adding indices, the building of the project tree was able to finish 50 times faster and the
-server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
+\par After adding indices, the building of the project tree was able to finish 50 times faster and
+the server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
 approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
-maximum number of clients that one node can handle will be >60000, without taking the web servers
-limits or content posting into consideration.
+maximum number of active clients that one node can handle will be >60000, without taking the web
+servers limits or content posting into consideration.
 
@@
 event that somebody, consciously or unconsciously, are leaking private CDN content from within the
-organisation using Perius the audit trail can be used to find which LDAP account that was
+organisation using Perius. The audit trail can then be used to find which LDAP account that was
 responsible for the leak.
@@
 other hand it was quite nice having the BSON documents for insertion as they were so similar 
-the accepted JSON format from the REST service. As Perius most common operations involve 
-modifying the project tree, a graph database should have been researched too.
+to the accepted JSON format from the REST service. As Perius most common operations involve 
+modifying project trees, a graph database should have been researched too.
 
@@
 predecessor the simple conclusion is that it definitely is possible. In the beginning of the project
-thoughts were on having every element being operated upon in a copy-on-write fashion but this was
+thoughts were on having every element being operated upon in a Copy-on-Write fashion but this was
 later narrowed down to only have the most important part, the files, as Copy-on-Write. This was due
@@
 the application would not always be in a global consistent state as there would not be any global
-locks.  This could theoretically cause some inconvenience for the user but in all real world tests
-no users have noticed it. The theoretical inconvenience for the user was a trade-off made so that
-the application could scale on the width almost endlessly, especially if the issue with building the
-project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
+locks.  This could theoretically cause some inconvenience for the user but in all real world
+observations no users have noticed it. The theoretical inconvenience for the user was a trade-off
+made so that the application could scale on the width almost endlessly, especially if the issue with
+building the project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
 

### 2016-02-24 tex/paper.tex
Added implementation methods

@@
 \\
-\fix Fix this section, move things here and fill out with missing details
+\par As this work was not mainly about the system architecture, but rather about evaluating the
+effects of using Copy-on-Write, the decisions taken for the different technologies used were simply
+made from what the author was comfortable with and knew would have little impact on the evaluation. 
+With this said there is no comparisons of for example different databases, protocols or API
+solutions in this paper.
+
+\par The long term scaling was done by load testing the back-end with Wrk and Apache Bench. With
+the help of these tools, the computationally expensive operations could be found and it could be
+determined what the upper bound for how many active users the system could handle on a node with
+similar hardware specifications.
+
+\par A low timeout was set to make sure that the system could handle operations faster than they
+queued up. This was to be certain that the system could sustain the number of active users during a
+longer time frame without having a queue which grows beyond a manageable size.
 

### 2016-02-25 tex/paper.tex
Added snapshot explanation

@@
 \usepackage{caption}
+\usepackage{subcaption}
 \usepackage{url}
@@
 \usepackage{alltt}
-\usepackage{subfig}
+%\usepackage{subfig}
 
@@
 made from what the author was comfortable with and knew would have little impact on the evaluation. 
-With this said there is no comparisons of for example different databases, protocols or API
+With that said there are no comparisons of for example different databases, protocols or API
 solutions in this paper.
@@
 
-\subsection{Snapshot functionality}
-\fix Structure to compare snapshot systems and conclude how Perius snapshot system was designed
-\subsubsection{Copy-on-Write}
-\label{sec:copy-on-write}
+\subsection{Copy-on-Write} \label{sec:copy-on-write}
+This section is about how Copy-on-Write was used in Perius and how its implementation of it differs
+from other systems that incorporate Copy-on-Write.
+
+\subsubsection{Snapshot Functionality}
+The Copy-on-Write concept is not only well suited for consistency requirements, it is also works 
+well for creating snapshots.
+
 To efficiently create snapshots of a system, Copy-on-Write can be used to make it possible to create
@@
 not removed, see Figure~\ref{fig:btrfs_tree}.
-
-As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does not
-implement transactions or locks a lot of different problems can occur when several clients are
-working on the same data set at the same time. Such problems could be race conditions or
-determining the happened-before relation. In this work this problem is solved by implementing
-Copy-on-Write.
-\fix Move last paragraph
 
@@
 
-\subsubsection{Comparison of Copy-on-Write system implementations}
+\subsubsection{Examples of Copy-on-Write system implementations}
 \subsubsubsection{BTRFS}
@@
 
-\fix Insert more systems here and add a comparison of the systems
-
-\subsubsection{Snapshot functionality of Perius}
-In Perius snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
+\subsubsection{Perius}
+As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does
+not implement transactions or locks a lot of different problems can occur when several clients are
+working on the same data set at the same time. Such problems could be race conditions or determining
+the happened-before relation. In this work this problem is solved by implementing Copy-on-Write on
+data with high consistency requirements.
+
+\begin{figure}[H] 
+    \centering
+    \begin{subfigure}{.3\textwidth}
+        \centering
+        \begin{tikzpicture}[nodes={draw, circle, fill=blue!20}, -]
+            \node{C1}
+                child{node{C2} 
+                    child{node{C3}} 
+                    child{node{C4} 
+                        child{node{M1} child{node{F1}}}
+                        child{node{M2} child{node{F2}}}
+                    }
+                };
+        \end{tikzpicture}
+        \caption{Original Tree}
+        \label{fig:orignal_tree}
+    \end{subfigure}\qquad
+    \begin{subfigure}{.4\textwidth}
+        \centering
+        \begin{tikzpicture}[nodes={draw, circle, fill=blue!20}, -, level 1/.style={sibling
+             distance=28mm}, level 2/.style={sibling distance=15mm}]
+             \node{C1}
+                 child{node{C2} 
+                     child{node{C3}} 
+                     child{node{C4} 
+                         child{node{M1} child{node{F1}}}
+                         child{node{M2} child{node{F2}}}
+                     }
+                 }
+                 child{node[fill=yellow!20]{C5} 
+                     child{node[fill=yellow!20]{C6}} 
+                     child{node[fill=yellow!20]{C7} 
+                         child{node[fill=yellow!20]{M3} child{node[fill=green!20]{F1}}}
+                         child{node[fill=yellow!20]{M4} child{node[fill=green!20]{F2}}}
+                     }
+                 };
+         \end{tikzpicture}
+         \caption{Tree with C2 Snapshot}
+         \label{fig:snapshot_tree}
+    \end{subfigure}
+    \caption{Snapshot Explanation\\ C - Container, M - Content, F - File}
+    \label{fig:snapshot_expl}
+\end{figure}
+
+In Perius, snapshots and clones are not taken in the fashion which Btrfs uses, which can be seen in
 Figure~\ref{fig:btrfs_tree}. As Perius does not have the tree structure pre-built and each node is
@@
 application. Instead this implementation makes a full copy of the meta-data of the tree, but still
-refers to the same binary files until they are changed.
+refers to the same binary files until the meta-data is changed to point to another file.
+
+Figure~\ref{fig:snapshot_expl} shows how a snapshot is performed on C2 and stored in C1, and how the
+resulting tree would look like. In Figure~\ref{fig:snapshot_tree} all of the yellow nodes are copies
+of meta-data that were in the left tree, for example C5 and C2 contain the same data (except for
+their newly generated IDs). The green nodes refer to files that have not been copied but are still
+referenced in the copied meta-data, this is due to the fact that a snapshot is not a write to the
+files and they are therefore not copied but remain intact.
 
@@
 \subsubsection{Copy-on-Write}
-\fix Rename section and move
-
 This implementation is far from as efficient as the other Copy-on-Write systems described in
@@
 in a fully consistent state. The loss would be that all of the database nodes would need to 
-constantly have to be fully aware of the state of the other nodes.
+constantly have to, in some sense, be fully aware of the state of the other nodes. Sharding could be
+used to minimise the communication of state between the nodes, but that results in less fault
+tolerance and possibly higher loads on some nodes than others due to that the data most likely wont
+be accessed uniformly.
 

### 2016-02-25 tex/paper.tex
Added references

@@
 
-\subsubsection{Conflict Resolution}
+\subsubsection{Conflict Resolution} \label{sec:conflict_res}
 If the persistent storage would be distributed the different database nodes would have to have rules
@@
 working on the same data set at the same time. Such problems could be race conditions or determining
-the happened-before relation. In this work this problem is solved by implementing Copy-on-Write on
-data with high consistency requirements.
+the happened-before relation~\cite{LAMPORT}. In this work the race condition problem is solved by 
+implementing Copy-on-Write on data with high consistency requirements. The happened-before relation
+is not fully solved, but suggestions of how to solve it is given in Section~\ref{sec:conflict_res}.
 
@@
 
-\fix Relate section more to comparison
-
 \subsection{Resulting system}

### 2016-02-26 tex/paper.tex
Added a bit of introduction

@@
 and then distributed over the globe when the public requests it. For secret content this can be a
-problem and an inconvenience, and that is what this thesis is about. This work examines ways of
+problem and an inconvenience, and that is what this thesis is about. This work examines a way of
 enforcing virtual access control on content and groups of content, in the form of containers and
-snapshots. A system was developed to make the underlying theory work in practice.
+snapshots. A system was developed using the Copy-on-Write principle to improve scalability and make
+it a lighter operation to take snapshots.
 
@@
 
-\fix Clarify problem definition and research questions
 \newpage
@@
 optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.
-
-\fix Polish and extend
 

### 2016-03-02 tex/paper.tex
Fixed grammar etc

@@
 \title{\textbf{Designing a Security Management Layer with Copy-on-Write for CDN Assets}}
-\author{Lukas Klingsbo}
+\author{Klingsbo, Lukas}
 
@@
 similar consistency requirements. The scalability of the system is tested and the technique is
-evaluated and the technique is shown to be quite efficient (scalable to >6000 simultaneous active
+evaluated. The technique is shown to be quite efficient (scalable to >6000 simultaneous active
 users per node), but a better solution is also suggested in the conclusion.\\
@@
 
-The research question that this report tries to answers is how and whether it is practically 
+The research question that this report tries to answer is how and whether it is practically 
 feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
-it can effect the systems scalability.
+it can effect the system's scalability.
 
@@
 This section gives the reader some background of the company where the thesis was written and
-implementation was made. It also gives some brief explanation of the core concepts which are 
+implementation was made. It also gives some brief explanations of the core concepts which are 
 exerted throughout this work.
@@
 \subsection{About Uprise}
-Uprise (formerly known as ESN) is a company based in Uppsala, Sweden. It is an EA studio focussing 
-on creating great gaming experiences, which means that they are not focussed on the actual 
-gameplay, which other EA studios like DICE are. Currently Uprise has a lot of focus on developing 
-companion apps and a new form of menu systems.
+Uprise (formerly known as ESN) is a company based in Uppsala, Sweden. It is an EA studio focusing on
+creating great gaming experiences, which means that they are not focused on the actual gameplay, in
+contrast to other EA studios like DICE. Currently Uprise has a lot of focus on developing companion
+apps and a new form of menu systems.
 
@@
 Its principle is that when processes or nodes share data in between each other, the data is not
-copied until one of the processes makes changes to it. This is an optimisation as the processes does
+copied until one of the processes makes changes to it. This is an optimisation as the processes do
 not have to send or copy all of the related data that is in memory, rather they only have to send
@@
 Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
-licence, which made more people contribute to the project. JPF is usually used for doing model
-checking of concurrent programs to easily find for example race conditions and dead locks.
+license, which made more people contribute to the project. JPF is usually used for doing model
+checking of concurrent programs to easily find for example race conditions and deadlocks.
 
@@
 The Copy-on-Write system is only used for files as all of the other objects in the system is
-basically just meta data and not classed as important, from a data integrity point of view, as the 
-actual files themselves.
+basically just metadata and not classed as important as the actual files themselves, from a data
+integrity point of view.
 
@@
         F   & $f_0\dots f_n$                & Files; files, images, videos\\ \hline
-        M   & $m_0\dots m_n$                & Content; Meta-data for files\\ \hline
+        M   & $m_0\dots m_n$                & Content; Metadata for files\\ \hline
         U   & $u_0\dots u_n$                & Users; registered users in the system\\ \hline
@@
         u \in U \text{ can read } c \in C & \Iff u \in A[u,c] \\
-        u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
-        u \in U \text{ can delete } m \in c & \Iff u \in A[u,c] \text{ and readonly } \notin c \\
+        u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and read-only } \notin c \\
+        u \in U \text{ can delete } m \in c & \Iff u \in A[u,c] \text{ and read-only } \notin c \\
         u \nexists U \text{ can delete } f \in F \\
@@
 and subsystem internal. The external sources could be another system calling the subsystem with
-faulty data or trying to make inaccurate calls to program functions, it could also be somebody
-trying to tamper with the exposed functions of the program. Threats that are internal could be a
-malicious part of the subsystem or simply an incorrect part of the subsystem, which does not behave
-according to specification.
+faulty data or trying to make inaccurate calls to program functions. Another external source could
+also be somebody trying to tamper with the exposed functions of the program. Threats that are
+internal could be a malicious part of the subsystem or simply an incorrect part of the subsystem,
+which does not behave according to its specification.
 
@@
 the API (See Section~\ref{sec:API}) and internal threats as incorrect implementation. As the server
-and its system are assumed, safe malicious subsystems are not considered.
+and its system are assumed safe, malicious subsystems are not considered.
 
@@
 \end{equation}
-If a user copies a content, the new content is not equals the old (because of its new ID), however
+If a user copies a content, the new content is not equal to the old (because of its new ID), however
 they both refer to the same file.
@@
 \end{equation}
-If a user creates a snapshot of a container the full container tree is re-created with new ids and
+If a user creates a snapshot of a container the full container tree is re-created with new IDs and
 its content still refers to the same files.
@@
 is that an entity is deleted meanwhile another is created or modified to have that entity as its
-parent and thus creating a separate tree that can not be reached from the root node of the project.
+parent, thus creating a separate tree that can not be reached from the root node of the project.
 These type of inconveniences should also be cleaned up by the garbage collector.
@@
 
-JPF was used to test the idea and state transitioning of the model. It's a very simplified
+JPF was used to test the idea and state transitioning of the model. It is a very simplified
 version of the real system that still contains all the important Copy-on-Write concepts and the 
@@
     }
-    \caption{High Level Entity Relationships}
+    \caption{High-Level Entity Relationships}
     \label{fig:relation}
@@
 \par \textbf{Content} \\
-Content is meta data about a file and is stored in a container, it is a form of virtual file.  The
+Content is metadata about a file and is stored in a container, it is a form of virtual file.  The
 content can refer to for example an image, video or binary blob.\\
@@
 A project is what is created to contain all content and containers related to a real project. Files
-and meta-data can only be changed within a project and the system can contain several projects 
+and metadata can only be changed within a project and the system can contain several projects 
 that have their virtual content and containers completely disjoint.\\
@@
 \par \textbf{Snapshot} \\
-A snapshot is a read-only container from the state which the container the was in when the snapshot
+A snapshot is a read-only container from the state which the container was in when the snapshot
 was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
@@
 \subsubsection{Execution}
-Java path finder was used to show that the model and plan of how to build the system was sound. The
+Java Path Finder was used to show that the model and plan of how to build the system was sound. The
 model was built in Java with the objective of being as reduced and simple as possible, without
@@
 Each collection in the persistent storage was emulated by using the built-in ConcurrentHashMap type.
-Each client was represented by a thread and each action taken by the client was randomised. The id
-hashes which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3 and
-each object had its own id, generated in the same fashion as the real implementation is using,
-randomly generated by the ObjectId class to minimise collisions that is.  Furthermore no locking or
-transactions were used and the threads were running fully concurrently and without any sleep
-statements.
-
-ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can't be
-iterated over concurrently.
+Each client was represented by a thread and each action taken by the client was randomised. The ID
+hash function which MongoDB is using for each entity was imported from the mongo-java-driver-2.13.3
+and each object had its own ID, generated in the same fashion as the real implementation. The ID's
+are randomly generated by the ObjectId class to minimise or completely avoid collisions.
+Furthermore no locking or transactions were used and the threads were running fully concurrently and
+without any sleep statements.
+
+ConcurrentHashMap had to be used in instead of the normal HashMap, as the normal HashMaps can not 
+be iterated over concurrently.
 
@@
 \subsubsection{Conflict Resolution} \label{sec:conflict_res}
-If the persistent storage would be distributed the different database nodes would have to have rules
-set up for conflict resolution as this model does not consider what to do if the same meta data is
+If the persistent storage would be distributed, the different database nodes would have to have rules
+set up for conflict resolution as this model does not consider what to do if the same metadata is
 modified on the nodes before they have time to synchronise. The easiest way to handle this is to
@@
 \subsubsection{The current system} \label{sec:current_system}
-Today a system called battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
+Today a system called Battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
 mostly images, to content delivery networks. The current system does not make use of the
@@
 content just wastes space on the CDN. The problem with this is that the current interface does not
-handle it very well as one file can not be shown or uploaded in two different places in Battle
-Binary's virtual filesystem.
+handle it very well as one file can not be shown or uploaded in two different places in 
+Battlebinary's virtual filesystem.
 
 \subsubsection{Problem description}
-As the current system does not offer proper security measurements, is lacking a lot of features that
-are needed and does not scale very well, a new system should be developed. This work is about
-examining a way of implementing Copy-on-Write in a high level system like this, which should solve
+The current system does not offer proper security measurements, is lacking a lot of features that
+are needed and does not scale very well, hence a new system are to be developed. This work is about
+examining a way of implementing Copy-on-Write in a high-level system like this, which should solve
 the scalability problem and make it possible to implement wanted features like snapshots, cloning
@@
 
-To clarify what is meant by Perius being high level software the stack in Figure~\ref{fig:stack} can
+To clarify what is meant by Perius being high-level software the stack in Figure~\ref{fig:stack} can
 be examined. Theoretically the Copy-on-Write features could be implemented in any of the separate
@@
 \subsubsection{Scala} 
-Scala is a multi-paradigm programming language. It most commonly runs on the JVM and as apart from
-Java it supports most functional programming features at the same time as it supports object
-oriented programming~\cite{SCALA}.
+Scala is a multi-paradigm programming language. It most commonly runs on the JVM and unlike
+Java it supports most functional programming features at the same time as it supports 
+object-oriented programming~\cite{SCALA}.
 
@@
 is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
-improve its possibility to scale, which can be a big down side to some applications containing the
+improve its possibility to scale, which can be a big downside to some applications containing the
 need for such logic.
@@
 This chapter introduces the different methods used to determine how the new system should be
-implemented, which DBMS it should use and how the estimation of long term scaling was done.
+implemented, which persistent storage it should use and how the estimation of long-term scaling 
+was done.
 \\
@@
 
-\par The long term scaling was done by load testing the back-end with Wrk and Apache Bench. With
+\par The long-term scaling was done by load testing the back-end with Wrk and Apache Bench. With
 the help of these tools, the computationally expensive operations could be found and it could be
@@
 \subsubsection{Snapshot Functionality}
-The Copy-on-Write concept is not only well suited for consistency requirements, it is also works 
+The Copy-on-Write concept is not only well-suited for consistency requirements, it is also works 
 well for creating snapshots.
@@
 \subsubsection{Full Copy}
-Full copy or deep copy, as opposed to copy-on-write, is a copy where everything is copied directly
+Full Copy or Deep Copy, as opposed to Copy-on-Write, is a copy where everything is copied directly
 and not only when an object is changed. This is easier to implement but is in most cases more
@@
 reference counter is not stored in the nodes themselves but rather in a separate data structure so
-that a nodes counter can be modified without modifying the node itself and therefore eludes the
+that a node's counter can be modified without modifying the node itself and therefore eludes the
 Copy-on-Write that would have to occur.
@@
 In the mid 80's when the development of the Mach kernel started, there was problems with that
-physically copying memory was too slow. Too minimise the copying of memory, Copy-on-Write was
+physically copying memory was too slow. To minimise the copying of memory, Copy-on-Write was
 implemented. It was implemented so that virtual copy operations could be done and so that tasks
@@
 trees would have to be merged when collisions occur, due to the non-blocking nature of the
-application. Instead this implementation makes a full copy of the meta-data of the tree, but still
-refers to the same binary files until the meta-data is changed to point to another file.
+application. Instead this implementation makes a full copy of the metadata of the tree, but still
+refers to the same binary files until the metadata is changed to point to another file.
 
@@
 resulting tree would look like. In Figure~\ref{fig:snapshot_tree} all of the yellow nodes are copies
-of meta-data that were in the left tree, for example C5 and C2 contain the same data (except for
+of metadata that were in the left tree, for example C5 and C2 contain the same data (except for
 their newly generated IDs). The green nodes refer to files that have not been copied but are still
-referenced in the copied meta-data, this is due to the fact that a snapshot is not a write to the
+referenced in the copied metadata, this is due to the fact that a snapshot is not a write to the
 files and they are therefore not copied but remain intact.
@@
 
-The service features a virtual file structure over the assets and meta-data that have been stored,
+The service features a virtual file structure over the assets and metadata that have been stored,
 as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
@@
 
-\subsubsection{Run-time Complexity of Operations} \label{sec:run_time}
+\subsubsection{Runtime Complexity of Operations} \label{sec:run_time}
 The runtime complexities of the operations in this section are quite intuitive, that is why they 
-are only explained, and not proven to fulfil the complexities which are claimed. 
+are only explained, and not proven to fulfill the complexities which are claimed. 
 %To prove them, methods like the master theorem~\cite{ALGORITHMS} or \dots could be used.
@@
         \hline
-        \textbf{Type} & \textbf{Id} & \textbf{Parent Id} \\
+        \textbf{Type} & \textbf{ID} & \textbf{Parent ID} \\
         \hline
@@
 this case MongoDB. A very simplified version of how the data is stored can be seen in
-Figure~\ref{fig:storage}, however in reality it not quite so table-like. One important point that 
+Figure~\ref{fig:storage}, however in reality it is not quite so table-like. One important point that
 differs in the example and in the actual storage is that GUIDs are used for identifiers and not
@@
 
-The data is not stored as a tree (see Example~\ref{fig:storage}, but rather as separate documents 
+The data is not stored as a tree (see Example~\ref{fig:storage}), but rather as separate documents 
 loosely referring to each other. How the different operations perform their actions and what 
@@
 When the back-end is handed a GET request it looks at what type of entity that is requested and
-returns the document with the correct id from the corresponding collection. As the collections are
-hash indexed on id this operation can be made in constant time, O(1) and thus $\theta(1)$. \\
+returns the document with the correct ID from the corresponding collection. As the collections are
+hash indexed on ID this operation can be made in constant time, O(1) and thus $\theta(1)$. \\
 
@@
 the entities not are stored as a tree in MongoDB. The tree is built layer by layer by fetching
-content with the same parent id, starting from the root of the project. Parent id's are also indexed
-by a hash index which makes it possible to get all nodes with the same parent id in constant time.
+content with the same parent ID, starting from the root of the project. Parent ID's are also indexed
+by a hash index which makes it possible to get all nodes with the same parent ID in constant time.
 This has to be done for every level with a different parent in the tree which makes the upper bound
@@
 \par \textbf{Snapshot creation} \\
-In the creation of a snapshot all meta data in the tree that the snapshot is taken of is duplicated,
+In the creation of a snapshot all metadata in the tree that the snapshot is taken of is duplicated,
 meanwhile the files remain the same, this results in n insertions where n is the number of
@@
 unique identifier. This could be seen as a deletion of the old document and an insertion of a new
-one, but with the same id, which makes it unnecessary to update the index for the unique id. Other 
+one, but with the same ID, which makes it unnecessary to update the index for the unique ID. Other 
 indices might have to be updated though, if for example the parent of the document is changed. 
@@
 computationally expensive as the tree is not stored directly in the database but has to be built
-from the id and parent id of each container and content, a simulated project tree of similar size to
+from the ID and parent ID of each container and content. A simulated project tree of similar size to
 the ones stored in the current management system reaches around 25KB in uncompressed size.
@@
 real bottleneck in the application. There are two solutions to fix this bottleneck, as the project
-tree wont be modified nearly as often as it is requested the first solution would be to cache the
+tree will not be modified nearly as often as it is requested the first solution would be to cache the
 results of the tree requests and invalidate those caches when the tree is modified. The second
@@
 tree from the database. This is left to implement if the need comes for it in the future. Even
-though this massive bottleneck is present, it wont be a problem in a small scale production
+though this massive bottleneck is present, it will not be a problem in a small scale production
 environment as the tree is fetched approximately once every 30 seconds by active users, which means
@@
 
-\par As a project tree is built by querying documents parent id an optimisation that could be done was
-to create an index of the documents parent ids. This drastically improved the size of the tree that
-the server was able to handle within the time out, going from being able to handle 4000 documents in 
+\par As a project tree is built by querying documents parent ID an optimisation that could be done was
+to create an index of the documents parent IDs. This drastically improved the size of the tree that
+the server was able to handle within the time-out, going from being able to handle 4000 documents in 
 one tree to at least 50000 documents, which the browser instead will have problems displaying, 
@@
 When requesting the project that the containers were posted to the web server timed out the request,
-as it took too long for the back-end to build the project tree. In this work, the web server wont 
+as it took too long for the back-end to build the project tree. In this work, the web server will
+not  
 be configured to accept longer data processing time, as trees as large as this for a single project
-wont be used in production.
+will not be used in production.
 
@@
 File uploads was not supported by WRK and therefore Apache Bench was used. As there is a lot of
-difference depending on the hard drive and file system when benchmarking uploads which are greater
-in size, which images usually are, no real conclusions can be drawn from this benchmark. It could
-perform about 200 POSTS/second with an image of 200KB, but it's unsure if the server could
-keep up that pace forever. If following the same example as earlier in the section with one POST per
-user every 30 seconds that would make the system able to serve 6000 active users.
+difference depending on the hard drive and filesystem when benchmarking uploads which are greater in
+size, which images usually are, no real conclusions can be drawn from this benchmark. It could
+perform about 200 POSTS/second with an image of 200KB, but it is unsure if the server could keep up
+that pace forever. If following the same example as earlier in the section with one POST per user
+every 30 seconds that would make the system able to serve 6000 active users.
 
@@
 serve content in a private manner. The first two are signed URLs and signed cookies, and the third
-is a mixture between one of the first two together with a range of IP addresses.~\cite{AWSPRIVATE}
+is a mixture between one of the first two together with a range of IP addresses~\cite{AWSPRIVATE}.
 These mechanisms are needed, as explained in Section~\ref{sec:current_system}, to ensure that assets
@@
 longer valid. In the custom policy it is possible to also specify the date when the asset should be
-made available, ranges of IP addresses (which is discussed more in a later paragraph) and inclusion
-of the base64 version of the policy in the URL~\cite{AWSSIGNED}. For custom policies it is also 
-possible to reuse the policy and have it refer to multiple assets.\\
+made available, ranges of IP addresses (which is discussed more in the later paragraph ``IP range
+restriction'') and inclusion of the base64 version of the policy in the URL~\cite{AWSSIGNED}. For
+custom policies it is also possible to reuse the policy and have it refer to multiple assets.\\
 
@@
 ~\cite{KLINGSBO} many applications that choose NoSQL databases as their persistent storage 
-actually don't need it. This is most likely true for Perius too, it would have been efficient 
+actually do not need it. This is most likely true for Perius too, it would have been efficient 
 enough generating the project trees from an indexed SQL database with foreign keys. On the 
@@
 the Copy-on-write system would have an easier time to scale on the width as no global locks or
-transactions has to be used, which with high probability would slow down the system, especially 
-when widely distributed. 
-The gain of having the locks or transactions would be that the system would always be
-in a fully consistent state. The loss would be that all of the database nodes would need to 
-constantly have to, in some sense, be fully aware of the state of the other nodes. Sharding could be
-used to minimise the communication of state between the nodes, but that results in less fault
-tolerance and possibly higher loads on some nodes than others due to that the data most likely wont
-be accessed uniformly.
+transactions has to be used, which with high probability would slow down the system, especially when
+widely distributed.
+The gain of having the locks or transactions would be that the system would
+always be in a fully consistent state. The loss would be that all of the database nodes would need
+to constantly have to, in some sense, be fully aware of the state of the other nodes.
+Sharding~\cite{SHARDING} could be used to minimise the communication of state between the nodes, but
+that results in less fault tolerance and possibly higher loads on some nodes than others due to that
+the data most likely will not be accessed uniformly.
 
@@
 \subsection{Conclusions}
-The goal of this thesis was to see whether it was feasible to use Copy-on-Write in a high level
+The goal of this thesis was to see whether it was feasible to use Copy-on-Write in a high-level
 application (See Figure~\ref{fig:stack}). As the implementation (see
@@
 made so that the application could scale on the width almost endlessly, especially if the issue with
-building the project trees is (as mentioned in Section~\ref{sec:load_testing}) solved.
+building the project trees is solved (as mentioned in Section~\ref{sec:load_testing}).
 
@@
 positive when deploying Perius in environments with high requirements for the number of users
-that the system should be able to handle as the level of consistency needed could be fully tweaked
-to make a trade off for how many users the application would work for.
+that the system should be able to handle. It would also be well suited if the level of consistency
+needs to be fully tweakable to make the trade off for how many users the application would work for
+and consistency of data.
 

### 2016-03-17 tex/paper.tex
Better Uprise description
Related related technologies to each other

@@
 creating great gaming experiences, which means that they are not focused on the actual gameplay, in
-contrast to other EA studios like DICE. Currently Uprise has a lot of focus on developing companion
-apps and a new form of menu systems.
+contrast to other EA studios like DICE. Historically Uprise has developed for example an
+application called BattleLog, which is a companion for the BattleField game series. What Uprise is
+currently working on is unfortunately classified.
 
@@
 \subsection{Interaction of Multiple Accesses} \label{sec:multiple_access}
-\fix Rename consequence?
-\\
 As a result of consequence~\ref{eq:contentupdate} and~\ref{eq:containerupdate} in
@@
 \subsection{JPF}
-\fix Not sure where to put this section
-
 JPF was used to test the idea and state transitioning of the model. It is a very simplified
@@
 \subsection{Related Technologies}
-\subsubsection{React}
+\subsubsection{React} \label{sec:react}
 React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
@@
 and then only update the browser's DOM according to the delta between the virtual DOMs~\cite{REACT}.
-React can be seen as the system for handling views in front-ends implementing a MVC
-(Model-View-Controller) architecture.
-
-\subsubsection{Reflux}
+React can be seen as the system for handling views in front-ends implementing a MVC-like
+(Model-View-Controller) architecture. In this work React is used as a part of building the
+front-end.
+
+\subsubsection{Reflux} \label{sec:reflux}
 Reflux~\cite{REFLUX} is an idea and a simple library of how to structure your application. It
 features a unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than
-for example Flux~\cite{FLUX}, when using a functional reactive programming style.
+for example Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used
+together with React (see Section~\ref{sec:react}) for this implementation to define a structure of
+the flow of data and state changes.
 
@@
 
-\subsubsection{Scala} 
+\subsubsection{Scala} \label{sec:scala}
 Scala is a multi-paradigm programming language. It most commonly runs on the JVM and unlike
 Java it supports most functional programming features at the same time as it supports 
-object-oriented programming~\cite{SCALA}.
-
-\subsubsection{REST} 
+object-oriented programming~\cite{SCALA}. Scala is used to program the back-end and the Spray
+toolkit (see Section~\ref{sec:spray}) is used with it to simplify the API.
+
+\subsubsection{REST} \label{sec:rest}
 REST stands for representational state transfer, it is an architectural idea for writing stateless
@@
 
-\subsubsection{MongoDB}
+\subsubsection{Spray} \label{sec:spray}
+Spray~\cite{CASBAH} is an open-source toolkit written in Scala and for Scala (see
+section~\ref{sec:scala}). It is used to make it easier to build REST/HTTP APIs (see
+section~\ref{sec:rest}) and those abilities is what it is used in the implementation (see
+section~\ref{sec:api}).
+
+\subsubsection{MongoDB} \label{sec:spray}
 MongoDB is a document-oriented database, which means that it does not have the concept of rows as
@@
 improve its possibility to scale, which can be a big downside to some applications containing the
-need for such logic.
+need for such logic. MongoDB is used as a persistent storage for the back-end to store all the
+projects and nodes in the tree structures and also the files uploaded to the system. Scala (see
+Section~\ref{sec:scala}) uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB.
 
@@
 
-\subsubsection{API}
-\label{sec:API}
+\subsubsection{API} \label{sec:API}
 In this work a RESTful API was implemented and used for back-end $\Iff$ front-end

### 2016-03-17 tex/paper.tex
Added response data to load testing section

@@
 
-REST was chosen as all operations can be defined as basic CRUD operations needs to be performed and
-because the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the
-standardised JSON format which is usually used by RESTful services~\cite{JSON}.
+REST (see Section~\ref{sec:rest} was chosen as all operations can be defined as basic CRUD
+operations needs to be performed and because the BSON format which is used in MongoDB is almost
+identical ~\cite{BSON} to the standardised JSON format which is usually used by RESTful
+services~\cite{JSON}.
 
@@
 Transfer/sec:     17.25MB
+
+Response data:
+200 OK
 \end{lstlisting}
@@
 Transfer/sec:      1.96MB
+
+Response data:
+[{
+  "id": "565cdae1898b798008105fd47",
+  "name": Test Project,
+  "readOnly": false,
+  "lastModified": 1456320024643
+}]
 \end{lstlisting}
@@
 Transfer/sec:     22.44MB
+
+Response data:
+Static test data for WRK not involving MongoDB 
+or filesystem access. 
+Lorem ipsum dolor sit amet.
 \end{lstlisting}
@@
 \begin{lstlisting}[label=TREEREQUEST,caption=Result of project tree requests]
-wrk -t10 -c10 -d10s http://perius:8000/projects/56a5f...
-
-Running 10s test @ http://perius:8000/projects/56a5f...
+wrk -t10 -c10 -d10s http://perius:8000/projects/56...
+
+Running 10s test @ http://perius:8000/projects/56...
   10 threads and 10 connections
@@
 Transfer/sec:      0.96MB
+
+Response data:
+{
+  "parent": "",
+  "name": "Test Project",
+  "readOnly": false,
+  "lastModified": 1456320024643,
+  "id": "56cdae1898b798008105fd47",
+  "containers": [{
+    "parent": "56cdae1898b798008105fd47",
+    "name": "Upload stuff",
+    "readOnly": false,
+    "lastModified": 0,
+    "id": "56cdae2298b798008105fd48",
+    "containers": [],
+    "content": [{
+      "fileId": "56cdae2d98b798008105fd49",
+      "parent": "56cdae2298b798008105fd48",
+      "size": 35245,
+      "readOnly": false,
+      "lastModified": 1456320045336,
+      "tags": [],
+      "cdnLinks": [],
+      "id": "56cdae2d98b798008105fd4b",
+      "filename": "newtree.png",
+      "dimensions": [1088, 369],
+      "originalId": "56cdae2d98b798008105fd4b"
+    }, ...]
+  }],
+  "content": [{
+    ...
+  }]
+}
 \end{lstlisting}
@@
 Transfer/sec:    174.17MB
+
+Response data:
+Same as in Listing 5
 \end{lstlisting}
@@
 Transfer/sec:      0.97MB
+
+Response data:
+Same as in Listing 5
 \end{lstlisting}
@@
 Transfer/sec:      1.23MB
+
+Response data:
+{
+  "id": "56cdae2298b798008105fd48",
+  "parent": "56cdae1898b798008105fd47",
+  "name": "Upload",
+  "readOnly": false,
+  "lastModified": 1456320045336,
+}
 \end{lstlisting}

### 2016-03-21 tex/paper.tex
Added invariant

@@
 what its creator intended ~\cite{BIBA}.
+\\
+The invariant for this model is the file. Which means that it shouldn't be possible to
+change a file though out the cycle of operation without creating a new file object with a new
+identifier. This is to ensure that files can not lose their integrity by simultaneous
+operations from different nodes or get mixed up when referred to from a content object.
 
 \subsection{Initial Assumptions}
-To create an integrity model, some initial assumptions have to be made about what the correct behaviour
-of the system is, which the model then can be shown to follow. In this work unintentional behaviour
-as the result of data modification is the main concern, which could be used for sabotage or simply be the
-effect unintentional unfortunate race conditions etc.
+To create an integrity model, some initial assumptions have to be made about what the correct
+behaviour of the system is, which the model then can be shown to follow. In this work unintentional
+behaviour as the result of data modification is the main concern, which could be used for sabotage
+or simply be the effect unintentional unfortunate race conditions etc.
 
@@
 request is simply the same amount of static payload as the second request but without involving any
-connections to MongoDB.
+connections to MongoDB. Most responses are JSON structures, except for the ones in
+Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}.\\
+
 \par
-The typical requests with Wrk in this work will look similar to this:
+The typical requests performed with Wrk in this work will look similar to this:
 \begin{lstlisting}[frame=single]

### 2016-03-21 tex/paper.tex
Rewrote eventual consistency

@@
 \begin{abstract}
-This work presents a technique for incorporating Copy-on-Write in a high level system on top of an
-unaware persistent storage. The system in this case is a security management system, but the
-technique and conclusions drawn are general enough to be applicable to any system with vaguely
-similar consistency requirements. The scalability of the system is tested and the technique is
-evaluated. The technique is shown to be quite efficient (scalable to >6000 simultaneous active
-users per node), but a better solution is also suggested in the conclusion.\\
+This work presents a technique for incorporating Copy-on-Write on top of an unaware persistent
+storage. The system in this case is a security management system, but the technique and conclusions
+drawn are general enough to be applicable to any system with vaguely similar consistency
+requirements. The scalability of the storage system is tested and the technique is evaluated. The
+technique is shown to be quite efficient (scalable to >6.000 simultaneous active users per node),
+but a better solution is also suggested in the conclusion.\\
 
@@
 \subsubsubsection{GUID}
-Global Unique Identifier - Usually in distributed environments normal incremented identifiers can
+Global Unique Identifier - In distributed environments normal incremented identifiers can
 not be used as there can be insertions on several nodes at the same time. A GUID is generated by a
@@
 \subsubsubsection{JPF}
-Java Path Finder - It was developed by NASA and in 2005 they released it under an open source
+Java Path Finder was developed by NASA and in 2005 they released it under an open source
 license, which made more people contribute to the project. JPF is usually used for doing model
@@
 \subsubsubsection{Eventual Consistency}
-The expression eventual consistency is a term used for describing how data is processed. It is often
-used in the context of database systems. Eventual consistency means that stored data is guaranteed 
-to be consistent in any one moment, but at some point the data will eventually converge to being 
-consistent~\cite{EVENTUAL}.
-
-The effect of this is that persistent storages, especially distributed databases, can
-operate faster as there is no need for locks and transactions for any operations. As no locks or
-transactions are used, operations on the same entities can be done before earlier operations have
-propagated to all nodes. This can mean that the system ends up in a state where there is a conflict
-that needs to be handled, for example two operations that wants to change the same data at once. To
-solve this the system needs to have defined which operation that will take precedence, or how the
-concurrent operations are mixed, and this is usually called conflict or sibling resolution.
+Eventual consistency is an informal guarantee that defines how the consistency of stored data can
+behave. In a system defined as using eventual consistency, access to stored data is not guaranteed
+to return the result of the latest modification. At some point, if modifications to the stored data
+stop, access to a data item will return the resulting data of the last modification of that
+item.~\cite{EVENTUAL}. It is often used in the context of database systems.
+
+The effect of the lack of a strong consistency requirement is that persistent storages, especially
+distributed databases, can operate faster since there is no need for locks or transactions.  This
+results in that operations on the same items can be made before earlier operations have propagated
+to all the distributed nodes.
+
+A system using eventual consistency can end up in a state where there is a conflict that needs to be
+handled. If for example two concurrent operations performed on two different servers changes the
+same data item, a conflict of how the resulting item should look like will occur when synchronising
+these servers. To solve the conflict the system needs to have a defined set of rules to be able to
+determine which operation that will take precedence, or how the concurrent operations are mixed, and
+this is usually called conflict or sibling resolution.
 
@@
 \\
-The invariant for this model is the file. Which means that it shouldn't be possible to
+The invariant for this model is the file. Which means that it should not be possible to
 change a file though out the cycle of operation without creating a new file object with a new

### 2016-03-21 tex/paper.tex
Fixed yellow comments

@@
 \begin{abstract}
-This work presents a technique for incorporating Copy-on-Write on top of an unaware persistent
+This work presents a technique for implementing Copy-on-Write on top of an unaware persistent
 storage. The system in this case is a security management system, but the technique and conclusions
@@
 \subsubsection{Abbreviations}
-\subsubsubsection{CDN}
+\begin{description}
+
+\item[CDN] \hfill \\
 Content Distribution/Delivery Network - Replicates content to several servers, usually spread out
@@
 
-\subsubsubsection{GUID}
+\item[GUID] \hfill \\
 Global Unique Identifier - In distributed environments normal incremented identifiers can
@@
 
-\subsubsubsection{JPF}
+\item[JPF] \hfill \\
 Java Path Finder was developed by NASA and in 2005 they released it under an open source
@@
 
+\end{description}
+
 \subsubsection{Terms}
-\subsubsubsection{Eventual Consistency}
+\begin{description}
+\item[Eventual Consistency] \hfill \\
 Eventual consistency is an informal guarantee that defines how the consistency of stored data can
@@
 these servers. To solve the conflict the system needs to have a defined set of rules to be able to
-determine which operation that will take precedence, or how the concurrent operations are mixed, and
-this is usually called conflict or sibling resolution.
-
-\subsubsubsection{Perius}
+determine which operation that will take precedence, or how the concurrent operations are mixed,
+which is usually called conflict or sibling resolution. %% Clock sync and check time as rule
+
+\item[Perius] \hfill \\
 Perius is the name of the implementation that was made during this project. It is an anagram of
@@
 
-\subsubsubsection{Snapshot}
+\item[Snapshot] \hfill \\
 A snapshot is a way to record the full state of a system at a specific time. The term comes from
@@
 can be used as backups meanwhile snapshots are not very effective means of backup in the case of
-data corruption. It is not effective against data corruption as snapshots usually still refer to 
-unchanged data that is still a part of the system~\cite{SNAPSHOT}.
-
+data corruption. Snapshots are not effective against data corruption as snapshots usually still
+refer to unchanged data that is still a part of the system~\cite{SNAPSHOT}.
+
+\end{description}
 \newpage
 \section{Model} \label{sec:model}
-The model for this work should show how the data can only be accessed or modified by authorized
+The model for this work shows how the data can only be accessed or modified by authorized
 users and how the defined integrity of the data is always kept in the Perius system.
@@
 
-\par \textbf{Content} \\
+\hfill \par \textbf{Content} \\
 Content is metadata about a file and is stored in a container, it is a form of virtual file.  The

### 2016-03-21 tex/paper.tex
Added branching section

@@
 
+\subsection{Branching}
+Branching was also implemented in Perius, but to not make the report too wide it was not included.
+The branching worked by the user selecting a container that the branch should be based on and a
+container in which the branch should be places. The user could then choose which containers and
+content that it wanted to update with the origin source of and which objects that should be deep
+copied to the new branch.
+
 \subsection{GET/POST Estimation}

### 2016-03-22 tex/paper.tex
Added tech stack figure
Changed to description for related tech

@@
 
-\subsection{Related Technologies}
-\subsubsection{React} \label{sec:react}
-React is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
+\subsection{Architecture and Technologies used}
+
+\begin{figure}[H] 
+    \centering{
+        \includegraphics[scale=0.8]{techstack.pdf}
+    }
+    \caption{Technology Stack}
+    \label{fig:tech_stack}
+\end{figure}
+
+\begin{description}
+\item[React]% \hfill \\ % \label{sec:react}
+is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
 the browser's normal DOM, this makes it able to efficiently update dynamic web pages after a change
@@
 
-\subsubsection{Reflux} \label{sec:reflux}
-Reflux~\cite{REFLUX} is an idea and a simple library of how to structure your application. It
-features a unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than
-for example Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used
-together with React (see Section~\ref{sec:react}) for this implementation to define a structure of
-the flow of data and state changes.
+% \hfill \\ % \label{sec:reflux}
+\item[Reflux]~\cite{REFLUX} 
+is an idea and a simple library of how to structure your application. It features a unidirectional
+dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example
+Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used together with
+React for this implementation to define a structure of the flow of data and state changes.
 
@@
 
-\subsubsection{Scala} \label{sec:scala}
-Scala is a multi-paradigm programming language. It most commonly runs on the JVM and unlike
-Java it supports most functional programming features at the same time as it supports 
-object-oriented programming~\cite{SCALA}. Scala is used to program the back-end and the Spray
-toolkit (see Section~\ref{sec:spray}) is used with it to simplify the API.
-
-\subsubsection{REST} \label{sec:rest}
-REST stands for representational state transfer, it is an architectural idea for writing stateless
+\item[Scala]% \hfill \\ %\label{sec:scala}
+is a multi-paradigm programming language. It most commonly runs on the JVM and unlike Java it
+supports most functional programming features at the same time as it supports object-oriented
+programming~\cite{SCALA}. Scala is used to program the back-end and the Spray toolkit is used with
+it to simplify the API.
+
+\item[REST]% \hfill \\ %\label{sec:rest}
+stands for representational state transfer, it is an architectural idea for writing stateless
 services. These services usually use URIs to identify specific resources and HTTP to modify or query
@@
 
-\subsubsection{Spray} \label{sec:spray}
-Spray~\cite{CASBAH} is an open-source toolkit written in Scala and for Scala (see
-section~\ref{sec:scala}). It is used to make it easier to build REST/HTTP APIs (see
-section~\ref{sec:rest}) and those abilities is what it is used in the implementation (see
-section~\ref{sec:api}).
-
-\subsubsection{MongoDB} \label{sec:spray}
-MongoDB is a document-oriented database, which means that it does not have the concept of rows as
+% \hfill \\ %\label{sec:spray}
+\item[Spray]~\cite{CASBAH} % Spray in casbah book
+is an open-source toolkit written in Scala and for Scala. It is used to make it easier to build
+REST/HTTP APIs and those abilities is what it is used in the implementation.
+
+\item[MongoDB]% \hfill \\ %\label{sec:spray}
+is a document-oriented database, which means that it does not have the concept of rows as
 normal relational databases has. Instead each entity in the database is stored as a document which
@@
 need for such logic. MongoDB is used as a persistent storage for the back-end to store all the
-projects and nodes in the tree structures and also the files uploaded to the system. Scala (see
-Section~\ref{sec:scala}) uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB.
-
+projects and nodes in the tree structures and also the files uploaded to the system. Scala uses a
+layer called Casbah~\cite{CASBAH} to interface with MongoDB.
+
+\end{description}
 \subsection{Methods for determining\\implementation details}

### 2016-03-22 tex/paper.tex
Added some tech stack description

@@
 
-\subsection{Architecture and Technologies used}
+\subsection{Architecture and Technologies used} \label{sec:tech}
+This section introduces the different technologies used in the implementation. The section also
+contains motivations and explanations why certain technologies were chosen and briefly outlines the
+software architecture.
 
@@
 \end{figure}
+Figure~\ref{fig:tech_stack} shows how the different technologies used are related. In the Figure, 
+the technologies that are indented are built on top of its parent. The arrows that can be seen is
+the flow of data after a request from the front-end has been made. The arrows are unidirectional,
+but the response follows the same path back to the requester.
 
 \begin{description}
-\item[React]% \hfill \\ % \label{sec:react}
+
+\item[Javascript]
+\fix Add description
+
+\item[SuperAgent]
+\fix Add citation to SuperAgent
+is a small HTTP request library which can be used from client-side Javascript or as a node.js
+module. It is used to form the requests to the REST API from the front-end, which can be seen in
+Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
+
+\item[React]
 is a JavaScript library for building user interfaces. React uses both its own virtual DOM and
@@
 
+In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
+communication, which can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
+
+REST was chosen as all operations that needs to be performed can be defined as basic CRUD operations
+and because the BSON format which is used in MongoDB is almost identical ~\cite{BSON} to the
+standardised JSON format which is usually used by RESTful services~\cite{JSON}.
+
+\hfill \\
+\textbf{REST Endpoints} \hfill \\
+For the front-end to communicate with the back-end, a RESTful service is implemented.
+The following endpoints were configured:
+
+\begin{itemize}
+  \item projects
+      \subitem GET - list all projects
+      \subitem POST - create new project
+  \item projects/\{id\}
+      \subitem GET - get specific project
+      \subitem PUT - update existing project
+      \subitem DELETE - delete existing project
+  \item projects/\{id\}/content
+      \subitem POST - create new content in a specific project 
+  \item projects/\{id\}/content/\{id\}
+      \subitem GET - get specific content in a specific project
+      \subitem PUT - update existing content in a specific project
+      \subitem DELETE - delete existing content in a specific project
+
+  \item projects/\{id\}/snapshots
+      \subitem POST - create new snapshot in a specific project
+
+  \item projects/\{id\}/containers
+      \subitem POST - create new container in a specific project 
+  \item projects/\{id\}/containers/\{id\}
+      \subitem GET - get specific container in a specific project
+      \subitem PUT - update existing container
+      \subitem DELETE - delete existing container
+\end{itemize}
+
+As can be seen several expected endpoints are missing, this is intentional as the operations missing
+can be performed in a more efficient way. Such endpoint is for example \textit{GET 
+projects/\{id\}/containers} as all containers exist in \textit{GET projects/\{id\}} and the interface
+should present a file structure where both content and containers are shown.
+
 % \hfill \\ %\label{sec:spray}
@@
 need for such logic. MongoDB is used as a persistent storage for the back-end to store all the
-projects and nodes in the tree structures and also the files uploaded to the system. Scala uses a
-layer called Casbah~\cite{CASBAH} to interface with MongoDB.
+projects and nodes in the tree structures and also the files uploaded to the system. The Scala logic
+code uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB, which can be seen in
+Figure~\ref{fig:tech_stack} (Arrow 4 and 5).
+
+\par MongoDB was chosen as the persistent storage because of its quick lookups and because of its
+internal storage format called BSON, which is very similar to JSON which the API is using. As the
+formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
+core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
+scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
+migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
+and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
+
+\par All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
+according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
+that backups of the Perius system is easily done through a database backup, no separate files 
+needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
+ranges of a file, although that advantage is not needed in the Perius implementation. For
+scalability this could be used to retrieve different parts of a file from different servers, 
+normal load balancing would probably work better in a system like Perius where no extremely large 
+files are expected to be stored.
+
+\par The disadvantage of using the GridFS approach is that when using a non-distributed database 
+it is slower to read and write to the database than reading or writing directly to a filesystem. 
+Another disadvantage is that to access the files it is needed to go through the database layer 
+in some way, instead of accessing the filesystem directly.
 
@@
 \subsubsection{Perius}
-As the persistent storage, used in this implementation (Section~\ref{sec:persistent_storage}), does
+As the persistent storage, used in this implementation (Section~\ref{sec:tech}), does
 not implement transactions or locks a lot of different problems can occur when several clients are
@@
 \subsubsection{Persistent storage} \label{sec:persistent_storage}
-\subsubsubsection{MongoDB}
-MongoDB was chosen as the persistent storage because of its quick lookups and because of its
-internal storage format called BSON, which is very similar to JSON which the API is using. As the
-formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
-core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
-scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
-migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
-and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
-
-All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
-according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
-that backups of the Perius system is easily done through a database backup, no separate files 
-needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
-ranges of a file, although that advantage is not needed in the Perius implementation. For
-scalability this could be used to retrieve different parts of a file from different servers, 
-normal load balancing would probably work better in a system like Perius where no extremely large 
-files are expected to be stored.
-
-The disadvantage of using the GridFS approach is that when using a non-distributed database 
-it is slower to read and write to the database than reading or writing directly to a filesystem. 
-Another disadvantage is that to access the files it is needed to go through the database layer 
-in some way, instead of accessing the filesystem directly.
-
-\subsubsection{API} \label{sec:API}
-In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
-communication.
-
-REST (see Section~\ref{sec:rest} was chosen as all operations can be defined as basic CRUD
-operations needs to be performed and because the BSON format which is used in MongoDB is almost
-identical ~\cite{BSON} to the standardised JSON format which is usually used by RESTful
-services~\cite{JSON}.
-
-\subsubsubsection{REST Endpoints}
-For the front-end to communicate with the back-end, a RESTful service is implemented.
-The following endpoints were configured:
-
-\begin{itemize}
-  \item projects
-      \subitem GET - list all projects
-      \subitem POST - create new project
-  \item projects/\{id\}
-      \subitem GET - get specific project
-      \subitem PUT - update existing project
-      \subitem DELETE - delete existing project
-  \item projects/\{id\}/content
-      \subitem POST - create new content in a specific project 
-  \item projects/\{id\}/content/\{id\}
-      \subitem GET - get specific content in a specific project
-      \subitem PUT - update existing content in a specific project
-      \subitem DELETE - delete existing content in a specific project
-
-  \item projects/\{id\}/snapshots
-      \subitem POST - create new snapshot in a specific project
-
-  \item projects/\{id\}/containers
-      \subitem POST - create new container in a specific project 
-  \item projects/\{id\}/containers/\{id\}
-      \subitem GET - get specific container in a specific project
-      \subitem PUT - update existing container
-      \subitem DELETE - delete existing container
-\end{itemize}
-
-As can be seen several expected endpoints are missing, this is intentional as the operations missing
-can be performed in a more efficient way. Such endpoint is for example \textit{GET 
-projects/\{id\}/containers} as all containers exist in \textit{GET projects/\{id\}} and the interface
-should present a file structure where both content and containers are shown.
-
 \newpage

### 2016-03-23 tex/paper.tex
Added javascript description

@@
 \item[Javascript]
-\fix Add description
+is a programming language mainly focussed on running code in web browsers~\cite{JAVASCRIPT}. In this
+implementation the ECMAScript 6 (ES6) language specification is used, which adds a lot of syntactic
+sugar compared to what usually is called normal JavaScript. An example of such syntactic sugar is
+lambda expressions. React and Reflux are both built with JavaScript in mind and is used on top of
+JavaScript is this implementation, as can be seen in Figure~\ref{fig:tech_stack}.
 
 \item[SuperAgent]
-\fix Add citation to SuperAgent
 is a small HTTP request library which can be used from client-side Javascript or as a node.js
-module. It is used to form the requests to the REST API from the front-end, which can be seen in
-Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
+module~\cite{SUPERAGENT}. It is used to form the requests to the REST API from the front-end, which
+can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
 
@@
 
-% \hfill \\ % \label{sec:reflux}
-\item[Reflux]~\cite{REFLUX} 
-is an idea and a simple library of how to structure your application. It features a unidirectional
-dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example
+\item[Reflux] 
+is an idea and a simple library of how to structure your application~\cite{REFLUX}. It features a
+unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example
 Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used together with
@@
 
-\item[Scala]% \hfill \\ %\label{sec:scala}
+\item[Scala]
 is a multi-paradigm programming language. It most commonly runs on the JVM and unlike Java it
@@
 
-\item[REST]% \hfill \\ %\label{sec:rest}
+\item[REST]
 stands for representational state transfer, it is an architectural idea for writing stateless

### 2016-03-23 tex/paper.tex
Updated the introduction

@@
 %}}}
-\title{\textbf{Designing a Security Management Layer with Copy-on-Write for CDN Assets}}
+\title{\textbf{Perius: A Security Management Layer with Copy-on-Write for CDN Assets}}
 \author{Klingsbo, Lukas}
@@
 Network to be able to scale to a larger user base. The commercial Content Distribution Networks are
-usually fairly easy to use, the content that is to be used in a project is usually simply uploaded
-and then distributed over the globe when the public requests it. For secret content this can be a
-problem and an inconvenience, and that is what this thesis is about. This work examines a way of
-enforcing virtual access control on content and groups of content, in the form of containers and
-snapshots. A system was developed using the Copy-on-Write principle to improve scalability and make
-it a lighter operation to take snapshots.
+usually fairly easy to use. The content that is to be used in a project is uploaded and then
+distributed over the globe when a client is trying to access the content. For secret content this
+can be a problem and an inconvenience, and that is what this thesis is about.
+
+When looking at the game industry, secret content could be anything from a picture related to a yet
+unreleased game to an internal video only intended for employees. Some secret content needs to stay
+secret and some content needs to get its security restraints changed during time. For example when a
+game is in a closed alpha stage, the content should only be accessible by the people working on it.
+But when that game later moves into an open beta phase (open for public), the content needs to be
+publicly accessible.
+
+As large projects most often are divided into different dividable parts, a concept called containers
+is introduced in this work. Containers can contain more containers and content and when
+modifying the security settings for a container it is applied recursively to all its content and
+sub-containers.
+
+This work examines a way of enforcing virtual access control on content and groups of content, in
+the form of containers and snapshots to be able to handle the secrecy requirements easily. A system
+was developed using the Copy-on-Write principle to improve scalability and make it a lighter
+operation to take snapshots of parts of the system.
 
@@
 \subsection{Load Testing} \label{sec:load_testing}
-Wrk was used to load test the back-end. It is a multi-threaded benchmarking tool for HTTP which can
-create large loads. The testing was done locally on a virtual server having 6 CPU cores and 48GB of
-RAM. The results that will be analysed are throughputs performed without breaking the set timeout 
-of the server, as that will give a fair judgement of how well the system can perform.
+Wrk and Apache Bench was used to load test the back-end. Wrk is a multi-threaded benchmarking tool
+for HTTP which can create large loads and Apache Bench has similar capabilities, but is better
+suited for file uploads than Wrk. The testing was done locally on a virtual server having 6 CPU
+cores and 48GB of RAM. The results that will be analysed are the throughput of each request
+performed without breaking the set timeout of the server, as that will give a fair judgement of how
+well the system can perform.
 
@@
 To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
-requests} are performed. The first one tries to maximise the number of requests that spray-can (HTTP
-server) can handle with the given specifications and therefore the server just returns 200 (OK). The
-second request requests the simplest API call which involves getting all currently stored projects
-from the MongoDB instance. As this involves the server answering with some payload, the third
-request is simply the same amount of static payload as the second request but without involving any
-connections to MongoDB. Most responses are JSON structures, except for the ones in
-Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}.\\
-
-\par
+requests} are performed. Most responses are JSON structures, except for the ones in
+Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}, which can be seen in the Response data part
+of the listings.\\
+
 The typical requests performed with Wrk in this work will look similar to this:
@@
 \end{minipage}
+
+\par
+The first request (Listing~\ref{OKREQUEST}) tries to maximise the number of requests that spray-can
+(HTTP server) can handle with the given specifications and therefore the server just returns 200
+(OK).
 
@@
 \end{minipage}
-
+\end{custommargins}
+
+\par
+The second request (Listing~\ref{DBREQUEST}) performs the simplest API call which involves getting
+all currently stored projects from the MongoDB instance. This request does not yield any extra
+computation in the back-end, apart from converting the MongoDB format to the API's JSON result
+format.
+
+\newpage
+\begin{custommargins}{0cm}{-2cm}
 \begin{minipage}{\linewidth-1cm}
@@
 or filesystem access. 
-Lorem ipsum dolor sit amet.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 \end{lstlisting}
@@
 
-Listing~\ref{OKREQUEST} shows that the HTTP server can answer about 100K requests/s when not
-involving any payload other than status code 200 (OK). When comparing that to the request which
-involved the server responding with a small (100 Byte) payload (Listing~\ref{STATICREQUEST}) one can
-see that it is about 10\% slower to send some more data, \textasciitilde90K requests/s. However when
-comparing that to the requests that needed database access it is obvious that the HTTP server can
-handle all the load that it needs to, as Listing~\ref{DBREQUEST} shows that to fetch all documents
-in a collection (in this case only one) is dramatically slower, the throughput was only
-\textasciitilde7000 requests/s.
-
-\par These are very simple requests, to evaluate how much load the application can handle in its
-current state, a more complex but common request has to be analysed. The most complicated and still
-common request that the system is receiving is requests of a full project tree. This request is
-computationally expensive as the tree is not stored directly in the database but has to be built
-from the ID and parent ID of each container and content. A simulated project tree of similar size to
-the ones stored in the current management system reaches around 25KB in uncompressed size.
+\par
+As the second request (Listing~\ref{DBREQUEST}) involves the server answering with some payload that
+is generated from the persistent storage, the third request (Listing~\ref{STATICREQUEST}) is simply
+the same amount of static payload as the second request but without involving any connections to
+MongoDB. This was done to be able to estimate what impact the round trip time to MongoDB and the
+conversions to the response format had on the throughput.
 
@@
 real bottleneck in the application. There are two solutions to fix this bottleneck, as the project
-tree will not be modified nearly as often as it is requested the first solution would be to cache the
-results of the tree requests and invalidate those caches when the tree is modified. The second
+tree will not be modified nearly as often as it is requested, the first solution would be to cache
+the results of the tree requests and invalidate those caches when the tree is modified. The second
 solution would be to make a more computationally feasible way of building and fetching the whole
@@
 content that the link referred to would be open to anyone.
+
+\subsubsection{Load Testing}
+Section~\ref{sec:load_testing}, specifically Listing~\ref{OKREQUEST}, shows that the HTTP server can
+answer about 100K requests/s when not involving any payload other than status code 200 (OK). When
+comparing that to the request which involved the server responding with a small (100 Byte) payload
+(Listing~\ref{STATICREQUEST}) one can see that it is about 10\% slower to send some more data,
+\textasciitilde90K requests/s. However when comparing that to the requests that needed database
+access it is obvious that the HTTP server can handle all the load that it needs to, as
+Listing~\ref{DBREQUEST} shows that to fetch all documents in a collection (in this case only one) is
+dramatically slower, the throughput was only \textasciitilde7000 requests/s.
+
+\par These are very simple requests, to evaluate how much load the application can handle in its
+current state, a more complex but common request has to be analysed. The most complicated and still
+common request that the system is receiving is requests of a full project tree. This request is
+computationally expensive as the tree is not stored directly in the database but has to be built
+from the ID and parent ID of each container and content. A simulated project tree of similar size to
+the ones stored in the current management system reaches around 25KB in uncompressed size.
+
+\par From the load testing a bottleneck of the application was found, namely building the project
+trees. Three solutions for this was presented and two were tested. The first was to cache the
+results of computationally heavy requests. The second was to simply make the operations more
+computationally efficient, which was not tried. The third presented solution was to add indices to
+keys that were often queried in MongoDB. The third solution yielded the best results and made Perius
+being able to serve 2100 requests/s, which involved building the project tree. A combination of
+solution one and three could possibly make Perius able to serve even more requests per second.
 

### 2016-03-23 tex/paper.tex
Reorganised findings section

@@
 As large projects most often are divided into different dividable parts, a concept called containers
-is introduced in this work. Containers can contain more containers and content and when
+is introduced in this work. Containers can contain more containers and content, and when
 modifying the security settings for a container it is applied recursively to all its content and
@@
 \subsection{Findings}
-\subsubsection{Scalability}
-When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
-the application has no limits of how much load and users it can handle as the hardware and nodes 
-can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
-implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
-support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
-is offered by ReactiveMongo is not needed as the load will not reach the peak, as discussed in 
-Section~\ref{sec:GET_REQUESTS}, for what Casbah can handle on a single server.
+This section presents the relevant findings derived from the implementation and testing of it.
 
@@
 content that the link referred to would be open to anyone.
+
+\subsubsection{Scalability}
+When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
+the application has no limits of how much load and users it can handle as the hardware and nodes 
+can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
+implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
+support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
+is offered by ReactiveMongo is not needed as the load will not reach the peak, as discussed in 
+Section~\ref{sec:GET_REQUESTS}, for what Casbah can handle on a single server.
 

### 2016-03-29 tex/paper.tex
Added subsection introductions

@@
 \subsection{Prior Work}
+The \textit{Prior Work} section goes through the related work that has already been made within the
+area and which this thesis is based on.
+
 \subsubsection{Copy-on-Write}
@@
 \subsection{Elements of the Model}
+In this section we show the elements of the model, identify the different sets and explain how they
+should be interpreted. This is done in the same way as in \textit{Secure Computer Systems: Mathematical
+Foundations}~\cite{BLP1} by Bell \& LaPadula.
+
 \begin{center}
@@
 \subsection{Findings}
-\fix Add more findings
+This section summarises the findings from the work on the model.
+
 \subsubsection{Garbage Collection}
@@
 
+\fix Add more findings
 \newpage 
@@
 
-\subsection{Background}
+\subsection{Background} \label{sec:background}
+Section~\ref{sec:background} explains the problem at hand, in relation to the system currently 
+used.
+
 \subsubsection{The current system} \label{sec:current_system}
@@
 \subsection{Resulting system}
+The \textit{Resulting System} section explains what the implementation looks like and how its
+different operations work, in the terms of efficiency.
+
 \subsubsection{Perius}
@@
 
-\subsubsection{Persistent storage} \label{sec:persistent_storage}
+%\subsubsection{Persistent storage} \label{sec:persistent_storage}
 \newpage
@@
 \subsection{Security of the system}
-\subsubsection{Authorization}
-The authorization of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
+This section declares for the software security features that were used in Perius, in the terms of
+how the system and its assets are protected internally and externally.
+
+\subsubsection{Authentication}
+The authentication of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
 is mainly focussed at being deployed in internal networks which usually has an LDAP service enabled. 
@@
 \subsection{Findings}
-This section presents the relevant findings derived from the implementation and testing of it.
+This section presents the relevant findings derived from the implementation of Perius and the 
+testing of it.
 
@@
 \section{Discussion}
+This chapter contains informal discussions about topics that were not covered by the rest of the
+report but that are still significant enough to mention.
+
 \subsection{Persistent Storage}
@@
 \section{Summary}
+This section covers the conclusions that could be drawn from this work and how the implementation 
+could, and will be, extended in the future.
+
 \subsection{Conclusions}

### 2016-03-29 tex/paper.tex
Added Findings and Prior work

@@
 %}}}
-\title{\textbf{Perius: A Security Management Layer with Copy-on-Write for CDN Assets}}
+\title{\textbf{Providing Access Control and Copy-on-Write for Content Distribution Network Assets}}
 \author{Klingsbo, Lukas}
@@
 optimisations can be done to simplify that structure~\cite{COPYONWRITE2}.
+
+\subsubsection{Snapshots}
+Several filesystems~\cite{BTRFS}\cite{ZFS} use snapshot techniques to make it possible for a user
+or system to roll back the filesystems state, to a state that the filesystem had at the time of
+snapshot creation. This can be useful for example when having a stable system and then doing a
+system update, if the system does not respond well to the update the filesystem can be rolled
+back to its previous stable state, provided that a snapshot was created before the update procedure.
 
@@
 
-\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label=JPFRESULT,caption=Results of JPF run]%,float,floatplacement=H]
 elapsed time:       14:26:53
@@
 providers offer (see Section~\ref{sec:private_content}) to make sure that the content only is
-accessible from where it should be. In Battlebinary the content was secured by adding a part of the
-hash of the file (see Section~\ref{sec:current_system}) to the filename that it was uploaded with, 
-which resulted in a URL on the CDN that was not guessable but if the link somehow leaked the 
-content that the link referred to would be open to anyone.
-
+accessible from where it should be. In Perius it is possible to have three different modes for a CDN
+asset; public to all, only accessible from certain IP ranges or only accessible by clients using a
+signed URL or cookie. In Battlebinary the content was secured by adding a part of the hash of the
+file (see Section~\ref{sec:current_system}) to the filename that it was uploaded with, which
+resulted in a URL on the CDN that was not guessable but if the link somehow leaked the content that
+the link referred to would be open to anyone.
+
+\newgeometry{top=4cm}
 \subsubsection{Scalability}
+Perius uses MongoDB as its persistent storage and all computationally expensive operations involve
+doing calls to it. Scala currently has two popular drivers for interaction with MongoDB, namely
+Casbah and ReactiveMongo.
+
 When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking, 
@@
 \par These are very simple requests, to evaluate how much load the application can handle in its
-current state, a more complex but common request has to be analysed. The most complicated and still
+current state, a more complex but common request was analysed. The most complicated and still
 common request that the system is receiving is requests of a full project tree. This request is
@@
 
+\restoregeometry
 \newpage

### 2016-04-01 tex/paper.tex
Playing around with the model

@@
 \usepackage{geometry}
+\usepackage{extarrows}
+\usepackage{mathpartir}
 
@@
 
+This work has taken a lot inspiration from the way BTRFS~\cite{BTRFS} handles snapshots, which is
+further explained in Section~\ref{sec:btrfs}.
+
 \newpage
@@
         u \in U \text{ can delete } m \in c & \Iff u \in A[u,c] \text{ and read-only } \notin c \\
-        u \nexists U \text{ can delete } f \in F \\
-        \forall c \in C, \quad \exists u \in U & \mid a \in A[u,c] 
+        \forall c \in C, \nexists u \in U & \text{ can delete } f \in c
     \end{split}
@@
         m' & = u, read(m) \land \\
-        x  & = f \in m' \land \\
-        f' & = u, change(x) \land \\
+        f  & = f \in m' \land \\
+        f' & = u, change(f) \land \\
         u, & write(f') \land \\
@@
 \end{equation}
+\begin{equation}
+    \begin{split}
+        (F, f \in c) \xrightarrow{update(f', c)} (\{f, f'\} \subset F, f' \in c) 
+    \end{split}
+\end{equation}
+\inferrule
+{f \in c \xrightarrow{update(f', c)} f' \in c }
+{c \cap \{f\} \\ c \cup \{f'\} \\ F \cup \{f, f'\}}
+
+\vspace{2em}
+
+\inferrule
+{f \in c \\ update(f', c) \\ f' \in c}
+{F \cup \{f, f'\}}
+
+\vspace{2em}
+
+\inferrule
+{f \in F \\ [fileId = f.id, \dots] \in C \\ f \neq f'} 
+{F \cup \{f\} \xrightarrow{write(fileId, f'.id)} F \cup \{f, f'\}}
+
+\vspace{2em}
 If a user wants to update a file in a content, the file is copied and the original is intact.
@@
 
-\fix Add more findings
+\subsubsection{JPF Results}
+Although the JPF model checking results did not return any errors or showed any possible misbehaviours
+of the model, it should not be considered proven. The results however strongly indicate that there are no
+major flaws in it. But as the model will be more complicated once fully implemented as an application
+there might be flaws in the implementation details.
+
 \newpage 
@@
 \subsubsection{Examples of Copy-on-Write system implementations}
-\subsubsubsection{BTRFS}
+\subsubsubsection{BTRFS} \label{sec:btrfs}
 Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do

### 2016-04-06 tex/paper.tex
Updated the equations

@@
 \section{Model} \label{sec:model}
-The model for this work shows how the data can only be accessed or modified by authorized
-users and how the defined integrity of the data is always kept in the Perius system.
+This section describes how different operations can be expected to effect the data in the system.
+The model for this work also shows how the data can only be accessed or modified by
+authorized users and how the defined integrity of the data is always kept in the Perius system.
 
@@
 
-In this section the consequences of different operations are stated in an informal mathematical
-manner. These rules are what should be expected to be followed by the implementation in
+In this section the consequences, side effects and direct effects, of different operations are
+stated in an informal mathematical manner, similar to how inference rules are composed. These rules
+are what should be expected to be followed by the implementation in
 Section~\ref{sec:implementation}.
@@
         u, copy(o)            & User u copies object o and the copy gets a new ID\\ \hline
-        u, change(o)          & User u locally changes object o \\ \hline
-        u, modify(o', o)      & User u globally modifies object o based on object o'\\ \hline
+        %u, change(o)          & User u locally changes object o \\ \hline
+        u, update(o, o')      & User u globally modifies object o based on object o'\\ \hline
+        %u, modify(o', o)      & User u globally modifies object o based on object o'\\ \hline
         u, delete(o)          & User u deletes object o\\ \hline
@@
 
+To describe a content object \textit{m} containing fileId \textit{f}, the following syntax is used:
+$[fileId = f.id, \dots] \in M$, the same syntax is also applied to different types of objects
+throughout the explanation of the model.
+
+%\begin{equation} \label{eq:fileupdate}
+%    \begin{split}
+%        m' & = u, read(m) \land \\
+%        f  & = f \in m' \land \\
+%        f' & = u, change(f) \land \\
+%        u, & write(f') \land \\
+%        u, & modify(f', m) \Implies f \in F \land f' \in F
+%    \end{split}
+%\end{equation}
+%\begin{equation}
+%    \begin{split}
+%        (F, f \in c) \xrightarrow{update(f', c)} (\{f, f'\} \subset F, f' \in c) 
+%    \end{split}
+%\end{equation}
+
+\vspace{2em}
+\textbf{File Update}
 \begin{equation} \label{eq:fileupdate}
+    \inferrule
+    {f \in F \\ [fileId = f.id, parent = c \dots] \in M \\ A[u,c] \in A} 
+    {F \cup \{f\} \xrightarrow{u, update(fileId, f'.id)} F \cup \{f, f'\}}
+\end{equation}
+\vspace{1em}
+
+If a user \textit{u} with access rights to a parent container \textit{c} of content \textit{m},
+updates a file reference \textit{f} in content \textit{m}, the original file stays intact.
+
+%\begin{equation} \label{eq:contentupdate}
+%    \begin{split}
+%        m' & = u, read(m) \land \\
+%        m'' & = u, change(m') \land \\
+%        m''' & = u, modify(m'', m) \Implies \\
+%        m''' & \in M \land m \not \in M
+%    \end{split}
+%\end{equation}
+
+\vspace{2em}
+\textbf{Content Update}
+\begin{equation} \label{eq:contentupdate}
+    \inferrule
+    {m \in M \\ m \in c \\ A[u,c] \in A}
+    {M \cup \{m\} \xrightarrow{u, update(m, m')} M \cup \{m'\}}
+\end{equation}
+\vspace{1em}
+
+If a user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
+updates that content, the content is directly changed, without any Copy-on-Write.
+
+%\begin{equation} \label{eq:containerupdate}
+%    \begin{split}
+%        c' & = u, read(c) \land \\
+%        c'' & = u, change(c') \land \\
+%        c''' & = u, modify(c'', c) \Implies \\
+%        c''' & \in C \land c \not \in C
+%    \end{split}
+%\end{equation}
+\vspace{2em}
+\textbf{Container Update}
+\begin{equation} \label{eq:containerupdate}
+    \inferrule
+    {c \in C \\ A[u,c] \in A}
+    {C \cup \{c\} \xrightarrow{u, update(c, c')} C \cup \{m'\}}
+\end{equation}
+\vspace{1em}
+
+If a user \textit{u} with access rights to a container \textit{c} updates that container, the
+container is directly changed, without any Copy-on-Write.
+
+%\begin{equation} \label{eq:copy}
+%    \begin{split}
+%        & f' = f \in m \\
+%        & m' = u, copy(m) \Implies \\
+%        & m' \neq m \land \\
+%        & f' \in m' \land f' \in m
+%    \end{split}
+%\end{equation}
+\vspace{2em}
+\textbf{Copy}
+\begin{equation} \label{eq:copy}
+    \inferrule
+    {m \in M \\ c \in C \\ m \in c \\ A[u,c] \in A}
+    {M \cup \{m\} \xrightarrow{u, copy(m')} M \cup \{m, m'\} \land m.id \neq m'}
+\end{equation}
+\vspace{1em}
+
+If a user \textit{u} with access rights to the parent (source and destination) container \textit{c}
+of content \textit{m} copies that content, the new content \textit{m'} is not equal to the old
+(because of its new ID), however they both contain the same metadata and refer to the same file.
+
+\vspace{2em}
+\textbf{Delete}
+\begin{equation} \label{eq:delete}
+    \inferrule
+    {m \in M \\ f \in F \\ f \in m \\ c \in C \\ m \in c \\ A[u,c] \in A}
+    {M \cup \{m\} \land F \cup \{f\} \xrightarrow{u, delete(m)} M \cup \{\} \land F \cup \{f\}}
+\end{equation}
+\vspace{1em}
+
+If a user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
+deletes that content, the file \textit{f} referred to in the content is not deleted.
+
+\vspace{2em}
+\textbf{Snapshot}
+\begin{equation} \label{eq:snapshot}
     \begin{split}
-        m' & = u, read(m) \land \\
-        f  & = f \in m' \land \\
-        f' & = u, change(f) \land \\
-        u, & write(f') \land \\
-        u, & modify(f', m) \Implies f \in F \land f' \in F
+        [C', M'] = u, snapshot(tree) \Implies \\
+        \forall c \in tree, u, copy(c) \in C \land \\
+        \forall m \in tree, u, copy(m) \in M \land \\
+        (\forall f \in tree \Implies f \in F)
     \end{split}
 \end{equation}
-\begin{equation}
-    \begin{split}
-        (F, f \in c) \xrightarrow{update(f', c)} (\{f, f'\} \subset F, f' \in c) 
-    \end{split}
+
+To explain the non-trivial snapshot operation it is first defined in \ref{eq:snapshot}. The snapshot
+operation is performed upon a tree of objects and results in two new sets (C' and M'), one with
+containers and one with content, the referred files remain the same as before the snapshot.
+
+\vspace{2em}
+\begin{equation} \label{eq:snapshot2}
+    \inferrule
+    {c \in C \\ A[u,c] \in A}
+    {M \land C \xrightarrow{[C', M'] = u, snapshot(c)} M \cup M' \land C \cup C'}
 \end{equation}
-\inferrule
-{f \in c \xrightarrow{update(f', c)} f' \in c }
-{c \cap \{f\} \\ c \cup \{f'\} \\ F \cup \{f, f'\}}
-
-\vspace{2em}
-
-\inferrule
-{f \in c \\ update(f', c) \\ f' \in c}
-{F \cup \{f, f'\}}
-
-\vspace{2em}
-
-\inferrule
-{f \in F \\ [fileId = f.id, \dots] \in C \\ f \neq f'} 
-{F \cup \{f\} \xrightarrow{write(fileId, f'.id)} F \cup \{f, f'\}}
-
-\vspace{2em}
-If a user wants to update a file in a content, the file is copied and the original is intact.
-
-\begin{equation} \label{eq:contentupdate}
-    \begin{split}
-        m' & = u, read(m) \land \\
-        m'' & = u, change(m') \land \\
-        m''' & = u, modify(m'', m) \Implies \\
-        m''' & \in M \land m \not \in M
-    \end{split}
-\end{equation}
-If a user reads content and then writes to it, the content is directly changed.
-
-\begin{equation} \label{eq:containerupdate}
-    \begin{split}
-        c' & = u, read(c) \land \\
-        c'' & = u, change(c') \land \\
-        c''' & = u, modify(c'', c) \Implies \\
-        c''' & \in C \land c \not \in C
-    \end{split}
-\end{equation}
-If a user reads a container and then writes to it, the container is directly changed.
-
-\begin{equation} \label{eq:copy}
-    \begin{split}
-        & f' = f \in m \\
-        & m' = u, copy(m) \Implies \\
-        & m' \neq m \land \\
-        & f' \in m' \land f' \in m
-    \end{split}
-\end{equation}
-If a user copies a content, the new content is not equal to the old (because of its new ID), however
-they both refer to the same file.
-
-\begin{equation} \label{eq:snapshot}
-    \begin{split}
-        s = u, snapshot(c') \Implies \\
-        \forall c \in c', u, copy(c) \in s \land \\
-        \forall m \in c', u, copy(m) \in s \land \\
-        (\forall f \in c' \Implies f \in s)
-    \end{split}
-\end{equation}
-If a user creates a snapshot of a container the full container tree is re-created with new IDs and
-its content still refers to the same files.
-
-\begin{equation} \label{eq:delete}
-    \begin{split}
-        f' = f \in (u, read(c)),  \\
-        u, delete(c) \Implies \\
-        c \not \in C \land f' \in F
-    \end{split}
-\end{equation}
-If a user deletes content, the file referred to in the content is not deleted.
+If a user creates a snapshot of a container and the full container tree is re-created with new IDs,
+but its content still refers to the same files.
+
+%\begin{equation} \label{eq:delete}
+%    \begin{split}
+%        f' = f \in (u, read(c)),  \\
+%        u, delete(c) \Implies \\
+%        c \not \in C \land f' \in F
+%    \end{split}
+%\end{equation}
 
@@
 \hfill \\
-\textbf{REST Endpoints} \hfill \\
+\textbf{REST Endpoints} \label{sec:API} \hfill \\
 For the front-end to communicate with the back-end, a RESTful service is implemented.

### 2016-04-06 tex/paper.tex
Rewrote access rules

@@
 \usepackage{mathpartir}
+\usepackage{centernot}
 
@@
 \subsection{Access rights}
-The \textit{a} notation is meant to symbolise an access token, and if \textit{a} exists in the
-requested entity of the matrix the corresponding user to that entity has full access to the
-corresponding object.
+The \textit{A[u,c]} notation is symbolising an access token, and if \textit{A[u,c]} exists in
+the access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and
+its children.
 
@@
     \begin{split}
-        u \in U \text{ can read } c \in C & \Iff u \in A[u,c] \\
-        u \in U \text{ can write } c \in C & \Iff u \in A[u,c] \text{ and read-only } \notin c \\
-        u \in U \text{ can delete } m \in c & \Iff u \in A[u,c] \text{ and read-only } \notin c \\
-        \forall c \in C, \nexists u \in U & \text{ can delete } f \in c
+        A[u,c] \in A \Rightarrow & u \text{ can read } c \\
+        A[u,c] \in A \land \text{ read-only } \notin c \Rightarrow & u \text{ can write to } c \\
+        A[u,c] \in A \land \text{ read-only } \notin c \Rightarrow & u \text{ can delete } c \\
+        A[u,c] \in A \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
     \end{split}
 \end{equation}
-
-\fix The last one means that no users can delete actual files, probably a bit confusingly written, 
-or even wrong?
 

### 2016-04-06 tex/paper.tex
Added description of model elements

@@
 \subsection{Approach}
-As this work only presents an informal model of how the system is designed it can not be regarded as
+As this work only presents an informal model of how the system is designed, it can not be regarded as
 a proof for the actual implementation of the system to be flawless. The model should however give a 
@@
 
-The Copy-on-Write system is only used for files as all of the other objects in the system is
+The Copy-on-Write principle is only used for files as all of the other objects in the system is
 basically just metadata and not classed as important as the actual files themselves, from a data
@@
 
+\begin{description}
+\item[Container] \hfill \\
+    A container is what would normally be called a folder in popular speech and in the terms of
+    filesystems. A container can contain content objects and more containers, this recursive
+    definition results in containers representing a virtual tree structure.
+
+\item[File] \hfill \\
+    A file is an actual file, in the same sense as in a file system. A file could be for example a
+    video or image.
+
+\item[Content] \hfill \\
+    A content object contains a pointer to a file and metadata about that file. The pointer to the
+    file within the object can change to a different file, but the content object still keeps the
+    same metadata as before.
+
+\item[User] \hfill \\
+    A user is a registered user in the system.
+
+\item[Access Matrix] \hfill \\
+    The access matrix contains what containers that all the users in the system have access to. If
+    A[u,c] exists in the access matrix, user \textit{u} has access to container \textit{c}.
+
+\end{description}
+
 \subsection{Access rights}
+Access is only handled on containers and if a user has access to a container it also has access to
+all of that containers children.
 The \textit{A[u,c]} notation is symbolising an access token, and if \textit{A[u,c]} exists in
@@
 throughout the explanation of the model.
+
+If the user is irrelevant for the rule described, it is left out from the action, for example u,
+read(o) = read(o).
 
@@
     {m \in M \\ c \in C \\ m \in c \\ A[u,c] \in A}
-    {M \cup \{m\} \xrightarrow{u, copy(m')} M \cup \{m, m'\} \land m.id \neq m'}
+    {M \cup \{m\} \xrightarrow{u, copy(m')} M \cup \{m, m'\} \land m.id \neq m'.id}
 \end{equation}
@@
 \vspace{2em}
-\textbf{Delete}
+\textbf{Content Deletion}
 \begin{equation} \label{eq:delete}
@@
     \begin{split}
-        [C', M'] = u, snapshot(tree) \Implies \\
-        \forall c \in tree, u, copy(c) \in C \land \\
-        \forall m \in tree, u, copy(m) \in M \land \\
-        (\forall f \in tree \Implies f \in F)
+        [C', M'] := snapshot(tree) & \Implies \\
+        \forall c \in tree, copy(c) \in & C' \\
+        \forall m \in tree, copy(m) \in & M' \\
+        \forall f \in tree, f \in & F
     \end{split}
@@
     {c \in C \\ A[u,c] \in A}
-    {M \land C \xrightarrow{[C', M'] = u, snapshot(c)} M \cup M' \land C \cup C'}
+    {M, C \xrightarrow{[C', M'] := u, snapshot(c)} M \cup M',  C \cup C'}
 \end{equation}

### 2016-04-06 tex/paper.tex
Added explanations and changed to description latex type

@@
 
-\subsection{Elements of the Model}
+\subsection{Sets and Elements of the Model}\label{sec:elements}
 In this section we show the elements of the model, identify the different sets and explain how they
@@
 
-\hfill \par \textbf{Content} \\
-Content is metadata about a file and is stored in a container, it is a form of virtual file.  The
-content can refer to for example an image, video or binary blob.\\
-
-\par \textbf{Project} \\
-A project is what is created to contain all content and containers related to a real project. Files
+All the entities in Figure~\ref{fig:relation} is described in Section~\ref{sec:elements}, except for
+projects, which is explained below. \\
+\vspace{0.5em}
+
+A \textbf{Project} is what is created to contain all content and containers related to a real project. Files
 and metadata can only be changed within a project and the system can contain several projects 
-that have their virtual content and containers completely disjoint.\\
-
-\par \textbf{Container} \\
-A container is a virtual folder within a project which can contain content and other containers.\\
-
-\par \textbf{Snapshot} \\
-A snapshot is a read-only container from the state which the container was in when the snapshot
-was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
-Snapshots are by default stored as siblings to the container which they were made from, but they can
-be contained by any container.\\
-
-\par \textbf{File} \\
-A file refers to an actual physical file. Files are stored in the database to make backup,
-deployment and migration easier.\\
-
-\par \textbf{User} \\
-A user is the structure that handles people who have been granted access to the system. Access to
-the system is handled by a separate service, like LDAP.\\
+that have their own virtual content and containers completely disjoint from the other projects.\\
 
@@
 
+\subsubsection{Entities of the Implementation}
+This section explains the different types of entities, and their interaction, that are handled by
+Perius. They are explained similarly to how they are explained for the model in
+Section~\ref{sec:elements}, but with more implementation specific details.
+
+\begin{description}
+\item[Content] \hfill \\
+Content is metadata about a file and is stored in a container, it is a form of virtual file.  The
+content can refer to for example an image, video or binary blob.\\
+
+\item[Project] \hfill \\
+A project is what is created to contain all content and containers related to a real project. Files
+and metadata can only be changed within a project and the system can contain several projects 
+that have their virtual content and containers completely disjoint. Each project is handled as a
+separate database in MongoDB.\\
+
+\item[Container] \hfill \\
+A container is a virtual folder within a project which can contain content and other containers.\\
+
+\item[Snapshot] \hfill \\
+A snapshot is a read-only container from the state which the container was in when the snapshot
+was created.  A snapshot can not be updated and can only be deleted from the root of the snapshot.
+Snapshots are by default stored as siblings to the container which they were made from, but they can
+be contained by any container.\\
+
+\item[File] \hfill \\
+A file refers to an actual physical file. Files are stored in the database to make backup,
+deployment and migration easier.\\
+
+\item[User] \hfill \\
+A user is the structure that handles people who have been granted access to the system. Access to
+the system is handled by a separate service, like LDAP.\\
+
+\end{description}
 \subsubsection{Runtime Complexity of Operations} \label{sec:run_time}
 The runtime complexities of the operations in this section are quite intuitive, that is why they 
-are only explained, and not proven to fulfill the complexities which are claimed. 
+are only explained, and not proven to fulfil the complexities which are claimed. 
 %To prove them, methods like the master theorem~\cite{ALGORITHMS} or \dots could be used.

### 2016-04-07 tex/paper.tex
Added sections to the presentation

@@
 The model for this work also shows how the data can only be accessed or modified by
-authorized users and how the defined integrity of the data is always kept in the Perius system.
+authorized users and how the defined integrity of the data is always kept.
 

### 2016-04-11 tex/paper.tex
Changed access rights definition

@@
         U   & $u_0\dots u_n$                & Users; registered users in the system\\ \hline
-        A   & $A[u_0,c_0]\dots A[u_n, c_n]$ & Access matrix; describes what containers users have access to\\ \hline
+        A   & $A[u_0,c_0]\dots A[u_n, c_n]$ & Access matrix; access rights \\ \hline
     \end{tabular}
@@
     The access matrix contains what containers that all the users in the system have access to. If
-    A[u,c] exists in the access matrix, user \textit{u} has access to container \textit{c}.
+    A[u,c] is true in the access matrix, user \textit{u} has access to container \textit{c}.
 
@@
 all of that containers children.
-The \textit{A[u,c]} notation is symbolising an access token, and if \textit{A[u,c]} exists in
-the access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and
-its children.
+The \textit{A[u,c]} notation is symbolising an access token. If \textit{A[u,c]} is true in the 
+access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and its
+children. If it is false or if the entry does not exist in the access matrix, the user \textit{u}
+does not have any access rights for \textit{c}.
 
@@
     \begin{split}
-        A[u,c] \in A \Rightarrow & u \text{ can read } c \\
-        A[u,c] \in A \land \text{ read-only } \notin c \Rightarrow & u \text{ can write to } c \\
-        A[u,c] \in A \land \text{ read-only } \notin c \Rightarrow & u \text{ can delete } c \\
-        A[u,c] \in A \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
+        A[u,c] \Rightarrow & u \text{ can read } c \\
+        A[u,c] \land \text{ read-only } \notin c \Rightarrow & u \text{ can write to } c \\
+        A[u,c] \land \text{ read-only } \notin c \Rightarrow & u \text{ can delete } c \\
+        A[u,c] \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
     \end{split}
@@
     \inferrule
-    {f \in F \\ [fileId = f.id, parent = c \dots] \in M \\ A[u,c] \in A} 
+    {f \in F \\ [fileId = f.id, parent = c \dots] \in M \\ A[u,c]} 
     {F \cup \{f\} \xrightarrow{u, update(fileId, f'.id)} F \cup \{f, f'\}}
@@
     \inferrule
-    {m \in M \\ m \in c \\ A[u,c] \in A}
+    {m \in M \\ m \in c \\ A[u,c]}
     {M \cup \{m\} \xrightarrow{u, update(m, m')} M \cup \{m'\}}
@@
     \inferrule
-    {c \in C \\ A[u,c] \in A}
+    {c \in C \\ A[u,c]}
     {C \cup \{c\} \xrightarrow{u, update(c, c')} C \cup \{m'\}}
@@
     \inferrule
-    {m \in M \\ c \in C \\ m \in c \\ A[u,c] \in A}
+    {m \in M \\ c \in C \\ m \in c \\ A[u,c]}
     {M \cup \{m\} \xrightarrow{u, copy(m')} M \cup \{m, m'\} \land m.id \neq m'.id}
@@
     \inferrule
-    {m \in M \\ f \in F \\ f \in m \\ c \in C \\ m \in c \\ A[u,c] \in A}
+    {m \in M \\ f \in F \\ f \in m \\ c \in C \\ m \in c \\ A[u,c]}
     {M \cup \{m\} \land F \cup \{f\} \xrightarrow{u, delete(m)} M \cup \{\} \land F \cup \{f\}}
@@
     \inferrule
-    {c \in C \\ A[u,c] \in A}
+    {c \in C \\ A[u,c]}
     {M, C \xrightarrow{[C', M'] := u, snapshot(c)} M \cup M',  C \cup C'}

### 2016-04-12 tex/paper.tex
Removed unnecessary f \in F
Changed equations according to Johannes spec

@@
         U   & $u_0\dots u_n$                & Users; registered users in the system\\ \hline
-        A   & $A[u_0,c_0]\dots A[u_n, c_n]$ & Access matrix; access rights \\ \hline
+        A   & $A[u_0,c_0]\dots A[u_n, c_n]$ & Boolean Access matrix; access rights \\ \hline
+    \end{tabular}
+\end{center}
+
+\begin{center}
+    \begin{tabular}{ | l | l | l | p{5cm} |}
+        \hline
+        \textbf{Set} & \textbf{Relation} & \textbf{Type} \\ \hline
+        \textbf{C} & $\textbf{C} \supseteq C \ni c$ & $\textbf{C}=P_{fin}(P_{fin}(C \cup M))$\\ \hline
+        \textbf{M} & $\textbf{M} \supseteq M \ni m$ & $\textbf{M}=P_{fin}(M)$\\ \hline
+        \textbf{F} & $\textbf{F} \supseteq F \ni f$ & $\textbf{F}=P_{fin}(F)$\\ \hline
+    \end{tabular}
+\end{center}
+
+The bold notation (\textbf{C}, \textbf{M}, \textbf{F}) refers to the sets containing all possible
+members of that type.
+
+\begin{center}
+    \begin{tabular}{ | l | l | p{5cm} |}
+        \hline
+        \textbf{Element} & \textbf{Type} \\ \hline
+        c & $C \cup M$ \\ \hline
+        m & $\{id: GUID, file: f\}$ \\ \hline
+        f & File \\ \hline
     \end{tabular}
@@
         A[u,c] \Rightarrow & u \text{ can read } c \\
-        A[u,c] \land \text{ read-only } \notin c \Rightarrow & u \text{ can write to } c \\
-        A[u,c] \land \text{ read-only } \notin c \Rightarrow & u \text{ can delete } c \\
+        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can write to } c \\
+        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can delete } c \\
         A[u,c] \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
@@
         \textbf{Action} & \textbf{Semantics} \\ \hline
-        u, read(o)            & User u reads object o\\ \hline
-        u, write(o)           & User u writes object o\\ \hline
-        u, copy(o)            & User u copies object o and the copy gets a new ID\\ \hline
-        %u, change(o)          & User u locally changes object o \\ \hline
-        u, update(o, o')      & User u globally modifies object o based on object o'\\ \hline
-        %u, modify(o', o)      & User u globally modifies object o based on object o'\\ \hline
-        u, delete(o)          & User u deletes object o\\ \hline
-        u, snapshot(o)        & User u takes a snapshot of object o \\ \hline
+        u, read(m, f)    & User \textit{u} reads \textit{f} from \textit{m}\\ \hline
+        u, create(m, c)  & User \textit{u} creates \textit{m} in \textit{c}\\ \hline
+        u, update(m, f)  & User \textit{u} updates \textit{f} with \textit{m}\\ \hline
+        u, delete(o)     & User \textit{u} deletes \textit{o}\\ \hline
+        u, snapshot(o)   & User \textit{u} takes a snapshot of \textit{o} \\ \hline
     \end{tabular}
@@
 
-%\begin{equation} \label{eq:fileupdate}
-%    \begin{split}
-%        m' & = u, read(m) \land \\
-%        f  & = f \in m' \land \\
-%        f' & = u, change(f) \land \\
-%        u, & write(f') \land \\
-%        u, & modify(f', m) \Implies f \in F \land f' \in F
-%    \end{split}
-%\end{equation}
-%\begin{equation}
-%    \begin{split}
-%        (F, f \in c) \xrightarrow{update(f', c)} (\{f, f'\} \subset F, f' \in c) 
-%    \end{split}
-%\end{equation}
+\vspace{2em}
+\textbf{File Read}
+\vspace{1em}
+\begin{equation} \label{eq:fileread}
+    \inferrule
+    {(\textbf{*}) \\ C \ni c \ni m \\ A[u,c]} 
+    {(C, F) \xrightarrow{u, read(m,f)} (C, F)}
+\end{equation}
+\vspace{1em}
+
+A user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
+can read a file \textit{f} in content \textit{m}.
+
+These predicates are used in several places and are therefore named \textit{*}.
+
+\vspace{2em}
+\textbf{File Creation}
+\vspace{1em}
+\begin{equation} \label{eq:filecreate}
+    \inferrule
+    {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
+    {(C \cup \{c'\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
+\end{equation}
+\vspace{1em}
+
+A user \textit{u} with access rights to a non-readonly container \textit{c} can create a content
+\textit{m}, containing a file \textit{f}, in container \textit{c}.
 
@@
 \textbf{File Update}
+\vspace{1em}
 \begin{equation} \label{eq:fileupdate}
     \inferrule
-    {f \in F \\ [fileId = f.id, parent = c \dots] \in M \\ A[u,c]} 
-    {F \cup \{f\} \xrightarrow{u, update(fileId, f'.id)} F \cup \{f, f'\}}
+    {\textbf{*} \\ c' = c \cup \{m'\} \\ m' = \{id = m.id, file = f'\} \\ \neg readOnly(c)} 
+    {(C \cup \{c \cup \{m\}\} , F) \xrightarrow{u, update(m, f')} (C \cup \{c \cup \{m'\}\}, F
+    \cup \{f'\}))}
 \end{equation}
@@
 
-If a user \textit{u} with access rights to a parent container \textit{c} of content \textit{m},
-updates a file reference \textit{f} in content \textit{m}, the original file stays intact.
-
-%\begin{equation} \label{eq:contentupdate}
-%    \begin{split}
-%        m' & = u, read(m) \land \\
-%        m'' & = u, change(m') \land \\
-%        m''' & = u, modify(m'', m) \Implies \\
-%        m''' & \in M \land m \not \in M
-%    \end{split}
-%\end{equation}
+If a user \textit{u} with access rights to a parent container \textit{c} (which is not read only) of
+content \textit{m}, updates a file \textit{f} in content \textit{m}, the original file stays
+intact as an effect of Copy-on-Write.
+
+\vspace{2em}
+\textbf{File Deletion}
+\newline
+
+Files can not be deleted by any user, however all the references to the file can
+be deleted, see Consequence \ref{eq:delete}.
 
@@
 \textbf{Content Update}
+\vspace{1em}
 \begin{equation} \label{eq:contentupdate}
     \inferrule
-    {m \in M \\ m \in c \\ A[u,c]}
-    {M \cup \{m\} \xrightarrow{u, update(m, m')} M \cup \{m'\}}
+    {\textbf{*} \\ \neg readOnly(c)}
+    {(C \cup \{c \cup \{m\}\}, F) \xrightarrow{u, update(m, m')} ((C \cup \{c \cup \{m'\}\}, F)}
 \end{equation}
@@
 
-%\begin{equation} \label{eq:containerupdate}
-%    \begin{split}
-%        c' & = u, read(c) \land \\
-%        c'' & = u, change(c') \land \\
-%        c''' & = u, modify(c'', c) \Implies \\
-%        c''' & \in C \land c \not \in C
-%    \end{split}
-%\end{equation}
 \vspace{2em}
 \textbf{Container Update}
+\vspace{1em}
 \begin{equation} \label{eq:containerupdate}
     \inferrule
-    {c \in C \\ A[u,c]}
-    {C \cup \{c\} \xrightarrow{u, update(c, c')} C \cup \{m'\}}
+    {\textbf{*} \\ \neg readOnly(c)}
+    {(C \cup \{c\}, F) \xrightarrow{u, update(c, c')} (C \cup \{c'\}, F)}
 \end{equation}
@@
 
-%\begin{equation} \label{eq:copy}
-%    \begin{split}
-%        & f' = f \in m \\
-%        & m' = u, copy(m) \Implies \\
-%        & m' \neq m \land \\
-%        & f' \in m' \land f' \in m
-%    \end{split}
-%\end{equation}
 \vspace{2em}
 \textbf{Copy}
+\vspace{1em}
 \begin{equation} \label{eq:copy}
     \inferrule
-    {m \in M \\ c \in C \\ m \in c \\ A[u,c]}
-    {M \cup \{m\} \xrightarrow{u, copy(m')} M \cup \{m, m'\} \land m.id \neq m'.id}
+    {\textbf{*} \\ m' = \{id = \text{"fresh"}, file = m.file\} \\ \neg readOnly(c)}
+    {(C \cup \{c \cup \{m\}\}, F) \xrightarrow{u, create(c, m')} (C \cup \{c \cup \{m, m'\}\}, F)}
 \end{equation}
@@
 of content \textit{m} copies that content, the new content \textit{m'} is not equal to the old
-(because of its new ID), however they both contain the same metadata and refer to the same file.
+(because of its new ID), however they both contain the same file.
 
@@
 \textbf{Content Deletion}
+\vspace{1em}
 \begin{equation} \label{eq:delete}
     \inferrule
-    {m \in M \\ f \in F \\ f \in m \\ c \in C \\ m \in c \\ A[u,c]}
-    {M \cup \{m\} \land F \cup \{f\} \xrightarrow{u, delete(m)} M \cup \{\} \land F \cup \{f\}}
+    {\textbf{*} \\ \neg readOnly(c)}
+    {(C \cup \{m\}, F) \xrightarrow{u, delete(m)} (C, F)}
 \end{equation}
@@
 \textbf{Snapshot}
+\vspace{1em}
 \begin{equation} \label{eq:snapshot}
@@
     \inferrule
-    {c \in C \\ A[u,c]}
-    {M, C \xrightarrow{[C', M'] := u, snapshot(c)} M \cup M',  C \cup C'}
+    {\textbf{*} \\ \neg readOnly(c)}
+    {(C, F) \xrightarrow{[C', M'] := u, snapshot(c)} (C \cup C' \cup M', F)}
 \end{equation}
@@
 but its content still refers to the same files.
-
-%\begin{equation} \label{eq:delete}
-%    \begin{split}
-%        f' = f \in (u, read(c)),  \\
-%        u, delete(c) \Implies \\
-%        c \not \in C \land f' \in F
-%    \end{split}
-%\end{equation}
 
@@
 
-\fix Picture of the newest front-end
 \begin{figure}[H] 
     \centering{
-        \includegraphics[scale=0.4]{frontend.png}
+        \includegraphics[scale=0.35]{frontend.png}
     }

### 2016-04-13 tex/paper.tex
Changed some effect to affect

@@
 feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
-it can effect the system's scalability.
+it can affect the system's scalability.
 
@@
 \section{Model} \label{sec:model}
-This section describes how different operations can be expected to effect the data in the system.
+This section describes how different operations can be expected to affect the data in the system.
 The model for this work also shows how the data can only be accessed or modified by
@@
 change a file though out the cycle of operation without creating a new file object with a new
-identifier. This is to ensure that files can not lose their integrity by simultaneous
+identifier. This is to ensure that files can not lose their integrity by concurrent
 operations from different nodes or get mixed up when referred to from a content object.
@@
 behaviour as the result of data modification is the main concern, which could be used for sabotage
-or simply be the effect unintentional unfortunate race conditions etc.
+or simply be the effect of unintentional race conditions etc.
 
@@
 To describe a content object \textit{m} containing fileId \textit{f}, the following syntax is used:
-$[fileId = f.id, \dots] \in M$, the same syntax is also applied to different types of objects
-throughout the explanation of the model.
+$\{id = \dots , file = f\} \in M$, the same syntax is also applied to different types of objects
+throughout the explanation of the model. The \"fresh\" keyword applied to id means that it is a new
+globally unique identifier.
 
@@
 
-These predicates are used in several places and are therefore named \textit{*}.
+The predicates in (\ref{eq:fileread}) are used in several places and are therefore named \textit{*}.
 
@@
 As no comparison with solutions that were not Copy-on-Write based were made, no scientific
-conclusions can be drawn from how its effect was on the scalability of the system. However some
+conclusions can be drawn from what its effect was on the scalability of the system. However some
 intuitively drawn conclusions can be suggested. If comparing to a system which uses a SQL database

### 2016-04-19 tex/paper.tex
Updated presentation with equation
Updated thesis title

@@
 %}}}
-\title{\textbf{Providing Access Control and Copy-on-Write for Content Distribution Network Assets}}
+\title{\textbf{Access Control for Content Distribution Network Assets}}
 \author{Klingsbo, Lukas}
@@
 \keywords{CDN, Copy-on-Write, Eventual Consistency, MongoDB, Persistent Storage, Scalability, 
-Software Security}
+Software Security, Access Control}
 \end{abstract}
@@
 feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
-it can affect the system's scalability.
+it can effect the system's scalability.
 
@@
 \section{Model} \label{sec:model}
-This section describes how different operations can be expected to affect the data in the system.
+This section describes how different operations can be expected to effect the data in the system.
 The model for this work also shows how the data can only be accessed or modified by
@@
     {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
-    {(C \cup \{c'\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
+    {(C \cup \{c\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
 \end{equation}

### 2016-04-19 tex/paper.tex
Added changes to runtime

@@
 consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
-the set of files, even though its modification process is the last one to write to the content.
+the set of files, even ough its modification process is the last one to write to the content.
 This is necessary as a file can be referenced from several content and the removal process can not
@@
 returns the document with the correct ID from the corresponding collection. As the collections are
-hash indexed on ID this operation can be made in constant time, O(1) and thus $\theta(1)$. \\
+hash indexed on ID this operation can be made in constant time, $O(1)$ and thus $\theta(1)$. \\
 
@@
 by a hash index which makes it possible to get all nodes with the same parent ID in constant time.
-This has to be done for every level with a different parent in the tree which makes the upper bound
-O(n) if all of the subtrees only have one node each. The tight bound can not be calculated as the
-trees balance is not determined by an algorithm but rather by a user, but from observation one could
-argue that the trees usually have several nodes for each parent node in the tree, which then would
-give a logarithmic tight bound, $\theta(log(n))$.\\
+This has to be done for every level with a different parent. The worst case for building the tree is
+thus $\theta(h)$ where h is the height of the tree and all of the subtrees only have one node each.
+The average case can not be calculated as the tree's balance is not determined by an algorithm but
+rather by a user, but from observation one could argue that the trees usually have several nodes for
+each parent node in the tree, which then would give a logarithmic average case, $\theta(log(n))$.\\
 
@@
 Creating a new entity in the system is a constant time operation for containers and content as they
-only have to be inserted as new documents in MongoDB, which is a O(1)
+only have to be inserted as new documents in MongoDB, which is a $O(1)$
 operation~\cite{MONGOPERFORMANCE}. Even though the insertion itself has constant run-time the index
@@
 accessible in constant time. When inserting files they need to be chunked into pieces and inserted
-into GridFS, this results in n/s insertions where n is the size of the file and s is the chunk size.
-The insertion of files can then be regarded as O(n) if the size of the file is regarded, but if n is
-regarded as the number of files in the system the upper bound is constant as the insertion for files
-is made in the same way as for other documents.\\
+into GridFS, this results in $c/s$ insertions where $c$ is the size of the file and s is the chunk size.
+The insertion of files can then be regarded as $\theta(c)$ if the size of the file is regarded, but
+if $c$ is regarded as the number of files in the system the upper bound is constant as the insertion
+for files is made in the same way as for other documents.\\
 
@@
 In the creation of a snapshot all metadata in the tree that the snapshot is taken of is duplicated,
-meanwhile the files remain the same, this results in n insertions where n is the number of
-containers and content entities in the subtree. A snapshot insertions runtime is therefore
-$\theta(n)$ and O(n). \\
+meanwhile the files remain the same, this results in $n$ insertions where $n$ is the number of
+containers and content entities in the subtree. A snapshots insertion runtime is therefore
+always $\theta(n)$. \\
 
@@
 the entity has children all those children has to be deleted too which will be done in the same
-manner as the project tree is built and therefore results in the same runtime, O(n) as an upper
-bound, where n is the number of entities in the subtree.\\
+manner as the project tree is built and therefore results in the same runtime, $\theta(n)$ as worst
+case and $\theta(log(n))$ as average (with the same argumentation as in \textit{GET project tree}),
+where $n$ is the number of entities in the subtree.\\
 
@@
 indices might have to be updated though, if for example the parent of the document is changed. 
-As both deletion and insertion are constant time operations, update is also O(1).\\
+As both deletion and insertion are constant time operations, update is also $O(1)$.\\
 

### 2016-05-11 tex/paper.tex
Changed according to reviewer comments

@@
 \section{Introduction}
-Developing large projects containing static content usually involves using a Content Distribution
-Network to be able to scale to a larger user base. The commercial Content Distribution Networks are
-usually fairly easy to use. The content that is to be used in a project is uploaded and then
-distributed over the globe when a client is trying to access the content. For secret content this
-can be a problem and an inconvenience, and that is what this thesis is about.
+The development of large projects using static content usually involves the use of a Content
+Distribution Network to be able to scale to a larger user base. Commercial Content Distribution
+Networks are usually fairly easy to use. The content that is to be used in a project is uploaded and
+then distributed over the globe when a client is trying to access the content. For secret content
+this can be a problem and an inconvenience due to the lack of tools available to handle groups of
+content and their security settings, and that is what this thesis is about.
 
@@
 
-As large projects most often are divided into different dividable parts, a concept called containers
-is introduced in this work. Containers can contain more containers and content, and when
-modifying the security settings for a container it is applied recursively to all its content and
-sub-containers.
-
-This work examines a way of enforcing virtual access control on content and groups of content, in
-the form of containers and snapshots to be able to handle the secrecy requirements easily. A system
-was developed using the Copy-on-Write principle to improve scalability and make it a lighter
-operation to take snapshots of parts of the system.
-
-The research question that this report tries to answer is how and whether it is practically 
-feasible to use Copy-on-Write for a high-level system like the one that is implemented and how 
-it can effect the system's scalability.
+As large projects most often are divided into parts, a concept called containers is introduced in
+this work. Containers can contain more containers and content, and when modifying the security
+settings for a container it is applied recursively to all its content and sub-containers.
+
+This work examines a way of enforcing access control on content and groups of content, in the form
+of containers and snapshots. By being able to control the secrecy settings on a container instead of
+separately modifying the settings for each individual content it becomes easier to release content
+to be accessible to the public or different groups during the progression of the project. To manage
+different versions or releases of the project, the snapshot functionality can be used to handle
+access control by taking a snapshot with the correct security settings at each release.
+
+A system was developed using the Copy-on-Write principle to improve scalability and make it a
+lighter operation to take snapshots of parts of the system.
+
+The research question that this report tries to answer is how and whether it is practically feasible
+to use Copy-on-Write for a high-level storage system and how it can affect the scalability of the
+system.
 
@@
 \section{Background}
-This section gives the reader some background of the company where the thesis was written and
-implementation was made. It also gives some brief explanations of the core concepts which are 
-exerted throughout this work.
+This section gives the reader some background of the company where the thesis was written and the
+implementation was made. It also gives some brief explanations of the core concepts which are used
+throughout this work.
 
@@
 Several filesystems~\cite{BTRFS}\cite{ZFS} use snapshot techniques to make it possible for a user
-or system to roll back the filesystems state, to a state that the filesystem had at the time of
+or system to roll back the system's state, to a state that the filesystem had at the time of
 snapshot creation. This can be useful for example when having a stable system and then doing a
@@
 
-This work has taken a lot inspiration from the way BTRFS~\cite{BTRFS} handles snapshots, which is
-further explained in Section~\ref{sec:btrfs}.
+This work has taken a lot inspiration from the way the BTRFS filesystem~\cite{BTRFS} handles
+snapshots, which is further explained in Section~\ref{sec:btrfs}.
 
@@
 \subsection{Related Terminology}
-In this section concepts and abbreviations, that are recurring throughout the paper, which the
+In this section concepts and abbreviations that are recurring throughout the paper and which the
 reader needs to be familiar with are explained.
@@
 \item[Eventual Consistency] \hfill \\
-Eventual consistency is an informal guarantee that defines how the consistency of stored data can
-behave. In a system defined as using eventual consistency, access to stored data is not guaranteed
-to return the result of the latest modification. At some point, if modifications to the stored data
-stop, access to a data item will return the resulting data of the last modification of that
-item.~\cite{EVENTUAL}. It is often used in the context of database systems.
+Eventual consistency~\cite{EVENTUAL} is an informal guarantee that defines how the consistency of
+stored data can behave. In a system defined as using eventual consistency, access to stored data is
+not guaranteed to return the result of the latest modification. At some point, if modifications to
+the stored data stop, access to a data item will return the resulting data of the last modification
+of that item. It is often used in the context of database systems.
 
@@
         \textbf{Set} & \textbf{Relation} & \textbf{Type} \\ \hline
-        \textbf{C} & $\textbf{C} \supseteq C \ni c$ & $\textbf{C}=P_{fin}(P_{fin}(C \cup M))$\\ \hline
-        \textbf{M} & $\textbf{M} \supseteq M \ni m$ & $\textbf{M}=P_{fin}(M)$\\ \hline
-        \textbf{F} & $\textbf{F} \supseteq F \ni f$ & $\textbf{F}=P_{fin}(F)$\\ \hline
+        \textbf{C} & $\textbf{C} \supseteq C \ni c$ & $\textbf{C}=P_\mathit{fin}(P_\mathit{fin}(C \cup M))$\\ \hline
+        \textbf{M} & $\textbf{M} \supseteq M \ni m$ & $\textbf{M}=P_\mathit{fin}(M)$\\ \hline
+        \textbf{F} & $\textbf{F} \supseteq F \ni f$ & $\textbf{F}=P_\mathit{fin}(F)$\\ \hline
     \end{tabular}
@@
         c & $C \cup M$ \\ \hline
-        m & $\{id: GUID, file: f\}$ \\ \hline
+        m & $\{\mathit{id}: \mathit{GUID}, \mathit{file}: f\}$ \\ \hline
         f & File \\ \hline
@@
     The access matrix contains what containers that all the users in the system have access to. If
-    A[u,c] is true in the access matrix, user \textit{u} has access to container \textit{c}.
+    $A[u,c]$ is true in the access matrix, user \textit{u} has access to container \textit{c}.
 
@@
 all of that containers children.
-The \textit{A[u,c]} notation is symbolising an access token. If \textit{A[u,c]} is true in the 
+The $A[u,c]$ notation is symbolising an access token. If $A[u,c]$ is true in the 
 access matrix \textit{A}, user \textit{u} has full access to the container object \textit{c} and its
@@
     \begin{split}
-        A[u,c] \Rightarrow & u \text{ can read } c \\
-        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can write to } c \\
-        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow & u \text{ can delete } c \\
-        A[u,c] \land f \in c \centernot\Rightarrow & u \text{ can delete } f 
+        A[u,c] \Rightarrow{}& u \text{ can read } c \\
+        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow{}& u \text{ can write to } c \\
+        A[u,c] \land \neg \text{ read-only(c) } \Rightarrow{}& u \text{ can delete } c \\
+        A[u,c] \land f \in c \centernot\Rightarrow{}& u \text{ can delete } f 
     \end{split}
@@
 The invariant for this model is the file. Which means that it should not be possible to
-change a file though out the cycle of operation without creating a new file object with a new
+change a file through out the cycle of operation without creating a new file object with a new
 identifier. This is to ensure that files can not lose their integrity by concurrent
@@
 \subsection{Integrity Threats}
-According to Biba et. al~\cite{BIBA} one can consider two threat sources, namely subsystem external
+According to Biba et al.~\cite{BIBA} one can consider two threat sources, namely subsystem external
 and subsystem internal. The external sources could be another system calling the subsystem with
@@
 also be somebody trying to tamper with the exposed functions of the program. Threats that are
-internal could be a malicious part of the subsystem or simply an incorrect part of the subsystem,
-which does not behave according to its specification.
+internal could be a malicious part of the subsystem or simply an incorrect part of the subsystem
+that does not behave according to its specification.
 
@@
 
-\subsection{Consequences of modification} \label{sec:conseq}
-
-In this section the consequences, side effects and direct effects, of different operations are
+\subsection{Consequences of Operations} \label{sec:conseq}
+
+In this section the preconditions and effects of different operations are
 stated in an informal mathematical manner, similar to how inference rules are composed. These rules
-are what should be expected to be followed by the implementation in
-Section~\ref{sec:implementation}.
+should be followed by the implementation in Section~\ref{sec:implementation}.
 
@@
 
+\vspace{1em}
+The following predicates are used in several places and are therefore named \textit{*} to avoid
+unnecessary repetition.
+\begin{equation} \label{eq:fileread}
+    \inferrule
+    {(\textbf{*}) \\ C \ni c \ni m \\ $A[u,c]$} 
+    {}
+\end{equation}
+\vspace{1em}
+
 \vspace{2em}
@@
     \inferrule
-    {(\textbf{*}) \\ C \ni c \ni m \\ A[u,c]} 
+    {(\textbf{*})} 
     {(C, F) \xrightarrow{u, read(m,f)} (C, F)}
@@
 A user \textit{u} with access rights to a parent container \textit{c} of content \textit{m}
-can read a file \textit{f} in content \textit{m}.
-
-The predicates in (\ref{eq:fileread}) are used in several places and are therefore named \textit{*}.
+can read a file~\(f\) in content \textit{m}.
 
@@
     \inferrule
-    {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
+    {C \ni c \\ $A[u,c]$ \\ m \centernot \in c \\ \neg readOnly(c)} 
     {(C \cup \{c\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
@@
 
-A user \textit{u} with access rights to a non-readonly container \textit{c} can create a content
-\textit{m}, containing a file \textit{f}, in container \textit{c}.
+A user \textit{u} with access rights to a non-readonly container \textit{c} can create a
+content~\(m\), containing a file \textit{f}, in container \textit{c}.
 
@@
 
-If a user \textit{u} with access rights to the parent (source and destination) container \textit{c}
+If a user \textit{u} with access rights to the parent (source and destination) container~\(c\)
 of content \textit{m} copies that content, the new content \textit{m'} is not equal to the old
-(because of its new ID), however they both contain the same file.
+(because of its fresh ID), however they both contain the same file.
 
@@
     \begin{split}
-        [C', M'] := snapshot(tree) & \Implies \\
-        \forall c \in tree, copy(c) \in & C' \\
-        \forall m \in tree, copy(m) \in & M' \\
-        \forall f \in tree, f \in & F
+        [C', M'] := snapshot(tree) {}& \Implies \\
+        \forall c \in tree, copy(c) \in {}& C' \\
+        \forall m \in tree, copy(m) \in {}& M' \\
+        \forall f \in tree, f \in {}& F
     \end{split}
@@
 
-To explain the non-trivial snapshot operation it is first defined in \ref{eq:snapshot}. The snapshot
-operation is performed upon a tree of objects and results in two new sets (C' and M'), one with
-containers and one with content, the referred files remain the same as before the snapshot.
+The non-trivial snapshot operation makes a deep-copy by traversing the full tree or sub-tree of
+containers and content that is to be taken a snapshot of. As the snapshot operation is performed
+upon a tree of objects it results in two new sets (C' and M') which are then added to C and M
+respectively, the referred files in the content remain the same as before the snapshot.
 
@@
 solved by locks or transactions, but as updates are not based on each other, the trade-off for this
-inconsistency to scalability and speed is intentional. However, as can also be seen in
+inconsistency in favour of scalability and speed is intentional. However, as can also be seen in
 consequence~\ref{eq:fileupdate}, if a file is removed from a content that file is not removed from
@@
 \par
-There is also the chance to have content and containers that are not referenced from anywhere and
-result in being garbage in the system. This could happen as a result of consequence~\ref{eq:delete}
-in combination with any of the other consequences listed in Section~\ref{sec:conseq}. The basic idea
-is that an entity is deleted meanwhile another is created or modified to have that entity as its
-parent, thus creating a separate tree that can not be reached from the root node of the project.
-These type of inconveniences should also be cleaned up by the garbage collector.
+There is also the chance to have content and containers that are not referenced from anywhere, these
+containers and content are defined as garbage. This could happen as a result of
+consequence~\ref{eq:delete} in combination with any of the other consequences listed in
+Section~\ref{sec:conseq}. The basic idea is that an entity is deleted while another is created or
+modified to have that entity as its parent, thus creating a separate tree that can not be reached
+from the root container. These entities should be removed, for example by a garbage collector, or
+made sure to be deleted in the instance when they are no longer referenced.
 
 \subsection{JPF}
-JPF was used to test the idea and state transitioning of the model. It is a very simplified
+JPF was used to test the idea and state transitions of the model. It is a very simplified
 version of the real system that still contains all the important Copy-on-Write concepts and the 
@@
 \subsubsection{Garbage Collection}
-For the system to reach a clean state and the decided eventual consistency~\cite{KLINGSBO}, garbage
-collection will be needed. As of the conclusion in Section~\ref{sec:multiple_access} from Section
-~\ref{sec:conseq} there will be objects in the system that are unreferenced and can not be reached
-and should therefore be cleaned up for them not to cause negative performance effects on the system.
+For the system to reach a clean state and eventual consistency~\cite{KLINGSBO}, garbage collection
+or reference counters will be needed. The conclusion of Section~\ref{sec:multiple_access} is that
+there will be objects in the system that are unreferenced and can not be reached and they should
+therefore be cleaned up to not cause negative performance impact on the system.
 
@@
 
+\subsection{Access Control} \label{sec:disc_access_control}
+Full object-level access control was not implemented according to the model described in
+Section~\ref{sec:model}. The implementation only checks whether a user should have access to the
+system as a whole, and does not set, nor check, specific access rights to contents or containers.
+
 \newpage
 \section{Summary}
-This section covers the conclusions that could be drawn from this work and how the implementation 
-could, and will be, extended in the future.
+This section covers the conclusions that I draw from this work and how I plan to extend the
+implementation in the future.
 
@@
 the application would not always be in a global consistent state as there would not be any global
-locks.  This could theoretically cause some inconvenience for the user but in all real world
-observations no users have noticed it. The theoretical inconvenience for the user was a trade-off
-made so that the application could scale on the width almost endlessly, especially if the issue with
-building the project trees is solved (as mentioned in Section~\ref{sec:load_testing}).
+locks. This could theoretically cause some inconvenience for the user where last write
+wins~\cite{LASTWRITEWINS} on meta data, but in all real world observations no users have noticed it.
+The theoretical inconvenience for the user was a trade-off made so that the application could scale
+almost endlessly, especially if the issue with building the project trees is solved (as mentioned in
+Section~\ref{sec:load_testing}).
 
@@
 \par The conclusion that can be drawn from implementing Copy-on-Write as high up in the stack as it
-was in Perius is that the code would have to be largely rewritten for every application that wants
-to use a similar approach. The solution to this could be to use Copy-on-Write in the persistent
-storage instead of in the actual application, which also efficiently could handle conflict 
-resolution and every improvement made to it could be shared for every system that uses the same 
-type of system architecture and underlying persistent storage. On the other hand, by having all the
-Copy-on-Write specific code in the back-end instead of the persistent storage the choices of
-storage becomes wider as the database does not have to implement the named things. This could be
-positive when deploying Perius in environments with high requirements for the number of users
-that the system should be able to handle. It would also be well suited if the level of consistency
-needs to be fully tweakable to make the trade off for how many users the application would work for
-and consistency of data.
+was in Perius is that the persistent storage logic would have to be largely rewritten for every
+existing application that is to use a similar approach. The solution to this could be to use
+Copy-on-Write in the persistent storage instead of in the actual application. This approach could
+efficiently handle conflict resolution and every improvement made to it could be shared for every
+system that uses the same type of system architecture and underlying persistent storage. On the
+other hand, by having all the Copy-on-Write specific code in the back-end instead of the persistent
+storage the choices of storage becomes wider as the database does not have to implement
+Copy-on-Write and conflict resolution. This could be an advantage when deploying Perius in
+environments with high requirements for the number of users that the system should be able to
+handle. It would also be well suited if the level of consistency needs to be fully tweakable to make
+it possible to trade off consistency for scalability.
 
@@
 \subsubsection{Access Control}
-Full access control was not implemented according to the model described in ~\ref{sec:model}, it
-was only implemented to check whether a user should have access to the system as a whole or not, the
-implementation did not set or check any specific access rights to certain contents or containers.
+As access control was not fully implemented (See Section~\ref{sec:disc_access_control}) as planned 
+in the model, the implementation will need to be extended in the future.
 
 \subsubsection{Front-end Refactorization}
-The application could be made substantially more efficient by rewriting the front-end to update
-itself according to the REST response after a modification of the project tree, instead of 
-re-fetching the full project tree every time a change is made. The front-end should also be fixed to
-properly follow the Reflux streamlining.
+Currently the implementation fetches the full project tree every time a change is made. This results
+in increased latency, bandwidth usage and server load. We propose a refactorization of the front-end
+where it updates the state according to the REST response of each modification to the project tree
+instead of re-fetching it.
 
@@
 \end{document}
-
-%{{{
-
-%What to write about
-%Battle Binary
-%Security features of CDNn's
-%What is needed by the logic, like what happens after two consecutive deletes
-%
-%What is needed
-%* Security layers
-%* snapshots
-%* Virtual file structure
-%* Versioning of content
-%* Multi project support
-%* Auth and audit logs
-%* Users
-
-%Stuff to write about:
-%Modular design, every piece should be interchangeable
-%LDAP - why it was used as standard AUTH
-
-%Using copy-on-write as the sole update strategy has pros and cons. The upside is
-%that it is simple to guarantee operation atomicity, and data-structure integrity. The
-%downside is that performance relies on the ability to maintain large extents of free
-%contiguous disk areas. In addition, random updates to a f i le tend to fragment it,
-%destroying sequentiality. A good defragmentation algorithm is required; this is
-%described in Section 5
-
-%}}}

### 2016-05-13 tex/paper.tex
Fixed faulty font on access rights
Added the meta variable for changing containers

@@
 $\{id = \dots , file = f\} \in M$, the same syntax is also applied to different types of objects
-throughout the explanation of the model. The \"fresh\" keyword applied to id means that it is a new
+throughout the explanation of the model. The ``fresh'' keyword applied to id means that it is a new
 globally unique identifier.
 
-If the user is irrelevant for the rule described, it is left out from the action, for example u,
-read(o) = read(o).
+To explain a change of a container the notation $C[c]$ is used when removing or adding a nested
+container from the set. $C[\cdot]$ is a \textit{C} that can contain the meta variable $\cdot$ and
+$C[c]$ is $C[\cdot]$ where $\cdot$ has been replaced by \textit{c}.
 
@@
 unnecessary repetition.
-\begin{equation} \label{eq:fileread}
+\begin{equation} \label{eq:predicate_rules}
     \inferrule
-    {(\textbf{*}) \\ C \ni c \ni m \\ $A[u,c]$} 
+    {(\textbf{*}) \\ C \ni c \ni m \\ A[u,c]} 
     {}
 \end{equation}
+
 \vspace{1em}
-
-\vspace{2em}
 \textbf{File Read}
@@
     \inferrule
-    {C \ni c \\ $A[u,c]$ \\ m \centernot \in c \\ \neg readOnly(c)} 
-    {(C \cup \{c\}, F) \xrightarrow{u, create(m,c)} (C \cup \{c \cup \{m\}\}, F \cup \{m.file\})}
+    {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
+    {(C[c], F) \xrightarrow{u, create(m,c)} (C[c \cup \{m\}], F \cup \{m.file\})}
 \end{equation}
@@
     {\textbf{*} \\ c' = c \cup \{m'\} \\ m' = \{id = m.id, file = f'\} \\ \neg readOnly(c)} 
-    {(C \cup \{c \cup \{m\}\} , F) \xrightarrow{u, update(m, f')} (C \cup \{c \cup \{m'\}\}, F
-    \cup \{f'\}))}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, f')} ((C[c \cup \{m'\}], F \cup \{f'\}))}
 \end{equation}
@@
     {\textbf{*} \\ \neg readOnly(c)}
-    {(C \cup \{c \cup \{m\}\}, F) \xrightarrow{u, update(m, m')} ((C \cup \{c \cup \{m'\}\}, F)}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, m')} ((C[c \cup \{m'\}], F)}
 \end{equation}
@@
 \vspace{2em}
+\newpage
 \textbf{Container Update}
@@
     {\textbf{*} \\ \neg readOnly(c)}
-    {(C \cup \{c\}, F) \xrightarrow{u, update(c, c')} (C \cup \{c'\}, F)}
+    {(C[c], F) \xrightarrow{u, update(c, c')} (C[c'], F)}
 \end{equation}
@@
     {\textbf{*} \\ m' = \{id = \text{"fresh"}, file = m.file\} \\ \neg readOnly(c)}
-    {(C \cup \{c \cup \{m\}\}, F) \xrightarrow{u, create(c, m')} (C \cup \{c \cup \{m, m'\}\}, F)}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, create(c, m')} (C[c \cup \{m, m'\}], F)}
 \end{equation}
@@
     {\textbf{*} \\ \neg readOnly(c)}
-    {(C \cup \{m\}, F) \xrightarrow{u, delete(m)} (C, F)}
+    {(C[m], F) \xrightarrow{u, delete(m)} (C[], F)}
 \end{equation}
@@
     {\textbf{*} \\ \neg readOnly(c)}
-    {(C, F) \xrightarrow{[C', M'] := u, snapshot(c)} (C \cup C' \cup M', F)}
+    {(C, F) \xrightarrow{[C', M'] := u, snapshot(c)} (C[C' \cup M'], F)}
 \end{equation}
@@
 
+\newpage
 \subsection{Findings}
@@
 \hfill \\
+\newpage
 \textbf{REST Endpoints} \label{sec:API} \hfill \\
@@
 \newpage
+
+\newgeometry{top=4cm}
 \section{Summary}
@@
 
+\restoregeometry
+
 \subsection{Future work}

### 2016-05-23 tex/paper.tex
Changed some model stuff

@@
         \textbf{Action} & \textbf{Semantics} \\ \hline
-        u, read(m, f)    & User \textit{u} reads \textit{f} from \textit{m}\\ \hline
+        u, read(o)       & User \textit{u} reads \textit{o}\\ \hline
         u, create(m, c)  & User \textit{u} creates \textit{m} in \textit{c}\\ \hline
@@
         u, delete(o)     & User \textit{u} deletes \textit{o}\\ \hline
+        u, copy(o)       & User \textit{u} creates a copy of \textit{o}\\ \hline
         u, snapshot(o)   & User \textit{u} takes a snapshot of \textit{o} \\ \hline
@@
 $C[c]$ is $C[\cdot]$ where $\cdot$ has been replaced by \textit{c}.
-
-\vspace{1em}
-The following predicates are used in several places and are therefore named \textit{*} to avoid
-unnecessary repetition.
-\begin{equation} \label{eq:predicate_rules}
-    \inferrule
-    {(\textbf{*}) \\ C \ni c \ni m \\ A[u,c]} 
-    {}
-\end{equation}
 
@@
     \inferrule
-    {(\textbf{*})} 
-    {(C, F) \xrightarrow{u, read(m,f)} (C, F)}
+    {A[u,c] \\ C[c \cup \{m\}]} 
+    {(C[c], F) \xrightarrow{u, read(f)} (C[c], F)}
 \end{equation}
@@
     \inferrule
-    {C \ni c \\ A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
+    {A[u,c] \\ m \centernot \in c \\ \neg readOnly(c)} 
     {(C[c], F) \xrightarrow{u, create(m,c)} (C[c \cup \{m\}], F \cup \{m.file\})}
@@
     \inferrule
-    {\textbf{*} \\ c' = c \cup \{m'\} \\ m' = \{id = m.id, file = f'\} \\ \neg readOnly(c)} 
+    {A[u,c] \\ m' = \{id = m.id, file = f'\} \\ \neg readOnly(c)} 
     {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, f')} ((C[c \cup \{m'\}], F \cup \{f'\}))}
@@
     \inferrule
-    {\textbf{*} \\ \neg readOnly(c)}
+    {A[u,c] \\ \neg readOnly(c)}
     {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, m')} ((C[c \cup \{m'\}], F)}
@@
     \inferrule
-    {\textbf{*} \\ \neg readOnly(c)}
+    {A[u,c] \\ \neg readOnly(c)}
     {(C[c], F) \xrightarrow{u, update(c, c')} (C[c'], F)}
@@
     \inferrule
-    {\textbf{*} \\ m' = \{id = \text{"fresh"}, file = m.file\} \\ \neg readOnly(c)}
-    {(C[c \cup \{m\}], F) \xrightarrow{u, create(c, m')} (C[c \cup \{m, m'\}], F)}
+    {A[u,c] \\ m' = \{id = \text{"fresh"}, file = m.file\} \\ \neg readOnly(c)}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, copy(m)} (C[c \cup \{m, m'\}], F)}
 \end{equation}
@@
     \inferrule
-    {\textbf{*} \\ \neg readOnly(c)}
-    {(C[m], F) \xrightarrow{u, delete(m)} (C[], F)}
+    {A[u,c] \\ \neg readOnly(c)}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, delete(m)} (C[c \cup \{\}], F)}
 \end{equation}
@@
     \begin{split}
-        [C', M'] := snapshot(tree) {}& \Implies \\
-        \forall c \in tree, copy(c) \in {}& C' \\
-        \forall m \in tree, copy(m) \in {}& M' \\
-        \forall f \in tree, f \in {}& F
+        &snapshot(c) = \{c'.id = \textit{"fresh"}, snapshot(c') \mid c' \in c \} \cup \{ snapshot(m) \mid m \in c \}\\
+        &snapshot({id = i, file = f }) = \{ id = \text{"fresh"}, file = f \} 
     \end{split}
@@
 
-The non-trivial snapshot operation makes a deep-copy by traversing the full tree or sub-tree of
+The snapshot operation makes a deep-copy by traversing the full tree or sub-tree of
 containers and content that is to be taken a snapshot of. As the snapshot operation is performed
@@
     \inferrule
-    {\textbf{*} \\ \neg readOnly(c)}
-    {(C, F) \xrightarrow{[C', M'] := u, snapshot(c)} (C[C' \cup M'], F)}
+    {A[u,c] \\ \neg readOnly(c)}
+    {(C[c], F) \xrightarrow{u, copy(c)} (C[c \cup snapshot(c)], F)}
 \end{equation}

### 2016-05-25 tex/paper.tex
Small reviewer requested changes

@@
     \inferrule
-    {A[u,c] \\ C[c \cup \{m\}]} 
-    {(C[c], F) \xrightarrow{u, read(f)} (C[c], F)}
+    {A[u,c]]} 
+    {(C[c \cup \{m\}], F) \xrightarrow{u, read(f)} (C[c \cup \{m\}], F)}
 \end{equation}
@@
     {A[u,c] \\ m' = \{id = m.id, file = f'\} \\ \neg readOnly(c)} 
-    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, f')} ((C[c \cup \{m'\}], F \cup \{f'\}))}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, f')} (C[c \cup \{m'\}], F \cup \{f'\})}
 \end{equation}
@@
     {A[u,c] \\ \neg readOnly(c)}
-    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, m')} ((C[c \cup \{m'\}], F)}
+    {(C[c \cup \{m\}], F) \xrightarrow{u, update(m, m')} (C[c \cup \{m'\}], F)}
 \end{equation}
@@
     \begin{split}
-        &snapshot(c) = \{c'.id = \textit{"fresh"}, snapshot(c') \mid c' \in c \} \cup \{ snapshot(m) \mid m \in c \}\\
+        &snapshot(c) = \{snapshot(c') \mid c' \in c \} \cup \{ snapshot(m) \mid m \in c \}\\
         &snapshot({id = i, file = f }) = \{ id = \text{"fresh"}, file = f \} 
@@
     {A[u,c] \\ \neg readOnly(c)}
-    {(C[c], F) \xrightarrow{u, copy(c)} (C[c \cup snapshot(c)], F)}
+    {(C[c], F) \xrightarrow{u, copy(c)} (C[c, snapshot(c)], F)}
 \end{equation}

### 2016-06-01 tex/paper.tex
Changed implementation background and abstract

@@
 \begin{abstract}
-This work presents a technique for implementing Copy-on-Write on top of an unaware persistent
-storage. The system in this case is a security management system, but the technique and conclusions
-drawn are general enough to be applicable to any system with vaguely similar consistency
-requirements. The scalability of the storage system is tested and the technique is evaluated. The
-technique is shown to be quite efficient (scalable to >6.000 simultaneous active users per node),
-but a better solution is also suggested in the conclusion.\\
+This work presents a technique for implementing Copy-on-Write on top of a persistent storage. The
+method was developed from the needs of a security management system, but the technique and
+conclusions drawn are general enough to be applicable to any system with vaguely similar consistency
+requirements. The scalability of the security management system is tested and the technique is
+evaluated with load testing and model checking. The implementation is shown to be scalable for up to
+>6.000 simultaneous active users per node, but based on our experience a better solution is 
+suggested in the conclusion.\\
 
@@
 \subsubsection{The current system} \label{sec:current_system}
-Today a system called Battlebinary~\cite{BATTLEBINARY} is used for managing and uploading files,
-mostly images, to content delivery networks. The current system does not make use of the
+Today a system called Battlebinary~\cite{BATTLEBINARY} is used by Uprise for managing and uploading
+files, mostly images, to content delivery networks. The current system does not make use of the
 security features that the CDNs are offering, instead it uses a form of security by obscurity. When
-a file is uploaded to a CDN it is open for the public, but its filename is composed out of its
-original filename concatenated with a part of the MD5 hash of the content of the file, which makes
-it an extremely hard process to access the file on the CDN without access to the original file or a
-reference to the URI.
-
-In the current system you can only upload a file once as there will be a collision in the upload
-otherwise, as the old and the new file will have the same MD5 hash, which is by design as duplicate
+a file is uploaded to a CDN it is available to the public, but since its identifier is composed out
+of its original filename concatenated with a part of the MD5 hash of the content of the file. This
+makes it a very computationally heavy process to access the file on the CDN without access to the
+original file or a reference to the URI, as the identifier would have to be brute forced.
+
+In the current system you can only upload a file once, since there will be an identifier collision.
+This is due to the old and the new file having the same MD5 hash, which is by design as duplicate
 content just wastes space on the CDN. The problem with this is that the current interface does not
-handle it very well as one file can not be shown or uploaded in two different places in 
+handle it very well, since one file can not be shown or uploaded in two different places in
 Battlebinary's virtual filesystem.

### 2016-06-13 tex/paper.tex
Added half of the implementation fixed

@@
 \subsubsection{Problem description}
-The current system does not offer proper security measurements, is lacking a lot of features that
-are needed and does not scale very well, hence a new system are to be developed. This work is about
-examining a way of implementing Copy-on-Write in a high-level system like this, which should solve
-the scalability problem and make it possible to implement wanted features like snapshots, cloning
-and concurrent modifications of content.
+The current system does not offer proper security measurements as its mechanism for keeping private
+content out of the hands of the public is based on security by obscurity, which is further explained
+in Section~\ref{sec:current_system}.
+
+It is also lacking several features that are needed and does not scale very well, hence a new system
+is to be developed. Features that needs to be implemented are for example snapshots, cloning and
+proper support of concurrent modifications of content.
+
+This work is about examining a way of implementing Copy-on-Write in a high-level system, which
+should solve the scalability problem and make it possible to implement wanted features.
 
@@
 To clarify what is meant by Perius being high-level software the stack in Figure~\ref{fig:stack} can
-be examined. Theoretically the Copy-on-Write features could be implemented in any of the separate
-parts of the stack to support the features that Perius has to achieve, in this work only the second
-highest layer is considered, the Perius back-end.
+be examined. Theoretically the Copy-on-Write features could be implemented in any part of the stack,
+in this work only the second highest layer is considered, the Perius back-end.
 
@@
 lambda expressions. React and Reflux are both built with JavaScript in mind and is used on top of
-JavaScript is this implementation, as can be seen in Figure~\ref{fig:tech_stack}.
+JavaScript in this implementation.
 
@@
 is a small HTTP request library which can be used from client-side Javascript or as a node.js
-module~\cite{SUPERAGENT}. It is used to form the requests to the REST API from the front-end, which
-can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
+module~\cite{SUPERAGENT}. It is used to form the requests to the REST API from the front-end, arrow
+1 and 2 in Figure~\ref{fig:tech_stack}.
 
@@
 Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used together with
-React for this implementation to define a structure of the flow of data and state changes.
+React to define a structure of the flow of data and state changes.
 
@@
 supports most functional programming features at the same time as it supports object-oriented
-programming~\cite{SCALA}. Scala is used to program the back-end and the Spray toolkit is used with
-it to simplify the API.
+programming~\cite{SCALA}. Scala is used to program the back-end and the Spray toolkit is used to
+simplify the API.
 
@@
 In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
-communication, which can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).
+communication, arrow 1 and 2 in Figure~\ref{fig:tech_stack}.
 
@@
 As can be seen several expected endpoints are missing, this is intentional as the operations missing
-can be performed in a more efficient way. Such endpoint is for example \textit{GET 
-projects/\{id\}/containers} as all containers exist in \textit{GET projects/\{id\}} and the interface
+can be performed in a more efficient way. Such endpoint is for example GET 
+projects/\{id\}/containers as all containers exist in GET projects/\{id\} and the interface
 should present a file structure where both content and containers are shown.
@@
 \item[Spray]~\cite{CASBAH} % Spray in casbah book
-is an open-source toolkit written in Scala and for Scala. It is used to make it easier to build
-REST/HTTP APIs and those abilities is what it is used in the implementation.
+is an open-source toolkit written in Scala and for Scala. Spray is mainly used to make it easier to
+build REST/HTTP APIs.
+
+\item[Casbah]~\cite{CASBAH} % Spray in casbah book
+is a Scala toolkit and a java driver wrapper used to interface with MongoDB instances in a more
+Scala oriented way.
 
 \item[MongoDB]% \hfill \\ %\label{sec:spray}
-is a document-oriented database, which means that it does not have the concept of rows as
-normal relational databases has. Instead each entity in the database is stored as a document which
-is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
-improve its possibility to scale, which can be a big downside to some applications containing the
-need for such logic. MongoDB is used as a persistent storage for the back-end to store all the
-projects and nodes in the tree structures and also the files uploaded to the system. The Scala logic
-code uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB, which can be seen in
-Figure~\ref{fig:tech_stack} (Arrow 4 and 5).
+is a document-oriented database, which means that it does not have the concept of rows as normal
+relational databases has. Instead each entity in the database is stored as a document that is not
+fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks support for join operations to
+improve its possibility to scale, which can be a big disadvantage for applications having the need
+for such logic. MongoDB is used as a persistent storage for the back-end to store all the projects
+and nodes in the tree structures and also the files uploaded to the system. The Scala logic code
+uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB, arrow 4 and 5 in
+Figure~\ref{fig:tech_stack}.
 
@@
 core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
-scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
-migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
+scale in the future then it is very easy to distribute MongoDB and, if needed, the system can easily
+be migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
 and can therefore make the system scale even further~\cite{REACTIVEMONGO}.
@@
 \par All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files
-according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
-that backups of the Perius system is easily done through a database backup, no separate files 
-needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
-ranges of a file, although that advantage is not needed in the Perius implementation. For
-scalability this could be used to retrieve different parts of a file from different servers, 
-normal load balancing would probably work better in a system like Perius where no extremely large 
-files are expected to be stored.
+according to the size limit of MongoDB objects, which is currently 4MB. The advantage of storing
+files directly in the database rather than in the filesystem is that backups of can be done through
+a database backup which will contain both the meta-data stored in MongoDB and the files which
+results in that no separate files needs to be backed up.  Another advantage is that you can retrieve
+specific ranges of a file, although that feature is not needed in the Perius implementation. For
+scalability this could be used to retrieve different parts of a file from different servers, normal
+load balancing would probably work better in a system like Perius where no extremely large files are
+expected to be stored.
 
@@
 effects of using Copy-on-Write, the decisions taken for the different technologies used were simply
-made from what the author was comfortable with and knew would have little impact on the evaluation. 
-With that said there are no comparisons of for example different databases, protocols or API
-solutions in this paper.
+made from what the author was comfortable with and expected would have little impact on the
+evaluation.  With that said there are no comparisons of for example different databases, protocols
+or API solutions in this paper.
 
@@
 
-\par A low timeout was set to make sure that the system could handle operations faster than they
+\par A low timeout threshold was used to make sure that the system could handle operations faster than they
 queued up. This was to be certain that the system could sustain the number of active users during a
-longer time frame without having a queue which grows beyond a manageable size.
+longer time frame without having a queue that could grow beyond a manageable size.
 
@@
 \subsubsection{Snapshot Functionality}
-The Copy-on-Write concept is not only well-suited for consistency requirements, it is also works 
-well for creating snapshots.
+The Copy-on-Write concept is not only well-suited for consistency requirements, its basis also gives
+the possibility of creating snapshots.
 
@@
 reference counter is not stored in the nodes themselves but rather in a separate data structure so
-that a node's counter can be modified without modifying the node itself and therefore eludes the
+that a node's counter can be modified without modifying the node itself and therefore avoids the
 Copy-on-Write that would have to occur.
@@
 physically copying memory was too slow. To minimise the copying of memory, Copy-on-Write was
-implemented. It was implemented so that virtual copy operations could be done and so that tasks
-could share read-write memory~\cite{MACH}.
+implemented to enable the use of virtual copy operations and to create the possibility of tasks
+sharing read-write memory~\cite{MACH}.
 
 \subsubsection{Perius}
-As the persistent storage, used in this implementation (Section~\ref{sec:tech}), does
-not implement transactions or locks a lot of different problems can occur when several clients are
-working on the same data set at the same time. Such problems could be race conditions or determining
-the happened-before relation~\cite{LAMPORT}. In this work the race condition problem is solved by 
-implementing Copy-on-Write on data with high consistency requirements. The happened-before relation
-is not fully solved, but suggestions of how to solve it is given in Section~\ref{sec:conflict_res}.
+As the persistent storage used in this implementation (Section~\ref{sec:tech}) does not implement
+transactions or locks, several different problems can occur when multiple clients are working on the
+same data set at the same time. Such problems could be race conditions or determining the
+happened-before relation~\cite{LAMPORT}. Race conditions are problematic due to that when several
+users or nodes modify data they can start in one state and expect the end state of the data to be
+the same as if they would have been alone modifying the data, but instead if another user modifies
+the data concurrently the changes that the first user made can be overwritten. In this work the
+race condition problem is solved by implementing Copy-on-Write on data with high consistency
+requirements. The happened-before relation is a problem of ordering events, if it is possible to
+know in which order concurrent events happened it is easier to know what the outcome of the events
+should be. The happened-before relation is not fully solved in Perius, but suggestions of how to
+solve it is given in Section~\ref{sec:conflict_res}.
 
@@
 back-end written in Scala and a front-end written in Javascript (ES6), but they are both
-interchangeable. The back-end has a REST API running, which is how the front-end communicates with
-the back-end.
+independently replaceable. The back-end has a REST API running, which is how the front-end
+communicates with the back-end.
 
@@
 as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
-whole containers as well as individual files, audit and auth logging, multiple separate project 
-and a design which allows for an interchangeable persistent storage.
+whole containers as well as individual files, audit and auth logging, multiple separate projects 
+and a design which allows for a replaceable persistent storage.
 
@@
 This section explains the different types of entities, and their interaction, that are handled by
-Perius. They are explained similarly to how they are explained for the model in
-Section~\ref{sec:elements}, but with more implementation specific details.
+Perius. They are explained similarly to the model in Section~\ref{sec:elements}, but with more
+implementation-specific details.
 
@@
 all projects are considered. For a very small project that tree could look like the example in
-Figure~\ref{fig:project_tree}. This is however not what the actual storage of the data looks like.
+Figure~\ref{fig:project_tree}.
 
@@
 
-The actual storage of the data is represented by documents in a flat document storage, which is in
-this case MongoDB. A very simplified version of how the data is stored can be seen in
-Figure~\ref{fig:storage}, however in reality it is not quite so table-like. One important point that
-differs in the example and in the actual storage is that GUIDs are used for identifiers and not
-incremented numbers, as that would not work in distributed environments.
-
-The data is not stored as a tree (see Example~\ref{fig:storage}), but rather as separate documents 
-loosely referring to each other. How the different operations perform their actions and what 
-implications these actions have on the operations run-time complexities are not obvious and that 
-is what is discussed in this section.\\
+A very simplified version of how the data is stored can be seen in Figure~\ref{fig:storage}, however
+in reality it is not quite so table-like. The implementation is using GUIDs, instead of incremental
+numbers as shown in the figure, to work correctly in a distributed environments.
+
+As seen above, the data is not stored as a tree (see Example~\ref{fig:storage}), but rather as
+separate documents loosely referring to each other. How the different operations perform their
+actions and what implications these actions have on the run-time complexities are not obvious and
+that is what is discussed in this section.\\
+
+\begin{figure}[H] 
+    \centering
+    \renewcommand{\arraystretch}{1.3}
+    \begin{tabular}{|r|c|l|}
+        \hline
+        \textbf{Variable} & \textbf{Explanation} \\
+        \hline
+        \textit{n} & All content+containers \\
+        \textit{c} & All containers \\
+        \textit{m} & All content \\
+        \textit{f} & All files \\
+        \textit{s} & Size of a file \\
+        \textit{k} & Chunk size \\
+        \hline
+    \end{tabular}
+    \renewcommand{\arraystretch}{1.0}
+    \caption{Variables used in run-time complexities}
+    \label{fig:runtime-variables}
+\end{figure}
 
@@
 returns the document with the correct ID from the corresponding collection. As the collections are
-hash indexed on ID this operation can be made in constant time, $O(1)$ and thus $\theta(1)$. \\
+hash indexed on ID this operation can be made in constant time, $O(1)$ and thus $\Theta(1)$. \\
 
 \par \textbf{GET project tree} \\
-When requesting a project the full project tree with all entities except files has to be built as
+When requesting a project the full project tree with all entities except files has to be built since
 the entities not are stored as a tree in MongoDB. The tree is built layer by layer by fetching
@@
 by a hash index which makes it possible to get all nodes with the same parent ID in constant time.
-This has to be done for every level with a different parent. The worst case for building the tree is
-thus $\theta(h)$ where h is the height of the tree and all of the subtrees only have one node each.
-The average case can not be calculated as the tree's balance is not determined by an algorithm but
-rather by a user, but from observation one could argue that the trees usually have several nodes for
-each parent node in the tree, which then would give a logarithmic average case, $\theta(log(n))$.\\
+This has to be done for every level with each parent. 
+Since all of the containers have to be processed when building the tree, the runtime complexity will
+thus be $\Theta(c)$/
 
@@
 accessible in constant time. When inserting files they need to be chunked into pieces and inserted
-into GridFS, this results in $c/s$ insertions where $c$ is the size of the file and s is the chunk size.
-The insertion of files can then be regarded as $\theta(c)$ if the size of the file is regarded, but
-if $c$ is regarded as the number of files in the system the upper bound is constant as the insertion
+into GridFS, this results in $s/k$ insertions.
+The insertion of files can then be regarded as $\Theta(s)$ if the size of the file is regarded, but
+if $k$ is larger than the size of the files inserted, the upper bound is constant as the insertion
 for files is made in the same way as for other documents.\\
@@
 In the creation of a snapshot all metadata in the tree that the snapshot is taken of is duplicated,
-meanwhile the files remain the same, this results in $n$ insertions where $n$ is the number of
+meanwhile the files remain the same, this results in $t$ insertions where $t$ is the number of
 containers and content entities in the subtree. A snapshots insertion runtime is therefore
-always $\theta(n)$. \\
+$\Theta(n)$. \\
 
@@
 the entity has children all those children has to be deleted too which will be done in the same
-manner as the project tree is built and therefore results in the same runtime, $\theta(n)$ as worst
-case and $\theta(log(n))$ as average (with the same argumentation as in \textit{GET project tree}),
-where $n$ is the number of entities in the subtree.\\
+manner as the project tree is built and therefore results in the same runtime, $\Theta(n)$ for
+containers and $\Theta(1)$ for content.
 
@@
 To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
-requests} are performed. Most responses are JSON structures, except for the ones in
-Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}, which can be seen in the Response data part
+requests} are performed. Most responses are JSON structures (except for the ones in
+Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}) which can be seen in the Response data part
 of the listings.\\
@@
 (HTTP server) can handle with the given specifications and therefore the server just returns 200
-(OK).
+(OK). 100 000 requests per second is well over what we need to serve and with a 1.22ms latency we
+know that without practically any payload the system is responding fast.
 
@@
 computation in the back-end, apart from converting the MongoDB format to the API's JSON result
-format.
+format. From these requests it can clearly be deduced that the round trip time to fetch a document
+in the database is adding some latency, from the average of 1.22ms to 14.52ms with the database
+request.
 
@@
 MongoDB. This was done to be able to estimate what impact the round trip time to MongoDB and the
-conversions to the response format had on the throughput.
+conversions to the response format had on the throughput. This test still gave an average latency
+that was larger than for the smaller payload in Listing~\ref{OKREQUEST}, but still half of the
+latency compared to the request involving a MongoDB call.
 
@@
 The result in Listing~\ref{TREEREQUEST} shows an average of 40 requests/s, which means this is the
-real bottleneck in the application. There are two solutions to fix this bottleneck, as the project
+real bottleneck in the application. There are two solutions to fix this bottleneck. Since the project
 tree will not be modified nearly as often as it is requested, the first solution would be to cache
 the results of the tree requests and invalidate those caches when the tree is modified. The second
-solution would be to make a more computationally feasible way of building and fetching the whole
-tree from the database. This is left to implement if the need comes for it in the future. Even
+solution would be to make a more computationally feasible algorithm for building and fetching the whole
+tree from the database. This is will not be discussed further, but left as future work. Even
 though this massive bottleneck is present, it will not be a problem in a small scale production
-environment as the tree is fetched approximately once every 30 seconds by active users, which means
+environment where the tree is fetched approximately once every 30 seconds by active users, which means
 that the application still could support at least 1200 very active users.
@@
 With caching turned on, on a single back-end, the application can theoretically support several
-million active users which are not posting any content, this is simulated in 
+million active users that are not posting any content, this is simulated in
 Listing~\ref{CACHEREQUEST}.
@@
 
-\par As a project tree is built by querying documents parent ID an optimisation that could be done was
-to create an index of the documents parent IDs. This drastically improved the size of the tree that
-the server was able to handle within the time-out, going from being able to handle 4000 documents in 
-one tree to at least 50000 documents, which the browser instead will have problems displaying, 
-represented in the UI, without lag.
+\par As a project tree is built by querying documents parent ID an optimisation that could be done
+is to create an index of the documents parent IDs. This drastically improved the size of the tree
+that the server was able to handle within the time-out, going from being able to handle 4000
+documents in one tree to at least 50000 documents, which the browser instead will have problems
+displaying, represented in the UI, without lag.
 
@@
 approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
-maximum number of active clients that one node can handle will be >60000, without taking the web
-servers limits or content posting into consideration.
+maximum number of active clients that one node can handle will be >60000 clients at the same time, 
+without taking the web server limits or content posting into consideration.
 

### 2016-06-14 tex/paper.tex
Added the last changes according to the comments

@@
 a database backup which will contain both the meta-data stored in MongoDB and the files which
-results in that no separate files needs to be backed up.  Another advantage is that you can retrieve
-specific ranges of a file, although that feature is not needed in the Perius implementation. For
-scalability this could be used to retrieve different parts of a file from different servers, normal
-load balancing would probably work better in a system like Perius where no extremely large files are
-expected to be stored.
+results in that no separate files needs to be backed up.
 
@@
 As can be seen in Listing~\ref{CONTAINERPOSTREQUEST} the server can handle around 4000 container
-POST requests per second. The problem that this test resulted in was not the amount of containers
-that could be posted, but rather loading all those containers in the interface or API afterwards.
-When requesting the project that the containers were posted to the web server timed out the request,
-as it took too long for the back-end to build the project tree. In this work, the web server will
-not  
-be configured to accept longer data processing time, as trees as large as this for a single project
-will not be used in production.
+POST requests per second. A problem that was discovered during this test was not the amount of
+containers that could be posted, but rather loading all those containers in the interface or API
+afterwards. When requesting the project tree from the web server it timed out the request
+since it took too long for the back-end to build the tree. In this work the web server will not be
+configured to accept longer data processing time, due to the low probability of trees as large as
+this being used in production.
 
@@
 \subsubsection{Authentication}
-The authentication of users in the system is currently being handled by LDAP~\cite{LDAP}, as Perius 
-is mainly focussed at being deployed in internal networks which usually has an LDAP service enabled. 
-LDAP also makes it easier for the user to login as no separate account is needed for Perius and 
-thus the user can use the same account as for all the LDAP connected services on the internal 
-network.
+The authentication of users in the system is currently being handled by LDAP~\cite{LDAP}, since
+Perius mainly will be deployed in internal networks with an existing LDAP infrastructure.  LDAP also
+makes it easier for the user to login as no separate account is needed for Perius and thus the user
+can use the same account as for all the LDAP connected services on the internal network.
 
 \subsubsection{Audit logs}
-Audit logs are used in the system to keep track of which users that perform which actions. In the
-event that somebody, consciously or unconsciously, are leaking private CDN content from within the
-organisation using Perius. The audit trail can then be used to find which LDAP account that was
-responsible for the leak.
+Audit logs are used in the system to keep track of which users have perform which actions. In the
+event that somebody, consciously or unconsciously, are leaking private CDN content the audit trail
+can then be used to find which LDAP account that was responsible for the leak.
 
@@
 containers in Perius, which makes it possible to handle the security settings for all files in a
-group at once, without having to manually traverse through the tree in Perius.
+group at once, without having to manually traverse the tree in Perius.
 
@@
 IP addresses that can access the asset, this option can be fitting when for example an office or
-companies internet access is based on a set of public static IP addresses and thus restricts the
-content from ever being accessed outside of that controlled network.
+a restricted groups internet access is based on a set of public static IP addresses.
 
@@
 The greatest security improvement of Perius compared to its predecessor Battlebinary is not the
-security of the system itself but rather how Perius handles the contents security settings on the
+security of the system itself but rather how Perius handles the content security settings on the
 CDN providers. Perius uses several of the security features for private content that the CDN
@@
 When using the ReactiveMongo driver~\cite{REACTIVEMONGO}, which is asynchronous and non-blocking,
-the application has no limits of how much load and users it can handle as the hardware and nodes 
-can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used with the current 
-implementation, it is harder to scale to the enormous amounts of load which ReactiveMongo can 
-support as Casbah is synchronous and has blocking IO. For this work the kind of scalability which 
-is offered by ReactiveMongo is not needed as the load will not reach the peak, as discussed in 
+the application has no limits of how much load or how many users it can handle as the hardware and
+nodes can be scaled up linearly when needed. With Cashbah~\cite{CASBAH}, which is used in the
+current implementation, it is harder to scale to the enormous amounts of load that ReactiveMongo
+can support as Casbah is synchronous and has blocking IO. For this work the kind of scalability
+which is offered by ReactiveMongo is not needed as the load will not reach the peak, as discussed in 
 Section~\ref{sec:GET_REQUESTS}, for what Casbah can handle on a single server.
@@
 Listing~\ref{DBREQUEST} shows that to fetch all documents in a collection (in this case only one) is
-dramatically slower, the throughput was only \textasciitilde7000 requests/s.
+dramatically slower, the throughput was only $\approx$7000 requests/s.
 
@@
 \subsection{Persistent Storage}
-More research could have been done in the choosing of persistent storage, as mentioned in 
-\textit{``NoSQL: Moving from MapReduce Batch Jobs to Event-Driven Data Collection''}
-~\cite{KLINGSBO} many applications that choose NoSQL databases as their persistent storage 
-actually do not need it. This is most likely true for Perius too, it would have been efficient 
-enough generating the project trees from an indexed SQL database with foreign keys. On the 
-other hand it was quite nice having the BSON documents for insertion as they were so similar 
-to the accepted JSON format from the REST service. As Perius most common operations involve 
-modifying project trees, a graph database should have been researched too.
+More research could have been done in the choosing of persistent storage, as mentioned in many
+applications that choose NoSQL databases as their persistent storage actually do not need
+it~\cite{KLINGSBO}. This is most likely true for Perius too: it would have been efficient enough
+generating the project trees from an indexed SQL database with foreign keys. On the other hand it
+was quite nice having the BSON documents for insertion as they were so similar to the accepted JSON
+format from the REST service. As Perius most common operations involve modifying project trees, a
+graph database should have been researched too.
 
 \subsection{Copy-on-Write's effect on Scalability}
-As no comparison with solutions that were not Copy-on-Write based were made, no scientific
+Since no comparison with solutions that were not Copy-on-Write based were made, no scientific
 conclusions can be drawn from what its effect was on the scalability of the system. However some
 intuitively drawn conclusions can be suggested. If comparing to a system which uses a SQL database
-the Copy-on-write system would have an easier time to scale on the width as no global locks or
-transactions has to be used, which with high probability would slow down the system, especially when
-widely distributed.
+the Copy-on-write system would have an easier time to scale since no global locks or transactions
+has to be used, which with high probability would slow down the system, especially when widely
+distributed.
 The gain of having the locks or transactions would be that the system would
@@
 
+\subsection{Scalability}
+Since GridFS was used to store the files an unexplored option for making it easier to scale is to
+use GridFS feature of retrieving specific ranges of a file. For scalability this could be used to
+retrieve different parts of a file from different servers, but normal load balancing would probably
+work better in a system like Perius where no extremely large files are expected to be stored.
+
 \subsection{Branching}
-Branching was also implemented in Perius, but to not make the report too wide it was not included.
+Branching was also implemented in Perius, but to not make the report too broad it was not included.
 The branching worked by the user selecting a container that the branch should be based on and a
-container in which the branch should be places. The user could then choose which containers and
-content that it wanted to update with the origin source of and which objects that should be deep
+container in which the branch should be placed. The user could then choose which containers and
+content that it wanted to update with the origin and which objects that should be deep
 copied to the new branch.
@@
 where it updates the state according to the REST response of each modification to the project tree
-instead of re-fetching it.
+instead of fetching the whole tree again.
 
